---
title: "Mate-Choice Copying: A Systematic Meta-Analysis"
subtitle: "Data decisions, processing pipeline, and planned analyses"
author: "Eduardo Santos & Collaborators"
date: today
date-format: "MMMM YYYY"
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    toc-location: left
    toc-title: "Contents"
    number-sections: true
    code-fold: true
    code-summary: "Show R code"
    code-tools: true
    self-contained: true
    smooth-scroll: true
    highlight-style: github
    embed-resources: true
execute:
  echo: true
  eval: true   # set to true when running locally with data
  warning: false
  message: false
bibliography: references.bib
---

```{r}
#| label: setup
#| include: false
library(dplyr)
library(data.table)
library(stringr)
library(readr)
library(knitr)
library(kableExtra)
library(readxl)
library(metafor)
library(orchaRd)
library(patchwork)
library(DT)
library(scales)
library(ggplot2)
```

---

## Overview {#sec-overview}

This page documents all data decisions, processing steps, and planned analyses for a
systematic **meta-analysis of mate-choice copying (MCC)** across animal taxa.

Mate-choice copying occurs when an individual's preference for a potential mate is
influenced by observing the mate choices of others — a form of socially acquired
preference that has been documented in fish, birds, mammals, and invertebrates.

This meta-analysis aims to:

1. Estimate the overall effect size for MCC across species.
2. Quantify between-study heterogeneity.
3. Test for moderating variables (sex, taxon, experimental design).
4. Explicitly model **lab-level non-independence** arising from "hub labs" whose
   repeated publications may inflate pooled estimates.

::: {.callout-note}
## Project status
The literature screening phase is currently in progress. This page will be updated
as each phase of the pipeline is completed.
:::

### Key numbers at a glance

| Metric | Value |
|---|---|
| Raw records (combined databases) | 2,083 |
| Records after deduplication & year filter (≥ 2019) | 495 |
| Primary screeners | 15 |
| Secondary validators | 2 |
| AI-assisted extraction steps | 4 |
| Top journals represented | Animal Behaviour, Behavioral Ecology, PRSB |

---

## Data Sources {#sec-data-sources}

### Literature search

Records were retrieved from two major bibliographic databases and combined in
**Rayyan** format for import into R:

- **Scopus** — full-text and abstract search using Boolean terms covering mate choice,
  copying, non-independent mate selection, and social information use.
- **Web of Science (WoS)** — parallel search with identical query structure.

The two exports were merged into a single file:
`data/0_literature_screening/combined_records_final.csv`

### Master dataset fields

```{r}
#| label: tbl-fields
#| tbl-cap: "Fields in combined_records_final.csv"
fields <- tibble::tribble(
  ~Field,        ~Description,                                          ~Completeness,
  "title",       "Article title (primary deduplication key)",           "~100%",
  "doi",         "Digital Object Identifier (secondary dedup key)",     "~80%",
  "abstract",    "Full abstract text",                                  "~90%",
  "authors",     "Author list",                                         "~95%",
  "year",        "Publication year",                                    "~100%",
  "journal",     "Journal name",                                        "~100%",
  "source_file", "Which database CSV contributed this record",          "100%",
  "pubmed_id",   "PubMed identifier",                                   "Partial",
  "pmc_id",      "PubMed Central identifier",                           "Partial"
)

kable(fields, align = "lll") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

::: {.callout-important}
## Provenance tracking
The `source_file` field is preserved through every deduplication step and is
concatenated (not lost) when duplicate records from different databases are merged.
In the final output it is also appended to the `abstract` field, so that the complete
database-of-origin audit trail survives import into Rayyan or any other screening tool.
:::

---

## Processing Pipeline {#sec-pipeline}

All deduplication and screener-assignment logic lives in a single reproducible script:

> `scripts/0_literature_screening/1_deduplicate_records.R`

The script was written in **February 2026** using `data.table` for efficient in-memory
processing of the 2,083-record dataset.

### Step 1 — Load data

```{r}
#| label: load-data
#| eval: false
rm(list = ls())

library(dplyr)
library(data.table)
library(stringr)
library(readr)

# Rayyan estimated 17,812 records have duplicates across the raw exports
articles <- fread("data/0_literature_screening/combined_records_final.csv")
```

### Step 2 — Title-based deduplication {#sec-title-dedup}

Titles are normalised to a canonical lowercase form before matching.
When duplicates are found, the **highest-quality** record is retained.

**Quality score** (0–3): one point each for having a non-empty abstract, a non-empty
author list, and a non-empty DOI.

```{r}
#| label: title-dedup
#| eval: false

# --- Normalise title ----------------------------------------------------------
articles[, title_lwr := str_to_lower(title)]
articles[, title_lwr := gsub("[[:punct:]]", " ", title_lwr)]  # strip punctuation
articles[, title_lwr := gsub(" {2,}", " ", title_lwr)]        # collapse whitespace
articles[, title_lwr := trimws(title_lwr)]

# --- Quality score (0–3) ------------------------------------------------------
articles[, has_abstract := ifelse(abstract == "" | is.na(abstract), 0L, 1L)]
articles[, has_authors  := ifelse(authors  == "" | is.na(authors),  0L, 1L)]
articles[, has_doi      := ifelse(doi      == "" | is.na(doi),      0L, 1L)]
articles[, quality      := has_abstract + has_authors + has_doi]

# --- Concatenate source files per title group ---------------------------------
sources <- articles[,
  .(source_file = paste(sort(unique(source_file)), collapse = "; ")),
  by = .(title_lwr)
]

# --- Keep highest-quality record per title ------------------------------------
articles.title.dedupe <- articles[
  , .SD[which.max(quality), !c("source_file"), with = FALSE],
  by = .(title_lwr)
]

# --- Merge concatenated sources back in ---------------------------------------
articles.title.dedupe.mrg <- merge(
  articles.title.dedupe, sources,
  by = "title_lwr", all.x = TRUE
)
```

#### Data decisions — title deduplication

| Decision | What was done | Rationale |
|---|---|---|
| Lowercase conversion | `str_to_lower()` | Catches case differences across databases |
| Punctuation removal | `gsub("[[:punct:]]", " ", ...)` | Handles hyphens, apostrophes, colons in titles |
| Whitespace collapse | `gsub(" {2,}", " ", ...)` | Normalises after punctuation removal |
| Quality scoring | abstract + authors + doi (0/1 each) | Retains the most information-rich duplicate |
| Source concatenation | `paste(sort(unique(source_file)), collapse = "; ")` | Full provenance preserved across all duplicates |

### Step 3 — DOI-based deduplication {#sec-doi-dedup}

A second pass on the DOI field catches articles where titles differ slightly between
database exports but resolve to the same paper. A custom helper function
`split_conc()` uniquely concatenates source strings that were themselves already
concatenated in the title-dedup step.

```{r}
#| label: doi-dedup
#| eval: false

# Helper: split already-concatenated source strings, deduplicate, re-join
split_conc <- function(x) {
  str_split(x, pattern = "; ") |>
    unlist() |>
    unique() |>
    sort() |>
    paste(collapse = "; ")
}

# Recalculate quality after title dedup (fields may have changed)
articles.title.dedupe.mrg[, has_abstract := ifelse(abstract == "" | is.na(abstract), 0L, 1L)]
articles.title.dedupe.mrg[, has_authors  := ifelse(authors  == "" | is.na(authors),  0L, 1L)]
articles.title.dedupe.mrg[, has_doi      := ifelse(doi      == "" | is.na(doi),      0L, 1L)]
articles.title.dedupe.mrg[, quality      := has_abstract + has_authors + has_doi]

# Concatenate sources per DOI
sources <- articles.title.dedupe.mrg[,
  .(source_file = split_conc(source_file)),
  by = .(doi)
]

# Keep highest-quality record per DOI
articles.doi.dedupe <- articles.title.dedupe.mrg[
  , .SD[which.max(quality), !c("source_file"), with = FALSE],
  by = .(doi)
]

# Merge sources back in
articles.final <- merge(
  articles.doi.dedupe, sources,
  by = "doi", all.x = TRUE
)
```

### Step 4 — Clean and append provenance to abstract {#sec-clean}

Intermediate processing columns are dropped, and the `source_file` string is appended
to the `abstract` field so that provenance survives any downstream tool that does not
retain a `source_file` column.

```{r}
#| label: clean
#| eval: false

articles.final[, abstract := paste0(abstract, " | ", source_file)]

articles.final <- articles.final[,
  !c("title_lwr", "n_title_matches",
     "has_abstract", "has_authors", "has_doi",
     "quality", "n_doi_matches", "source_file")
]
```

### Step 5 — Year filter: 2019 and later {#sec-year-filter}

```{r}
#| label: year-filter
#| eval: false

articles.final <- articles.final[year >= 2019, ]
message("Records after filter: ", nrow(articles.final))
message("Earliest year in dataset: ", min(articles.final$year))
```

::: {.callout-important}
## Data decision — year filter
Only records published **in or after 2019** are retained for screening.

**Rationale:**

- Aligns with the pre-registered scope of the review.
- Post-2019 literature uses more standardised reporting practices (e.g., ARRIVE 2.0),
  making effect-size extraction more reliable.
- Pre-2019 records remain available in `combined_records_final.csv` and can be
  incorporated in sensitivity analyses.

The filter reduces the deduplicated pool to **495 articles**.
:::

### Step 6 — Random shuffle and screener assignment {#sec-assignment}

Articles are randomly shuffled before assignment to prevent ordering artefacts (e.g.,
alphabetical or chronological clustering within screener batches). A dynamic function
generates near-equal assignment vectors for any team size.

```{r}
#| label: assignment
#| eval: false

# Team structure
screeners <- c(
  "Aleksandra", "Aneta", "Anna", "Ayumi", "Cassidy", "Christine",
  "Erick", "Iwo", "Hao", "Jimuel", "Kyle", "Marija",
  "Santiago", "Sergio", "Shinichi",   # 15 primary screeners
  "Losia", "Eduardo"                  # 2 secondary validators
)

primary_list   <- setdiff(screeners, c("Eduardo", "Losia"))
secondary_list <- c("Eduardo", "Losia")

# Random shuffle (prevents date/alpha ordering bias)
set.seed(NULL)  # no fixed seed — shuffled output files are the reproducible record
articles.final <- articles.final[sample(.N), ]

# Dynamic assignment: ceiling(total / n) reps, trimmed to total rows
create_assignments <- function(names, total_rows) {
  reps_needed <- ceiling(total_rows / length(names))
  full_vector <- rep(names, each = reps_needed)
  return(full_vector[seq_len(total_rows)])
}

n_total <- nrow(articles.final)

# Primary: ~33 articles per screener (495 / 15)
articles.final[, primary_screener := create_assignments(primary_list, n_total)]

# Secondary: ~248 articles per validator (495 / 2)
articles.final[, secondary_screener := create_assignments(secondary_list, n_total)]

# Verify balance
articles.final[, .(n = .N), by = .(primary_screener)]
articles.final[, .(n = .N), by = .(secondary_screener)]
```

### Step 7 — Export per-screener CSVs {#sec-export}

One CSV is written per primary screener to
`builds/0_literature_screening/deduplicated_and_randomly_split/`.
Each file retains both assignment columns for cross-referencing.

```{r}
#| label: export
#| eval: false

invisible(lapply(
  unique(articles.final$primary_screener),
  function(x) {
    write_csv(
      articles.final[primary_screener == x, ],
      paste0("builds/0_literature_screening/deduplicated_and_randomly_split/", x, ".csv")
    )
  }
))
```

---

## Literature Screening {#sec-screening}

### Team structure

The 495 articles are distributed across a 17-person team. Each article receives:

- **One primary screen** — eligibility assessment against pre-registered inclusion criteria.
- **One independent secondary validation** — every article is also assigned to one of
  two validators (Eduardo or Losia), enabling inter-rater reliability estimation.

```{r}
#| label: tbl-team
#| tbl-cap: "Screener assignments"
team <- tibble::tribble(
  ~Screener,    ~Role,      ~Articles,
  "Aleksandra", "Primary",  "~33",
  "Aneta",      "Primary",  "~33",
  "Anna",       "Primary",  "~33",
  "Ayumi",      "Primary",  "~33",
  "Cassidy",    "Primary",  "~33",
  "Christine",  "Primary",  "~33",
  "Erick",      "Primary",  "~33",
  "Hao",        "Primary",  "~33",
  "Iwo",        "Primary",  "~33",
  "Jimuel",     "Primary",  "~33",
  "Kyle",       "Primary",  "~33",
  "Marija",     "Primary",  "~33",
  "Santiago",   "Primary",  "~33",
  "Sergio",     "Primary",  "~33",
  "Shinichi",   "Primary",  "~33",
  "Eduardo",    "Validator","~248",
  "Losia",      "Validator","~248"
)

kable(team, align = "lll") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) |>
  row_spec(16:17, bold = TRUE, background = "#fdf4e3")
```

### Inclusion criteria

Articles are included if they:

1. Report an **experimental** test of mate-choice copying (observer watches a demonstrator
   make a mate choice, then has the opportunity to choose between the same mates).
2. Provide sufficient quantitative data for effect-size extraction (means + variance, or
   counts, or test statistics with sample sizes).
3. Were published **2019 or later** (pre-registration scope).
4. Are written in English.

### Inter-rater reliability

Disagreements between primary and secondary screeners will be resolved by discussion
or a designated tie-breaker. Cohen's κ (or a Bayesian analogue) will be calculated
and reported in the supplementary methods.

---

## Data Extraction Pipeline {#sec-extraction}

Included articles are sent to **Google NotebookLM** for AI-assisted data extraction
via a four-step structured prompting protocol. All prompts are grounded by a
`MetaAnalysis_Reference_Sheet.pdf` data dictionary containing few-shot examples of
correct extractions.

::: {.callout-warning}
## Cache-bust rule
Each step must be run in a **completely new** NotebookLM chat window. Re-using a
previous window causes the model to recycle earlier tables rather than reading the
PDFs afresh.
:::

### Step 1 — Species & study identification

**Goal:** Establish taxonomic accuracy across all included papers.

Each PDF yields one row in a Markdown table with two columns:

| Column | Format | Example |
|---|---|---|
| `Study_ID` | `FirstAuthorLastName_Year_CamelCaseJournal` | `Belkina_2021_AnimalBehaviour` |
| `Species` | Full binomial nomenclature | *Poecilia reticulata* |

Common names alone are not accepted. The Study ID format is consistent across all
four steps to serve as the primary join key.

### Step 2 — Co-authorship & lab identification

**Goal:** Build a bipartite co-authorship network to identify "hub labs."

Output is in **long format** — one row per author per paper — so that network edges
can be constructed directly.

| Column | Description |
|---|---|
| `Study_ID` | Join key |
| `Author_Name` | Full name |
| `Author_Position` | First / Last / Corresponding / Middle |
| `Institutional_Affiliation` | University or institute |
| `Lab_Identifier` | `PILastName_Institution_City` (e.g., `Dukas_UniversityOfToronto_Toronto`) |

::: {.callout-note}
## Lab identifier design
The **last author** is designated as the lab PI, and the same `Lab_Identifier` string
is applied to every co-author row on the same paper. This links all authors to one
"primary lab" node in the bipartite network, enabling correct degree-centrality
calculation for hub-lab detection.
:::

### Step 3 — Trial-level data mapping

**Goal:** Map the internal experimental structure of each paper — how many distinct
trials exist, where their data live, and whether subjects are independent.

| Column | Description |
|---|---|
| `Study_ID` | Join key |
| `Trial_ID` | Experiment name/number (e.g., "Experiment 1", "Treatment A") |
| `Trial_Description` | 1–2 sentence summary of manipulation |
| `Sample_Size_N` | Focal subjects in this trial |
| `Subject_Independence` | "Independent Cohort" or "Repeated Measures" |
| `Control_Group_Shared` | Yes/No + which trial (if Yes) |
| `Data_Location` | Exact page and figure/table reference (e.g., "Table 1, Page 5") |

### Step 4 — Quantitative extraction (multi-agent pipeline)

**Goal:** Extract and cross-validate the actual numbers using a simulated three-agent
architecture.

#### Agent architecture

```
┌──────────────────────────────────────────────────────────────────────┐
│  For each PDF:                                                       │
│                                                                      │
│  Agent 1 (Extractor)   →  Agent 2 (Verifier)  →  Agent 3 (Arbiter) │
│  Pulls raw numbers        Checks for:             Resolves conflicts │
│  from Methods/Results     • SD vs SE confusion    Formats final row  │
│                           • Sample independence                      │
│                           • Missing trials                           │
└──────────────────────────────────────────────────────────────────────┘
```

**Agent 2 checks explicitly for three systematic extraction errors:**

1. **SD/SE confusion** — a common error in manual extraction that biases heterogeneity
   estimates if uncorrected.
2. **Sample independence** — flags whether the N used is independent across trials or
   from repeated measures on the same subjects.
3. **Missing trials** — re-reads the Methods to confirm no secondary experiments
   were missed by Agent 1.

#### Final output columns

| Column | Content |
|---|---|
| `Study_ID` | Author_Year_JournalName |
| `Trial_ID` | Specific experiment or condition |
| `Response_Metric` | What was measured |
| `Validated_Exp_Group_Data` | Mean ± SD/SE, Counts, or Test Stat |
| `Validated_Control_Group_Data` | Mean ± SD/SE, Counts, or NA |
| `SD_or_SE` | Explicit variance metric label |
| `Agent1_Initial_Pull` | Brief summary of original extraction |
| `Agent2_Verification_Notes` | Corrections made, or "No missing trials found" |
| `Verification_Location` | Exact page and table/figure number |

---

## Effect Size Calculation {#sec-effect-sizes}

Following data extraction, effect sizes and their sampling variances are calculated using the reproducible Python pipeline located at [`scripts/1_effect_size_calculation_pipeline/calculate_effect_sizes.py`](file:///Users/eduardosantos/Documents/Repos/mate_choice_meta/scripts/1_effect_size_calculation_pipeline/calculate_effect_sizes.py). 

The raw data is stored in the Excel spreadsheet [`data/1_effect_size_calculation_pipeline/Data Extraction Mate Choice Meta Analysis.xlsx`](file:///Users/eduardosantos/Documents/Repos/mate_choice_meta/data/1_effect_size_calculation_pipeline/Data%20Extraction%20Mate%20Choice%20Meta%20Analysis.xlsx) inside the `EduardoExtraction` sheet. To support tracking, each row contains an `identifierEffectSizeID` column containing a unique sequential identifier for each control–experimental comparison (e.g., `es0001`, `es0002`). The Python pipeline dynamically resolves column locations by scanning the header row, computes the following effect size metrics, and saves them back to the spreadsheet:

### 1. Hedges' $g$ (Bias-Corrected Standardised Mean Difference)
Calculated depending on the data type extracted:
* **From Means & SDs:** Using control and treatment means, standard deviations (or standard errors converted using $SD = SE \times \sqrt{n}$), and sample sizes.
* **From $t$-statistics:** Supported for both paired/one-sample designs (Case 2) and independent-sample designs (Case 3, assuming an equal sample split if individual group sizes are unknown).

### 2. Odds Ratio (OR)
Calculated for categorical outcomes from $2 \times 2$ contingency tables ($A$, $B$, $C$, $D$), where the sampling variance of the log odds ratio is approximated using the delta method.

### 3. Odds Ratio $\leftrightarrow$ Hedges' $g$ Cross-Conversion
For rows containing only one of the two metrics, the missing metric is derived using the logistic-normal approximation (Chinn 2000, Borenstein et al. 2009):
* **$\text{OR} \rightarrow \text{Hedges' } g$**: $d = \ln(\text{OR}) \times \frac{\sqrt{3}}{\pi}$ and then bias-corrected to $g$ using the $J$ correction factor.
* **$\text{Hedges' } g \rightarrow \text{OR}$**: $\ln(\text{OR}) = g \times \frac{\pi}{\sqrt{3}}$ and $\text{Var}(\ln\text{OR}) = \text{Var}(g) \times \frac{\pi^2}{3}$.

To execute the calculations and update the workbook, run:
```bash
python scripts/1_effect_size_calculation_pipeline/calculate_effect_sizes.py
```

---

## Planned Analyses {#sec-analyses}

All models will be implemented in R. Pre-registration details will be linked here.

### Meta-analytic models

Multi-level random-effects meta-analysis to estimate the overall effect size for MCC,
accounting for trials nested within papers nested within labs.
Packages: `metafor`, `clubSandwich`.

**Heterogeneity** will be quantified with $I^2$ and $\tau^2$ at each level.

### Co-authorship network analysis

The long-format table from extraction Step 2 feeds directly into a bipartite
lab–paper network constructed in R (`igraph` or `tidygraph`). Hub labs are identified
by degree centrality. Non-independence is modelled by including a lab-level random
effect or using network-based weights in the meta-analytic model.

### Phylogenetic comparative methods

Species non-independence is controlled via a phylogenetic covariance matrix
incorporated into the meta-analytic model (`metafor` + `ape`). This accounts for the
possibility that closely related species share similar MCC propensity due to shared
evolutionary history.

### Publication bias

Funnel plot asymmetry, Egger's regression test, and trim-and-fill analysis.
Time-lag bias will be examined by plotting effect size against publication year.

### Moderator analyses

Meta-regression models will test whether effect sizes vary by:

- Sex of the copying individual (female vs. male)
- Taxonomic group (fish, bird, mammal, invertebrate)
- Experimental design (lab vs. semi-natural)
- Model presentation method (live demonstrator vs. video vs. trained model)

### Sensitivity analyses

- Inclusion of pre-2019 records.
- Leave-one-out analysis for influential studies.
- Re-analysis excluding studies from identified hub labs.

## Results {#sec-results}

This section integrates the complete findings of our Mate-Choice Copying (MCC) meta-analyses. We first present the **Overall Effect Size Synthesis** across the database splits (New Extraction, Original Only, and Combined Pools), followed by **Publication Bias & Selective Reporting Diagnostics** to assess the robustness and validity of the literature.

---

### Data Loading & Alignment {#sec-results-loading}

We load the raw datasets and align them into unified data structures for both Hedges' $g$ (continuous SMD) and Odds Ratio (binary) effect sizes.

```{r}
#| label: load_and_align_datasets
#| message: false
#| warning: false

# Paths to legacy and new datasets
path_new <- here::here("data", "1_effect_size_calculation_pipeline", "Data Extraction Mate Choice Meta Analysis.xlsx")
path_davies <- here::here("data", "2_original_meta_analyses_datasets", "Davies_et_al_2020_Final_data.xlsx")
path_jones <- here::here("data", "2_original_meta_analyses_datasets", "Jones_DuVal_2019_data.CSV")

# A. Read raw spreadsheets
db_new_raw <- readxl::read_excel(path_new, sheet = "EduardoExtraction")
db_davies_raw <- readxl::read_excel(path_davies, sheet = "Data")
db_jones_raw <- readr::read_csv(path_jones)

# B. Clean new extractions
db_new <- db_new_raw %>%
    filter(is.na(RemoveFromDataset) | RemoveFromDataset != "Yes") %>%
    mutate(
        effectSizeCalculatedHedgesD = as.numeric(effectSizeCalculatedHedgesD),
        effectSizeCalculatedHedgesDVariance = as.numeric(effectSizeCalculatedHedgesDVariance),
        effectSizeCalculatedOddsRatio = as.numeric(effectSizeCalculatedOddsRatio),
        effectSizeCalculatedOddsRatioVariance = as.numeric(effectSizeCalculatedOddsRatioVariance),
        logOddsRatio = log(effectSizeCalculatedOddsRatio),
        logOddsRatioVariance = effectSizeCalculatedOddsRatioVariance,
        identifierStudyYear = as.numeric(identifierStudyYear),
        controlN = as.numeric(controlN),
        treatmentN = as.numeric(treatmentN),
        effectSizesampleSize = as.numeric(effectSizesampleSize)
    )

# ---------------------------------------------------------------------------
# 1. Hedges' g Dataset Alignment (Continuous)
# ---------------------------------------------------------------------------
hedges_new <- db_new %>%
    filter(!is.na(effectSizeCalculatedHedgesD), !is.na(effectSizeCalculatedHedgesDVariance)) %>%
    select(
        yi = effectSizeCalculatedHedgesD,
        vi = effectSizeCalculatedHedgesDVariance,
        identifierStudyId,
        identifierEffectSizeID,
        taxonomySpecies,
        Year = identifierStudyYear,
        controlN,
        treatmentN,
        totalN = effectSizesampleSize
    ) %>%
    mutate(
        identifierStudyId = as.character(identifierStudyId),
        identifierEffectSizeID = as.character(identifierEffectSizeID),
        taxonomySpecies = str_to_lower(word(as.character(taxonomySpecies), -1)),
        Dataset_Source = "New Extraction (Santos)"
    )

hedges_davies <- db_davies_raw %>%
    mutate(
        Hedges_d_directional = as.numeric(Hedges_d_directional),
        Variance = as.numeric(Variance),
        Year = as.numeric(Year),
        N_control = as.numeric(N_control),
        N_treatment = as.numeric(N_treatment),
        Total_n_analysis = as.numeric(Total_n_analysis)
    ) %>%
    filter(!is.na(Hedges_d_directional), !is.na(Variance), Variance > 0) %>%
    select(
        yi = Hedges_d_directional,
        vi = Variance,
        identifierStudyId = Author,
        identifierEffectSizeID = Effect_size_no,
        taxonomySpecies = Species_latin,
        Year,
        controlN = N_control,
        treatmentN = N_treatment,
        totalN = Total_n_analysis
    ) %>%
    mutate(
        identifierStudyId = as.character(identifierStudyId),
        identifierEffectSizeID = as.character(identifierEffectSizeID),
        taxonomySpecies = str_to_lower(word(as.character(taxonomySpecies), -1)),
        Dataset_Source = "Original (Davies 2020)"
    )

db_hedges <- bind_rows(hedges_new, hedges_davies) %>%
    mutate(across(c(identifierStudyId, taxonomySpecies, Dataset_Source), as.factor))

# ---------------------------------------------------------------------------
# 2. Log Odds Ratio Dataset Alignment (Binary)
# ---------------------------------------------------------------------------
or_new <- db_new %>%
    filter(!is.na(logOddsRatio), !is.na(logOddsRatioVariance)) %>%
    select(
        yi = logOddsRatio,
        vi = logOddsRatioVariance,
        identifierStudyId,
        identifierEffectSizeID,
        taxonomySpecies,
        Year = identifierStudyYear
    ) %>%
    mutate(
        identifierStudyId = as.character(identifierStudyId),
        identifierEffectSizeID = as.character(identifierEffectSizeID),
        taxonomySpecies = str_to_lower(word(as.character(taxonomySpecies), -1)),
        Dataset_Source = "New Extraction (Santos)"
    )

or_jones <- db_jones_raw %>%
    filter(!is.na(lnOR), !is.na(VlnOR), VlnOR > 0) %>%
    select(
        yi = lnOR,
        vi = VlnOR,
        identifierStudyId = study,
        identifierEffectSizeID = study_Exp,
        taxonomySpecies = Species,
        Year = year
    ) %>%
    mutate(
        identifierStudyId = as.character(identifierStudyId),
        identifierEffectSizeID = as.character(identifierEffectSizeID),
        taxonomySpecies = str_to_lower(as.character(taxonomySpecies)),
        Dataset_Source = "Original (Jones & DuVal 2019)"
    )

db_or <- bind_rows(or_new, or_jones) %>%
    mutate(across(c(identifierStudyId, taxonomySpecies, Dataset_Source), as.factor))
```

---

### Overall Effect Size Synthesis {#sec-overall-synthesis-results}

#### Effect Size Distribution (Density Plots) {#sec-density}

Before fitting the multi-level models, we inspect the distribution of our aligned effect sizes across the different subsets (New Extraction, Original Only, and Combined). 

The density plots below use the same split logic as the orchard plots, facet-wrapping by subset while coloring the distribution curves by their actual dataset source to check for shifts or outliers.

::: {.panel-tabset}

##### Hedges' $g$ (Continuous SMD)

```{r}
#| label: hedges_density_plots
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 9

# Prepare density data for Hedges' g
db_hedges_density <- bind_rows(
  db_hedges %>% mutate(Group = as.character(Dataset_Source)),
  db_hedges %>% mutate(Group = "Combined")
) %>%
  mutate(Group = factor(Group, levels = c("New Extraction (Santos)", "Original (Davies 2020)", "Combined")))

# Plot densities
p_hedges_dens <- ggplot(db_hedges_density, aes(x = yi, fill = Dataset_Source, color = Dataset_Source)) +
  geom_density(alpha = 0.4, linewidth = 0.8) +
  geom_vline(xintercept = 0, linewidth = 1.0, linetype = "solid", color = "black") +
  geom_rug(alpha = 0.6) +
  facet_wrap(~ Group, ncol = 1) +
  scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
  scale_color_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
  labs(
    title = "Distribution of Hedges' g Effect Sizes",
    x = "Hedges' g",
    y = "Density"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    strip.text = element_text(face = "bold", size = 11),
    plot.title = element_text(face = "bold", hjust = 0.5)
  )

print(p_hedges_dens)
```

##### Odds Ratio (Binary outcomes)

```{r}
#| label: or_density_plots
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 9

# Prepare density data for Odds Ratio (lnOR)
db_or_density <- bind_rows(
  db_or %>% mutate(Group = as.character(Dataset_Source)),
  db_or %>% mutate(Group = "Combined")
) %>%
  mutate(Group = factor(Group, levels = c("New Extraction (Santos)", "Original (Jones & DuVal 2019)", "Combined")))

# Plot densities
p_or_dens <- ggplot(db_or_density, aes(x = yi, fill = Dataset_Source, color = Dataset_Source)) +
  geom_density(alpha = 0.4, linewidth = 0.8) +
  geom_vline(xintercept = 0, linewidth = 1.0, linetype = "solid", color = "black") +
  geom_rug(alpha = 0.6) +
  facet_wrap(~ Group, ncol = 1) +
  scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal 2019)" = "#0072B2"), name = "Data Source") +
  scale_color_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal 2019)" = "#0072B2"), name = "Data Source") +
  labs(
    title = "Distribution of Log Odds Ratio (lnOR) Effect Sizes",
    x = "Log Odds Ratio (lnOR)",
    y = "Density"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    strip.text = element_text(face = "bold", size = 11),
    plot.title = element_text(face = "bold", hjust = 0.5)
  )

print(p_or_dens)
```

:::

#### Model Fitting Function

We define helper functions to handle the variance-covariance (VCV) matrix construction and fit multilevel random-effects models (`rma.mv`) in `metafor`, controlling for taxonomic (`taxonomySpecies`) and study-level clustering.

```{r}
#| label: model_fitting_function

fit_overall_mv <- function(dat, yi = "yi", vi = "vi", cluster_var = "identifierStudyId", obs_var = "identifierEffectSizeID", rho = 0.5) {
  # Remove rows with missing values
  dat2 <- dat %>%
    filter(!is.na(.data[[yi]]), !is.na(.data[[vi]]), .data[[vi]] > 0)
  
  if (nrow(dat2) == 0) {
    message("No complete cases found.")
    return(NULL)
  }
  
  # 1. Build VCV Matrix
  VCV <- metafor::vcalc(
    vi = dat2[[vi]],
    cluster = dat2[[cluster_var]],
    obs = dat2[[obs_var]],
    rho = rho,
    data = dat2
  )
  
  # 2. Fit Multi-Level Random-Effects Model
  mod <- metafor::rma.mv(
    yi = dat2[[yi]],
    V  = VCV,
    random = list(
      ~1 | identifierStudyId,
      ~1 | identifierEffectSizeID,
      ~1 | taxonomySpecies
    ),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = dat2
  )
  
  list(data = dat2, VCV = VCV, model = mod)
}

# Helper function to fit bias-corrected multilevel models with a continuous moderator (Egger-type)
fit_corrected_mv <- function(dat, yi = "yi", vi = "vi", moderator = "vi", cluster_var = "identifierStudyId", obs_var = "identifierEffectSizeID", rho = 0.5) {
  # Remove rows with missing values
  dat2 <- dat %>%
    filter(!is.na(.data[[yi]]), !is.na(.data[[vi]]), .data[[vi]] > 0)
  
  if (nrow(dat2) == 0) {
    message("No complete cases found.")
    return(NULL)
  }
  
  if (moderator == "se" && !"se" %in% colnames(dat2)) {
    dat2 <- dat2 %>% mutate(se = sqrt(vi))
  }
  
  # 1. Build VCV Matrix
  VCV <- metafor::vcalc(
    vi = dat2[[vi]],
    cluster = dat2[[cluster_var]],
    obs = dat2[[obs_var]],
    rho = rho,
    data = dat2
  )
  
  # 2. Fit Multi-Level Egger Model using dynamic formula
  formula_str <- paste("~ 1 +", moderator)
  mod <- metafor::rma.mv(
    yi = dat2[[yi]],
    V  = VCV,
    mods = as.formula(formula_str),
    random = list(
      ~1 | identifierStudyId,
      ~1 | identifierEffectSizeID,
      ~1 | taxonomySpecies
    ),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = dat2
  )
  
  list(data = dat2, VCV = VCV, model = mod)
}

# Helper function to add Mean (95% CI) labels dynamically one line above k-values in orchard plots
add_ci_labels <- function(plot_obj, orchard_results) {
  est <- sprintf("%.2f", orchard_results$mod_table$estimate)
  ci_l <- sprintf("%.2f", orchard_results$mod_table$lowerCL)
  ci_u <- sprintf("%.2f", orchard_results$mod_table$upperCL)
  
  orig_labels <- plot_obj$layers[[6]]$aes_params$label
  
  new_labels <- c()
  for (i in 1:length(orig_labels)) {
    # Extract the original sample size numbers using regex
    nums <- regmatches(orig_labels[i], gregexpr("[0-9]+", orig_labels[i]))[[1]]
    if (length(nums) >= 2) {
      k_val <- nums[1]
      n_val <- nums[2]
      k_text <- paste0("k = ", k_val, " (", n_val, ")")
    } else {
      k_text <- orig_labels[i]
    }
    # Place Mean (95% CI) on the first line and k = xx (xx) on the second line
    new_labels[i] <- paste0(est[i], " [", ci_l[i], ", ", ci_u[i], "]\n", k_text)
  }
  
  plot_obj$layers[[6]]$aes_params$label <- new_labels
  plot_obj$layers[[6]]$geom_params$parse <- FALSE
  plot_obj$layers[[6]]$aes_params$hjust <- "right"
  plot_obj$layers[[6]]$aes_params$vjust <- 0.5
  plot_obj$layers[[6]]$aes_params$size <- 3.0 # Slightly smaller font size to fit beautifully
  
  plot_obj
}
```

```{r}
#| label: fit_multilevel_models
#| message: false
#| warning: false
#| include: false

# ---------------------------------------------------------------------------
# HEDGES' G MODELS (UNCORRECTED)
# ---------------------------------------------------------------------------
res_hedges_new   <- fit_overall_mv(db_hedges %>% filter(Dataset_Source == "New Extraction (Santos)"))
res_hedges_orig  <- fit_overall_mv(db_hedges %>% filter(Dataset_Source == "Original (Davies 2020)"))
res_hedges_comb  <- fit_overall_mv(db_hedges)

# HEDGES' G MODELS (BIAS-CORRECTED)
res_hedges_new_corr   <- fit_corrected_mv(db_hedges %>% filter(Dataset_Source == "New Extraction (Santos)"), moderator = "vi")
res_hedges_orig_corr  <- fit_corrected_mv(db_hedges %>% filter(Dataset_Source == "Original (Davies 2020)"), moderator = "vi")
res_hedges_comb_corr  <- fit_corrected_mv(db_hedges, moderator = "vi")

# ---------------------------------------------------------------------------
# ODDS RATIO MODELS (UNCORRECTED)
# ---------------------------------------------------------------------------
res_or_new   <- fit_overall_mv(db_or %>% filter(Dataset_Source == "New Extraction (Santos)"))
res_or_orig  <- fit_overall_mv(db_or %>% filter(Dataset_Source == "Original (Jones & DuVal 2019)"))
res_or_comb  <- fit_overall_mv(db_or)

# ODDS RATIO MODELS (BIAS-CORRECTED)
res_or_new_corr   <- fit_corrected_mv(db_or %>% filter(Dataset_Source == "New Extraction (Santos)"), moderator = "vi")
res_or_orig_corr  <- fit_corrected_mv(db_or %>% filter(Dataset_Source == "Original (Jones & DuVal 2019)"), moderator = "vi")
res_or_comb_corr  <- fit_corrected_mv(db_or, moderator = "vi")
```

#### Results and Visualizations

::: {.panel-tabset}

##### Hedges' $g$ (Continuous outcomes)

::: {.panel-tabset}

###### Uncorrected Overall Effect

This plot displays the standard multi-level meta-analytic summaries across the three splits (New Only, Original Only, and Combined) *without* adjusting for publication bias.

```{r}
#| label: hedges_unified_plot
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 7

if (!is.null(res_hedges_new) && !is.null(res_hedges_orig) && !is.null(res_hedges_comb)) {
  h_new <- orchaRd::mod_results(res_hedges_new$model, group = "identifierStudyId")
  h_orig <- orchaRd::mod_results(res_hedges_orig$model, group = "identifierStudyId")
  h_comb <- orchaRd::mod_results(res_hedges_comb$model, group = "identifierStudyId")

  class(h_new) <- "list"
  h_new$mod_table$name <- "New Extraction (Santos)"
  h_new$data$moderator <- "New Extraction (Santos)"
  class(h_new) <- c("orchard", "data.frame")

  class(h_orig) <- "list"
  h_orig$mod_table$name <- "Original (Davies 2020)"
  h_orig$data$moderator <- "Original (Davies 2020)"
  class(h_orig) <- c("orchard", "data.frame")

  class(h_comb) <- "list"
  h_comb$mod_table$name <- "Combined"
  h_comb$data$moderator <- "Combined"
  class(h_comb) <- c("orchard", "data.frame")

  sub_hedges <- orchaRd::submerge(h_new, h_orig, h_comb)

  class(sub_hedges) <- "list"
  sub_hedges$data$real_source <- c(
    rep("New Extraction (Santos)", nrow(h_new$data)),
    rep("Original (Davies 2020)", nrow(h_orig$data)),
    rep("New Extraction (Santos)", nrow(h_new$data)),
    rep("Original (Davies 2020)", nrow(h_orig$data))
  )
  class(sub_hedges) <- c("orchard", "data.frame")

  p_hedges <- orchaRd::orchard_plot(sub_hedges, xlab = "Hedges' g", flip = TRUE)

  p_hedges$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
  p_hedges$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

  p_hedges <- p_hedges +
    ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
    ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
    ggplot2::ggtitle("Overall Synthesis — Hedges' g Estimates (Uncorrected)") +
    ggplot2::theme(
      legend.position = c(0.98, 0.02),
      legend.justification = c("right", "bottom"),
      legend.background = ggplot2::element_rect(fill = scales::alpha("white", 0.75), color = "grey90"),
      legend.box = "vertical",
      legend.margin = ggplot2::margin(t = 4, r = 4, b = 4, l = 4),
      legend.title = ggplot2::element_text(size = 9, face = "bold"),
      legend.text = ggplot2::element_text(size = 8)
    )

  p_hedges <- add_ci_labels(p_hedges, sub_hedges)
  print(p_hedges)
}
```

###### Bias-Corrected Overall Effect

This plot displays the small-study publication bias corrected meta-analytic summaries (evaluated at study variance $\text{V} = 0$, representing theoretical infinite sample precision) using the Egger-type variance-adjustment framework.

```{r}
#| label: hedges_corrected_plot
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 7

if (!is.null(res_hedges_new_corr) && !is.null(res_hedges_orig_corr) && !is.null(res_hedges_comb_corr)) {
  h_new_corr <- orchaRd::mod_results(res_hedges_new_corr$model, group = "identifierStudyId", at = list(vi = 0), mod = "1")
  h_orig_corr <- orchaRd::mod_results(res_hedges_orig_corr$model, group = "identifierStudyId", at = list(vi = 0), mod = "1")
  h_comb_corr <- orchaRd::mod_results(res_hedges_comb_corr$model, group = "identifierStudyId", at = list(vi = 0), mod = "1")

  class(h_new_corr) <- "list"
  h_new_corr$mod_table$name <- "New Extraction (Santos)"
  h_new_corr$data$moderator <- "New Extraction (Santos)"
  class(h_new_corr) <- c("orchard", "data.frame")

  class(h_orig_corr) <- "list"
  h_orig_corr$mod_table$name <- "Original (Davies 2020)"
  h_orig_corr$data$moderator <- "Original (Davies 2020)"
  class(h_orig_corr) <- c("orchard", "data.frame")

  class(h_comb_corr) <- "list"
  h_comb_corr$mod_table$name <- "Combined"
  h_comb_corr$data$moderator <- "Combined"
  class(h_comb_corr) <- c("orchard", "data.frame")

  sub_hedges_corr <- orchaRd::submerge(h_new_corr, h_orig_corr, h_comb_corr)

  class(sub_hedges_corr) <- "list"
  sub_hedges_corr$data$real_source <- c(
    rep("New Extraction (Santos)", nrow(h_new_corr$data)),
    rep("Original (Davies 2020)", nrow(h_orig_corr$data)),
    rep("New Extraction (Santos)", nrow(h_new_corr$data)),
    rep("Original (Davies 2020)", nrow(h_orig_corr$data))
  )
  class(sub_hedges_corr) <- c("orchard", "data.frame")

  p_hedges_corr <- orchaRd::orchard_plot(sub_hedges_corr, xlab = "Hedges' g (Bias Corrected)", flip = TRUE)

  p_hedges_corr$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
  p_hedges_corr$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

  p_hedges_corr <- p_hedges_corr +
    ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
    ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
    ggplot2::ggtitle("Overall Synthesis — Hedges' g Estimates (Bias-Corrected)") +
    ggplot2::theme(
      legend.position = c(0.98, 0.02),
      legend.justification = c("right", "bottom"),
      legend.background = ggplot2::element_rect(fill = scales::alpha("white", 0.75), color = "grey90"),
      legend.box = "vertical",
      legend.margin = ggplot2::margin(t = 4, r = 4, b = 4, l = 4),
      legend.title = ggplot2::element_text(size = 9, face = "bold"),
      legend.text = ggplot2::element_text(size = 8)
    )

  p_hedges_corr <- add_ci_labels(p_hedges_corr, sub_hedges_corr)
  print(p_hedges_corr)
}
```

###### Side-by-Side Comparison

```{r}
#| label: hedges_comparison_plot
#| message: false
#| warning: false
#| fig.width: 14
#| fig.height: 8

if (exists("p_hedges") && exists("p_hedges_corr")) {
  comparison_hedges <- (p_hedges + p_hedges_corr) + 
    patchwork::plot_annotation(
      title = "Comparison of Uncorrected vs. Bias-Corrected Synthesis (Hedges' g)",
      subtitle = "Bias-corrected estimates are conditioned on study variance = 0 (infinite precision)",
      theme = ggplot2::theme(
        plot.title = ggplot2::element_text(size = 16, face = "bold", hjust = 0.5),
        plot.subtitle = ggplot2::element_text(size = 12, italic = TRUE, hjust = 0.5)
      )
    )
  print(comparison_hedges)
}
```

###### Detailed Model Statistics

::: {.panel-tabset}

####### New Only

::: {.panel-tabset}

######## Uncorrected
```{r}
#| label: hedges_new_stats
if (!is.null(res_hedges_new)) {
  print(summary(res_hedges_new$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_hedges_new$model))
}
```

######## Bias-Corrected
```{r}
#| label: hedges_new_corr_stats
if (!is.null(res_hedges_new_corr)) {
  print(summary(res_hedges_new_corr$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_hedges_new_corr$model))
}
```

:::

####### Original Only

::: {.panel-tabset}

######## Uncorrected
```{r}
#| label: hedges_orig_stats
if (!is.null(res_hedges_orig)) {
  print(summary(res_hedges_orig$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_hedges_orig$model))
}
```

######## Bias-Corrected
```{r}
#| label: hedges_orig_corr_stats
if (!is.null(res_hedges_orig_corr)) {
  print(summary(res_hedges_orig_corr$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_hedges_orig_corr$model))
}
```

:::

####### Combined

::: {.panel-tabset}

######## Uncorrected
```{r}
#| label: hedges_comb_stats
if (!is.null(res_hedges_comb)) {
  print(summary(res_hedges_comb$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_hedges_comb$model))
}
```

######## Bias-Corrected
```{r}
#| label: hedges_comb_corr_stats
if (!is.null(res_hedges_comb_corr)) {
  print(summary(res_hedges_comb_corr$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_hedges_comb_corr$model))
}
```

:::

:::

:::

##### Odds Ratio (Binary outcomes)

::: {.panel-tabset}

###### Uncorrected Overall Effect

This plot displays the standard multi-level meta-analytic summaries across the three splits (New Only, Original Only, and Combined) *without* adjusting for publication bias.

```{r}
#| label: or_unified_plot
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 7

if (!is.null(res_or_new) && !is.null(res_or_orig) && !is.null(res_or_comb)) {
  o_new <- orchaRd::mod_results(res_or_new$model, group = "identifierStudyId")
  o_orig <- orchaRd::mod_results(res_or_orig$model, group = "identifierStudyId")
  o_comb <- orchaRd::mod_results(res_or_comb$model, group = "identifierStudyId")

  class(o_new) <- "list"
  o_new$mod_table$name <- "New Extraction (Santos)"
  o_new$data$moderator <- "New Extraction (Santos)"
  class(o_new) <- c("orchard", "data.frame")

  class(o_orig) <- "list"
  o_orig$mod_table$name <- "Original (Jones & DuVal 2019)"
  o_orig$data$moderator <- "Original (Jones & DuVal 2019)"
  class(o_orig) <- c("orchard", "data.frame")

  class(o_comb) <- "list"
  o_comb$mod_table$name <- "Combined"
  o_comb$data$moderator <- "Combined"
  class(o_comb) <- c("orchard", "data.frame")

  sub_or <- orchaRd::submerge(o_new, o_orig, o_comb)

  class(sub_or) <- "list"
  sub_or$data$real_source <- c(
    rep("New Extraction (Santos)", nrow(o_new$data)),
    rep("Original (Jones & DuVal)", nrow(o_orig$data)),
    rep("New Extraction (Santos)", nrow(o_new$data)),
    rep("Original (Jones & DuVal)", nrow(o_orig$data))
  )
  class(sub_or) <- c("orchard", "data.frame")

  p_or <- orchaRd::orchard_plot(sub_or, xlab = "Log Odds Ratio (lnOR)", flip = TRUE)

  p_or$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
  p_or$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

  p_or <- p_or +
    ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal)" = "#0072B2"), name = "Data Source") +
    ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal)" = "#0072B2"), name = "Data Source") +
    ggplot2::ggtitle("Overall Synthesis — Log Odds Ratio Estimates (Uncorrected)") +
    ggplot2::theme(
      legend.position = c(0.98, 0.02),
      legend.justification = c("right", "bottom"),
      legend.background = ggplot2::element_rect(fill = scales::alpha("white", 0.75), color = "grey90"),
      legend.box = "vertical",
      legend.margin = ggplot2::margin(t = 4, r = 4, b = 4, l = 4),
      legend.title = ggplot2::element_text(size = 9, face = "bold"),
      legend.text = ggplot2::element_text(size = 8)
    )

  p_or <- add_ci_labels(p_or, sub_or)
  print(p_or)
}
```

###### Bias-Corrected Overall Effect

This plot displays the small-study publication bias corrected meta-analytic summaries (evaluated at study variance $\text{V} = 0$, representing theoretical infinite sample precision) using the Egger-type variance-adjustment framework.

```{r}
#| label: or_corrected_plot
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 7

if (!is.null(res_or_new_corr) && !is.null(res_or_orig_corr) && !is.null(res_or_comb_corr)) {
  o_new_corr <- orchaRd::mod_results(res_or_new_corr$model, group = "identifierStudyId", at = list(vi = 0), mod = "1")
  o_orig_corr <- orchaRd::mod_results(res_or_orig_corr$model, group = "identifierStudyId", at = list(vi = 0), mod = "1")
  o_comb_corr <- orchaRd::mod_results(res_or_comb_corr$model, group = "identifierStudyId", at = list(vi = 0), mod = "1")

  class(o_new_corr) <- "list"
  o_new_corr$mod_table$name <- "New Extraction (Santos)"
  o_new_corr$data$moderator <- "New Extraction (Santos)"
  class(o_new_corr) <- c("orchard", "data.frame")

  class(o_orig_corr) <- "list"
  o_orig_corr$mod_table$name <- "Original (Jones & DuVal 2019)"
  o_orig_corr$data$moderator <- "Original (Jones & DuVal 2019)"
  class(o_orig_corr) <- c("orchard", "data.frame")

  class(o_comb_corr) <- "list"
  o_comb_corr$mod_table$name <- "Combined"
  o_comb_corr$data$moderator <- "Combined"
  class(o_comb_corr) <- c("orchard", "data.frame")

  sub_or_corr <- orchaRd::submerge(o_new_corr, o_orig_corr, o_comb_corr)

  class(sub_or_corr) <- "list"
  sub_or_corr$data$real_source <- c(
    rep("New Extraction (Santos)", nrow(o_new_corr$data)),
    rep("Original (Jones & DuVal)", nrow(o_orig_corr$data)),
    rep("New Extraction (Santos)", nrow(o_new_corr$data)),
    rep("Original (Jones & DuVal)", nrow(o_orig_corr$data))
  )
  class(sub_or_corr) <- c("orchard", "data.frame")

  p_or_corr <- orchaRd::orchard_plot(sub_or_corr, xlab = "Log Odds Ratio (lnOR) (Bias Corrected)", flip = TRUE)

  p_or_corr$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
  p_or_corr$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

  p_or_corr <- p_or_corr +
    ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal)" = "#0072B2"), name = "Data Source") +
    ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal)" = "#0072B2"), name = "Data Source") +
    ggplot2::ggtitle("Overall Synthesis — Log Odds Ratio Estimates (Bias-Corrected)") +
    ggplot2::theme(
      legend.position = c(0.98, 0.02),
      legend.justification = c("right", "bottom"),
      legend.background = ggplot2::element_rect(fill = scales::alpha("white", 0.75), color = "grey90"),
      legend.box = "vertical",
      legend.margin = ggplot2::margin(t = 4, r = 4, b = 4, l = 4),
      legend.title = ggplot2::element_text(size = 9, face = "bold"),
      legend.text = ggplot2::element_text(size = 8)
    )

  p_or_corr <- add_ci_labels(p_or_corr, sub_or_corr)
  print(p_or_corr)
}
```

###### Side-by-Side Comparison

```{r}
#| label: or_comparison_plot
#| message: false
#| warning: false
#| fig.width: 14
#| fig.height: 8

if (exists("p_or") && exists("p_or_corr")) {
  comparison_or <- (p_or + p_or_corr) + 
    patchwork::plot_annotation(
      title = "Comparison of Uncorrected vs. Bias-Corrected Synthesis (Odds Ratio)",
      subtitle = "Bias-corrected estimates are conditioned on study variance = 0 (infinite precision)",
      theme = ggplot2::theme(
        plot.title = ggplot2::element_text(size = 16, face = "bold", hjust = 0.5),
        plot.subtitle = ggplot2::element_text(size = 12, italic = TRUE, hjust = 0.5)
      )
    )
  print(comparison_or)
}
```

###### Detailed Model Statistics

::: {.panel-tabset}

####### New Only

::: {.panel-tabset}

######## Uncorrected
```{r}
#| label: or_new_stats
if (!is.null(res_or_new)) {
  print(summary(res_or_new$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_or_new$model))
}
```

######## Bias-Corrected
```{r}
#| label: or_new_corr_stats
if (!is.null(res_or_new_corr)) {
  print(summary(res_or_new_corr$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_or_new_corr$model))
}
```

:::

####### Original Only

::: {.panel-tabset}

######## Uncorrected
```{r}
#| label: or_orig_stats
if (!is.null(res_or_orig)) {
  print(summary(res_or_orig$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_or_orig$model))
}
```

######## Bias-Corrected
```{r}
#| label: or_orig_corr_stats
if (!is.null(res_or_orig_corr)) {
  print(summary(res_or_orig_corr$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_or_orig_corr$model))
}
```

:::

####### Combined

::: {.panel-tabset}

######## Uncorrected
```{r}
#| label: or_comb_stats
if (!is.null(res_or_comb)) {
  print(summary(res_or_comb$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_or_comb$model))
}
```

######## Bias-Corrected
```{r}
#| label: or_comb_corr_stats
if (!is.null(res_or_comb_corr)) {
  print(summary(res_or_comb_corr$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_or_comb_corr$model))
}
```

:::

:::

:::

:::

#### Sensitivity Analysis: Impact of the VCV Structure

To evaluate the robustness of our meta-analytic results, we conduct a sensitivity analysis comparing the main multilevel models (which incorporate a Variance-Covariance [VCV] matrix with an assumed within-study correlation $\rho = 0.5$) against simplified multilevel models that ignore this correlation (i.e., treating effect size variances as independent diagonal elements, $V = v_i$).

```{r}
#| label: vcv_sensitivity_analysis
#| message: false
#| warning: false

# 1. Define comparison function
compare_vcv_models <- function(dat, name, yi = "yi", vi = "vi", cluster_var = "identifierStudyId", obs_var = "identifierEffectSizeID", rho = 0.5) {
  dat2 <- dat %>%
    filter(!is.na(.data[[yi]]), !is.na(.data[[vi]]), .data[[vi]] > 0)
  
  if (nrow(dat2) == 0) return(NULL)
  
  # A. With VCV (rho = 0.5)
  VCV <- metafor::vcalc(
    vi = dat2[[vi]],
    cluster = dat2[[cluster_var]],
    obs = dat2[[obs_var]],
    rho = rho,
    data = dat2
  )
  mod_w_vcv <- metafor::rma.mv(
    yi = dat2[[yi]],
    V  = VCV,
    random = list(
      ~1 | identifierStudyId,
      ~1 | identifierEffectSizeID,
      ~1 | taxonomySpecies
    ),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = dat2
  )
  
  # B. Without VCV (diagonal V = vi)
  mod_no_vcv <- metafor::rma.mv(
    yi = dat2[[yi]],
    V  = dat2[[vi]],
    random = list(
      ~1 | identifierStudyId,
      ~1 | identifierEffectSizeID,
      ~1 | taxonomySpecies
    ),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = dat2
  )
  
  # Combine results
  data.frame(
    Analysis = name,
    Model_Type = c("With VCV (rho = 0.5)", "Without VCV (Independent)"),
    k = c(mod_w_vcv$k, mod_no_vcv$k),
    Estimate = c(mod_w_vcv$b[1], mod_no_vcv$b[1]),
    SE = c(mod_w_vcv$se[1], mod_no_vcv$se[1]),
    CI_Lower = c(mod_w_vcv$ci.lb[1], mod_no_vcv$ci.lb[1]),
    CI_Upper = c(mod_w_vcv$ci.ub[1], mod_no_vcv$ci.ub[1]),
    p_val = c(mod_w_vcv$pval[1], mod_no_vcv$pval[1])
  )
}

# 2. Run comparisons across all subsets
comp_hedges_new  <- compare_vcv_models(db_hedges %>% filter(Dataset_Source == "New Extraction (Santos)"), "Hedges' g (New Only)")
comp_hedges_orig <- compare_vcv_models(db_hedges %>% filter(Dataset_Source == "Original (Davies 2020)"), "Hedges' g (Original Only)")
comp_hedges_comb <- compare_vcv_models(db_hedges, "Hedges' g (Combined)")

comp_or_new  <- compare_vcv_models(db_or %>% filter(Dataset_Source == "New Extraction (Santos)"), "Log OR (New Only)")
comp_or_orig <- compare_vcv_models(db_or %>% filter(Dataset_Source == "Original (Jones & DuVal 2019)"), "Log OR (Original Only)")
comp_or_comb <- compare_vcv_models(db_or, "Log OR (Combined)")

# 3. Bind all comparisons together
vcv_comparison_results <- bind_rows(
  comp_hedges_new, comp_hedges_orig, comp_hedges_comb,
  comp_or_new, comp_or_orig, comp_or_comb
)
```

##### Sensitivity Comparison Table

The interactive table below compares overall pooled estimates, standard errors (SE), 95% Confidence Intervals, and p-values for both models across all subsets.

```{r}
#| label: vcv_comparison_table
#| echo: false

DT::datatable(
  vcv_comparison_results,
  options = list(
    pageLength = 12,
    dom = 't',
    columnDefs = list(list(className = 'dt-center', targets = "_all"))
  ),
  rownames = FALSE
) %>%
  DT::formatRound(columns = c("Estimate", "SE", "CI_Lower", "CI_Upper"), digits = 4) %>%
  DT::formatSignif(columns = c("p_val"), digits = 4)
```

##### Key Methodological Findings

1. **Marginal Shifts in Pooled Estimates:** Treating effect sizes as completely independent (Without VCV) causes very minor changes to the overall pooled estimates. For the combined dataset, the Hedges' $g$ estimate shifts from $0.1526$ to $0.1473$, while the Log Odds Ratio (lnOR) estimate shifts from $0.6147$ to $0.6530$. This indicates that the main directional conclusions of the meta-analyses are robust to VCV specification.
2. **Standard Error Shifts:** Omitting the VCV matrix and assuming independence leads to a *slight inflation* of standard errors and wider confidence intervals for the combined analyses:
   * **Hedges' $g$ (Combined):** SE increases from $0.0780$ to $0.0816$.
   * **Log Odds Ratio (Combined):** SE increases from $0.1388$ to $0.1494$.
   
   This demonstrates that accounting for known within-study covariance via `vcalc` yields more precise, correctly weighted estimates.
3. **Impact on Statistical Significance:** The marginal significance test for continuous outcomes (Hedges' $g$ Combined) is sensitive to this specification. With the VCV matrix, the effect size is marginally non-significant ($p = 0.0517$). When omitting the VCV matrix, the $p$-value increases further to $p = 0.0724$. Conversely, binary outcomes (Log OR Combined) remain highly statistically significant ($p < 0.0001$) under both models.

---

### Publication Bias & Selective Reporting Diagnostics {#sec-pub-bias-results-new}

To thoroughly evaluate the presence of publication bias and report-level selection in our Mate-Choice Copying (MCC) dataset, we employ three complementary diagnostic approaches (Visual Funnel Plots, Egger's Meta-Regression for small-study effects, and Time-Lag Meta-Regression for Decline Effects).

#### A. Hedges' $g$ (Continuous Outcomes)

##### 1. Visual Inspection (Funnel Plot)

We plot continuous Hedges' $g$ estimates against their precision (inverse of the standard error, $1/\text{SE}$) to check for visual asymmetry around the overall mean estimate.

```{r}
#| label: hedges_baseline_and_funnel
#| message: false
#| warning: false
#| fig.width: 9
#| fig.height: 7

# Intercept-only model
VCV_hedges <- metafor::vcalc(vi = vi, cluster = identifierStudyId, obs = identifierEffectSizeID, rho = 0.5, data = db_hedges)

ma_hedges <- metafor::rma.mv(
    yi = yi,
    V = VCV_hedges,
    random = list(~ 1 | identifierStudyId, ~ 1 | identifierEffectSizeID, ~ 1 | taxonomySpecies),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = db_hedges
)

# Funnel plot
metafor::funnel(
    ma_hedges,
    yaxis = "seinv",
    las = 1,
    digits = c(2, 1),
    xlab = "Standardised Hedges' g Effect Size",
    ylab = "Precision (Inverse of SE)",
    pch = 21,
    bg = alpha("black", 0.2),
    col = "black",
    cex = 1.2,
    main = "Funnel Plot for Continuous Outcomes (Hedges' g)"
)
```

##### 2. Egger's Meta-Regression (Small-Study Effects)

To perform Egger's regression on Hedges' $g$, we follow the two-step framework recommended by Nakagawa et al. [-@nakagawaMethodsTestingPublication2022; -@Corrigendum2022] and Stanley and Doucouliagos [-@stanleyMetaregressionAnalysisEconomics2012; -@stanleyMetaregressionApproximationsReduce2014]. We use the sampling standard error ($\text{SE} = \sqrt{v_i}$) as our precision proxy.

Because standard error model intercepts can be downwardly biased if they are significant, we apply a two-step correction: if the intercept of the standard error model is statistically significant, we re-fit the Egger's regression model using the sampling variance ($\text{V} = v_i$) as the moderator to obtain a less biased estimate of the adjusted overall meta-analytic mean.

```{r}
#| label: hedges_eggers_regression
#| message: false
#| warning: false

# Calculate standard error precision proxy
db_hedges_egger <- db_hedges %>%
    mutate(se = sqrt(vi)) %>%
    filter(!is.na(se))

# Re-estimate VCV on this subset
VCV_hedges_egger <- metafor::vcalc(vi = vi, cluster = identifierStudyId, obs = identifierEffectSizeID, rho = 0.5, data = db_hedges_egger)

# Step 1: Egger regression model with standard error as moderator
egger_hedges <- metafor::rma.mv(
    yi = yi,
    V = VCV_hedges_egger,
    mods = ~ 1 + se,
    random = list(~ 1 | identifierStudyId, ~ 1 | identifierEffectSizeID, ~ 1 | taxonomySpecies),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = db_hedges_egger
)

summary(egger_hedges)

# Step 2: Since the intercept is significant, re-fit using variance (vi)
egger_hedges_v <- metafor::rma.mv(
    yi = yi,
    V = VCV_hedges_egger,
    mods = ~ 1 + vi,
    random = list(~ 1 | identifierStudyId, ~ 1 | identifierEffectSizeID, ~ 1 | taxonomySpecies),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = db_hedges_egger
)

summary(egger_hedges_v)
```

We visualize the Egger regression using a bubble plot. Effect sizes are plotted as circles scaled by precision, with the predicted regression line showing the relationship.

```{r}
#| label: hedges_egger_bubble_plot
#| message: false
#| warning: false
#| fig.width: 9
#| fig.height: 6

p_egger_hedges <- orchaRd::bubble_plot(
    egger_hedges,
    mod = "se",
    group = "identifierStudyId",
    xlab = "Standard Error of Hedges' g (se)",
    ylab = "Effect size (Hedges' g)"
)

# Color points by real source
p_egger_hedges$layers[[1]]$data$real_source <- db_hedges_egger$Dataset_Source
p_egger_hedges$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
p_egger_hedges$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

p_egger_hedges <- p_egger_hedges +
    ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
    ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
    ggplot2::ggtitle("Egger's Meta-Regression (Hedges' g)") +
    ggplot2::theme_minimal()

print(p_egger_hedges)
```

> [!NOTE]
> The slope for standard error (`se`) is **highly significant** ($\beta = 4.3353$, $SE = 0.5786$, $t = 7.4929$, $p < 0.0001$), indicating extremely clear evidence of small-study effects. Studies with larger standard errors tend to report larger, more positive Hedges' $g$ values.
> 
> **Two-Step Correction:** Because the intercept of the standard error model is statistically significant ($p < 0.0001$), the intercept provides a downwardly biased estimate of the adjusted mean. Following the two-step recommendation of Nakagawa et al. (2022a) and Stanley (2012, 2014), we re-fit Egger's regression using the sampling variance (`vi`) as the moderator. In this corrected model, the adjusted overall meta-analytic mean is estimated at $\beta_0 = -0.2040$ ($SE = 0.1017$, $t = -2.0062$, $p = 0.0461$).

##### 3. Decline Effect (Time-Lag Bias)

We test whether effect sizes shrink over time by adding publication year (mean-centered) to our models while simultaneously controlling for study precision.

```{r}
#| label: hedges_time_lag
#| message: false
#| warning: false

# Mean-center publication year
db_hedges_decline <- db_hedges_egger %>%
    mutate(Year_c = Year - mean(Year, na.rm = TRUE))

# VCV for decline subset
VCV_hedges_decline <- metafor::vcalc(vi = vi, cluster = identifierStudyId, obs = identifierEffectSizeID, rho = 0.5, data = db_hedges_decline)

# Time-lag decline model controlling for standard error (se)
decline_hedges <- metafor::rma.mv(
    yi = yi,
    V = VCV_hedges_decline,
    mods = ~ 1 + Year_c + se,
    random = list(~ 1 | identifierStudyId, ~ 1 | identifierEffectSizeID, ~ 1 | taxonomySpecies),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = db_hedges_decline
)

summary(decline_hedges)
```

We visualize the Decline Effect using a bubble plot.

```{r}
#| label: hedges_decline_bubble_plot
#| message: false
#| warning: false
#| fig.width: 9
#| fig.height: 6

p_decline_hedges <- orchaRd::bubble_plot(
    decline_hedges,
    mod = "Year_c",
    group = "identifierStudyId",
    xlab = "Publication Year (centered)",
    ylab = "Effect size (Hedges' g)"
)

# Color points by real source
p_decline_hedges$layers[[1]]$data$real_source <- db_hedges_decline$Dataset_Source
p_decline_hedges$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
p_decline_hedges$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

p_decline_hedges <- p_decline_hedges +
    ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
    ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source")

# Relabel centered x-axis to actual calendar years
mean_year_hedges <- mean(db_hedges_egger$Year, na.rm = TRUE)
target_years_hedges <- c(2000, 2005, 2010, 2015, 2020, 2025)

p_decline_hedges_relabeled <- p_decline_hedges +
    ggplot2::scale_x_continuous(
        name = "Publication Year",
        breaks = target_years_hedges - mean_year_hedges,
        labels = target_years_hedges
    ) +
    ggplot2::ggtitle("Decline Effect / Time-Lag Bias (Hedges' g)") +
    ggplot2::theme_minimal()

print(p_decline_hedges_relabeled)
```

> [!NOTE]
> The publication year slope is **non-significant** ($\beta = -0.0009$, $SE = 0.0069$, $t = -0.1263$, $p = 0.8996$), indicating no statistically significant decline effect over time in continuous outcomes.

##### 4. Unified Diagnostic Dashboard (Hedges' $g$)

We assemble the continuous outcome diagnostics into a single dashboard.

```{r}
#| label: hedges_unified_dashboard
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 12

dashboard_hedges <- (p_egger_hedges / p_decline_hedges_relabeled) +
    patchwork::plot_annotation(
        title = "Publication Bias Diagnostic Dashboard — Hedges' g",
        tag_levels = "A"
    ) &
    ggplot2::theme(
        plot.title = ggplot2::element_text(size = 14, face = "bold", hjust = 0.5),
        plot.tag = ggplot2::element_text(size = 12, face = "bold")
    )

print(dashboard_hedges)
```

---

#### B. Odds Ratio (Binary Outcomes)

##### 1. Visual Inspection (Funnel Plot)

For Odds Ratio studies, we inspect the distribution of Log Odds Ratios against precision.

```{r}
#| label: or_baseline_and_funnel
#| message: false
#| warning: false
#| fig.width: 9
#| fig.height: 7

# Intercept-only model
VCV_or <- metafor::vcalc(vi = vi, cluster = identifierStudyId, obs = identifierEffectSizeID, rho = 0.5, data = db_or)

ma_or <- metafor::rma.mv(
    yi = yi,
    V = VCV_or,
    random = list(~ 1 | identifierStudyId, ~ 1 | identifierEffectSizeID, ~ 1 | taxonomySpecies),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = db_or
)

summary(ma_or)

# Funnel plot
metafor::funnel(
    ma_or,
    yaxis = "seinv",
    las = 1,
    digits = c(2, 1),
    xlab = "Standardised Log Odds Ratio (lnOR)",
    ylab = "Precision (Inverse of SE)",
    pch = 21,
    bg = alpha("black", 0.2),
    col = "black",
    cex = 1.2,
    main = "Funnel Plot for Binary Outcomes (Log Odds Ratio)"
)
```

##### 2. Egger's Meta-Regression (Small-Study Effects)

For Odds Ratio studies, we use the standard error ($\text{SE} = \sqrt{v_i}$) as our precision proxy in the two-step framework. Because the intercept of the standard error model is statistically significant ($p = 0.0495$), we perform Step 2 of the two-step correction, re-fitting the Egger's regression model using the sampling variance ($\text{V} = vi$) as the moderator.

```{r}
#| label: or_eggers_regression
#| message: false
#| warning: false

# Calculate standard error proxy
db_or_egger <- db_or %>%
    mutate(se = sqrt(vi)) %>%
    filter(!is.na(se))

# Re-estimate VCV on this subset
VCV_or_egger <- metafor::vcalc(vi = vi, cluster = identifierStudyId, obs = identifierEffectSizeID, rho = 0.5, data = db_or_egger)

# Step 1: Egger regression model with standard error as moderator
egger_or <- metafor::rma.mv(
    yi = yi,
    V = VCV_or_egger,
    mods = ~ 1 + se,
    random = list(~ 1 | identifierStudyId, ~ 1 | identifierEffectSizeID, ~ 1 | taxonomySpecies),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = db_or_egger
)

summary(egger_or)

# Step-2: Since the intercept is significant, re-fit using variance (vi)
egger_or_v <- metafor::rma.mv(
    yi = yi,
    V = VCV_or_egger,
    mods = ~ 1 + vi,
    random = list(~ 1 | identifierStudyId, ~ 1 | identifierEffectSizeID, ~ 1 | taxonomySpecies),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = db_or_egger
)

summary(egger_or_v)
```

We visualize the Egger regression for Odds Ratio using a bubble plot.

```{r}
#| label: or_egger_bubble_plot
#| message: false
#| warning: false
#| fig.width: 9
#| fig.height: 6

p_egger_or <- orchaRd::bubble_plot(
    egger_or,
    mod = "se",
    group = "identifierStudyId",
    xlab = "Standard Error of Log Odds Ratio (se)",
    ylab = "Effect size (lnOR)"
)

# Color points by real source
p_egger_or$layers[[1]]$data$real_source <- db_or_egger$Dataset_Source
p_egger_or$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
p_egger_or$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

p_egger_or <- p_egger_or +
    ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal 2019)" = "#0072B2"), name = "Data Source") +
    ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal 2019)" = "#0072B2"), name = "Data Source") +
    ggplot2::ggtitle("Egger's Meta-Regression (Log Odds Ratio)") +
    ggplot2::theme_minimal()

print(p_egger_or)
```

> [!NOTE]
> The slope for the standard error is **extremely significant** ($\beta = 2.1129$, $SE = 0.4471$, $t = 4.7257$, $p < 0.0001$), indicating highly pronounced small-study effects in binary mate-choice copying outcomes. Studies with higher standard errors (smaller precision) report significantly larger log odds ratios.
> 
> **Two-Step Correction:** Because the intercept in the standard error model is statistically significant ($p = 0.0495$), we re-fit the model using the sampling variance (`vi`) as the moderator to obtain a less biased estimate of the adjusted mean. In this corrected model, the adjusted overall log odds ratio is non-significant at $\beta_0 = 0.1139$ ($SE = 0.1627$, $t = 0.7003$, $p = 0.4847$).

##### 3. Decline Effect (Time-Lag Bias)

We test whether binary effect sizes shrink over time by adding centered publication year to our model while controlling for study standard error.

```{r}
#| label: or_time_lag
#| message: false
#| warning: false

# Mean-center publication year
db_or_decline <- db_or_egger %>%
    mutate(Year_c = Year - mean(Year, na.rm = TRUE))

# VCV for decline subset
VCV_or_decline <- metafor::vcalc(vi = vi, cluster = identifierStudyId, obs = identifierEffectSizeID, rho = 0.5, data = db_or_decline)

# Time-lag decline model
decline_or <- metafor::rma.mv(
    yi = yi,
    V = VCV_or_decline,
    mods = ~ 1 + Year_c + se,
    random = list(~ 1 | identifierStudyId, ~ 1 | identifierEffectSizeID, ~ 1 | taxonomySpecies),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = db_or_decline
)

summary(decline_or)
```

We visualize the Decline Effect for Odds Ratios using a bubble plot.

```{r}
#| label: or_decline_bubble_plot
#| message: false
#| warning: false
#| fig.width: 9
#| fig.height: 6

p_decline_or <- orchaRd::bubble_plot(
    decline_or,
    mod = "Year_c",
    group = "identifierStudyId",
    xlab = "Publication Year (centered)",
    ylab = "Effect size (lnOR)"
)

# Color points by real source
p_decline_or$layers[[1]]$data$real_source <- db_or_decline$Dataset_Source
p_decline_or$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
p_decline_or$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

p_decline_or <- p_decline_or +
    ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal 2019)" = "#0072B2"), name = "Data Source") +
    ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal 2019)" = "#0072B2"), name = "Data Source")

# Relabel centered x-axis to actual calendar years
mean_year_or <- mean(db_or_egger$Year, na.rm = TRUE)
target_years_or <- c(2000, 2005, 2010, 2015, 2020, 2025)

p_decline_or_relabeled <- p_decline_or +
    ggplot2::scale_x_continuous(
        name = "Publication Year",
        breaks = target_years_or - mean_year_or,
        labels = target_years_or
    ) +
    ggplot2::ggtitle("Decline Effect / Time-Lag Bias (Log Odds Ratio)") +
    ggplot2::theme_minimal()

print(p_decline_or_relabeled)
```

##### 4. Unified Diagnostic Dashboard (Odds Ratio)

We assemble the binary outcome diagnostics into a single dashboard.

```{r}
#| label: or_unified_dashboard
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 12

dashboard_or <- (p_egger_or / p_decline_or_relabeled) +
    patchwork::plot_annotation(
        title = "Publication Bias Diagnostic Dashboard — Log Odds Ratio",
        tag_levels = "A"
    ) &
    ggplot2::theme(
        plot.title = ggplot2::element_text(size = 14, face = "bold", hjust = 0.5),
        plot.tag = ggplot2::element_text(size = 12, face = "bold")
    )

print(dashboard_or)
```

---

#### C. Empirical Findings Summary Table

We compile the results of the small-study effect regressions and time-lag meta-regressions across both continuous and binary outcomes.

```{r}
#| label: summary_table_creation
#| echo: false

bias_summary <- data.frame(
    Effect_Size_Class = c("Hedges' g (Continuous)", "Hedges' g (Continuous)", "Log Odds Ratio (Binary)", "Log Odds Ratio (Binary)"),
    Diagnostic_Test = c("Egger's Regression (Small-Study)", "Time-Lag (Decline Effect)", "Egger's Regression (Small-Study)", "Time-Lag (Decline Effect)"),
    Moderator = c("Standard Error (se)", "Year_c (centered)", "Standard Error (se)", "Year_c (centered)"),
    Slope_Beta = c(egger_hedges$b[2], decline_hedges$b[2], egger_or$b[2], decline_or$b[2]),
    Std_Error = c(egger_hedges$se[2], decline_hedges$se[2], egger_or$se[2], decline_or$se[2]),
    t_z_value = c(egger_hedges$zval[2], decline_hedges$zval[2], egger_or$zval[2], decline_or$zval[2]),
    p_value = c(egger_hedges$pval[2], decline_hedges$pval[2], egger_or$pval[2], decline_or$pval[2]),
    Evidence_of_Bias = c("Highly Significant Bias (p < 0.0001)", "No Significant Bias (p = 0.8996)", "Highly Significant Bias (p < 0.0001)", "Significant Bias (p = 0.0104)")
)

DT::datatable(
    bias_summary,
    options = list(
        dom = "t",
        columnDefs = list(list(className = "dt-center", targets = "_all"))
    ),
    rownames = FALSE
) %>%
    DT::formatRound(columns = c("Slope_Beta", "Std_Error", "t_z_value"), digits = 4) %>%
    DT::formatSignif(columns = c("p_value"), digits = 4)
```

##### Main Takeaways

1. **Pervasive Small-Study Effects:** Both Hedges' $g$ and Log Odds Ratio datasets show statistically significant evidence of small-study effects. Smaller, less precise trials systematically report larger mate-choice copying effect sizes. This asymmetrical pattern is a classic signature of publication selection.
2. **Decline Effect in Binary Outcomes:** We found robust evidence of a decline effect in binary mate-choice copying studies ($\beta = -0.0282$, $p = 0.0104$). Over the last two decades, the strength of reported mate-choice copying has significantly decreased. This is typical in fields experiencing a "novelty effect," where initial breakthrough publications show large effect sizes, and subsequent rigorous replications or expanded testing reveal more moderate, realistic estimates.
3. **Continuous Outcomes Show More Stable Trends Over Time:** For Hedges' $g$ (continuous), although small-study effects are present, the publication year trend is non-significant ($\beta = -0.0009$, $p = 0.8996$). This suggests that the continuous reporting scales have been relatively more stable in magnitude over the history of the literature.

These diagnostic results suggest that while mate-choice copying is a genuine ecological phenomenon, the overall raw pooled estimates from both continuous and binary literatures may be moderately inflated due to selective reporting and publication biases. Future research should prioritize large-sample pre-registered replication studies to provide unbiased baselines.

---

## Repository Structure {#sec-repo}

```
mate_choice_meta/
├── index.qmd                          ← this document (source)
├── index.html                         ← rendered GitHub Pages output
├── mate_choice_meta.Rproj
├── data/
│   ├── 0_literature_screening/
│   │   └── combined_records_final.csv   (2,083 records, all years)
│   ├── 1_effect_size_calculation_pipeline/
│   │   └── Data Extraction Mate Choice Meta Analysis.xlsx (raw data & outputs)
│   └── 2_original_meta_analyses_datasets/
│       ├── Davies_et_al_2020_Final_data.xlsx
│       └── Jones_DuVal_2019_data.CSV
├── scripts/
│   ├── 0_literature_screening/
│   │   └── 1_deduplicate_records.R
│   └── 1_effect_size_calculation_pipeline/
│       ├── calculate_effect_sizes.py    (Hedges' g & OR calculator)
│       ├── overall_effect.qmd           (overall effect size multilevel model)
│       └── publication_bias.qmd         (publication bias diagnostics)
└── builds/
    ├── 0_literature_screening/
    │   └── deduplicated_and_randomly_split/
    │       ├── Aleksandra.csv  … Shinichi.csv  (15 primary screeners)
    │       ├── Eduardo.csv                     (validator)
    │       └── Losia.csv                       (validator)
    └── 1_effect_size_calculation_pipeline/
        ├── notebookLM_pipeline.md
        └── notebookLM_pipeline.html
```

### Rendering this document

To regenerate `index.html` after editing `index.qmd`, run from the repo root:

```bash
quarto render index.qmd
```

Requires [Quarto CLI](https://quarto.org/docs/get-started/) ≥ 1.4 and R with the
packages listed in the setup chunk.

---

## Reproducibility & Contact {#sec-reproducibility}

All data processing is fully scripted in R. The random shuffle used for screener
assignment is not seeded; instead, the exported per-screener CSV files in
`builds/0_literature_screening/deduplicated_and_randomly_split/` serve as the
reproducible record of the assignment actually used.

The AI extraction prompts are version-controlled in
`builds/1_effect_size_calculation_pipeline/notebookLM_pipeline.md` (version 1.1).

For questions about the data pipeline or extraction protocol, please open an issue
on the GitHub repository.

---

*Document generated with [Quarto](https://quarto.org). Last updated: `r format(Sys.Date(), "%B %Y")`.*
