---
title: "Literature Screening & Processing Pipeline"
subtitle: "Methodological framework and protocol for mate-choice copying"
author: "Eduardo S.A. Santos & Collaborators"
format:
  html:
    toc: true
    toc-depth: 3
    theme: cosmo
    self-contained: true
execute:
  echo: true
  eval: false
bibliography: references.bib
---

```{r}
#| label: setup
#| include: false
#| eval: true
library(dplyr)
library(data.table)
library(stringr)
library(readr)
library(knitr)
library(kableExtra)
library(readxl)
library(metafor)
library(orchaRd)
library(patchwork)
library(DT)
library(scales)
library(ggplot2)
```

---

## Overview {#sec-overview}

This page documents all data decisions, processing steps, and planned analyses for the
systematic review and **meta-analysis of mate-choice copying (MCC) update** across animal taxa.

Mate-choice copying occurs when an individual's preference for a potential mate is
influenced by observing the mate choices of others — a form of socially acquired
preference that has been documented in fish, birds, mammals, and invertebrates.

This meta-analysis aims to:

1. Estimate the overall effect size for MCC across species.
2. Quantify between-study heterogeneity.
3. Test for moderating variables (sex, taxon, experimental design).

::: {.callout-note}
## Project status
As of this version, the literature screening phase is concluded. This page has been updated
with each phase of the pipeline.
:::

### Key numbers at a glance

| Metric | Value |
|---|---|
| Raw records (combined databases) | 2,083 |
| Records after deduplication & year filter (≥ 2019) | 495 |
| Primary screeners | 15 |
| Secondary validators | 2 |
| AI-assisted extraction steps | 4 |
| Top journals represented | Animal Behaviour, Behavioral Ecology, PRSB |

---

## Data Sources {#sec-data-sources}

### Literature search

Records were retrieved from two major bibliographic databases and combined in
**Rayyan** format for import into R:

- **Scopus** — full-text and abstract search using Boolean terms covering mate choice,
  copying, non-independent mate selection, and social information use.
- **Web of Science (WoS)** — parallel search with identical query structure.

The two exports were merged into a single file:
`data/0_literature_screening/combined_records_final.csv`

### Master dataset fields

```{r}
#| label: tbl-fields
#| tbl-cap: "Fields in combined_records_final.csv"
#| eval: true
fields <- tibble::tribble(
    ~Field,        ~Description,                                          ~Completeness,
    "title",       "Article title (primary deduplication key)",           "~100%",
    "doi",         "Digital Object Identifier (secondary dedup key)",     "~80%",
    "abstract",    "Full abstract text",                                  "~90%",
    "authors",     "Author list",                                         "~95%",
    "year",        "Publication year",                                    "~100%",
    "journal",     "Journal name",                                        "~100%",
    "source_file", "Which database CSV contributed this record",          "100%",
    "pubmed_id",   "PubMed identifier",                                   "Partial",
    "pmc_id",      "PubMed Central identifier",                           "Partial"
)

kable(fields, align = "lll") |>
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

::: {.callout-important}
## Provenance tracking
The `source_file` field is preserved through every deduplication step and is
concatenated (not lost) when duplicate records from different databases are merged.
In the final output it is also appended to the `abstract` field, so that the complete
database-of-origin audit trail survives import into Rayyan or any other screening tool.
:::

---

## Processing Pipeline {#sec-pipeline}

All deduplication and screener-assignment logic lives in a single reproducible script:

> `scripts/0_literature_screening/1_deduplicate_records.R`

The script was written in **February 2026** using `data.table` for efficient in-memory
processing of the 2,083-record dataset.

### Step 1 — Load data

```{r}
#| label: load-data
#| eval: false
rm(list = ls())

library(dplyr)
library(data.table)
library(stringr)
library(readr)

# Rayyan estimated 17,812 records have duplicates across the raw exports
articles <- fread("data/0_literature_screening/combined_records_final.csv")
```

### Step 2 — Title-based deduplication {#sec-title-dedup}

Titles are normalised to a canonical lowercase form before matching.
When duplicates are found, the **highest-quality** record is retained.

**Quality score** (0–3): one point each for having a non-empty abstract, a non-empty
author list, and a non-empty DOI.

```{r}
#| label: title-dedup
#| eval: false
# --- Normalise title ----------------------------------------------------------
articles[, title_lwr := str_to_lower(title)]
articles[, title_lwr := gsub("[[:punct:]]", " ", title_lwr)] # strip punctuation
articles[, title_lwr := gsub(" {2,}", " ", title_lwr)] # collapse whitespace
articles[, title_lwr := trimws(title_lwr)]

# --- Quality score (0–3) ------------------------------------------------------
articles[, has_abstract := ifelse(abstract == "" | is.na(abstract), 0L, 1L)]
articles[, has_authors := ifelse(authors == "" | is.na(authors), 0L, 1L)]
articles[, has_doi := ifelse(doi == "" | is.na(doi), 0L, 1L)]
articles[, quality := has_abstract + has_authors + has_doi]

# --- Concatenate source files per title group ---------------------------------
sources <- articles[,
    .(source_file = paste(sort(unique(source_file)), collapse = "; ")),
    by = .(title_lwr)
]

# --- Keep highest-quality record per title ------------------------------------
articles.title.dedupe <- articles[
    , .SD[which.max(quality), !c("source_file"), with = FALSE],
    by = .(title_lwr)
]

# --- Merge concatenated sources back in ---------------------------------------
articles.title.dedupe.mrg <- merge(
    articles.title.dedupe, sources,
    by = "title_lwr", all.x = TRUE
)
```

#### Data decisions — title deduplication

| Decision | What was done | Rationale |
|---|---|---|
| Lowercase conversion | `str_to_lower()` | Catches case differences across databases |
| Punctuation removal | `gsub("[[:punct:]]", " ", ...)` | Handles hyphens, apostrophes, colons in titles |
| Whitespace collapse | `gsub(" {2,}", " ", ...)` | Normalises after punctuation removal |
| Quality scoring | abstract + authors + doi (0/1 each) | Retains the most information-rich duplicate |
| Source concatenation | `paste(sort(unique(source_file)), collapse = "; ")` | Full provenance preserved across all duplicates |

### Step 3 — DOI-based deduplication {#sec-doi-dedup}

A second pass on the DOI field catches articles in which titles differ slightly between
database exports but resolve to the same paper. A custom helper function
`split_conc()` uniquely concatenates source strings that were themselves already
concatenated in the title-dedup step.

```{r}
#| label: doi-dedup
#| eval: false
# Helper: split already-concatenated source strings, deduplicate, re-join
split_conc <- function(x) {
    str_split(x, pattern = "; ") |>
        unlist() |>
        unique() |>
        sort() |>
        paste(collapse = "; ")
}

# Recalculate quality after title dedup (fields may have changed)
articles.title.dedupe.mrg[, has_abstract := ifelse(abstract == "" | is.na(abstract), 0L, 1L)]
articles.title.dedupe.mrg[, has_authors := ifelse(authors == "" | is.na(authors), 0L, 1L)]
articles.title.dedupe.mrg[, has_doi := ifelse(doi == "" | is.na(doi), 0L, 1L)]
articles.title.dedupe.mrg[, quality := has_abstract + has_authors + has_doi]

# Concatenate sources per DOI
sources <- articles.title.dedupe.mrg[,
    .(source_file = split_conc(source_file)),
    by = .(doi)
]

# Keep highest-quality record per DOI
articles.doi.dedupe <- articles.title.dedupe.mrg[
    , .SD[which.max(quality), !c("source_file"), with = FALSE],
    by = .(doi)
]

# Merge sources back in
articles.final <- merge(
    articles.doi.dedupe, sources,
    by = "doi", all.x = TRUE
)
```

### Step 4 — Clean and append provenance to abstract {#sec-clean}

Intermediate processing columns are dropped, and the `source_file` string is appended
to the `abstract` field so that provenance survives any downstream tool that does not
retain a `source_file` column.

```{r}
#| label: clean
#| eval: false
articles.final[, abstract := paste0(abstract, " | ", source_file)]

articles.final <- articles.final[
    ,
    !c(
        "title_lwr", "n_title_matches",
        "has_abstract", "has_authors", "has_doi",
        "quality", "n_doi_matches", "source_file"
    )
]
```

### Step 5 — Year filter: 2019 and later {#sec-year-filter}

```{r}
#| label: year-filter
#| eval: false
articles.final <- articles.final[year >= 2019, ]
message("Records after filter: ", nrow(articles.final))
message("Earliest year in dataset: ", min(articles.final$year))
```

::: {.callout-important}
## Data decision — year filter
Only records published **in or after 2019** are retained for screening.

**Rationale:**

- Aligns with the pre-registered scope of the review.
- Pre-2019 records remain available in `combined_records_final.csv` and can be
  incorporated in sensitivity analyses.

The filter reduces the deduplicated pool to **495 articles**.
:::

### Step 6 — Random shuffle and screener assignment {#sec-assignment}

Articles are randomly shuffled before assignment to prevent ordering artefacts (e.g.,
alphabetical or chronological clustering within screener batches). A dynamic function
generates near-equal assignment vectors for any team size.

```{r}
#| label: assignment
#| eval: false
# Team structure
screeners <- c(
    "Aleksandra", "Aneta", "Anna", "Ayumi", "Cassidy", "Christine",
    "Erick", "Iwo", "Hao", "Jimuel", "Kyle", "Marija",
    "Santiago", "Sergio", "Shinichi", # 15 primary screeners
    "Losia", "Eduardo" # 2 secondary validators
)

primary_list <- setdiff(screeners, c("Eduardo", "Losia"))
secondary_list <- c("Eduardo", "Losia")

# Random shuffle (prevents date/alpha ordering bias)
set.seed(NULL) # no fixed seed — shuffled output files are the reproducible record
articles.final <- articles.final[sample(.N), ]

# Dynamic assignment: ceiling(total / n) reps, trimmed to total rows
create_assignments <- function(names, total_rows) {
    reps_needed <- ceiling(total_rows / length(names))
    full_vector <- rep(names, each = reps_needed)
    return(full_vector[seq_len(total_rows)])
}

n_total <- nrow(articles.final)

# Primary: ~33 articles per screener (495 / 15)
articles.final[, primary_screener := create_assignments(primary_list, n_total)]

# Secondary: ~248 articles per validator (495 / 2)
articles.final[, secondary_screener := create_assignments(secondary_list, n_total)]

# Verify balance
articles.final[, .(n = .N), by = .(primary_screener)]
articles.final[, .(n = .N), by = .(secondary_screener)]
```

### Step 7 — Export per-screener CSVs {#sec-export}

One CSV is written per primary screener to
`builds/0_literature_screening/deduplicated_and_randomly_split/`.
Each file retains both assignment columns for cross-referencing.

```{r}
#| label: export
#| eval: false
invisible(lapply(
    unique(articles.final$primary_screener),
    function(x) {
        write_csv(
            articles.final[primary_screener == x, ],
            paste0("builds/0_literature_screening/deduplicated_and_randomly_split/", x, ".csv")
        )
    }
))
```

---

## Literature Screening {#sec-screening}

### Team structure

The 495 articles are distributed across a 17-person team. Each article receives:

- **One primary screen** — eligibility assessment against pre-registered inclusion criteria.
- **One independent secondary validation** — every article is also assigned to one of
  two validators (Eduardo or Losia), enabling inter-rater reliability estimation.

```{r}
#| label: tbl-team
#| tbl-cap: "Screener assignments"
#| eval: true
team <- tibble::tribble(
    ~Screener, ~Role, ~Articles,
    "Aleksandra", "Primary", "~33",
    "Aneta", "Primary", "~33",
    "Anna", "Primary", "~33",
    "Ayumi", "Primary", "~33",
    "Cassidy", "Primary", "~33",
    "Christine", "Primary", "~33",
    "Erick", "Primary", "~33",
    "Hao", "Primary", "~33",
    "Iwo", "Primary", "~33",
    "Jimuel", "Primary", "~33",
    "Kyle", "Primary", "~33",
    "Marija", "Primary", "~33",
    "Santiago", "Primary", "~33",
    "Sergio", "Primary", "~33",
    "Shinichi", "Primary", "~33",
    "Eduardo", "Validator", "~248",
    "Losia", "Validator", "~248"
)

kable(team, align = "lll") |>
    kable_styling(bootstrap_options = c("striped", "hover", "condensed")) |>
    row_spec(16:17, bold = TRUE, background = "#fdf4e3")
```

### Inclusion criteria

Articles are included if they:

1. Report an **experimental** test of mate-choice copying (observer watches a demonstrator
   make a mate choice, then has the opportunity to choose between the same mates).
2. Provide sufficient quantitative data for effect-size extraction (means + variance, or
   counts, or test statistics with sample sizes).
3. Were published **2019 or later** (pre-registration scope).
4. Are written in English.

### Inter-rater reliability

Disagreements between primary and secondary screeners will be resolved by discussion
or a designated tie-breaker.

---

## Data Extraction Pipeline {#sec-extraction}

Included articles are sent to **Google NotebookLM** for AI-assisted data extraction
via a four-step structured prompting protocol. All prompts are grounded by a
`MetaAnalysis_Reference_Sheet.pdf` data dictionary containing few-shot examples of
correct extractions.

::: {.callout-warning}
## Cache-bust rule
Each step must be run in a **completely new** NotebookLM chat window. Re-using a
previous window causes the model to recycle earlier tables rather than reading the
PDFs afresh.
:::

### Step 1 — Species & study identification

**Goal:** Establish taxonomic accuracy across all included papers.

Each PDF yields one row in a Markdown table with two columns:

| Column | Format | Example |
|---|---|---|
| `Study_ID` | `FirstAuthorLastName_Year_CamelCaseJournal` | `Belkina_2021_AnimalBehaviour` |
| `Species` | Full binomial nomenclature | *Poecilia reticulata* |

Common names alone are not accepted. The Study ID format is consistent across all
four steps to serve as the primary join key.

### Step 2 — Co-authorship & lab identification

**Goal:** Build a bipartite co-authorship network to identify "hub labs."

Output is in **long format** — one row per author per paper — so that network edges
can be constructed directly.

| Column | Description |
|---|---|
| `Study_ID` | Join key |
| `Author_Name` | Full name |
| `Author_Position` | First / Last / Corresponding / Middle |
| `Institutional_Affiliation` | University or institute |
| `Lab_Identifier` | `PILastName_Institution_City` (e.g., `Dukas_UniversityOfToronto_Toronto`) |

::: {.callout-note}
## Lab identifier design
The **last author** is designated as the lab PI, and the same `Lab_Identifier` string
is applied to every co-author row on the same paper. This links all authors to one
"primary lab" node in the bipartite network, enabling correct degree-centrality
calculation for hub-lab detection.
:::

### Step 3 — Trial-level data mapping

**Goal:** Map the internal experimental structure of each paper — how many distinct
trials exist, where their data live, and whether subjects are independent.

| Column | Description |
|---|---|
| `Study_ID` | Join key |
| `Trial_ID` | Experiment name/number (e.g., "Experiment 1", "Treatment A") |
| `Trial_Description` | 1–2 sentence summary of manipulation |
| `Sample_Size_N` | Focal subjects in this trial |
| `Subject_Independence` | "Independent Cohort" or "Repeated Measures" |
| `Control_Group_Shared` | Yes/No + which trial (if Yes) |
| `Data_Location` | Exact page and figure/table reference (e.g., "Table 1, Page 5") |

### Step 4 — Quantitative extraction (multi-agent pipeline)

**Goal:** Extract and cross-validate the actual numbers using a simulated three-agent
architecture.

#### Agent architecture

We utilize a three-agent quantitative extraction pipeline:

*   **Agent 1 (Extractor):** Pulls raw statistical data (means, variances, counts, and test statistics) directly from the Methods and Results text and tables.
*   **Agent 2 (Verifier):** Cross-checks calculations, verifies sample independence, identifies any SD vs. SE confusion, and scans the text to ensure no secondary trials were missed.
*   **Agent 3 (Arbiter):** Resolves any discrepancies between Agent 1 and Agent 2, standardizes reporting, and outputs the final clean data row.

**Agent 2 checks explicitly for three systematic extraction errors:**

1. **SD/SE confusion** — a common error in manual extraction that biases heterogeneity
   estimates if uncorrected.
2. **Sample independence** — flags whether the N used is independent across trials or
   from repeated measures on the same subjects.
3. **Missing trials** — re-reads the Methods to confirm no secondary experiments
   were missed by Agent 1.

#### Final output columns

| Column | Content |
|---|---|
| `Study_ID` | Author_Year_JournalName |
| `Trial_ID` | Specific experiment or condition |
| `Response_Metric` | What was measured |
| `Validated_Exp_Group_Data` | Mean ± SD/SE, Counts, or Test Stat |
| `Validated_Control_Group_Data` | Mean ± SD/SE, Counts, or NA |
| `SD_or_SE` | Explicit variance metric label |
| `Agent1_Initial_Pull` | Brief summary of original extraction |
| `Agent2_Verification_Notes` | Corrections made, or "No missing trials found" |
| `Verification_Location` | Exact page and table/figure number |

---

## Effect Size Calculation {#sec-effect-sizes}

Following data extraction, effect sizes and their sampling variances are calculated using the reproducible Python pipeline located at [scripts/1_effect_size_calculation_pipeline/calculate_effect_sizes.py](file:///Users/eduardosantos/Documents/Repos/mate_choice_meta/scripts/1_effect_size_calculation_pipeline/calculate_effect_sizes.py). 

The raw data is stored in the Excel spreadsheet [data/1_effect_size_calculation_pipeline/Data Extraction Mate Choice Meta Analysis.xlsx](file:///Users/eduardosantos/Documents/Repos/mate_choice_meta/data/1_effect_size_calculation_pipeline/Data%20Extraction%20Mate%20Choice%20Meta%20Analysis.xlsx) inside the `EduardoExtraction` sheet. To support tracking, each row contains an `identifierEffectSizeID` column containing a unique sequential identifier for each control–experimental comparison (e.g., `es0001`, `es0002`). The Python pipeline dynamically resolves column locations by scanning the header row, computes the following effect size metrics, and saves them back to the spreadsheet:

### 1. Hedges' $g$ (Bias-Corrected Standardised Mean Difference)
Calculated depending on the data type extracted:

* **From Means & SDs:** Using control and treatment means, standard deviations (or standard errors converted using $SD = SE \times \sqrt{n}$), and sample sizes.
* **From $t$-statistics:** Supported for both paired/one-sample designs (Case 2, in the Python script) and independent-sample designs (Case 3, assuming an equal sample split if individual group sizes are unknown).

### 2. Odds Ratio (OR)
Calculated for categorical outcomes from $2 \times 2$ contingency tables ($A$, $B$, $C$, $D$), where the sampling variance of the log odds ratio is approximated using the delta method.

### 3. Odds Ratio $\leftrightarrow$ Hedges' $g$ Cross-Conversion
For rows containing only one of the two metrics, the missing metric is derived using the logistic-normal approximation [@chinnSimpleMethodConverting2000; @borensteinIntroductionMetaanalysis2011]:

* **$\text{OR} \rightarrow \text{Hedges' } g$**: $d = \ln(\text{OR}) \times \frac{\sqrt{3}}{\pi}$ and then bias-corrected to $g$ using the $J$ correction factor.
* **$\text{Hedges' } g \rightarrow \text{OR}$**: $\ln(\text{OR}) = g \times \frac{\pi}{\sqrt{3}}$ and $\text{Var}(\ln\text{OR}) = \text{Var}(g) \times \frac{\pi^2}{3}$.

To execute the calculations and update the workbook, run:
```bash
python scripts/1_effect_size_calculation_pipeline/calculate_effect_sizes.py
```

---

## Planned Analyses {#sec-analyses}

All models will be implemented in R. Our pre-registration details are available in Santos et al. [-@santosMatechoiceCopyingNonhuman2026].

### Meta-analytic models

Multi-level random-effects meta-analysis to estimate the overall effect size for MCC, accounting for trials nested within papers nested within labs.

Package: `metafor`.

**Heterogeneity** will be quantified with $I^2$ and $\tau^2$ at each level.

### Phylogenetic comparative methods

Species non-independence is controlled via a phylogenetic covariance matrix
incorporated into the meta-analytic model (`metafor` + `ape`). This accounts for the possibility that closely related species share similar MCC propensity due to shared evolutionary history.

### Publication bias

Funnel plot asymmetry, and Egger's regression test to assess publication bias.
Time-lag bias will be examined by plotting effect size against publication year.

### Moderator analyses

Meta-regression models will test whether effect sizes vary by:

- Sex of the copying individual (female vs. male)
- Taxonomic group (fish, bird, mammal, invertebrate)
- Experimental design (before-and-after vs. no pre-test)
- Copying mechanism (individual vs. generalized)

### Sensitivity analyses

- Inclusion of pre-2019 records.
- Leave-one-out analysis for influential studies.
