% =====================================================================
% PROPOSED MANUSCRIPT ADDITIONS — exploratory moderator analyses
% Drafted to accompany the online-supplement chapter "moderator_analysis".
% These are SNIPPETS to paste into main.tex; they are not compiled here.
% All numbers are taken from the rendered supplement chapter.
% =====================================================================


% ---------------------------------------------------------------------
% (1) RESULTS — new subsection.
% Place immediately AFTER the existing subsection
% "Uni-moderator analyses of taxonomic group".
% ---------------------------------------------------------------------
\subsection{Exploratory uni-moderator analyses of extracted moderators}

Beyond taxonomic group, we extracted four candidate moderators during data
collection: the copying mechanism (generalized \textit{vs.} individual), the
modality of the social cue (visual, chemical, or acoustic), the virginity of the
observer, and the relative age of the demonstrator. These were not part of our
pre-registered analyses and several of their levels are sparsely sampled, so we
treat them as exploratory and report them in full in the online supplement.
The relative age of the demonstrator showed no variation in our extraction (all
``same-age'') and could not be analyzed. The remaining three moderators were fitted
as single-moderator multilevel models on the new extraction (both scales) and,
where the coding could be harmonized with an original dataset, on the combined
data. No moderator explained a detectable share of the heterogeneity on either
scale (all Wald $Q_M$, $p > 0.2$; total $I^2$ remained above 80\%). Two contrasts
are nonetheless worth noting. First, the strength of copying did not differ
between the two copying mechanisms: generalized and individual copying were
essentially identical in the combined Hedges' \textit{g} data (0.44 \textit{vs.}
0.44; Wald $Q_M$ = 0.003, $p$ = 0.96), and in the new extraction the generalized
estimate was, if anything, marginally larger (Hedges' \textit{g} = 0.26
\textit{vs.} 0.16; log OR = 0.48 \textit{vs.} 0.30; both $p > 0.5$). Second,
whereas Jones and DuVal~\cite{jonesMechanismsSocialInfluence2019} reported
stronger copying among virgin observers, our new extraction showed the opposite
ordering (Hedges' \textit{g} = 0.09 for virgin \textit{vs.} 0.36 for non-virgin
observers; Wald $Q_M$ = 1.61, $p$ = 0.22), and the difference all but disappeared
when the new data were pooled with the original binary data (OR = 1.75
\textit{vs.} 1.91; $p$ = 0.72). Neither contrast was statistically distinct on
either scale. Full diagnostics, model summaries, between-level contrasts, and
orchard plots are provided in the online supplement.


% ---------------------------------------------------------------------
% (2) DISCUSSION — replacement for the candidate-moderators sentence in the
% "Knowledge gaps and future opportunities" subsection.
%
% REPLACE the existing sentence:
%   "We collected several candidate moderators during extraction---the copying
%    mechanism (generalized versus individual), the modality of the social cue
%    (visual, chemical, or acoustic), the virginity of the observer, and the
%    relative age of the demonstrator---and we encourage future syntheses,
%    ideally pre-registered, to test these explicitly as the dataset continues
%    to grow."
% WITH the following:
% ---------------------------------------------------------------------
We collected several candidate moderators during extraction---the copying
mechanism (generalized versus individual), the modality of the social cue
(visual, chemical, or acoustic), the virginity of the observer, and the relative
age of the demonstrator---and explored them in post-hoc, non-pre-registered
uni-moderator analyses (online supplement). None explained a detectable share of
the heterogeneity, which is consistent with the variation residing at
finer-grained or interactive scales than our coding captured, and with several
levels (the chemical and acoustic cues in particular) being too sparsely sampled
to support inference. One result nonetheless deserves emphasis: the updated
evidence did not reproduce---and, in the new studies, tended to reverse---the
stronger copying among virgin observers reported by Jones and
DuVal~\cite{jonesMechanismsSocialInfluence2019}. We therefore encourage future
syntheses, ideally pre-registered, to test these moderators explicitly as the
dataset continues to grow, treating even the conditionality signals inherited
from the original meta-analyses as hypotheses to be re-examined rather than
settled results.


% ---------------------------------------------------------------------
% (3) DISCUSSION — OPTIONAL caveat in the social-learning subsection.
% In the paragraph beginning "Theory does not predict that animals should copy
% indiscriminately...", the sentence currently reads:
%   "Jones and DuVal found exactly such conditionality, with stronger copying
%    among inexperienced (virgin) females, in free-living settings, and when
%    social information ran counter to an innate preference;"
% Consider appending the parenthetical caveat shown here so the cited virginity
% effect is not presented as uncontested by our own update:
% ---------------------------------------------------------------------
Jones and DuVal~\cite{jonesMechanismsSocialInfluence2019} found exactly such
conditionality, with stronger copying among inexperienced (virgin) females
(although, as noted below and in the online supplement, our updated data did not
reproduce this particular virginity effect), in free-living settings, and when
social information ran counter to an innate preference;


% ---------------------------------------------------------------------
% (4) DISCUSSION — copying mechanism and its evolutionary consequences.
% This is the biologically most consequential of the exploratory contrasts.
% Suggested home: the subsection "Mate-choice copying as social learning, and
% its consequences for sexual selection" (it speaks directly to whether the
% transmissible form of copying is robustly expressed). Add as a short paragraph,
% e.g. after the paragraph on the phylogenetic distribution of the effect.
% ---------------------------------------------------------------------
One distinction among our exploratory moderators carries particular weight for
these evolutionary consequences: whether observers copy in a \textit{generalized}
or an \textit{individual} fashion. Only generalized copying---the acquisition of a
preference for a male phenotype or trait rather than for one specific
individual---can propagate through a population, bias the targets of sexual
selection, and seed the cultural transmission of mating preferences; individual
copying, by contrast, ends with the demonstrated male. The evolutionary reading of
mate-choice copying therefore rests on the assumption that the generalized,
transmissible form is itself robustly expressed. Our data support that assumption:
generalized copying was no weaker than individual copying---the two were
indistinguishable in the combined dataset (Hedges' \textit{g} = 0.44 \textit{vs.}
0.44) and the generalized estimate was marginally the larger of the two in the new
extraction (Hedges' \textit{g} = 0.26 \textit{vs.} 0.16). Although these contrasts
are exploratory and not statistically distinct, the absence of any deficit in
generalized copying is reassuring for the premise, central to interpreting
mate-choice copying as a driver of sexual selection and cultural evolution, that
the population-level, transmissible form of the behaviour is expressed at least as
strongly as its individual-level counterpart. This makes a mechanism-explicit,
pre-registered test of the generalized--individual distinction a high priority for
future work.
