---
title: "Moderator Analysis: Explaining the Heterogeneity"
subtitle: "Meta-Analytic Update of Mate-Choice Copying (MCC)"
author: "Eduardo S. A. Santos & Collaborators"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    theme: cosmo
---

```{r}
#| label: setup
#| include: false
# Ensure a CRAN mirror is set so package installation works in non-interactive renders
if (is.null(getOption("repos")) || getOption("repos")["CRAN"] %in% c(NA, "@CRAN@")) {
    options(repos = c(CRAN = "https://cloud.r-project.org"))
}
if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
if (!requireNamespace("orchaRd", quietly = TRUE)) {
    pacman::p_load_gh("daniel1noble/orchaRd")
}
pacman::p_load(
    readxl,
    tidyverse,
    metafor,
    orchaRd,
    patchwork,
    DT,
    scales,
    emmeans
)

# Number formatter (cleans negative zeros), matching overall_effect.qmd
fmt <- function(x, digits = 2) {
    res <- sprintf(paste0("%.", digits, "f"), x)
    zero_pattern <- paste0("-0.", paste0(rep("0", digits), collapse = ""))
    res[res == zero_pattern] <- paste0("0.", paste0(rep("0", digits), collapse = ""))
    res
}

# Inline p-value formatter
pstr <- function(p) {
    if (length(p) == 0 || is.na(p)) "p = NA" else if (p < 0.0001) "p < 0.0001" else paste0("p = ", fmt(p, 3))
}

# Significance-word helper
sigw <- function(p, yes = "statistically significant", no = "non-significant") {
    if (length(p) == 0 || is.na(p)) "indeterminate" else if (p < 0.05) yes else no
}

# Helper (identical to overall_effect.qmd) that rewrites orchard k-labels to
# "estimate [CI lower, CI upper]\nk = K (studies)". Wrapped in tryCatch by callers.
add_ci_labels <- function(plot_obj, orchard_results) {
    est <- sprintf("%.2f", orchard_results$mod_table$estimate)
    ci_l <- sprintf("%.2f", orchard_results$mod_table$lowerCL)
    ci_u <- sprintf("%.2f", orchard_results$mod_table$upperCL)

    orig_labels <- plot_obj$layers[[6]]$aes_params$label
    new_labels <- c()
    for (i in seq_along(orig_labels)) {
        nums <- regmatches(orig_labels[i], gregexpr("[0-9]+", orig_labels[i]))[[1]]
        if (length(nums) >= 2) {
            k_text <- paste0("k = ", nums[1], " (", nums[2], ")")
        } else {
            k_text <- orig_labels[i]
        }
        new_labels[i] <- paste0(est[i], " [", ci_l[i], ", ", ci_u[i], "]\n", k_text)
    }
    plot_obj$layers[[6]]$aes_params$label <- new_labels
    plot_obj$layers[[6]]$geom_params$parse <- FALSE
    plot_obj$layers[[6]]$aes_params$hjust <- "right"
    plot_obj$layers[[6]]$aes_params$vjust <- 0.5
    plot_obj$layers[[6]]$aes_params$size <- 3.0
    plot_obj
}

# Save a figure as a deliverable (PNG + PDF) in outputs/ (matches overall_effect.qmd).
save_mod_fig <- function(plot_obj, basename, width = 10, height = 5) {
    if (is.null(plot_obj)) return(invisible())
    tryCatch({
        d <- here::here("outputs", "1_effect_size_calculation_pipeline")
        dir.create(d, recursive = TRUE, showWarnings = FALSE)
        ggplot2::ggsave(file.path(d, paste0(basename, ".png")), plot_obj, width = width, height = height, dpi = 300, bg = "white")
        ggplot2::ggsave(file.path(d, paste0(basename, ".pdf")), plot_obj, width = width, height = height)
    }, error = function(e) warning("Could not save ", basename, ": ", conditionMessage(e)))
}
```

## Why a Moderator Analysis? {#sec-motivation}

The overall synthesis ([Heterogeneity Decomposition](overall_effect.qmd#sec-heterogeneity)) showed that the updated evidence base is dominated by **persistent, largely unexplained heterogeneity**. Total $I^2$ was high across every model, yet very little of that variance sat at the *study* or *phylogenetic* level. Instead it accumulated at the **within-study (observation)** and **non-phylogenetic species** levels — exactly where experimental and ecological distinctions among individual trials live. The disagreement among mate-choice copying estimates is therefore not mainly a matter of some laboratories producing systematically larger effects, nor of copying being concentrated in particular lineages; it points to **genuine biological and methodological moderators** that the intercept-only models could not capture.

During extraction we collected several candidate moderators that are plausible sources of this within-study/within-species variation:

* **Copying mechanism** — *generalized* (copying a *type* of male) versus *individual* (copying a *specific* demonstrated male);
* **Modality of the social cue** — the sensory channel of the demonstration (visual, chemical, or acoustic);
* **Virginity of the observer** — whether the choosing individual was sexually naïve;
* **Relative age of the demonstrator** — the age of the demonstrator relative to the observer.

This chapter does two things, in order:

1. **Diagnostics (Part 1).** Before fitting anything, we audit each candidate moderator *in the new extraction* — how its levels are populated, how much is missing, and how many studies/species back each level — and we check whether the coding can be reconciled with the original meta-analyses. This determines **which** moderators are analysable and **on which dataset** (new-only versus combined).
2. **Uni-moderator models (Part 2).** For the moderators that pass the diagnostic, we fit single-moderator multilevel models using the same `rma.mv` machinery, omnibus $Q_M$ test, and partial-$I^2$ heterogeneity decomposition used elsewhere in this workflow. We report two things for each: (i) whether the moderator explains a meaningful share of the *heterogeneity* (the $Q_M$ / $I^2$ question), and (ii) the **direction and magnitude of the differences between levels** ([§Biologically Meaningful Level Differences](#sec-biological)) — because even when a contrast is not statistically distinct, its sign can speak to specific hypotheses raised in the main text (most notably whether virgin observers copy more, as the original synthesis reported).

> These analyses are **exploratory**. With the dataset at its current size, several moderator levels are thin, and we have not pre-registered these tests. We report them to map the most promising directions for future, ideally pre-registered, syntheses — not as confirmatory claims.

---

## Data Loading & Moderator Construction {#sec-data}

We load the new extraction and derive the four candidate moderators from the extraction columns. We *also* load the moderator columns from the two original datasets (Davies et al. 2020 for Hedges' $g$; Jones & DuVal 2019 for the log odds ratio) so we can later test whether any moderator can be harmonised into a **combined** analysis. Each effect size in the new extraction carries **both** a Hedges' $d$ and an odds ratio, so the new-extraction moderator coverage is identical across the two effect-size scales.

```{r}
#| label: load_data
#| message: false
#| warning: false
path_new    <- here::here("data", "1_effect_size_calculation_pipeline", "Data Extraction Mate Choice Meta Analysis.xlsx")
path_davies <- here::here("data", "original_meta_analyses_datasets", "Davies_et_al_2020_Final_data.xlsx")
path_jones  <- here::here("data", "original_meta_analyses_datasets", "Jones_DuVal_2019_data.CSV")

db_new_raw <- readxl::read_excel(path_new, sheet = "EduardoExtraction")
db_davies  <- readxl::read_excel(path_davies, sheet = "Data")
db_jones   <- readr::read_csv(path_jones, show_col_types = FALSE)

# ---------------------------------------------------------------------------
# New extraction: filter excluded rows, coerce effect sizes, derive moderators.
# ---------------------------------------------------------------------------
# Modality recode: the extraction column `experimentalTypeStimulus` records the
# *type of demonstration*, which maps onto a sensory channel as follows —
#   Copulation / Association / Presence of eggs in nest -> Visual
#   Odour stimuli                                        -> Chemical
#   Acoustic stimuli                                     -> Acoustic
# `demo_type` keeps the two well-populated *visual* sub-types (Copulation vs
# Association) for a focused, better-powered sensitivity contrast.
db_new <- db_new_raw %>%
    filter(is.na(RemoveFromDataset) | RemoveFromDataset != "Yes") %>%
    mutate(
        across(c(
            effectSizeCalculatedHedgesD, effectSizeCalculatedHedgesDVariance,
            effectSizeCalculatedOddsRatio, effectSizeCalculatedOddsRatioVariance
        ), as.numeric),
        logOddsRatio = log(effectSizeCalculatedOddsRatio),
        logOddsRatioVariance = effectSizeCalculatedOddsRatioVariance,
        stim_raw = str_squish(experimentalTypeStimulus),
        mech = factor(str_squish(experimentalTypeCopyingMechanism), levels = c("Generalized", "Individual")),
        modality = factor(dplyr::case_when(
            stim_raw %in% c("Copulation", "Association", "Presence of eggs in nest") ~ "Visual",
            stim_raw %in% c("Odour stimuli")    ~ "Chemical",
            stim_raw %in% c("Acoustic stimuli") ~ "Acoustic",
            TRUE ~ NA_character_
        ), levels = c("Visual", "Chemical", "Acoustic")),
        demo_type = factor(dplyr::case_when(
            stim_raw == "Copulation"  ~ "Copulation",
            stim_raw == "Association" ~ "Association",
            TRUE ~ NA_character_
        ), levels = c("Association", "Copulation")),
        virgin = factor(str_squish(experimentalTypeVirgin), levels = c("No", "Yes")),
        demo_age = factor(dplyr::na_if(str_squish(experimentalModelAge), "NA"))
    )

# Tidy per-scale new-extraction frames (yi, vi, IDs, species epithet, moderators)
make_new <- function(yi_col, vi_col) {
    db_new %>%
        filter(!is.na(.data[[yi_col]]), !is.na(.data[[vi_col]]), .data[[vi_col]] > 0) %>%
        transmute(
            yi = .data[[yi_col]],
            vi = .data[[vi_col]],
            identifierStudyId = as.character(identifierStudyId),
            identifierEffectSizeID = as.character(identifierEffectSizeID),
            taxonomySpecies = str_to_lower(word(as.character(taxonomySpecies), -1)),
            mech, modality, demo_type, virgin, demo_age,
            Dataset_Source = "New Extraction (Santos)"
        )
}
new_hedges <- make_new("effectSizeCalculatedHedgesD", "effectSizeCalculatedHedgesDVariance")
new_or     <- make_new("logOddsRatio", "logOddsRatioVariance")

# ---------------------------------------------------------------------------
# Original datasets: extract ONLY the moderator columns we may be able to
# harmonise, with their native codings, for the diagnostic + combined models.
#   * Davies (Hedges' g): `Generalised_or_individual`  -> copying mechanism
#   * Jones  (log OR)    : `Virgin` (Y/N)              -> observer virginity
# ---------------------------------------------------------------------------
davies_hedges <- db_davies %>%
    mutate(yi = as.numeric(Hedges_d_directional), vi = as.numeric(Variance)) %>%
    filter(!is.na(yi), !is.na(vi), vi > 0) %>%
    transmute(
        yi, vi,
        identifierStudyId = as.character(Author),
        identifierEffectSizeID = as.character(Effect_size_no),
        taxonomySpecies = str_to_lower(word(as.character(Species_latin), -1)),
        mech = dplyr::recode(str_squish(Generalised_or_individual), "Generalised" = "Generalized"),
        Dataset_Source = "Original (Davies 2020)"
    ) %>%
    mutate(mech = factor(mech, levels = c("Generalized", "Individual")))

jones_or <- db_jones %>%
    filter(!is.na(lnOR), !is.na(VlnOR), VlnOR > 0) %>%
    transmute(
        yi = lnOR, vi = VlnOR,
        identifierStudyId = as.character(study),
        identifierEffectSizeID = as.character(study_Exp),
        taxonomySpecies = str_to_lower(as.character(Species)),
        virgin = dplyr::recode(str_squish(Virgin), "Y" = "Yes", "N" = "No"),
        Dataset_Source = "Original (Jones & DuVal 2019)"
    ) %>%
    mutate(virgin = factor(virgin, levels = c("No", "Yes")))

# Combined frames for the two harmonisable moderators
comb_hedges_mech <- bind_rows(
    new_hedges %>% select(yi, vi, identifierStudyId, identifierEffectSizeID, taxonomySpecies, mech, Dataset_Source),
    davies_hedges %>% select(yi, vi, identifierStudyId, identifierEffectSizeID, taxonomySpecies, mech, Dataset_Source)
)
comb_or_virgin <- bind_rows(
    new_or %>% select(yi, vi, identifierStudyId, identifierEffectSizeID, taxonomySpecies, virgin, Dataset_Source),
    jones_or %>% select(yi, vi, identifierStudyId, identifierEffectSizeID, taxonomySpecies, virgin, Dataset_Source)
)
```

---

## Part 1 — Moderator Diagnostics {#sec-diagnostics}

### Candidate moderators and their extraction columns

```{r}
#| label: moderator_dictionary
#| echo: false
mod_dict <- tibble::tribble(
    ~Moderator,                  ~`Extraction column`,                  ~`Levels (as analysed)`,
    "Copying mechanism",         "experimentalTypeCopyingMechanism",    "Generalized / Individual",
    "Social-cue modality",       "experimentalTypeStimulus (recoded)",  "Visual / Chemical / Acoustic",
    "Observer virginity",        "experimentalTypeVirgin",              "No / Yes",
    "Demonstrator relative age", "experimentalModelAge",                "Same-age / (other)"
)
DT::datatable(
    mod_dict,
    caption = "The four candidate moderators flagged in the discussion and the new-extraction columns they come from.",
    options = list(dom = "t", columnDefs = list(list(className = "dt-left", targets = "_all"))),
    rownames = FALSE
)
```

### Coverage in the new extraction {#sec-coverage}

For each candidate moderator we tabulate, per level, the number of effect sizes (`k`), the number of distinct studies and species backing that level, and the percentage of effect sizes with a missing value. A moderator is only worth modelling if it **varies**, if **each level is backed by more than one study** (so the moderator is not perfectly confounded with a single study), and if **missingness is low**.

```{r}
#| label: coverage_table
#| echo: false
# Coverage is computed on the new-extraction Hedges' frame; because every new
# effect size has both a Hedges' d and an OR, the OR frame is identical.
coverage_one <- function(dat, var, label) {
    total <- nrow(dat)
    dat %>%
        mutate(.lvl = .data[[var]]) %>%
        mutate(.lvl = forcats::fct_na_value_to_level(factor(.lvl), level = "(missing)")) %>%
        group_by(Level = .lvl) %>%
        summarise(
            k_es = dplyr::n(),
            studies = dplyr::n_distinct(identifierStudyId),
            species = dplyr::n_distinct(taxonomySpecies),
            .groups = "drop"
        ) %>%
        mutate(Moderator = label, `% of ES` = round(100 * k_es / total, 1)) %>%
        select(Moderator, Level, k_es, studies, species, `% of ES`)
}

coverage_tbl <- dplyr::bind_rows(
    coverage_one(new_hedges, "mech", "Copying mechanism"),
    coverage_one(new_hedges, "modality", "Social-cue modality"),
    coverage_one(new_hedges, "demo_type", "Demonstration sub-type (visual)"),
    coverage_one(new_hedges, "virgin", "Observer virginity"),
    coverage_one(new_hedges, "demo_age", "Demonstrator relative age")
) %>%
    rename(`k (effect sizes)` = k_es, Studies = studies, Species = species)

DT::datatable(
    coverage_tbl,
    caption = "New-extraction coverage of each candidate moderator: effect sizes, backing studies and species, and percentage of effect sizes per level. Identical across the Hedges' g and log OR scales (every new effect size carries both).",
    options = list(dom = "t", pageLength = 20, columnDefs = list(list(className = "dt-center", targets = c(2, 3, 4, 5)))),
    rownames = FALSE
)
```

```{r}
#| label: coverage_inline
#| include: false
# Pull a few numbers for the narrative
get_cov <- function(modlabel, level, col) {
    v <- coverage_tbl[[col]][coverage_tbl$Moderator == modlabel & coverage_tbl$Level == level]
    if (length(v) == 0) NA else v[1]
}
n_age_levels <- dplyr::n_distinct(na.omit(new_hedges$demo_age))
```

### The social-cue modality recode {#sec-modality-recode}

The extraction column `experimentalTypeStimulus` records the *form of the demonstration* rather than a sensory label directly, so we map it onto a modality:

| `experimentalTypeStimulus` | Sensory modality |
|---|---|
| Copulation, Association, Presence of eggs in nest | **Visual** |
| Odour stimuli | **Chemical** |
| Acoustic stimuli | **Acoustic** |

The coverage table makes the central limitation plain: the new extraction is **overwhelmingly visual**, with the *Chemical* and *Acoustic* levels each resting on a handful of effect sizes from very few studies. A three-level modality model can be fitted, but its non-visual contrasts are effectively confounded with one or two studies/species and must be read with extreme caution. We therefore also define `demo_type` — *Copulation* versus *Association*, the two well-populated visual sub-types — as a better-powered sensitivity contrast within the dominant modality.

### Harmonisation with the original meta-analyses {#sec-harmonisation}

The four moderators were coded under a single protocol in the **new extraction**, but the original datasets recorded only some of them, and not always on a compatible scale. The table below records, for each moderator, whether a compatible variable exists in Davies et al. 2020 (Hedges' $g$) and Jones & DuVal 2019 (log OR), and the resulting verdict.

```{r}
#| label: harmonisation_table
#| echo: false
harmon <- tibble::tribble(
    ~Moderator,                  ~`New extraction`,        ~`Davies 2020 (Hedges' g)`,                         ~`Jones & DuVal 2019 (log OR)`,                       ~Verdict,
    "Copying mechanism",         "Generalized / Individual","Generalised_or_individual — same dichotomy",        "not recorded",                                        "Combinable on the Hedges' g scale (new + Davies)",
    "Social-cue modality",       "Visual / Chemical / Acoustic","Demonstration_type — a behavioural-context code, not a sensory modality", "not recorded",                  "New-only (constructs do not match)",
    "Observer virginity",        "No / Yes",               "not recorded",                                      "Virgin (Y/N) — same construct",                       "Combinable on the log OR scale (new + Jones)",
    "Demonstrator relative age", "Same-age / (other)",     "not recorded",                                      "not recorded",                                        "Not analysable (no variation; see below)"
)
DT::datatable(
    harmon,
    caption = "Cross-dataset availability and coding of each candidate moderator, with the harmonisation verdict.",
    options = list(dom = "t", columnDefs = list(list(className = "dt-left", targets = "_all"))),
    rownames = FALSE
)
```

### Diagnostic verdict {#sec-verdict}

Putting the coverage and harmonisation evidence together:

* **Copying mechanism (Generalized vs Individual)** — well balanced in the new extraction (`r get_cov("Copying mechanism", "Generalized", "k (effect sizes)")` vs `r get_cov("Copying mechanism", "Individual", "k (effect sizes)")` effect sizes, each from many studies) and coded identically by Davies. **Analysable new-only on both scales, and combinable on the Hedges' $g$ scale.**
* **Observer virginity (No vs Yes)** — both levels well populated (`r get_cov("Observer virginity", "No", "k (effect sizes)")` vs `r get_cov("Observer virginity", "Yes", "k (effect sizes)")`) and coded identically by Jones & DuVal. **Analysable new-only on both scales, and combinable on the log OR scale.**
* **Social-cue modality (Visual/Chemical/Acoustic)** — analysable **new-only**, but the non-visual levels are sparse and study-confounded; treated as exploratory, with the Copulation-vs-Association visual contrast as a sensitivity check. The original "demonstration type" codes a different construct and cannot be merged.
* **Demonstrator relative age** — **dropped**: the new extraction records `r n_age_levels` non-missing level(s) ("Same-age"), i.e. essentially **no variation**, so it cannot enter a model. The original datasets carry no compatible variable.

**Bottom line.** Three moderators (mechanism, modality, virginity) can be tested on the **new dataset** for both effect-size scales. Two of them can additionally be tested on a **combined dataset** where the construct harmonises cleanly — mechanism on Hedges' $g$ (new + Davies) and virginity on log OR (new + Jones). Demonstrator age is not analysable. Part 2 implements exactly these models.

---

## Part 2 — Uni-Moderator Models {#sec-models}

### Modelling approach

Each model is a single-moderator multilevel meta-analysis fitted with `metafor::rma.mv()`, mirroring the [taxonomic-moderator model](overall_effect.qmd#sec-taxonomic-moderator):

* **Dependency.** A variance–covariance matrix from `vcalc()` ($\rho = 0.5$) handles multiple effect sizes within a study.
* **Random effects.** Study (`identifierStudyId`), residual effect-size (`identifierEffectSizeID`), and a non-phylogenetic species intercept (`taxonomySpecies`). The phylogenetic term is **omitted**: the heterogeneity decomposition showed it carries negligible variance, and dropping it keeps these smaller moderator models identifiable (the same choice made for the taxonomic model).
* **Heteroscedasticity with a fallback cascade.** We first attempt a heteroscedastic structure that lets the **residual (within-study) variance differ by moderator level** (`~ moderator | effect-size`, `struct = "HCS"`, `rho = 0`), trying several optimisers; if it errors we fall back to a homoscedastic structure. This is the `orchaRd` 2.0 recommendation for meta-regression with heterogeneous variances [@nakagawaOrchaRd2Visualising2023].
* **Inference.** The moderator is intercept-coded (`mods = ~ moderator`); per-level means, CIs, and prediction intervals come from `orchaRd::mod_results()` (marginalised means via `emmeans`), and the among-level test is the model's omnibus $Q_M$. Heterogeneity is decomposed with partial $I^2$ from `orchaRd::i2_ml()`.

```{r}
#| label: fit_function
#| message: false
#| warning: false
# Generalised heteroscedastic single-moderator fitter (adapted from
# fit_taxon_hetero_mv in overall_effect.qmd). Returns the chosen model plus a
# cascade log so the reader can see which structure was actually used.
fit_mod_hetero <- function(dat, group_var, group_levels, yi = "yi", vi = "vi",
                           cluster_var = "identifierStudyId",
                           obs_var = "identifierEffectSizeID", rho = 0.5) {
    dat2 <- dat %>%
        filter(!is.na(.data[[yi]]), !is.na(.data[[vi]]), .data[[vi]] > 0, !is.na(.data[[group_var]]))
    if (nrow(dat2) == 0) return(NULL)
    dat2[[group_var]] <- droplevels(factor(dat2[[group_var]], levels = group_levels))

    VCV <- metafor::vcalc(vi = dat2[[vi]], cluster = dat2[[cluster_var]], obs = dat2[[obs_var]], rho = rho, data = dat2)
    mods_f <- stats::as.formula(paste0("~ ", group_var))
    obs_f  <- stats::as.formula(paste0("~ ", group_var, " | ", obs_var))

    opt_grid <- list(
        "nlminb"            = list(optimizer = "nlminb", iter.max = 10000, eval.max = 10000),
        "optim/BFGS"        = list(optimizer = "optim", optmethod = "BFGS", maxit = 10000),
        "optim/L-BFGS-B"    = list(optimizer = "optim", optmethod = "L-BFGS-B", maxit = 10000),
        "optim/Nelder-Mead" = list(optimizer = "optim", optmethod = "Nelder-Mead", maxit = 50000, reltol = 1e-9)
    )
    build <- function(cand, ctrl) {
        args <- c(
            list(yi = dat2[[yi]], V = VCV, mods = mods_f, random = cand$random,
                 test = "t", dfs = "contain", method = "REML", sparse = TRUE, data = dat2),
            cand$args
        )
        if (!is.null(cand$struct)) args$struct <- cand$struct
        if (!is.null(ctrl)) args$control <- ctrl
        do.call(metafor::rma.mv, args)
    }
    candidates <- list(
        list(spec = "Heteroscedastic (residual only, per level; HCS, rho = 0)",
             random = list(~ 1 | identifierStudyId, obs_f, ~ 1 | taxonomySpecies),
             struct = "HCS", args = list(rho = 0), het = TRUE),
        list(spec = "Homoscedastic (single variance per component)",
             random = list(~ 1 | identifierStudyId, ~ 1 | identifierEffectSizeID, ~ 1 | taxonomySpecies),
             struct = NULL, args = list(), het = FALSE)
    )
    fit_once <- function(cand, ctrl) {
        warns <- character(0); t0 <- Sys.time()
        mod <- tryCatch(
            withCallingHandlers(build(cand, ctrl),
                warning = function(w) { warns[[length(warns) + 1L]] <<- conditionMessage(w); invokeRestart("muffleWarning") }),
            error = function(e) structure(list(msg = conditionMessage(e)), class = "mod_fit_error")
        )
        list(mod = mod, warns = warns, secs = as.numeric(difftime(Sys.time(), t0, units = "secs")))
    }
    attempt <- function(cand) {
        if (!isTRUE(cand$het)) {
            r <- fit_once(cand, NULL); failed <- inherits(r$mod, "mod_fit_error")
            return(list(model = if (failed) NULL else r$mod, failed = failed,
                        error = if (failed) r$mod$msg else NA_character_, warnings = r$warns,
                        secs = r$secs, optimizer = "default (nlminb)"))
        }
        msgs <- character(0); total <- 0
        for (onm in names(opt_grid)) {
            r <- fit_once(cand, opt_grid[[onm]]); total <- total + r$secs
            if (!inherits(r$mod, "mod_fit_error")) {
                return(list(model = r$mod, failed = FALSE, error = NA_character_,
                            warnings = r$warns, secs = total, optimizer = onm))
            }
            msgs <- c(msgs, paste0(onm, ": ", r$mod$msg))
        }
        list(model = NULL, failed = TRUE, error = paste(msgs, collapse = " || "),
             warnings = character(0), secs = total, optimizer = NA_character_)
    }
    diag_rows <- list(); chosen <- NULL; chosen_spec <- NA_character_
    for (i in seq_along(candidates)) {
        cand <- candidates[[i]]
        if (!is.null(chosen)) {
            diag_rows[[i]] <- tibble::tibble(stage = i, structure = cand$spec, status = "not attempted",
                optimizer = NA_character_, seconds = NA_real_, n_warnings = NA_integer_,
                detail = "an earlier structure succeeded")
            next
        }
        a <- attempt(cand)
        status <- if (a$failed) "ERROR" else if (length(a$warnings) > 0) "OK (with warnings)" else "OK"
        detail <- if (a$failed) a$error else if (length(a$warnings) > 0) paste(a$warnings, collapse = " || ") else "clean fit"
        diag_rows[[i]] <- tibble::tibble(stage = i, structure = cand$spec, status = status,
            optimizer = a$optimizer, seconds = round(a$secs, 2), n_warnings = length(a$warnings), detail = detail)
        if (!a$failed) { chosen <- a$model; chosen_spec <- cand$spec }
    }
    diagnostics <- dplyr::bind_rows(diag_rows)
    if (is.null(chosen)) {
        warning("All structures failed for moderator '", group_var, "'.")
        return(list(data = dat2, VCV = VCV, model = NULL, spec = "ALL FAILED",
                    diagnostics = diagnostics, group_var = group_var))
    }
    list(data = dat2, VCV = VCV, model = chosen, spec = chosen_spec,
         diagnostics = diagnostics, group_var = group_var)
}

# Omnibus among-level test (QM tests all non-intercept coefficients jointly).
among_test <- function(res) {
    if (is.null(res) || is.null(res$model)) return(NULL)
    list(QM = res$model$QM, QMp = res$model$QMp, QMdf = res$model$QMdf)
}
```

```{r}
#| label: fit_all_models
#| message: false
#| warning: false
# Moderator registry: each entry is fitted on both new-extraction scales.
mod_specs <- list(
    mech      = list(var = "mech",      levels = c("Generalized", "Individual"),
                     label = "Copying mechanism",
                     palette = c("Generalized" = "#E69F00", "Individual" = "#56B4E9")),
    modality  = list(var = "modality",  levels = c("Visual", "Chemical", "Acoustic"),
                     label = "Social-cue modality",
                     palette = c("Visual" = "#009E73", "Chemical" = "#CC79A7", "Acoustic" = "#D55E00")),
    demo_type = list(var = "demo_type", levels = c("Association", "Copulation"),
                     label = "Demonstration sub-type (visual)",
                     palette = c("Association" = "#0072B2", "Copulation" = "#E69F00")),
    virgin    = list(var = "virgin",    levels = c("No", "Yes"),
                     label = "Observer virginity",
                     palette = c("No" = "#0072B2", "Yes" = "#D55E00"))
)

# Fit every moderator on the new-extraction Hedges' g and log OR frames.
fits_new <- list()
for (key in names(mod_specs)) {
    s <- mod_specs[[key]]
    fits_new[[key]] <- list(
        hedges = fit_mod_hetero(new_hedges, s$var, s$levels),
        or     = fit_mod_hetero(new_or,     s$var, s$levels)
    )
}

# Combined-dataset extensions (harmonisable moderators only).
fit_comb_mech_hedges  <- fit_mod_hetero(comb_hedges_mech, "mech",   c("Generalized", "Individual"))
fit_comb_virgin_or    <- fit_mod_hetero(comb_or_virgin,   "virgin", c("No", "Yes"))

# Persist all fits so they can be inspected without re-rendering.
tryCatch({
    d <- here::here("outputs", "1_effect_size_calculation_pipeline")
    dir.create(d, recursive = TRUE, showWarnings = FALSE)
    saveRDS(list(new = fits_new, comb_mech_hedges = fit_comb_mech_hedges,
                 comb_virgin_or = fit_comb_virgin_or, saved_at = Sys.time()),
            file.path(d, "moderator_model_fits.rds"))
}, error = function(e) warning("Could not save moderator_model_fits.rds: ", conditionMessage(e)))
```

### Reporting helpers

```{r}
#| label: report_helpers
#| message: false
#| warning: false
# Partial I^2 (long form) for a fitted moderator model.
comp_labels <- c(
    Total                  = "Total",
    identifierStudyId      = "Study",
    identifierEffectSizeID = "Effect size (residual)",
    taxonomySpecies        = "Species (non-phylogenetic)"
)
# orchaRd::i2_ml() does not support the heteroscedastic (HCS) structure that the
# cascade prefers, so for the partial-I² *decomposition* we compute i2_ml on a
# homoscedastic refit of the same data (one variance per random effect). This is
# exactly how the overall synthesis reports partial I² (from the homoscedastic
# intercept models); the heteroscedastic model is retained for the means/Q_M.
i2_homosc_refit <- function(res) {
    if (is.null(res) || is.null(res$model) || is.null(res$data)) return(NULL)
    d <- res$data
    m <- tryCatch(
        metafor::rma.mv(
            yi = d$yi, V = res$VCV,
            mods = stats::as.formula(paste0("~ ", res$group_var)),
            random = list(~ 1 | identifierStudyId, ~ 1 | identifierEffectSizeID, ~ 1 | taxonomySpecies),
            test = "t", method = "REML", sparse = TRUE, data = d
        ),
        error = function(e) NULL
    )
    if (is.null(m)) return(NULL)
    tryCatch(orchaRd::i2_ml(m), error = function(e) NULL)
}

get_i2_long <- function(res, analysis) {
    if (is.null(res) || is.null(res$model)) return(NULL)
    v <- tryCatch(orchaRd::i2_ml(res$model), error = function(e) NULL)
    if (is.null(v)) v <- i2_homosc_refit(res)
    if (is.null(v)) return(NULL)
    tibble::tibble(Analysis = analysis, Component = sub("^I2_", "", names(v)), I2 = round(as.numeric(v), 1)) %>%
        mutate(Component = dplyr::recode(Component, !!!comp_labels))
}

# Total I^2: orchaRd on the fitted model, else on the homoscedastic refit (i2_ml
# does not support the HCS structure), else a Higgins typical-variance fallback.
total_i2 <- function(res) {
    if (is.null(res) || is.null(res$model)) return(NA_real_)
    i2 <- tryCatch(orchaRd::i2_ml(res$model), error = function(e) NULL)
    if (is.null(i2)) i2 <- i2_homosc_refit(res)
    if (!is.null(i2) && any(grepl("Total", names(i2)))) return(as.numeric(i2[grep("Total", names(i2))[1]]))
    m <- res$model; vi <- res$data$vi; wi <- 1 / vi; k <- length(vi)
    s2 <- (k - 1) * sum(wi) / (sum(wi)^2 - sum(wi^2))
    vc <- sum(c(m$sigma2, m$tau2, m$gamma2), na.rm = TRUE)
    100 * vc / (vc + s2)
}

# Per-level marginalised means table.
level_means <- function(res, analysis) {
    if (is.null(res) || is.null(res$model)) return(NULL)
    mr <- orchaRd::mod_results(res$model, mod = res$group_var, group = "identifierStudyId")
    mr$mod_table %>%
        transmute(Analysis = analysis, Level = name,
                  Estimate = round(estimate, 3),
                  `CI lower` = round(lowerCL, 3), `CI upper` = round(upperCL, 3),
                  `PI lower` = round(lowerPR, 3), `PI upper` = round(upperPR, 3))
}

# Between-level contrasts: each non-intercept model coefficient is the
# difference between that level and the reference level, on the effect-size
# scale. A statistically "distinct" difference would have a CI excluding zero;
# a biologically suggestive one is read from the sign and magnitude even when the
# CI overlaps zero. For log OR models we also report the multiplicative odds
# ratio (exp(difference)) so the contrast is interpretable on the OR scale.
level_contrasts <- function(res, analysis, is_or = FALSE) {
    if (is.null(res) || is.null(res$model)) return(NULL)
    m <- res$model
    nm <- rownames(m$b)
    keep <- !grepl("intrcpt|intercept", nm, ignore.case = TRUE)
    if (!any(keep)) return(NULL)
    ref <- levels(droplevels(factor(res$data[[res$group_var]], levels = levels(res$data[[res$group_var]]))))[1]
    out <- tibble::tibble(
        Analysis = analysis,
        Contrast = paste0(sub(res$group_var, "", nm[keep]), " − ", ref),
        Difference = round(as.numeric(m$b)[keep], 3),
        `CI lower` = round(m$ci.lb[keep], 3),
        `CI upper` = round(m$ci.ub[keep], 3),
        p = round(m$pval[keep], 3)
    )
    if (is_or) {
        out$`Odds ratio` <- round(exp(out$Difference), 2)
        out <- dplyr::relocate(out, `Odds ratio`, .after = Difference)
    }
    out
}

# Orchard plot for a fitted moderator model.
plot_mod_orchard <- function(res, palette, xlab, title) {
    if (is.null(res) || is.null(res$model)) return(NULL)
    mr <- orchaRd::mod_results(res$model, mod = res$group_var, group = "identifierStudyId")
    oc <- orchaRd::orchard_plot(mr, xlab = xlab, angle = 0, g = TRUE, k = TRUE)
    oc <- tryCatch(add_ci_labels(oc, mr), error = function(e) { warning("add_ci_labels failed: ", conditionMessage(e)); oc })
    oc +
        ggplot2::scale_colour_manual(values = palette, name = NULL, aesthetics = c("colour", "fill")) +
        ggplot2::ggtitle(title) +
        ggplot2::theme(plot.title = ggplot2::element_text(face = "bold", hjust = 0.5), legend.position = "none")
}

# Compact omnibus + I2 one-liner for a model.
qm_line <- function(res, scale_label) {
    a <- among_test(res); i2 <- total_i2(res)
    if (is.null(a)) return(paste0("*", scale_label, ":* model did not fit."))
    paste0("*", scale_label, ":* total $I^2$ = ", fmt(i2, 1), "%; omnibus $Q_M$ = ",
           fmt(a$QM, 2), " (", pstr(a$QMp), "), a **", sigw(a$QMp), "** difference among levels.")
}
```

### New-extraction models {#sec-new-models}

Each tab below presents one moderator, with the Hedges' $g$ and log odds ratio models side by side: the omnibus test and total heterogeneity, the per-level means, the partial-$I^2$ decomposition, the fit-cascade log, and the orchard plots.

::: {.panel-tabset}

```{r}
#| label: new_model_panels
#| echo: false
#| results: asis
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 9
for (key in names(mod_specs)) {
    s <- mod_specs[[key]]
    rh <- fits_new[[key]]$hedges
    ro <- fits_new[[key]]$or

    cat("\n\n#### ", s$label, "\n\n")
    cat(qm_line(rh, "Hedges' g"), "\n\n")
    cat(qm_line(ro, "Log odds ratio"), "\n\n")

    # Per-level means
    means_tbl <- dplyr::bind_rows(level_means(rh, "Hedges' g"), level_means(ro, "Log OR"))
    if (!is.null(means_tbl) && nrow(means_tbl) > 0) {
        print(htmltools::tagList(DT::datatable(
            means_tbl,
            caption = paste0("Marginalised mean effect per level — ", s$label, " (new extraction)."),
            options = list(dom = "t", pageLength = 12, columnDefs = list(list(className = "dt-center", targets = "_all"))),
            rownames = FALSE
        )))
    }

    # Between-level contrasts (difference vs reference level, with CI)
    contr_tbl <- dplyr::bind_rows(
        level_contrasts(rh, "Hedges' g", is_or = FALSE),
        level_contrasts(ro, "Log OR", is_or = TRUE)
    )
    if (!is.null(contr_tbl) && nrow(contr_tbl) > 0) {
        print(htmltools::tagList(DT::datatable(
            contr_tbl,
            caption = paste0("Between-level contrasts (difference from the reference level, on the effect-size scale) — ", s$label, " (new extraction). A CI excluding zero would indicate a statistically distinct difference."),
            options = list(dom = "t", pageLength = 12, columnDefs = list(list(className = "dt-center", targets = "_all"))),
            rownames = FALSE
        )))
    }

    # Partial I2
    i2_tbl <- dplyr::bind_rows(get_i2_long(rh, "Hedges' g"), get_i2_long(ro, "Log OR"))
    if (!is.null(i2_tbl) && nrow(i2_tbl) > 0 && "Component" %in% names(i2_tbl)) {
        i2_tbl <- tidyr::pivot_wider(i2_tbl, names_from = Component, values_from = I2)
    }
    if (!is.null(i2_tbl) && nrow(i2_tbl) > 0) {
        print(htmltools::tagList(DT::datatable(
            i2_tbl,
            caption = paste0("Partial I² (%) by random effect — ", s$label, " (new extraction; decomposition from a homoscedastic refit, since i2_ml does not support the HCS structure)."),
            options = list(dom = "t", columnDefs = list(list(className = "dt-center", targets = "_all"))),
            rownames = FALSE
        )))
    }

    # Fit cascade log
    diag_tbl <- dplyr::bind_rows(
        if (!is.null(rh)) dplyr::mutate(rh$diagnostics, Analysis = "Hedges' g"),
        if (!is.null(ro)) dplyr::mutate(ro$diagnostics, Analysis = "Log OR")
    )
    if (!is.null(diag_tbl) && nrow(diag_tbl) > 0) {
        diag_tbl <- diag_tbl %>% dplyr::select(Analysis, structure, status, optimizer, seconds, n_warnings)
        print(htmltools::tagList(DT::datatable(
            diag_tbl,
            caption = paste0("Fit-cascade log — ", s$label, ". The first OK row per analysis is the structure used."),
            options = list(dom = "t", columnDefs = list(list(className = "dt-left", targets = "_all"))),
            rownames = FALSE
        )))
    }

    # Orchard plots (stacked)
    ph <- plot_mod_orchard(rh, s$palette, "Hedges' g", paste0(s$label, " — Hedges' g"))
    po <- plot_mod_orchard(ro, s$palette, "Log Odds Ratio (lnOR)", paste0(s$label, " — log OR"))
    save_mod_fig(ph, paste0("moderator_orchard_", key, "_hedges"))
    save_mod_fig(po, paste0("moderator_orchard_", key, "_or"))
    if (!is.null(ph) && !is.null(po)) {
        fig <- (ph / po) + patchwork::plot_annotation(
            tag_levels = "a", tag_prefix = "(", tag_suffix = ")",
            title = paste0("Mate-choice copying by ", tolower(s$label), " (new extraction)"),
            theme = ggplot2::theme(plot.title = ggplot2::element_text(size = 14, face = "bold", hjust = 0.5)))
        save_mod_fig(fig, paste0("moderator_orchard_", key), width = 10, height = 9)
        print(fig)
    } else {
        if (!is.null(ph)) print(ph)
        if (!is.null(po)) print(po)
    }
    cat("\n\n")
}
```

:::

### Combined-dataset extensions {#sec-combined-models}

Where the construct harmonises cleanly with an original dataset, we re-fit the moderator on the **combined** data. This roughly triples the sample and lets us ask whether the (null) new-extraction patterns hold once the legacy effect sizes are folded in.

* **Copying mechanism — Hedges' $g$ (new + Davies 2020):** `r nrow(comb_hedges_mech)` effect sizes.
* **Observer virginity — log OR (new + Jones & DuVal 2019):** `r nrow(comb_or_virgin)` effect sizes.

```{r}
#| label: combined_models_report
#| echo: false
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 5
# Means + I2 tables
comb_means <- dplyr::bind_rows(
    level_means(fit_comb_mech_hedges, "Mechanism — Hedges' g (new + Davies)"),
    level_means(fit_comb_virgin_or,   "Virginity — log OR (new + Jones)")
)
DT::datatable(
    comb_means,
    caption = "Combined-dataset moderator models: marginalised mean effect per level.",
    options = list(dom = "t", pageLength = 12, columnDefs = list(list(className = "dt-center", targets = "_all"))),
    rownames = FALSE
)
```

```{r}
#| label: combined_models_contrasts
#| echo: false
#| message: false
#| warning: false
comb_contr <- dplyr::bind_rows(
    level_contrasts(fit_comb_mech_hedges, "Mechanism — Hedges' g (new + Davies)", is_or = FALSE),
    level_contrasts(fit_comb_virgin_or,   "Virginity — log OR (new + Jones)",     is_or = TRUE)
)
DT::datatable(
    comb_contr,
    caption = "Combined-dataset moderator models: between-level contrast (difference from the reference level, with CI).",
    options = list(dom = "t", pageLength = 12, columnDefs = list(list(className = "dt-center", targets = "_all"))),
    rownames = FALSE
)
```

```{r}
#| label: combined_models_i2
#| echo: false
#| message: false
#| warning: false
comb_i2 <- dplyr::bind_rows(
    get_i2_long(fit_comb_mech_hedges, "Mechanism — Hedges' g (new + Davies)"),
    get_i2_long(fit_comb_virgin_or,   "Virginity — log OR (new + Jones)")
)
if (nrow(comb_i2) > 0 && "Component" %in% names(comb_i2)) {
    comb_i2 <- tidyr::pivot_wider(comb_i2, names_from = Component, values_from = I2)
}
DT::datatable(
    comb_i2,
    caption = "Combined-dataset moderator models: partial I² (%) by random effect (decomposition from a homoscedastic refit).",
    options = list(dom = "t", columnDefs = list(list(className = "dt-center", targets = "_all"))),
    rownames = FALSE
)
```

```{r}
#| label: combined_models_orchards
#| echo: false
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 9
p_comb_mech <- plot_mod_orchard(fit_comb_mech_hedges, mod_specs$mech$palette,
                                "Hedges' g", "Copying mechanism — Hedges' g (new + Davies 2020)")
p_comb_virg <- plot_mod_orchard(fit_comb_virgin_or, mod_specs$virgin$palette,
                                "Log Odds Ratio (lnOR)", "Observer virginity — log OR (new + Jones & DuVal 2019)")
save_mod_fig(p_comb_mech, "moderator_orchard_combined_mech_hedges")
save_mod_fig(p_comb_virg, "moderator_orchard_combined_virgin_or")
if (!is.null(p_comb_mech) && !is.null(p_comb_virg)) {
    fig_comb <- (p_comb_mech / p_comb_virg) + patchwork::plot_annotation(
        tag_levels = "a", tag_prefix = "(", tag_suffix = ")",
        title = "Combined-dataset moderator models",
        theme = ggplot2::theme(plot.title = ggplot2::element_text(size = 14, face = "bold", hjust = 0.5)))
    save_mod_fig(fig_comb, "moderator_orchard_combined", width = 10, height = 9)
    print(fig_comb)
}
```

```{r}
#| label: combined_inline
#| include: false
qm_mech <- among_test(fit_comb_mech_hedges)
qm_virg <- among_test(fit_comb_virgin_or)
i2_mech <- total_i2(fit_comb_mech_hedges)
i2_virg <- total_i2(fit_comb_virgin_or)
qmp <- function(a) if (is.null(a) || is.null(a$QMp)) NA_real_ else as.numeric(a$QMp)
qms <- function(a) if (is.null(a) || is.null(a$QM)) NA_real_ else as.numeric(a$QM)
```

In the combined data the conclusions are unchanged. For **copying mechanism** (Hedges' $g$), total heterogeneity is `r fmt(i2_mech, 1)`% and the among-level test is `r pstr(qmp(qm_mech))` ($Q_M$ = `r fmt(qms(qm_mech), 2)`) — a **`r sigw(qmp(qm_mech))`** difference between generalized and individual copying. For **observer virginity** (log OR), total heterogeneity is `r fmt(i2_virg, 1)`% and the among-level test is `r pstr(qmp(qm_virg))` ($Q_M$ = `r fmt(qms(qm_virg), 2)`), again a **`r sigw(qmp(qm_virg))`** difference between virgin and non-virgin observers.

---

## Biologically Meaningful Level Differences {#sec-biological}

A non-significant $Q_M$ tells us a moderator does not *explain the heterogeneity*, but it does not tell us whether the **direction and magnitude** of the level differences are biologically interesting. Because the point estimates speak to specific hypotheses raised in the main text, we read them here as well — keeping firmly in mind that every contrast CI below overlaps zero, so these are **directional signals, not demonstrated effects**.

```{r}
#| label: biological_inline
#| include: false
mean_of <- function(res, lvl) {
    if (is.null(res) || is.null(res$model)) return(NA_real_)
    mr <- orchaRd::mod_results(res$model, mod = res$group_var, group = "identifierStudyId")
    v <- mr$mod_table$estimate[mr$mod_table$name == lvl]
    if (length(v) == 0) NA_real_ else v[1]
}
# Virginity — new extraction (both scales) and combined OR
vh_no  <- mean_of(fits_new$virgin$hedges, "No");  vh_yes <- mean_of(fits_new$virgin$hedges, "Yes")
vo_no  <- mean_of(fits_new$virgin$or, "No");      vo_yes <- mean_of(fits_new$virgin$or, "Yes")
vc_no  <- mean_of(fit_comb_virgin_or, "No");      vc_yes <- mean_of(fit_comb_virgin_or, "Yes")
# Demonstration sub-type (visual): copulation vs association
dt_assoc <- mean_of(fits_new$demo_type$hedges, "Association")
dt_cop   <- mean_of(fits_new$demo_type$hedges, "Copulation")
# Mechanism — combined
mc_gen <- mean_of(fit_comb_mech_hedges, "Generalized")
mc_ind <- mean_of(fit_comb_mech_hedges, "Individual")
# Modality — new
mo_vis <- mean_of(fits_new$modality$hedges, "Visual")
mo_che <- mean_of(fits_new$modality$hedges, "Chemical")
mo_aco <- mean_of(fits_new$modality$hedges, "Acoustic")
```

### Observer virginity — a signal that runs *against* the original meta-analysis {#sec-bio-virginity}

This is the most noteworthy contrast. Jones & DuVal [@jonesMechanismsSocialInfluence2019] reported — and the present manuscript's discussion repeats as support for the "copy when inexperienced" social-learning strategy — that **virgin (sexually naïve) observers copy more strongly**. Our new extraction shows the **opposite ordering**: virgin observers copied *less* than non-virgins (Hedges' $g$ = `r fmt(vh_yes, 2)` for virgins vs `r fmt(vh_no, 2)` for non-virgins; log OR = `r fmt(vo_yes, 2)` vs `r fmt(vo_no, 2)`). The difference is not statistically distinct (CIs overlap zero), but the *sign* is reversed relative to the classic result. When the new data are pooled with the original Jones & DuVal binary data, the contrast all but disappears (non-virgin OR = `r fmt(exp(vc_no), 2)` vs virgin OR = `r fmt(exp(vc_yes), 2)`).

The takeaway is not that the original finding is wrong, but that the **updated, broader evidence base no longer reproduces the virgin-copies-more pattern**, and the new studies, if anything, point the other way. This is precisely the kind of context-dependence the discussion invokes, and it argues for treating the virginity-conditionality claim with more caution than the original synthesis implied.

### Demonstration cue: copulation versus mere association {#sec-bio-cue}

Within the dominant *visual* modality, restricting to the two well-populated cue types shows copying trends **stronger when the social cue is a copulation than mere association** (Hedges' $g$ = `r fmt(dt_cop, 2)` vs `r fmt(dt_assoc, 2)`). This is the biologically expected direction — a completed copulation is a less ambiguous, higher-value social signal than spatial proximity alone — and it offers a concrete, testable prediction for future work, even though the contrast is not significant here.

### Copying mechanism: generalized copying is not the weaker form {#sec-bio-mechanism}

The introduction motivates *generalized* copying (forming a preference for a phenotype rather than an individual) as the route by which mate-choice copying could drive the cultural transmission of preferences and shape sexual selection. Reassuringly for that argument, generalized copying is **no weaker than individual copying** in the combined data (Hedges' $g$ = `r fmt(mc_gen, 2)` generalized vs `r fmt(mc_ind, 2)` individual) — the evolutionarily consequential form of the behaviour is at least as strong as the individual-recognition form.

### Social-cue modality: a sampling gap, not a result {#sec-bio-modality}

The modality comparison (visual $g$ = `r fmt(mo_vis, 2)`, chemical = `r fmt(mo_che, 2)`, acoustic = `r fmt(mo_aco, 2)`) cannot be read biologically: the chemical and acoustic levels rest on a handful of effect sizes from one to three studies and are confounded with particular species. Its real value is as a **map of where the evidence is missing** — the mate-choice-copying literature is overwhelmingly visual, even though the field's own inclusion criteria explicitly admit chemical and acoustic cues. Broadening sensory sampling is a clear opportunity.

---

## Heterogeneity After Moderators {#sec-heterogeneity-after}

The point of this exercise was to see whether any single moderator could absorb a meaningful share of the heterogeneity that the intercept models left unexplained. The table below collects the total $I^2$ and omnibus $Q_M$ for every model fitted in this chapter.

```{r}
#| label: heterogeneity_summary_table
#| echo: false
summarise_fit <- function(res, analysis) {
    if (is.null(res) || is.null(res$model)) {
        return(tibble::tibble(Analysis = analysis, k = NA_integer_, `Total I² (%)` = NA_real_,
                              `Q_M` = NA_real_, `Q_M p` = NA_real_, Structure = "did not fit"))
    }
    a <- among_test(res)
    tibble::tibble(
        Analysis = analysis,
        k = res$model$k,
        `Total I² (%)` = round(total_i2(res), 1),
        `Q_M` = round(a$QM, 2),
        `Q_M p` = signif(a$QMp, 3),
        Structure = res$spec
    )
}

het_summary <- dplyr::bind_rows(
    summarise_fit(fits_new$mech$hedges,      "Mechanism — Hedges' g (new)"),
    summarise_fit(fits_new$mech$or,          "Mechanism — log OR (new)"),
    summarise_fit(fits_new$modality$hedges,  "Modality — Hedges' g (new)"),
    summarise_fit(fits_new$modality$or,      "Modality — log OR (new)"),
    summarise_fit(fits_new$demo_type$hedges, "Demonstration sub-type — Hedges' g (new)"),
    summarise_fit(fits_new$demo_type$or,     "Demonstration sub-type — log OR (new)"),
    summarise_fit(fits_new$virgin$hedges,    "Virginity — Hedges' g (new)"),
    summarise_fit(fits_new$virgin$or,        "Virginity — log OR (new)"),
    summarise_fit(fit_comb_mech_hedges,      "Mechanism — Hedges' g (new + Davies)"),
    summarise_fit(fit_comb_virgin_or,        "Virginity — log OR (new + Jones)")
)

DT::datatable(
    het_summary,
    caption = "Total heterogeneity and omnibus moderator test for every uni-moderator model in this chapter. High total I² alongside a non-significant Q_M means the moderator does not explain the heterogeneity.",
    options = list(dom = "t", pageLength = 12, columnDefs = list(list(className = "dt-center", targets = c(1, 2, 3, 4)))),
    rownames = FALSE
) %>%
    DT::formatStyle("Q_M p", fontWeight = DT::styleInterval(0.05, c("bold", "normal")))
```

### Interpretation {#sec-interpretation}

Across every model, the pattern is the same: **total heterogeneity remains high and no single moderator's omnibus test is significant.** None of copying mechanism, social-cue modality, demonstration sub-type, or observer virginity reduces the heterogeneity to a degree that the among-level test can detect, on either effect-size scale, and the two combined-dataset extensions confirm this once the legacy data are added. The marginalised level means overlap substantially, and partial $I^2$ stays concentrated at the residual effect-size and species levels — precisely the components the overall synthesis flagged as carrying the bulk of the variance.

This is a genuinely informative null. It tells us that the heterogeneity is **not** explained by the most obvious experimental and ecological axes available in the current data, and it sharpens the case made in the discussion: the moderators that matter are likely finer-grained (specific paradigms, demonstrator quality, ecological context) or interactive, and the dataset is not yet large or balanced enough — particularly for the non-visual modalities and for any moderator absent from the legacy datasets — to resolve them. We therefore present these as **hypothesis-generating** results and echo the discussion's recommendation that future, ideally pre-registered, syntheses test these moderators explicitly as the evidence base grows.

That said, the null heterogeneity result and the [directional level differences](#sec-biological) are two different readings of the same models, and the second is where the biological interest lies. The standout is **observer virginity**: the updated data do not reproduce — and trend against — the virgin-copies-more pattern of the original synthesis, a nuance worth carrying into the discussion even though the contrast is not statistically distinct.

::: {.callout-note}
## Caveats
These models are exploratory and were not pre-registered. Several levels (Chemical and Acoustic modality in particular) rest on very few studies and are confounded with individual species, so their estimates are unstable. Absence of a significant $Q_M$ is *not* evidence that a moderator has no effect — only that the current data cannot resolve one.
:::

---

## Session Info

```{r}
#| label: session_info
sessionInfo()
```
