---
title: "Overall Effect Size Calculation & Synthesis"
subtitle: "Meta-Analytic Update of Mate-Choice Copying (MCC)"
author: "Eduardo S. A. Santos & Collaborators"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    theme: cosmo
    self-contained: true
---

```{r}
#| label: setup
#| include: false
# Ensure a CRAN mirror is set so package installation works in non-interactive renders
if (is.null(getOption("repos")) || getOption("repos")["CRAN"] %in% c(NA, "@CRAN@")) {
    options(repos = c(CRAN = "https://cloud.r-project.org"))
}
# Ensure pacman is loaded to manage dependencies
if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
# orchaRd is hosted on GitHub (not CRAN); install from GitHub if missing.
# We require a recent build that ships `pub_bias_plot()` (Yang et al. 2024
# bias-robust correction), so we (re)install from GitHub if the function is
# not yet available in the locally installed version.
if (!requireNamespace("orchaRd", quietly = TRUE) ||
    !("pub_bias_plot" %in% getNamespaceExports("orchaRd"))) {
    pacman::p_load_gh("daniel1noble/orchaRd")
}
pacman::p_load(
    readxl,
    tidyverse,
    metafor,
    orchaRd,
    patchwork,
    DT,
    scales,
    clubSandwich,
    emmeans
)

# Helper function to format numbers to a given precision and clean negative zeros
fmt <- function(x, digits = 2) {
    res <- sprintf(paste0("%.", digits, "f"), x)
    zero_pattern <- paste0("-0.", paste0(rep("0", digits), collapse = ""))
    res[res == zero_pattern] <- paste0("0.", paste0(rep("0", digits), collapse = ""))
    res
}

# Inline p-value formatter (reports "p < 0.0001" for very small p-values)
pstr <- function(p) {
    if (length(p) == 0 || is.na(p)) "p = NA" else if (p < 0.0001) "p < 0.0001" else paste0("p = ", fmt(p, 3))
}

# Significance-word helper: returns `yes` when p < 0.05, else `no`
sigw <- function(p, yes = "statistically significant", no = "non-significant") {
    if (length(p) == 0 || is.na(p)) "indeterminate" else if (p < 0.05) yes else no
}
```

## Introduction & Project Update

This protocol implements the overall effect size synthesis for the updated Mate-Choice Copying (MCC) meta-analysis. 

Specifically, we integrate:

1. **The New Extractions:** Our newly extracted and verified dataset containing recent trials.
2. **Original Continuous Dataset (@daviesMetaanalysisFactorsInfluencing2020):** Standardised mean differences (Hedges' $d$) analyzing continuous metrics.
3. **Original Binary Dataset (@jonesMechanismsSocialInfluence2019):** Log Odds Ratio ($\ln\text{OR}$) analyzing binary choices.

The R script aligns and synthesizes these three sources into unified datasets, allowing comparison of meta-analytic summaries across three key subsets:

* **New Extractions Only:** Eduardo's new dataset.
* **Original Only:** Legacy data from the original publications.
* **Combined Dataset:** A complete synthesis of old and new trials to estimate the updated global effect sizes.

---

## Data Loading and Alignment

We read the three datasets, perform data type conversions, extract species epithets for taxonomic control, and align columns into unified Hedges' $g$ and Odds Ratio datasets.

```{r}
#| label: load_and_align_datasets
#| message: false
#| warning: false
# ---------------------------------------------------------------------------
# 1. Load Datasets
# ---------------------------------------------------------------------------

# Paths to legacy and new datasets
path_new <- here::here("data", "1_effect_size_calculation_pipeline", "Data Extraction Mate Choice Meta Analysis.xlsx")
path_davies <- here::here("data", "original_meta_analyses_datasets", "Davies_et_al_2020_Final_data.xlsx")
path_jones <- here::here("data", "original_meta_analyses_datasets", "Jones_DuVal_2019_data.CSV")

# A. Read new extraction dataset (EduardoExtraction sheet)
db_new_raw <- readxl::read_excel(path_new, sheet = "EduardoExtraction")

# B. Read Davies et al. 2020 (Original continuous dataset)
db_davies_raw <- readxl::read_excel(path_davies, sheet = "Data")

# C. Read Jones & DuVal 2019 (Original binary dataset)
db_jones_raw <- readr::read_csv(path_jones)

# ---------------------------------------------------------------------------
# 1b. Species-name reconciliation + phylogenetic correlation matrices
# ---------------------------------------------------------------------------
# The phylogenetic correlation matrices (built in scripts/2_phylogeny/
# build_phylogeny.R) are indexed by FULL species binomials. The mappings below
# reproduce that script's reconciliation so the `species_phylo` grouping in the
# effect-size data matches the tree tip labels (and matrix dimnames) exactly.
recode_new    <- c("Taeniopygia guttata castanotis" = "Taeniopygia guttata")
recode_davies <- c("Schizocosa" = "Schizocosa ocreata")

# Jones & DuVal store only the species EPITHET; rebuild full binomials from
# Family + epithet (identical to build_phylogeny.R).
jones_binomial <- tibble::tribble(
    ~family,            ~epithet,        ~species_phylo,
    "Adrianichthyidae", "latipes",       "Oryzias latipes",
    "Blenniidae",       "nitidus",       "Rhabdoblennius nitidus",
    "Gasterosteidae",   "aculeatus",     "Gasterosteus aculeatus",
    "Gobiidae",         "minutu",        "Pomatoschistus minutus",
    "Percidae",         "olmstedi",      "Etheostoma olmstedi",
    "Poeciliidae",      "latipinna",     "Poecilia latipinna",
    "Poeciliidae",      "nigrofasciata", "Limia nigrofasciata",
    "Poeciliidae",      "perugiae",      "Limia perugiae",
    "Poeciliidae",      "reticulata",    "Poecilia reticulata",
    "Pomacentridae",    "leucogaster",   "Amblyglyphidodon leucogaster",
    "Lycosidae",        "ocreata",       "Schizocosa ocreata",
    "Estrildidae",      "guttata",       "Taeniopygia guttata",
    "Icteridae",        "ater",          "Molothrus ater",
    "Muscicapidae",     "hypoleuca",     "Ficedula hypoleuca",
    "Drosophilidae",    "melanogaster",  "Drosophila melanogaster",
    "Drosophilidae",    "serrata",       "Drosophila serrata",
    "Ocypodidae",       "mjoebergi",     "Uca mjoebergi"
)

# Phylogenetic correlation matrices (full binomials as dimnames). Loaded
# defensively: if the phylogeny build has not been run, these are NULL and the
# models gracefully fall back to a non-phylogenetic species random effect only.
read_phylo <- function(f) {
    p <- here::here("outputs", "2_phylogeny", f)
    if (file.exists(p)) readRDS(p) else NULL
}
phylo_cor_new    <- read_phylo("phylo_cor_new_extraction.rds") # new-extraction tree
phylo_cor_hedges <- read_phylo("phylo_cor_hedges_model.rds")   # combined Hedges' g species
phylo_cor_or     <- read_phylo("phylo_cor_or_model.rds")       # combined log OR species

# ---------------------------------------------------------------------------
# 2. Process and Filter New Extraction
# ---------------------------------------------------------------------------
db_new <- db_new_raw %>%
    filter(is.na(RemoveFromDataset) | RemoveFromDataset != "Yes") %>%
    mutate(
        effectSizeCalculatedHedgesD = as.numeric(effectSizeCalculatedHedgesD),
        effectSizeCalculatedHedgesDVariance = as.numeric(effectSizeCalculatedHedgesDVariance),
        effectSizeCalculatedOddsRatio = as.numeric(effectSizeCalculatedOddsRatio),
        effectSizeCalculatedOddsRatioVariance = as.numeric(effectSizeCalculatedOddsRatioVariance),
        logOddsRatio = log(effectSizeCalculatedOddsRatio),
        logOddsRatioVariance = effectSizeCalculatedOddsRatioVariance
    )

# ---------------------------------------------------------------------------
# 3. Align Datasets for Hedges' g (Continuous SMD Analysis)
# ---------------------------------------------------------------------------

# A. Hedges' g from New Extraction
hedges_new <- db_new %>%
    filter(!is.na(effectSizeCalculatedHedgesD), !is.na(effectSizeCalculatedHedgesDVariance)) %>%
    select(
        yi = effectSizeCalculatedHedgesD,
        vi = effectSizeCalculatedHedgesDVariance,
        identifierStudyId,
        identifierEffectSizeID,
        taxonomySpecies
    ) %>%
    mutate(
        identifierStudyId = as.character(identifierStudyId),
        identifierEffectSizeID = as.character(identifierEffectSizeID),
        # Full binomial for the phylogeny (matches tree tip labels / matrix dimnames)
        species_phylo = dplyr::recode(str_squish(as.character(taxonomySpecies)), !!!recode_new),
        # Extract species epithet (last word) and make lowercase for cross-dataset mapping
        taxonomySpecies = str_to_lower(word(as.character(taxonomySpecies), -1)),
        Dataset_Source = "New Extraction (Santos)"
    )

# B. Hedges' g from Davies et al. 2020 (Original)
hedges_davies <- db_davies_raw %>%
    mutate(
        Hedges_d_directional = as.numeric(Hedges_d_directional),
        Variance = as.numeric(Variance)
    ) %>%
    filter(!is.na(Hedges_d_directional), !is.na(Variance), Variance > 0) %>%
    select(
        yi = Hedges_d_directional,
        vi = Variance,
        identifierStudyId = Author,
        identifierEffectSizeID = Effect_size_no,
        taxonomySpecies = Species_latin
    ) %>%
    mutate(
        identifierStudyId = as.character(identifierStudyId),
        identifierEffectSizeID = as.character(identifierEffectSizeID),
        # Full binomial for the phylogeny (matches tree tip labels / matrix dimnames)
        species_phylo = dplyr::recode(str_squish(as.character(taxonomySpecies)), !!!recode_davies),
        # Extract species epithet (last word) and make lowercase
        taxonomySpecies = str_to_lower(word(as.character(taxonomySpecies), -1)),
        Dataset_Source = "Original (Davies 2020)"
    )

# C. Unified Hedges' g Dataset
db_hedges <- bind_rows(hedges_new, hedges_davies) %>%
    mutate(across(c(identifierStudyId, taxonomySpecies, species_phylo, Dataset_Source), as.factor))

# ---------------------------------------------------------------------------
# 4. Align Datasets for Odds Ratio (Binary Analysis)
# ---------------------------------------------------------------------------

# A. Log Odds Ratio from New Extraction
or_new <- db_new %>%
    filter(!is.na(logOddsRatio), !is.na(logOddsRatioVariance)) %>%
    select(
        yi = logOddsRatio,
        vi = logOddsRatioVariance,
        identifierStudyId,
        identifierEffectSizeID,
        taxonomySpecies
    ) %>%
    mutate(
        identifierStudyId = as.character(identifierStudyId),
        identifierEffectSizeID = as.character(identifierEffectSizeID),
        # Full binomial for the phylogeny (matches tree tip labels / matrix dimnames)
        species_phylo = dplyr::recode(str_squish(as.character(taxonomySpecies)), !!!recode_new),
        taxonomySpecies = str_to_lower(word(as.character(taxonomySpecies), -1)),
        Dataset_Source = "New Extraction (Santos)"
    )

# B. Log Odds Ratio from Jones & DuVal 2019 (Original)
or_jones <- db_jones_raw %>%
    filter(!is.na(lnOR), !is.na(VlnOR), VlnOR > 0) %>%
    mutate(
        family  = str_squish(as.character(Family)),
        epithet = str_squish(as.character(Species))
    ) %>%
    # Rebuild full binomial (species_phylo) from Family + epithet
    left_join(jones_binomial, by = c("family", "epithet")) %>%
    select(
        yi = lnOR,
        vi = VlnOR,
        identifierStudyId = study,
        identifierEffectSizeID = study_Exp,
        taxonomySpecies = Species,
        species_phylo
    ) %>%
    mutate(
        identifierStudyId = as.character(identifierStudyId),
        identifierEffectSizeID = as.character(identifierEffectSizeID),
        taxonomySpecies = str_to_lower(as.character(taxonomySpecies)),
        Dataset_Source = "Original (Jones & DuVal 2019)"
    )

# Sanity check: every Jones epithet should map to a full binomial
if (any(is.na(or_jones$species_phylo))) {
    warning(
        "Unmapped Jones epithets (phylogeny): ",
        paste(unique(or_jones$taxonomySpecies[is.na(or_jones$species_phylo)]), collapse = ", ")
    )
}

# C. Unified Log Odds Ratio Dataset
db_or <- bind_rows(or_new, or_jones) %>%
    mutate(across(c(identifierStudyId, taxonomySpecies, species_phylo, Dataset_Source), as.factor))
```

---

## Study Inclusion & Exclusion {#sec-inclusion}

The following tables document which effect sizes from the new extraction are included in the meta-analytic models and which have been removed, with the reason for each exclusion.

### Excluded Effect Sizes

```{r}
#| label: excluded_studies_table
#| echo: false
db_excluded <- db_new_raw %>%
    filter(!is.na(RemoveFromDataset) & RemoveFromDataset == "Yes") %>%
    transmute(
        `Effect Size ID` = identifierEffectSizeID,
        `Study` = identifierStudyId,
        `Year` = identifierStudyYear,
        `Exclusion Reason` = ExclusionReason
    ) %>%
    arrange(`Effect Size ID`)

DT::datatable(
    db_excluded,
    caption = "Effect sizes removed from the meta-analytic models and their reasons for exclusion.",
    options = list(
        dom = "t",
        pageLength = 20,
        columnDefs = list(list(className = "dt-left", targets = "_all"))
    ),
    rownames = FALSE
)
```

### Included Effect Sizes

```{r}
#| label: included_studies_table
#| echo: false
db_included <- db_new %>%
    transmute(
        `Effect Size ID` = identifierEffectSizeID,
        `Study` = identifierStudyId,
        `Year` = identifierStudyYear,
        `Species` = taxonomySpecies,
        `Data Type` = effectSizeDataTypeForCalculation,
        `Hedges' g` = round(effectSizeCalculatedHedgesD, 3),
        `OR` = round(effectSizeCalculatedOddsRatio, 3)
    ) %>%
    arrange(`Effect Size ID`)

DT::datatable(
    db_included,
    caption = "Effect sizes included in the meta-analytic models.",
    options = list(
        pageLength = 25,
        columnDefs = list(list(className = "dt-left", targets = "_all"))
    ),
    rownames = FALSE
)
```

---

## Dataset Composition Summary {#sec-composition}

Before inspecting the effect size distributions, we summarize how the aligned datasets are composed. The first table reports the number of studies (`k`) and effect sizes contributed by each original source and by our new extraction. The second table summarizes the number of effect sizes per species, ordered from most to least common.

```{r}
#| label: build_summary_tables
#| message: false
#| warning: false
# Combine both aligned effect-size datasets, tagging the effect-size class.
# Note: the New Extraction (Santos) contributes to BOTH the Hedges' g and the
# Odds Ratio datasets; n_distinct() on study IDs therefore avoids double-counting
# studies while total effect sizes are summed across classes.
db_summary_all <- bind_rows(
    db_hedges %>% mutate(EffectType = "Hedges' g"),
    db_or %>% mutate(EffectType = "Log Odds Ratio")
) %>%
    mutate(across(c(identifierStudyId, taxonomySpecies, Dataset_Source, EffectType), as.character))
```

### Studies & Effect Sizes by Source

```{r}
#| label: source_summary_table
#| echo: false
# Per-source counts of studies and effect sizes, broken down by effect-size class
source_summary <- db_summary_all %>%
    group_by(`Data Source` = Dataset_Source) %>%
    summarise(
        `Studies (k)` = n_distinct(identifierStudyId),
        `Hedges' g (ES)` = sum(EffectType == "Hedges' g"),
        `Log OR (ES)` = sum(EffectType == "Log Odds Ratio"),
        `Total Effect Sizes` = n(),
        `Species` = n_distinct(taxonomySpecies),
        .groups = "drop"
    ) %>%
    arrange(`Data Source`)

# Append an overall TOTAL row (unique studies/species pooled across all sources)
source_total <- tibble(
    `Data Source` = "TOTAL",
    `Studies (k)` = n_distinct(db_summary_all$identifierStudyId),
    `Hedges' g (ES)` = sum(db_summary_all$EffectType == "Hedges' g"),
    `Log OR (ES)` = sum(db_summary_all$EffectType == "Log Odds Ratio"),
    `Total Effect Sizes` = nrow(db_summary_all),
    `Species` = n_distinct(db_summary_all$taxonomySpecies)
)

source_summary_tbl <- bind_rows(source_summary, source_total)

DT::datatable(
    source_summary_tbl,
    caption = "Number of studies and effect sizes contributed by each data source.",
    options = list(
        dom = "t",
        pageLength = 10,
        columnDefs = list(list(className = "dt-center", targets = "_all"))
    ),
    rownames = FALSE
)
```

### Effect Sizes per Species

```{r}
#| label: species_summary_table
#| echo: false
# Effect sizes per species (combining both effect-size classes), most to least common
species_summary <- db_summary_all %>%
    group_by(Species = taxonomySpecies) %>%
    summarise(
        `Studies (k)` = n_distinct(identifierStudyId),
        `Hedges' g (ES)` = sum(EffectType == "Hedges' g"),
        `Log OR (ES)` = sum(EffectType == "Log Odds Ratio"),
        `Total Effect Sizes` = n(),
        .groups = "drop"
    ) %>%
    arrange(desc(`Total Effect Sizes`), Species)

DT::datatable(
    species_summary,
    caption = "Number of effect sizes per species, ordered from most to least common.",
    options = list(
        pageLength = 15,
        order = list(list(3, "desc")),
        columnDefs = list(list(className = "dt-center", targets = "_all"))
    ),
    rownames = FALSE
)
```

In total, the aligned data comprise `r nrow(db_summary_all)` effect sizes from `r n_distinct(db_summary_all$identifierStudyId)` studies across `r n_distinct(db_summary_all$taxonomySpecies)` species. The most-studied species is *`r str_to_title(species_summary$Species[1])`* (`r species_summary$"Total Effect Sizes"[1]` effect sizes from `r species_summary$"Studies (k)"[1]` studies).

---

## Effect Size Distribution (Density Plots) {#sec-density}

Before fitting the multi-level models, we inspect the distribution of our aligned effect sizes across the different subsets (New Extraction, Original Only, and Combined). 

The density plots below use the same split logic as the orchard plots in the result section, facet-wrapping by subset while coloring the distribution curves by their actual dataset source to check for shifts or outliers.

::: {.panel-tabset}

### Hedges' $g$ (Continuous SMD)

```{r}
#| label: hedges_density_plots
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 9
# Prepare density data for Hedges' g
db_hedges_density <- bind_rows(
    db_hedges %>% mutate(Group = as.character(Dataset_Source)),
    db_hedges %>% mutate(Group = "Combined")
) %>%
    mutate(Group = factor(Group, levels = c("New Extraction (Santos)", "Original (Davies 2020)", "Combined")))

# Plot densities
p_hedges_dens <- ggplot(db_hedges_density, aes(x = yi, fill = Dataset_Source, color = Dataset_Source)) +
    geom_density(alpha = 0.4, linewidth = 0.8) +
    geom_vline(xintercept = 0, linewidth = 1.0, linetype = "solid", color = "black") +
    geom_rug(alpha = 0.6) +
    facet_wrap(~Group, ncol = 1) +
    scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
    scale_color_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
    labs(
        title = "Distribution of Hedges' g Effect Sizes",
        x = "Hedges' g",
        y = "Density"
    ) +
    theme_minimal(base_size = 12) +
    theme(
        legend.position = "bottom",
        strip.text = element_text(face = "bold", size = 11),
        plot.title = element_text(face = "bold", hjust = 0.5)
    )

print(p_hedges_dens)
```

### Odds Ratio (Binary outcomes)

```{r}
#| label: or_density_plots
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 9
# Prepare density data for Odds Ratio (lnOR)
db_or_density <- bind_rows(
    db_or %>% mutate(Group = as.character(Dataset_Source)),
    db_or %>% mutate(Group = "Combined")
) %>%
    mutate(Group = factor(Group, levels = c("New Extraction (Santos)", "Original (Jones & DuVal 2019)", "Combined")))

# Plot densities
p_or_dens <- ggplot(db_or_density, aes(x = yi, fill = Dataset_Source, color = Dataset_Source)) +
    geom_density(alpha = 0.4, linewidth = 0.8) +
    geom_vline(xintercept = 0, linewidth = 1.0, linetype = "solid", color = "black") +
    geom_rug(alpha = 0.6) +
    facet_wrap(~Group, ncol = 1) +
    scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal 2019)" = "#0072B2"), name = "Data Source") +
    scale_color_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal 2019)" = "#0072B2"), name = "Data Source") +
    labs(
        title = "Distribution of Log Odds Ratio (lnOR) Effect Sizes",
        x = "Log Odds Ratio (lnOR)",
        y = "Density"
    ) +
    theme_minimal(base_size = 12) +
    theme(
        legend.position = "bottom",
        strip.text = element_text(face = "bold", size = 11),
        plot.title = element_text(face = "bold", hjust = 0.5)
    )

print(p_or_dens)
```

:::

---

## Model Fitting Function

This helper function handles the variance–covariance (VCV) matrix construction and fits a multilevel random-effects model using `rma.mv()` in `metafor`. 

* **Variance–Covariance:** We construct a VCV matrix via `vcalc()` to handle dependencies among multiple effect sizes within the same study (assuming $r = 0.5$ for nested measurements).
* **Random intercepts** are modeled for:

  * Study (`identifierStudyId`)
  * Individual trial/effect size (`identifierEffectSizeID`)
  * Species (`taxonomySpecies`) — the *non-phylogenetic* species effect, capturing species-specific deviations not explained by shared ancestry
  * Phylogeny (`species_phylo`) — a *phylogenetic* random effect supplied with a species-by-species correlation matrix (`R = list(species_phylo = phylo_cor)`) derived from the Open Tree of Life topology with Grafen branch lengths (see `scripts/2_phylogeny/build_phylogeny.R`). This partitions the among-species variance into a component attributable to phylogenetic relatedness and a residual species component.

The phylogenetic term is added to the **uncorrected** models only (for the New, Original, and Combined datasets); the small-study bias-corrected (Egger-type) models retain the original structure. When a phylogenetic correlation matrix is unavailable, the fit gracefully falls back to the non-phylogenetic structure.

```{r}
#| label: model_fitting_function
fit_overall_mv <- function(dat, yi = "yi", vi = "vi", cluster_var = "identifierStudyId", obs_var = "identifierEffectSizeID", rho = 0.5, phylo_cor = NULL, phylo_var = "species_phylo") {
    # Remove rows with missing values
    dat2 <- dat %>%
        filter(!is.na(.data[[yi]]), !is.na(.data[[vi]]), .data[[vi]] > 0)

    if (nrow(dat2) == 0) {
        message("No complete cases found.")
        return(NULL)
    }

    # 1. Build VCV Matrix
    VCV <- metafor::vcalc(
        vi = dat2[[vi]],
        cluster = dat2[[cluster_var]],
        obs = dat2[[obs_var]],
        rho = rho,
        data = dat2
    )

    # 2. Assemble random-effects structure. Always include study, effect-size,
    #    and a (non-phylogenetic) species intercept. If a phylogenetic
    #    correlation matrix is supplied AND every species level is present in it,
    #    add a phylogenetic random effect with R = subset-of-correlation-matrix.
    random_list <- list(
        ~ 1 | identifierStudyId,
        ~ 1 | identifierEffectSizeID,
        ~ 1 | taxonomySpecies
    )
    R_arg <- NULL
    used_phylo <- FALSE

    if (!is.null(phylo_cor) && phylo_var %in% colnames(dat2)) {
        sp <- unique(as.character(dat2[[phylo_var]]))
        missing_sp <- setdiff(sp, rownames(phylo_cor))
        if (length(missing_sp) == 0) {
            # Subset the correlation matrix to the species actually present so the
            # dimnames match the grouping levels exactly. Also drop any stale
            # (unused) factor levels on the grouping column so metafor's level
            # set matches the R-matrix dimnames.
            dat2[[phylo_var]] <- factor(as.character(dat2[[phylo_var]]))
            Csub <- phylo_cor[sp, sp, drop = FALSE]
            random_list <- c(random_list, list(stats::as.formula(paste0("~ 1 | ", phylo_var))))
            R_arg <- stats::setNames(list(Csub), phylo_var)
            used_phylo <- TRUE
        } else {
            warning(
                "Phylogenetic random effect skipped (species absent from correlation matrix): ",
                paste(missing_sp, collapse = ", ")
            )
        }
    }

    # 3. Fit Multi-Level Random-Effects Model
    mod <- metafor::rma.mv(
        yi = dat2[[yi]],
        V = VCV,
        random = random_list,
        R = R_arg,
        test = "t",
        method = "REML",
        sparse = TRUE,
        data = dat2
    )

    list(data = dat2, VCV = VCV, model = mod, used_phylo = used_phylo)
}

# Helper function to fit bias-corrected multilevel models with a continuous moderator (Egger-type)
fit_corrected_mv <- function(dat, yi = "yi", vi = "vi", moderator = "vi", cluster_var = "identifierStudyId", obs_var = "identifierEffectSizeID", rho = 0.5) {
    # Remove rows with missing values
    dat2 <- dat %>%
        filter(!is.na(.data[[yi]]), !is.na(.data[[vi]]), .data[[vi]] > 0)

    if (nrow(dat2) == 0) {
        message("No complete cases found.")
        return(NULL)
    }

    if (moderator == "se" && !"se" %in% colnames(dat2)) {
        dat2 <- dat2 %>% mutate(se = sqrt(vi))
    }

    # 1. Build VCV Matrix
    VCV <- metafor::vcalc(
        vi = dat2[[vi]],
        cluster = dat2[[cluster_var]],
        obs = dat2[[obs_var]],
        rho = rho,
        data = dat2
    )

    # 2. Fit Multi-Level Egger Model using dynamic formula
    formula_str <- paste("~ 1 +", moderator)
    mod <- metafor::rma.mv(
        yi = dat2[[yi]],
        V = VCV,
        mods = as.formula(formula_str),
        random = list(
            ~ 1 | identifierStudyId,
            ~ 1 | identifierEffectSizeID,
            ~ 1 | taxonomySpecies
        ),
        test = "t",
        method = "REML",
        sparse = TRUE,
        data = dat2
    )

    list(data = dat2, VCV = VCV, model = mod)
}

# Helper function to add Mean (95% CI) labels dynamically one line above k-values in orchard plots
add_ci_labels <- function(plot_obj, orchard_results) {
    est <- sprintf("%.2f", orchard_results$mod_table$estimate)
    ci_l <- sprintf("%.2f", orchard_results$mod_table$lowerCL)
    ci_u <- sprintf("%.2f", orchard_results$mod_table$upperCL)

    orig_labels <- plot_obj$layers[[6]]$aes_params$label

    new_labels <- c()
    for (i in 1:length(orig_labels)) {
        # Extract the original sample size numbers using regex
        nums <- regmatches(orig_labels[i], gregexpr("[0-9]+", orig_labels[i]))[[1]]
        if (length(nums) >= 2) {
            k_val <- nums[1]
            n_val <- nums[2]
            k_text <- paste0("k = ", k_val, " (", n_val, ")")
        } else {
            k_text <- orig_labels[i]
        }
        # Place Mean (95% CI) on the first line and k = xx (xx) on the second line
        new_labels[i] <- paste0(est[i], " [", ci_l[i], ", ", ci_u[i], "]\n", k_text)
    }

    plot_obj$layers[[6]]$aes_params$label <- new_labels
    plot_obj$layers[[6]]$geom_params$parse <- FALSE
    plot_obj$layers[[6]]$aes_params$hjust <- "right"
    plot_obj$layers[[6]]$aes_params$vjust <- 0.5
    plot_obj$layers[[6]]$aes_params$size <- 3.0 # Slightly smaller font size to fit better

    plot_obj
}
```

---

## Multi-Level Synthesis Results

We fit the multilevel meta-analysis across all three database splits (New Only, Original Only, and Combined) for both effect size classes.

```{r}
#| label: fit_multilevel_models
#| message: false
#| warning: false
# ---------------------------------------------------------------------------
# HEDGES' G MODELS (UNCORRECTED) — now incorporating phylogenetic relatedness
# ---------------------------------------------------------------------------
# New-only uses the new-extraction tree; Original-only and Combined use the
# combined Hedges' g tree (Davies species are a subset of the combined tree).
res_hedges_new <- fit_overall_mv(db_hedges %>% filter(Dataset_Source == "New Extraction (Santos)"), phylo_cor = phylo_cor_new)
res_hedges_orig <- fit_overall_mv(db_hedges %>% filter(Dataset_Source == "Original (Davies 2020)"), phylo_cor = phylo_cor_hedges)
res_hedges_comb <- fit_overall_mv(db_hedges, phylo_cor = phylo_cor_hedges)

# HEDGES' G MODELS (BIAS-CORRECTED)
res_hedges_new_corr <- fit_corrected_mv(db_hedges %>% filter(Dataset_Source == "New Extraction (Santos)"), moderator = "vi")
res_hedges_orig_corr <- fit_corrected_mv(db_hedges %>% filter(Dataset_Source == "Original (Davies 2020)"), moderator = "vi")
res_hedges_comb_corr <- fit_corrected_mv(db_hedges, moderator = "vi")

# ---------------------------------------------------------------------------
# ODDS RATIO MODELS (UNCORRECTED) — now incorporating phylogenetic relatedness
# ---------------------------------------------------------------------------
# New-only uses the new-extraction tree; Original-only and Combined use the
# combined log OR tree (Jones species are a subset of the combined tree).
res_or_new <- fit_overall_mv(db_or %>% filter(Dataset_Source == "New Extraction (Santos)"), phylo_cor = phylo_cor_new)
res_or_orig <- fit_overall_mv(db_or %>% filter(Dataset_Source == "Original (Jones & DuVal 2019)"), phylo_cor = phylo_cor_or)
res_or_comb <- fit_overall_mv(db_or, phylo_cor = phylo_cor_or)

# ODDS RATIO MODELS (BIAS-CORRECTED)
res_or_new_corr <- fit_corrected_mv(db_or %>% filter(Dataset_Source == "New Extraction (Santos)"), moderator = "vi")
res_or_orig_corr <- fit_corrected_mv(db_or %>% filter(Dataset_Source == "Original (Jones & DuVal 2019)"), moderator = "vi")
res_or_comb_corr <- fit_corrected_mv(db_or, moderator = "vi")
```

---

## Results and Visualizations

::: {.panel-tabset}

### Hedges' $g$ (Continuous outcomes)

::: {.panel-tabset}

#### Uncorrected Overall Effect

This plot displays the standard multi-level meta-analytic summaries across the three splits (New Only, Original Only, and Combined) *without* adjusting for publication bias. Points represent individual effect sizes colored by their data source.

```{r}
#| label: hedges_unified_plot
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 7
if (!is.null(res_hedges_new) && !is.null(res_hedges_orig) && !is.null(res_hedges_comb)) {
    # 1. Extract S3 result objects using mod_results
    h_new <- orchaRd::mod_results(res_hedges_new$model, group = "identifierStudyId")
    h_orig <- orchaRd::mod_results(res_hedges_orig$model, group = "identifierStudyId")
    h_comb <- orchaRd::mod_results(res_hedges_comb$model, group = "identifierStudyId")

    # 2. Clone S3 result objects to lists for manual editing
    class(h_new) <- "list"
    h_new$mod_table$name <- "New Extraction (Santos)"
    h_new$data$moderator <- "New Extraction (Santos)"
    class(h_new) <- c("orchard", "data.frame")

    class(h_orig) <- "list"
    h_orig$mod_table$name <- "Original (Davies 2020)"
    h_orig$data$moderator <- "Original (Davies 2020)"
    class(h_orig) <- c("orchard", "data.frame")

    class(h_comb) <- "list"
    h_comb$mod_table$name <- "Combined"
    h_comb$data$moderator <- "Combined"
    class(h_comb) <- c("orchard", "data.frame")

    # 3. Merge model results
    sub_hedges <- orchaRd::submerge(h_new, h_orig, h_comb)

    # 4. Add point-level real source mapping
    class(sub_hedges) <- "list"
    sub_hedges$data$real_source <- c(
        rep("New Extraction (Santos)", nrow(h_new$data)),
        rep("Original (Davies 2020)", nrow(h_orig$data)),
        rep("New Extraction (Santos)", nrow(h_new$data)),
        rep("Original (Davies 2020)", nrow(h_orig$data))
    )
    class(sub_hedges) <- c("orchard", "data.frame")

    # 4. Generate Plot
    p_hedges <- orchaRd::orchard_plot(sub_hedges, xlab = "Hedges' g", flip = TRUE)

    # 5. Customize color mapping for points
    p_hedges$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
    p_hedges$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

    # 6. Apply themes & custom palette (colorblind friendly orange/blue)
    p_hedges <- p_hedges +
        ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
        ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
        ggplot2::ggtitle("Overall Synthesis — Hedges' g Estimates (Uncorrected)") +
        ggplot2::theme(
            legend.position = c(0.98, 0.02),
            legend.justification = c("right", "bottom"),
            legend.background = ggplot2::element_rect(fill = scales::alpha("white", 0.75), color = "grey90"),
            legend.box = "vertical",
            legend.margin = ggplot2::margin(t = 4, r = 4, b = 4, l = 4),
            legend.title = ggplot2::element_text(size = 9, face = "bold"),
            legend.text = ggplot2::element_text(size = 8)
        )

    # Add Mean (95% CI) and k labels
    p_hedges <- add_ci_labels(p_hedges, sub_hedges)

    print(p_hedges)
}
```

#### Bias-Corrected Overall Effect

This plot displays the small-study publication bias corrected meta-analytic summaries (evaluated at $\text{V} = 0$, representing the theoretical estimate for a study of infinite sample size/precision) using the two-step Egger-type meta-regression framework. Points represent the same raw effect sizes.

```{r}
#| label: hedges_corrected_plot
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 7
if (!is.null(res_hedges_new_corr) && !is.null(res_hedges_orig_corr) && !is.null(res_hedges_comb_corr)) {
    # 1. Extract S3 result objects using mod_results at vi = 0
    h_new_corr <- orchaRd::mod_results(res_hedges_new_corr$model, group = "identifierStudyId", at = list(vi = 0), mod = "1")
    h_orig_corr <- orchaRd::mod_results(res_hedges_orig_corr$model, group = "identifierStudyId", at = list(vi = 0), mod = "1")
    h_comb_corr <- orchaRd::mod_results(res_hedges_comb_corr$model, group = "identifierStudyId", at = list(vi = 0), mod = "1")

    # 2. Clone S3 result objects to lists for manual editing
    class(h_new_corr) <- "list"
    h_new_corr$mod_table$name <- "New Extraction (Santos)"
    h_new_corr$data$moderator <- "New Extraction (Santos)"
    class(h_new_corr) <- c("orchard", "data.frame")

    class(h_orig_corr) <- "list"
    h_orig_corr$mod_table$name <- "Original (Davies 2020)"
    h_orig_corr$data$moderator <- "Original (Davies 2020)"
    class(h_orig_corr) <- c("orchard", "data.frame")

    class(h_comb_corr) <- "list"
    h_comb_corr$mod_table$name <- "Combined"
    h_comb_corr$data$moderator <- "Combined"
    class(h_comb_corr) <- c("orchard", "data.frame")

    # 3. Merge model results
    sub_hedges_corr <- orchaRd::submerge(h_new_corr, h_orig_corr, h_comb_corr)

    # 4. Add point-level real source mapping
    class(sub_hedges_corr) <- "list"
    sub_hedges_corr$data$real_source <- c(
        rep("New Extraction (Santos)", nrow(h_new_corr$data)),
        rep("Original (Davies 2020)", nrow(h_orig_corr$data)),
        rep("New Extraction (Santos)", nrow(h_new_corr$data)),
        rep("Original (Davies 2020)", nrow(h_orig_corr$data))
    )
    class(sub_hedges_corr) <- c("orchard", "data.frame")

    # 4. Generate Plot
    p_hedges_corr <- orchaRd::orchard_plot(sub_hedges_corr, xlab = "Hedges' g (Bias Corrected)", flip = TRUE)

    # 5. Customize color mapping for points
    p_hedges_corr$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
    p_hedges_corr$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

    # 6. Apply themes & custom palette
    p_hedges_corr <- p_hedges_corr +
        ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
        ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
        ggplot2::ggtitle("Overall Synthesis — Hedges' g Estimates (Bias-Corrected)") +
        ggplot2::theme(
            legend.position = c(0.98, 0.02),
            legend.justification = c("right", "bottom"),
            legend.background = ggplot2::element_rect(fill = scales::alpha("white", 0.75), color = "grey90"),
            legend.box = "vertical",
            legend.margin = ggplot2::margin(t = 4, r = 4, b = 4, l = 4),
            legend.title = ggplot2::element_text(size = 9, face = "bold"),
            legend.text = ggplot2::element_text(size = 8)
        )

    # Add Mean (95% CI) and k labels
    p_hedges_corr <- add_ci_labels(p_hedges_corr, sub_hedges_corr)

    print(p_hedges_corr)
}
```

#### Side-by-Side Comparison

We combine the uncorrected and small-study bias corrected orchard plots side-by-side using `patchwork` to directly evaluate the shifts in pooled estimates and prediction intervals when publication selection is accounted for.

```{r}
#| label: hedges_comparison_plot
#| message: false
#| warning: false
#| fig.width: 14
#| fig.height: 8
if (exists("p_hedges") && exists("p_hedges_corr")) {
    comparison_hedges <- (p_hedges + p_hedges_corr) +
        patchwork::plot_annotation(
            title = "Comparison of Uncorrected vs. Bias-Corrected Synthesis (Hedges' g)",
            subtitle = "Bias-corrected estimates are conditioned on study variance = 0 (infinite precision)",
            theme = ggplot2::theme(
                plot.title = ggplot2::element_text(size = 16, face = "bold", hjust = 0.5),
                plot.subtitle = ggplot2::element_text(size = 12, italic = TRUE, hjust = 0.5)
            )
        )
    print(comparison_hedges)
}
```

#### Detailed Model Statistics

Here are the text outputs and fit statistics for the three database splits. Each split displays the uncorrected and small-study bias-corrected (Egger's variance-adjusted) model summaries:

::: {.panel-tabset}

##### New Only

::: {.panel-tabset}

###### Uncorrected
```{r}
#| label: hedges_new_stats
if (!is.null(res_hedges_new)) {
    print(summary(res_hedges_new$model))
    cat("\nHeterogeneity (I^2):\n")
    print(orchaRd::i2_ml(res_hedges_new$model))
}
```

###### Bias-Corrected
```{r}
#| label: hedges_new_corr_stats
if (!is.null(res_hedges_new_corr)) {
    print(summary(res_hedges_new_corr$model))
    cat("\nHeterogeneity (I^2):\n")
    print(orchaRd::i2_ml(res_hedges_new_corr$model))
}
```

:::

##### Original Only

::: {.panel-tabset}

###### Uncorrected
```{r}
#| label: hedges_orig_stats
if (!is.null(res_hedges_orig)) {
    print(summary(res_hedges_orig$model))
    cat("\nHeterogeneity (I^2):\n")
    print(orchaRd::i2_ml(res_hedges_orig$model))
}
```

###### Bias-Corrected
```{r}
#| label: hedges_orig_corr_stats
if (!is.null(res_hedges_orig_corr)) {
    print(summary(res_hedges_orig_corr$model))
    cat("\nHeterogeneity (I^2):\n")
    print(orchaRd::i2_ml(res_hedges_orig_corr$model))
}
```

:::

##### Combined

::: {.panel-tabset}

###### Uncorrected
```{r}
#| label: hedges_comb_stats
if (!is.null(res_hedges_comb)) {
    print(summary(res_hedges_comb$model))
    cat("\nHeterogeneity (I^2):\n")
    print(orchaRd::i2_ml(res_hedges_comb$model))
}
```

###### Bias-Corrected
```{r}
#| label: hedges_comb_corr_stats
if (!is.null(res_hedges_comb_corr)) {
    print(summary(res_hedges_comb_corr$model))
    cat("\nHeterogeneity (I^2):\n")
    print(orchaRd::i2_ml(res_hedges_comb_corr$model))
}
```

:::

:::

:::

### Odds Ratio (Binary outcomes)

::: {.panel-tabset}

#### Uncorrected Overall Effect

This plot displays the standard multi-level meta-analytic summaries across the three splits (New Only, Original Only, and Combined) *without* adjusting for publication bias. Points represent individual log odds ratios colored by their data source.

```{r}
#| label: or_unified_plot
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 7
if (!is.null(res_or_new) && !is.null(res_or_orig) && !is.null(res_or_comb)) {
    # 1. Extract S3 result objects using mod_results
    o_new <- orchaRd::mod_results(res_or_new$model, group = "identifierStudyId")
    o_orig <- orchaRd::mod_results(res_or_orig$model, group = "identifierStudyId")
    o_comb <- orchaRd::mod_results(res_or_comb$model, group = "identifierStudyId")

    # 2. Clone S3 result objects to lists for manual editing
    class(o_new) <- "list"
    o_new$mod_table$name <- "New Extraction (Santos)"
    o_new$data$moderator <- "New Extraction (Santos)"
    class(o_new) <- c("orchard", "data.frame")

    class(o_orig) <- "list"
    o_orig$mod_table$name <- "Original (Jones & DuVal 2019)"
    o_orig$data$moderator <- "Original (Jones & DuVal 2019)"
    class(o_orig) <- c("orchard", "data.frame")

    class(o_comb) <- "list"
    o_comb$mod_table$name <- "Combined"
    o_comb$data$moderator <- "Combined"
    class(o_comb) <- c("orchard", "data.frame")

    # 3. Merge model results
    sub_or <- orchaRd::submerge(o_new, o_orig, o_comb)

    # 4. Add point-level real source mapping
    class(sub_or) <- "list"
    sub_or$data$real_source <- c(
        rep("New Extraction (Santos)", nrow(o_new$data)),
        rep("Original (Jones & DuVal)", nrow(o_orig$data)),
        rep("New Extraction (Santos)", nrow(o_new$data)),
        rep("Original (Jones & DuVal)", nrow(o_orig$data))
    )
    class(sub_or) <- c("orchard", "data.frame")

    # 4. Generate Plot
    p_or <- orchaRd::orchard_plot(sub_or, xlab = "Log Odds Ratio (lnOR)", flip = TRUE)

    # 5. Customize color mapping for points
    p_or$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
    p_or$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

    # 6. Apply themes & custom palette (colorblind friendly orange/blue)
    p_or <- p_or +
        ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal)" = "#0072B2"), name = "Data Source") +
        ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal)" = "#0072B2"), name = "Data Source") +
        ggplot2::ggtitle("Overall Synthesis — Log Odds Ratio Estimates (Uncorrected)") +
        ggplot2::theme(
            legend.position = c(0.98, 0.02),
            legend.justification = c("right", "bottom"),
            legend.background = ggplot2::element_rect(fill = scales::alpha("white", 0.75), color = "grey90"),
            legend.box = "vertical",
            legend.margin = ggplot2::margin(t = 4, r = 4, b = 4, l = 4),
            legend.title = ggplot2::element_text(size = 9, face = "bold"),
            legend.text = ggplot2::element_text(size = 8)
        )

    # Add Mean (95% CI) and k labels
    p_or <- add_ci_labels(p_or, sub_or)

    print(p_or)
}
```

#### Bias-Corrected Overall Effect

This plot displays the small-study publication bias corrected meta-analytic summaries (evaluated at $\text{V} = 0$, representing the theoretical estimate for a study of infinite sample size/precision) using the two-step Egger-type meta-regression framework. Points represent the same raw effect sizes.

```{r}
#| label: or_corrected_plot
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 7
if (!is.null(res_or_new_corr) && !is.null(res_or_orig_corr) && !is.null(res_or_comb_corr)) {
    # 1. Extract S3 result objects using mod_results at vi = 0
    o_new_corr <- orchaRd::mod_results(res_or_new_corr$model, group = "identifierStudyId", at = list(vi = 0), mod = "1")
    o_orig_corr <- orchaRd::mod_results(res_or_orig_corr$model, group = "identifierStudyId", at = list(vi = 0), mod = "1")
    o_comb_corr <- orchaRd::mod_results(res_or_comb_corr$model, group = "identifierStudyId", at = list(vi = 0), mod = "1")

    # 2. Clone S3 result objects to lists for manual editing
    class(o_new_corr) <- "list"
    o_new_corr$mod_table$name <- "New Extraction (Santos)"
    o_new_corr$data$moderator <- "New Extraction (Santos)"
    class(o_new_corr) <- c("orchard", "data.frame")

    class(o_orig_corr) <- "list"
    o_orig_corr$mod_table$name <- "Original (Jones & DuVal 2019)"
    o_orig_corr$data$moderator <- "Original (Jones & DuVal 2019)"
    class(o_orig_corr) <- c("orchard", "data.frame")

    class(o_comb_corr) <- "list"
    o_comb_corr$mod_table$name <- "Combined"
    o_comb_corr$data$moderator <- "Combined"
    class(o_comb_corr) <- c("orchard", "data.frame")

    # 3. Merge model results
    sub_or_corr <- orchaRd::submerge(o_new_corr, o_orig_corr, o_comb_corr)

    # 4. Add point-level real source mapping
    class(sub_or_corr) <- "list"
    sub_or_corr$data$real_source <- c(
        rep("New Extraction (Santos)", nrow(o_new_corr$data)),
        rep("Original (Jones & DuVal)", nrow(o_orig_corr$data)),
        rep("New Extraction (Santos)", nrow(o_new_corr$data)),
        rep("Original (Jones & DuVal)", nrow(o_orig_corr$data))
    )
    class(sub_or_corr) <- c("orchard", "data.frame")

    # 4. Generate Plot
    p_or_corr <- orchaRd::orchard_plot(sub_or_corr, xlab = "Log Odds Ratio (lnOR) (Bias Corrected)", flip = TRUE)

    # 5. Customize color mapping for points
    p_or_corr$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
    p_or_corr$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

    # 6. Apply themes & custom palette
    p_or_corr <- p_or_corr +
        ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal)" = "#0072B2"), name = "Data Source") +
        ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal)" = "#0072B2"), name = "Data Source") +
        ggplot2::ggtitle("Overall Synthesis — Log Odds Ratio Estimates (Bias-Corrected)") +
        ggplot2::theme(
            legend.position = c(0.98, 0.02),
            legend.justification = c("right", "bottom"),
            legend.background = ggplot2::element_rect(fill = scales::alpha("white", 0.75), color = "grey90"),
            legend.box = "vertical",
            legend.margin = ggplot2::margin(t = 4, r = 4, b = 4, l = 4),
            legend.title = ggplot2::element_text(size = 9, face = "bold"),
            legend.text = ggplot2::element_text(size = 8)
        )

    # Add Mean (95% CI) and k labels
    p_or_corr <- add_ci_labels(p_or_corr, sub_or_corr)

    print(p_or_corr)
}
```

#### Side-by-Side Comparison

We combine the uncorrected and small-study bias corrected orchard plots side-by-side using `patchwork` to directly evaluate the shifts in pooled estimates and prediction intervals when publication selection is accounted for.

```{r}
#| label: or_comparison_plot
#| message: false
#| warning: false
#| fig.width: 14
#| fig.height: 8
if (exists("p_or") && exists("p_or_corr")) {
    comparison_or <- (p_or + p_or_corr) +
        patchwork::plot_annotation(
            title = "Comparison of Uncorrected vs. Bias-Corrected Synthesis (Odds Ratio)",
            subtitle = "Bias-corrected estimates are conditioned on study variance = 0 (infinite precision)",
            theme = ggplot2::theme(
                plot.title = ggplot2::element_text(size = 16, face = "bold", hjust = 0.5),
                plot.subtitle = ggplot2::element_text(size = 12, italic = TRUE, hjust = 0.5)
            )
        )
    print(comparison_or)
}
```

#### Detailed Model Statistics

Here are the text outputs and fit statistics for the three database splits. Each split displays the uncorrected and small-study bias-corrected (Egger's variance-adjusted) model summaries:

::: {.panel-tabset}

##### New Only

::: {.panel-tabset}

###### Uncorrected
```{r}
#| label: or_new_stats
if (!is.null(res_or_new)) {
    print(summary(res_or_new$model))
    cat("\nHeterogeneity (I^2):\n")
    print(orchaRd::i2_ml(res_or_new$model))
}
```

###### Bias-Corrected
```{r}
#| label: or_new_corr_stats
if (!is.null(res_or_new_corr)) {
    print(summary(res_or_new_corr$model))
    cat("\nHeterogeneity (I^2):\n")
    print(orchaRd::i2_ml(res_or_new_corr$model))
}
```

:::

##### Original Only

::: {.panel-tabset}

###### Uncorrected
```{r}
#| label: or_orig_stats
if (!is.null(res_or_orig)) {
    print(summary(res_or_orig$model))
    cat("\nHeterogeneity (I^2):\n")
    print(orchaRd::i2_ml(res_or_orig$model))
}
```

###### Bias-Corrected
```{r}
#| label: or_orig_corr_stats
if (!is.null(res_or_orig_corr)) {
    print(summary(res_or_orig_corr$model))
    cat("\nHeterogeneity (I^2):\n")
    print(orchaRd::i2_ml(res_or_orig_corr$model))
}
```

:::

##### Combined

::: {.panel-tabset}

###### Uncorrected
```{r}
#| label: or_comb_stats
if (!is.null(res_or_comb)) {
    print(summary(res_or_comb$model))
    cat("\nHeterogeneity (I^2):\n")
    print(orchaRd::i2_ml(res_or_comb$model))
}
```

###### Bias-Corrected
```{r}
#| label: or_comb_corr_stats
if (!is.null(res_or_comb_corr)) {
    print(summary(res_or_comb_corr$model))
    cat("\nHeterogeneity (I^2):\n")
    print(orchaRd::i2_ml(res_or_comb_corr$model))
}
```

:::

:::

:::

:::

---

## Combined Correction Comparison {#sec-correction-comparison}

The tabs above present the uncorrected and bias-corrected estimates on separate plots. To make the comparison easier to read, we produce **two figures** — one for Hedges' $g$ and one for the log odds ratio — using the same `orchaRd::pub_bias_plot()` approach as the publication-bias chapter. Each figure stacks the three model levels (New Extraction, Original, Combined) as rows; within each row a single orchard trunk shows the uncorrected multilevel mean drawn over its raw effect sizes **once**, with the two corrected means overlaid on top. The raw points are coloured by data source and the legend is boxed in the corner, matching the side-by-side comparison plots above.

The three estimates shown per subplot are:

* **Uncorrected** (the trunk) — the multilevel `rma.mv` mean.
* **Bias-Corrected (Nakagawa et al. [-@nakagawaMethodsTestingPublication2022]); purple overlay** — the intercept of the multilevel meta-regression with sampling variance ($v_i$) as a moderator, evaluated at $v_i = 0$ (the effect a hypothetical infinitely-precise study would report).
* **Bias-Robust (Yang et al. [-@yangRobustPointVariance2024]); green overlay** — a fixed-/common-effect model with cluster-robust (CR2) variance estimation clustered by study.

```{r}
#| label: fit_bias_robust_splits
#| message: false
#| warning: false
# Yang et al. (2024) bias-robust estimate (fixed-effect + CR2 cluster-robust),
# fitted for each database split. Mirrors the construction in publication_bias.qmd
# but applied to New, Original, and Combined subsets.
fit_bias_robust <- function(dat, yi = "yi", vi = "vi", cluster_var = "identifierStudyId") {
    dat2 <- dat %>% filter(!is.na(.data[[yi]]), !is.na(.data[[vi]]), .data[[vi]] > 0)
    if (nrow(dat2) == 0) return(NULL)
    tryCatch({
        fe <- metafor::rma(yi = dat2[[yi]], vi = dat2[[vi]], method = "FE", test = "t")
        metafor::robust(fe, cluster = dat2[[cluster_var]], clubSandwich = TRUE)
    }, error = function(e) {
        warning("Bias-robust fit failed: ", conditionMessage(e)); NULL
    })
}

rob_hedges_new  <- fit_bias_robust(db_hedges %>% filter(Dataset_Source == "New Extraction (Santos)"))
rob_hedges_orig <- fit_bias_robust(db_hedges %>% filter(Dataset_Source == "Original (Davies 2020)"))
rob_hedges_comb <- fit_bias_robust(db_hedges)

rob_or_new  <- fit_bias_robust(db_or %>% filter(Dataset_Source == "New Extraction (Santos)"))
rob_or_orig <- fit_bias_robust(db_or %>% filter(Dataset_Source == "Original (Jones & DuVal 2019)"))
rob_or_comb <- fit_bias_robust(db_or)
```

```{r}
#| label: correction_comparison_builder
#| message: false
#| warning: false
# ---------------------------------------------------------------------------
# Subplot builder: one orchard trunk per model level, styled to match the
# `hedges_comparison_plot` aesthetics — raw points coloured by data source
# (colourblind orange/blue) with the boxed corner legend.
#
# The two corrected means are overlaid by hand (rather than via
# orchaRd::pub_bias_plot(), whose colours are hardcoded red/blue and whose text
# sits on top of the data) so that we can:
#   * use overlay colours that stand out from the orange/blue raw points
#     (green = Bias-Robust / Yang 2024, purple = Bias-Corrected / Nakagawa 2022);
#   * right-align the estimate text on the far side of the panel, kept inside
#     the plot box;
#   * share a common x-axis (effect-size) range across the stacked panels.
# ---------------------------------------------------------------------------
source_palette <- c(
    "New Extraction (Santos)"       = "#D55E00",
    "Original (Davies 2020)"        = "#0072B2",
    "Original (Jones & DuVal 2019)" = "#0072B2"
)

# Overlay colours — chosen to be distinct from the orange/blue raw points and
# colourblind-distinguishable (Okabe-Ito green + a strong violet).
yang_col <- "#009E73"  # Bias-Robust (Yang et al. 2024)
naka_col <- "#7B2D8E"  # Bias-Corrected (Nakagawa et al. 2022)

# Extract the corrected meta-analytic mean + 95% CI from an (intercept-only or
# vi-moderated) rma / rma.mv model — same intercept used by orchaRd::get_ints_dat.
corrected_est <- function(model) {
    b <- tryCatch(as.numeric(model$b["intrcpt", ]),
                  error = function(e) as.numeric(model$b[1]))
    data.frame(name = "Intrcpt", pred = b,
               ci.lb = model$ci.lb[1], ci.ub = model$ci.ub[1])
}

# src: either a single data-source label (recycled) or a per-point character
# vector matching the order of the model's effect sizes (used for Combined).
# xlim: optional c(lo, hi) shared effect-size range for axis alignment; when
# supplied, the estimate text is right-aligned to xlim[2] (inside the box).
make_pub_bias_subplot <- function(unc_res, corr_res, rob_model, src, xlab, title, xlim = NULL) {
    if (is.null(unc_res) || is.null(unc_res$model) ||
        is.null(corr_res) || is.null(corr_res$model) || is.null(rob_model)) {
        return(NULL)
    }
    # Single-trunk orchard result; attach the true per-point data source.
    h <- orchaRd::mod_results(unc_res$model, group = "identifierStudyId")
    if (length(src) == 1) src <- rep(src, nrow(h$data))
    class(h) <- "list"
    h$data$real_source <- src
    class(h) <- c("orchard", "data.frame")

    # k/g labels to the left so the right margin is free for the estimate text.
    # Widen point.size range so bubble area tracks precision more visibly.
    # trunk.size = 1 -> point estimate drawn at size 4 (matches the overlay diamonds).
    oc <- orchaRd::orchard_plot(h, xlab = xlab, flip = TRUE, k.pos = "left",
                                point.size = c(0.8, 10), trunk.size = 1)

    # Recolour the raw points (layer 1) by data source — as in hedges_comparison_plot.
    oc$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
    oc$layers[[1]]$mapping$fill   <- ggplot2::aes(fill = real_source)$fill

    # Corrected estimates to overlay on the single uncorrected trunk.
    yang <- corrected_est(rob_model)       # Bias-Robust   (Yang 2024)
    naka <- corrected_est(corr_res$model)  # Bias-Corrected (Nakagawa 2022)

    # Common effect-size range (falls back to this panel's data if not supplied).
    if (is.null(xlim)) {
        rng <- range(c(h$data$yi, yang$ci.lb, yang$ci.ub, naka$ci.lb, naka$ci.ub),
                     na.rm = TRUE, finite = TRUE)
        pad <- diff(rng) * 0.08
        xlim <- c(rng[1] - pad, rng[2] + pad)
    }
    txt_y <- xlim[2]  # right edge (effect-size axis is horizontal after flip)

    oc <- oc +
        # Bias-Robust (Yang) — nudged below the trunk. Black CI bar; coloured
        # diamond with a black border so it stands out against the bubbles.
        ggplot2::geom_linerange(data = yang,
            ggplot2::aes(x = name, ymin = ci.lb, ymax = ci.ub),
            colour = "black", linewidth = 1.2,
            position = ggplot2::position_nudge(x = -0.08), inherit.aes = FALSE) +
        ggplot2::geom_point(data = yang,
            ggplot2::aes(x = name, y = pred),
            fill = yang_col, colour = "black", shape = 23, size = 4, stroke = 0.9,
            position = ggplot2::position_nudge(x = -0.08), inherit.aes = FALSE) +
        # Bias-Corrected (Nakagawa) — nudged above the trunk. Black CI bar;
        # coloured diamond with a black border.
        ggplot2::geom_linerange(data = naka,
            ggplot2::aes(x = name, ymin = ci.lb, ymax = ci.ub),
            colour = "black", linewidth = 1.2,
            position = ggplot2::position_nudge(x = 0.08), inherit.aes = FALSE) +
        ggplot2::geom_point(data = naka,
            ggplot2::aes(x = name, y = pred),
            fill = naka_col, colour = "black", shape = 23, size = 4, stroke = 0.9,
            position = ggplot2::position_nudge(x = 0.08), inherit.aes = FALSE) +
        # Estimate text, right-aligned to the right edge so it stays in the box.
        # x positions mirror each diamond's nudge: Yang below (x<1), Naka above (x>1).
        ggplot2::annotate("text", x = 0.66, y = txt_y, hjust = 1, vjust = 0.5,
            colour = yang_col, size = 3.4, fontface = "bold",
            label = sprintf("Bias-Robust (Yang): %.2f [%.2f, %.2f]",
                            yang$pred, yang$ci.lb, yang$ci.ub)) +
        ggplot2::annotate("text", x = 1.34, y = txt_y, hjust = 1, vjust = 0.5,
            colour = naka_col, size = 3.4, fontface = "bold",
            label = sprintf("Bias-Corrected (Nakagawa): %.2f [%.2f, %.2f]",
                            naka$pred, naka$ci.lb, naka$ci.ub)) +
        ggplot2::scale_colour_manual(values = source_palette, name = "Data Source") +
        ggplot2::scale_fill_manual(values = source_palette, name = "Data Source") +
        ggplot2::ggtitle(title) +
        ggplot2::coord_flip(ylim = xlim) +
        ggplot2::theme(
            legend.position      = c(0.98, 0.02),
            legend.justification = c("right", "bottom"),
            legend.background    = ggplot2::element_rect(fill = scales::alpha("white", 0.75), color = "grey90"),
            legend.box           = "vertical",
            legend.margin        = ggplot2::margin(t = 4, r = 4, b = 4, l = 4),
            legend.title         = ggplot2::element_text(size = 9, face = "bold"),
            legend.text          = ggplot2::element_text(size = 8),
            plot.title           = ggplot2::element_text(face = "bold", hjust = 0.5),
            # Panel identifier (a)/(b)/(c) in the top-left corner of the box.
            plot.tag             = ggplot2::element_text(size = 13, face = "bold", hjust = 0, vjust = 1),
            plot.tag.position    = c(0.015, 0.985)
        )

    oc
}

# Shared effect-size range across the three panels of a stack, padded for the
# right-aligned estimate text. Includes raw points + all corrected-estimate CIs.
compute_stack_xlim <- function(db, corr_models, rob_models) {
    ci_vals <- function(m) if (is.null(m)) numeric(0) else c(m$ci.lb[1], m$ci.ub[1])
    vals <- c(db$yi,
              unlist(lapply(corr_models, function(r) if (is.null(r)) numeric(0) else ci_vals(r$model))),
              unlist(lapply(rob_models, ci_vals)))
    vals <- vals[is.finite(vals)]
    rng <- range(vals)
    pad <- diff(rng) * 0.08
    c(rng[1] - pad, rng[2] + pad)
}

# Per-point source vectors for the Combined trunks (new rows then original rows,
# matching bind_rows() order — same assumption as hedges_comparison_plot).
n_h_new  <- nrow(orchaRd::mod_results(res_hedges_new$model,  group = "identifierStudyId")$data)
n_h_orig <- nrow(orchaRd::mod_results(res_hedges_orig$model, group = "identifierStudyId")$data)
src_h_comb <- c(rep("New Extraction (Santos)", n_h_new), rep("Original (Davies 2020)", n_h_orig))

n_o_new  <- nrow(orchaRd::mod_results(res_or_new$model,  group = "identifierStudyId")$data)
n_o_orig <- nrow(orchaRd::mod_results(res_or_orig$model, group = "identifierStudyId")$data)
src_o_comb <- c(rep("New Extraction (Santos)", n_o_new), rep("Original (Jones & DuVal 2019)", n_o_orig))
```

### Hedges' $g$ (Continuous outcomes)

```{r}
#| label: correction_comparison_hedges
#| message: false
#| warning: false
#| fig.width: 9
#| fig.height: 12
# Shared x-axis range so the three Hedges panels align for easy comparison.
hedges_xlim <- compute_stack_xlim(
    db_hedges,
    list(res_hedges_new_corr, res_hedges_orig_corr, res_hedges_comb_corr),
    list(rob_hedges_new, rob_hedges_orig, rob_hedges_comb)
)

ph_new  <- make_pub_bias_subplot(res_hedges_new,  res_hedges_new_corr,  rob_hedges_new,  "New Extraction (Santos)", "Hedges' g", "New Extraction", xlim = hedges_xlim)
ph_orig <- make_pub_bias_subplot(res_hedges_orig, res_hedges_orig_corr, rob_hedges_orig, "Original (Davies 2020)",  "Hedges' g", "Original",       xlim = hedges_xlim)
ph_comb <- make_pub_bias_subplot(res_hedges_comb, res_hedges_comb_corr, rob_hedges_comb, src_h_comb,                "Hedges' g", "Combined",       xlim = hedges_xlim)

fig_hedges <- (ph_orig / ph_new / ph_comb) +
    patchwork::plot_annotation(
        tag_levels = "a", tag_prefix = "(", tag_suffix = ")",
        title = "Hedges' g — Uncorrected mean with Bias-Corrected & Bias-Robust overlays",
        subtitle = "Overlays: purple = Bias-Corrected (Nakagawa 2022), green = Bias-Robust (Yang 2024).",
        theme = ggplot2::theme(
            plot.title    = ggplot2::element_text(size = 15, face = "bold", hjust = 0.5),
            plot.subtitle = ggplot2::element_text(size = 11, face = "italic", hjust = 0.5)
        )
    )

print(fig_hedges)
```

### Odds Ratio (Binary outcomes)

```{r}
#| label: correction_comparison_or
#| message: false
#| warning: false
#| fig.width: 9
#| fig.height: 12
# Shared x-axis range so the three OR panels align for easy comparison.
or_xlim <- compute_stack_xlim(
    db_or,
    list(res_or_new_corr, res_or_orig_corr, res_or_comb_corr),
    list(rob_or_new, rob_or_orig, rob_or_comb)
)

po_new  <- make_pub_bias_subplot(res_or_new,  res_or_new_corr,  rob_or_new,  "New Extraction (Santos)",       "Log Odds Ratio (lnOR)", "New Extraction", xlim = or_xlim)
po_orig <- make_pub_bias_subplot(res_or_orig, res_or_orig_corr, rob_or_orig, "Original (Jones & DuVal 2019)", "Log Odds Ratio (lnOR)", "Original",       xlim = or_xlim)
po_comb <- make_pub_bias_subplot(res_or_comb, res_or_comb_corr, rob_or_comb, src_o_comb,                      "Log Odds Ratio (lnOR)", "Combined",       xlim = or_xlim)

fig_or <- (po_orig / po_new / po_comb) +
    patchwork::plot_annotation(
        tag_levels = "a", tag_prefix = "(", tag_suffix = ")",
        title = "Log Odds Ratio — Uncorrected mean with Bias-Corrected & Bias-Robust overlays",
        subtitle = "Overlays: purple = Bias-Corrected (Nakagawa 2022), green = Bias-Robust (Yang 2024).",
        theme = ggplot2::theme(
            plot.title    = ggplot2::element_text(size = 15, face = "bold", hjust = 0.5),
            plot.subtitle = ggplot2::element_text(size = 11, face = "italic", hjust = 0.5)
        )
    )

print(fig_or)
```

---

## Taxonomic Moderator (Heteroscedastic Model) {#sec-taxonomic-moderator}

The combined synthesis pools effect sizes across very different organisms — insects and other arthropods, ray-finned fishes, and the remaining vertebrates (birds and mammals). Two questions follow naturally: does the *average* strength of mate-choice copying differ among these groups, and is the *amount of heterogeneity* itself group-specific? To answer both we fit a single `rma.mv` model with **taxonomic group as a categorical moderator** and a **heteroscedastic** random-effects structure that lets the variance components differ across groups.

Three design choices define this model:

* **Taxonomic grouping.** Each species (full binomial) is assigned to **Arthropods**, **Fish**, or **Other Vertebrates** by genus. The genus uniquely determines the class for every species in these data, so the mapping is unambiguous (verified in the table below).
* **No phylogenetic term.** The heterogeneity decomposition shows the phylogenetic component contributes very little variance, so it is dropped here; the species random intercept (`taxonomySpecies`) still absorbs among-species differences within each group.
* **Heteroscedasticity (`struct = "HCS"`, `rho = 0`).** The **residual (within-study) effect-size** variance is allowed to take a separate value in each taxonomic group, over a single pooled between-study variance, rather than forcing one common residual variance. This follows the `orchaRd` 2.0 recommendation for meta-regression with heterogeneous variances [@nakagawaOrchaRd2Visualising2023]: the heteroscedastic random term is written as an `~ moderator | unit` formula (`~ tax_group | identifierEffectSizeID`) with a heteroscedastic compound-symmetry structure and the between-level covariance fixed to zero (`rho = 0`), since each effect size belongs to exactly one taxonomic group and a cross-group covariance is not identifiable. (A fuller structure that *also* made the between-study variance group-specific was considered but dropped: its per-group between-study variances collapse to zero for the groups with few studies, so it is over-parameterised for these data.)
* **Marginalised means.** The moderator is fitted **with an intercept** (`mods = ~ tax_group`) and the per-group means, confidence intervals, and prediction intervals are obtained as **marginalised means** through `orchaRd::mod_results()` (which calls `emmeans` internally), again as recommended in the `orchaRd` 2.0 workflow. The overall test for differences among groups is the model's omnibus moderator test ($Q_M$).

```{r}
#| label: taxon_classify
#| message: false
#| warning: false
# ---------------------------------------------------------------------------
# Genus-based taxonomic classification of the combined-dataset species.
# Classification keys on the FULL binomial (`species_phylo`); the genus alone
# unambiguously determines the class for every species in these datasets
# (cross-checked against the higher-taxon cascade in build_phylogeny.R).
# ---------------------------------------------------------------------------
arthropod_genera <- c("Drosophila", "Enchenopa", "Schizocosa", "Uca")
fish_genera <- c(
    "Amblyglyphidodon", "Brachyrhaphis", "Danio", "Etheostoma", "Gambusia",
    "Gasterosteus", "Limia", "Oryzias", "Poecilia", "Pomatoschistus",
    "Porichthys", "Rhabdoblennius", "Syngnathus"
)
othervert_genera <- c("Coturnix", "Dama", "Ficedula", "Molothrus", "Mus", "Rattus", "Taeniopygia")

tax_levels  <- c("Arthropods", "Fish", "Other Vertebrates")
tax_palette <- c("Arthropods" = "#E69F00", "Fish" = "#56B4E9", "Other Vertebrates" = "#009E73")

classify_taxon <- function(binomial) {
    genus <- stringr::word(as.character(binomial), 1)
    grp <- dplyr::case_when(
        genus %in% arthropod_genera ~ "Arthropods",
        genus %in% fish_genera      ~ "Fish",
        genus %in% othervert_genera ~ "Other Vertebrates",
        TRUE                        ~ NA_character_
    )
    factor(grp, levels = tax_levels)
}

# Add the taxonomic group to the combined Hedges' g and log OR datasets.
db_hedges_tax <- db_hedges %>% mutate(tax_group = classify_taxon(species_phylo))
db_or_tax     <- db_or     %>% mutate(tax_group = classify_taxon(species_phylo))

# Defensive check: every species should be classified.
unmapped_tax <- sort(unique(as.character(c(
    db_hedges_tax$species_phylo[is.na(db_hedges_tax$tax_group)],
    db_or_tax$species_phylo[is.na(db_or_tax$tax_group)]
))))
if (length(unmapped_tax) > 0) {
    warning("Unclassified species (no taxonomic group): ", paste(unmapped_tax, collapse = ", "))
}
```

### Species-to-Group Mapping (verification)

The table below lists every species in the combined datasets, its full binomial, the species level used as a random effect (the epithet), and the taxonomic group it was assigned to, with the number of effect sizes it contributes to each effect-size class. This is the audit the analysis is built on — each row should look correct.

```{r}
#| label: taxon_check_table
#| echo: false
tax_map_check <- dplyr::bind_rows(
    db_hedges_tax %>% transmute(
        species_phylo = as.character(species_phylo),
        epithet = as.character(taxonomySpecies), tax_group, Class = "Hedges' g"
    ),
    db_or_tax %>% transmute(
        species_phylo = as.character(species_phylo),
        epithet = as.character(taxonomySpecies), tax_group, Class = "Log OR"
    )
) %>%
    dplyr::count(tax_group, species_phylo, epithet, Class, name = "n") %>%
    tidyr::pivot_wider(names_from = Class, values_from = n, values_fill = 0) %>%
    arrange(tax_group, species_phylo) %>%
    rename(
        `Taxonomic group` = tax_group,
        `Species (full binomial)` = species_phylo,
        `Species level (epithet)` = epithet
    )

DT::datatable(
    tax_map_check,
    caption = "Each combined-dataset species, its assigned taxonomic group, and the number of effect sizes it contributes to the Hedges' g and log OR classes. Use this to confirm the categorical variable matches the original species.",
    options = list(dom = "tip", pageLength = 40, columnDefs = list(list(className = "dt-center", targets = "_all"))),
    rownames = FALSE
)
```

### Model Fitting

```{r}
#| label: fit_taxon_hetero_models
#| message: false
#| warning: false
# ---------------------------------------------------------------------------
# Heteroscedastic taxonomic-moderator model, following the orchaRd 2.0
# heterogeneous-variance workflow (Nakagawa et al. 2023, vignette Example 4).
#
# * The moderator is fitted WITH an intercept (`mods = ~ tax_group`); per-group
#   means / CIs / PIs are obtained downstream as marginalised means via
#   orchaRd::mod_results() (emmeans), and the among-group test is the model's
#   omnibus moderator test (QM).
# * The heterogeneous (residual) variance is specified as a `~ tax_group | unit`
#   random term with struct = "HCS" and the between-level covariance fixed to
#   zero (rho = 0) — exactly the structure used in the orchaRd documentation.
#   With each unit belonging to a single group this estimates one residual
#   variance per group with no cross-group covariance.
# * PRIMARY structure: group-specific RESIDUAL variance
#   (`~ tax_group | identifierEffectSizeID`) over a single pooled between-study
#   variance (`~ 1 | identifierStudyId`) and a homoscedastic species intercept
#   (`~ 1 | taxonomySpecies`). NO phylogenetic term. (The fuller both-level
#   structure was dropped: its per-group between-study variances collapse to
#   zero for the thin groups, i.e. it is over-parameterised for these data.)
# * Each heteroscedastic structure is tried with several optimisers (nlminb,
#   optim/BFGS, optim/L-BFGS-B, then a high-iteration Nelder-Mead) before being
#   declared a failure; only if none converge does the cascade fall back to a
#   simpler structure. The vignette's default Nelder-Mead alone hits its
#   iteration cap on these data (convergence = 1), so robust optimisers go first.
# ---------------------------------------------------------------------------
fit_taxon_hetero_mv <- function(dat, group_var = "tax_group", yi = "yi", vi = "vi",
                                cluster_var = "identifierStudyId",
                                obs_var = "identifierEffectSizeID", rho = 0.5) {
    dat2 <- dat %>%
        filter(!is.na(.data[[yi]]), !is.na(.data[[vi]]), .data[[vi]] > 0, !is.na(.data[[group_var]]))
    if (nrow(dat2) == 0) {
        message("No complete cases found.")
        return(NULL)
    }
    dat2[[group_var]] <- droplevels(factor(dat2[[group_var]], levels = tax_levels))

    VCV <- metafor::vcalc(vi = dat2[[vi]], cluster = dat2[[cluster_var]], obs = dat2[[obs_var]], rho = rho, data = dat2)

    mods_f  <- stats::as.formula(paste0("~ ", group_var))                       # intercept-coded
    obs_f   <- stats::as.formula(paste0("~ ", group_var, " | ", obs_var))       # heteroscedastic residual

    # Optimisers tried (in order) for the heteroscedastic structures. The default
    # Nelder-Mead used in the orchaRd vignette hits its iteration cap on these
    # data (convergence = 1), so we lead with the more robust derivative-based
    # optimisers and only then fall back to a high-iteration Nelder-Mead.
    opt_grid <- list(
        "nlminb"            = list(optimizer = "nlminb", iter.max = 10000, eval.max = 10000),
        "optim/BFGS"        = list(optimizer = "optim", optmethod = "BFGS", maxit = 10000),
        "optim/L-BFGS-B"    = list(optimizer = "optim", optmethod = "L-BFGS-B", maxit = 10000),
        "optim/Nelder-Mead" = list(optimizer = "optim", optmethod = "Nelder-Mead", maxit = 50000, reltol = 1e-9)
    )

    # Build one rma.mv call for a candidate structure + optimiser control.
    build <- function(cand, ctrl) {
        args <- c(
            list(yi = dat2[[yi]], V = VCV, mods = mods_f, random = cand$random,
                 test = "t", dfs = "contain", method = "REML", sparse = TRUE, data = dat2),
            cand$args
        )
        if (!is.null(cand$struct)) args$struct <- cand$struct
        if (!is.null(ctrl))        args$control <- ctrl
        do.call(metafor::rma.mv, args)
    }

    # Candidate structures, tried in order. The cascade stops at the FIRST one
    # that fits without error; warnings do NOT stop the cascade but are recorded.
    # PRIMARY = residual-only HCS (orchaRd Example 4 default): a single pooled
    # between-study variance plus a group-specific residual variance. The fuller
    # both-level structure was dropped because its per-group between-study
    # variances collapse to the zero boundary for the thin groups (Other
    # Vertebrates, Arthropods) — i.e. it is over-parameterised for these data.
    candidates <- list(
        list(spec = "Heteroscedastic (residual only, per group; HCS, rho = 0)",
             random = list(~ 1 | identifierStudyId, obs_f, ~ 1 | taxonomySpecies),
             struct = "HCS", args = list(rho = 0), het = TRUE),
        list(spec = "Homoscedastic (single variance per component)",
             random = list(~ 1 | identifierStudyId, ~ 1 | identifierEffectSizeID, ~ 1 | taxonomySpecies),
             struct = NULL, args = list(), het = FALSE)
    )

    # Attempt one candidate. Heteroscedastic structures cycle through opt_grid
    # until one converges; the homoscedastic structure uses the metafor default.
    # Errors AND warnings are captured (warnings do not abort).
    fit_once <- function(cand, ctrl) {
        warns <- character(0)
        t0 <- Sys.time()
        mod <- tryCatch(
            withCallingHandlers(
                build(cand, ctrl),
                warning = function(w) {
                    warns[[length(warns) + 1L]] <<- conditionMessage(w)
                    invokeRestart("muffleWarning")
                }
            ),
            error = function(e) structure(list(msg = conditionMessage(e)), class = "tax_fit_error")
        )
        list(mod = mod, warns = warns, secs = as.numeric(difftime(Sys.time(), t0, units = "secs")))
    }

    attempt <- function(cand) {
        if (!isTRUE(cand$het)) {
            r <- fit_once(cand, NULL)
            failed <- inherits(r$mod, "tax_fit_error")
            return(list(model = if (failed) NULL else r$mod, failed = failed,
                        error = if (failed) r$mod$msg else NA_character_,
                        warnings = r$warns, secs = r$secs, optimizer = "default (nlminb)"))
        }
        msgs <- character(0)
        total <- 0
        for (onm in names(opt_grid)) {
            r <- fit_once(cand, opt_grid[[onm]])
            total <- total + r$secs
            if (!inherits(r$mod, "tax_fit_error")) {
                return(list(model = r$mod, failed = FALSE, error = NA_character_,
                            warnings = r$warns, secs = total, optimizer = onm))
            }
            msgs <- c(msgs, paste0(onm, ": ", r$mod$msg))
        }
        list(model = NULL, failed = TRUE, error = paste(msgs, collapse = " || "),
             warnings = character(0), secs = total, optimizer = NA_character_)
    }

    diag_rows <- list()
    chosen <- NULL
    chosen_spec <- NA_character_
    for (i in seq_along(candidates)) {
        cand <- candidates[[i]]
        if (!is.null(chosen)) {
            diag_rows[[i]] <- tibble::tibble(
                stage = i, structure = cand$spec, status = "not attempted", optimizer = NA_character_,
                seconds = NA_real_, n_warnings = NA_integer_, detail = "an earlier structure succeeded"
            )
            next
        }
        a <- attempt(cand)
        status <- if (a$failed) "ERROR" else if (length(a$warnings) > 0) "OK (with warnings)" else "OK"
        detail <- if (a$failed) a$error else if (length(a$warnings) > 0) paste(a$warnings, collapse = " || ") else "clean fit"
        diag_rows[[i]] <- tibble::tibble(
            stage = i, structure = cand$spec, status = status, optimizer = a$optimizer,
            seconds = round(a$secs, 2), n_warnings = length(a$warnings), detail = detail
        )
        if (!a$failed) {
            chosen <- a$model
            chosen_spec <- cand$spec
        }
    }
    diagnostics <- dplyr::bind_rows(diag_rows)

    if (is.null(chosen)) {
        warning("All taxonomic-moderator structures failed to fit.")
        return(list(data = dat2, VCV = VCV, model = NULL, spec = "ALL FAILED", diagnostics = diagnostics))
    }
    list(data = dat2, VCV = VCV, model = chosen, spec = chosen_spec, diagnostics = diagnostics)
}

res_tax_hedges <- fit_taxon_hetero_mv(db_hedges_tax)
res_tax_or     <- fit_taxon_hetero_mv(db_or_tax)

# Omnibus among-group test. With the intercept-coded moderator, metafor's built-in
# QM tests all non-intercept coefficients jointly — i.e. exactly whether the group
# means differ — so it is used directly (no custom contrast matrix needed).
among_group_test <- function(res) {
    if (is.null(res) || is.null(res$model)) {
        return(NULL)
    }
    list(QM = res$model$QM, QMp = res$model$QMp, QMdf = res$model$QMdf)
}
qm_tax_hedges <- among_group_test(res_tax_hedges)
qm_tax_or     <- among_group_test(res_tax_or)

# Total I^2 (orchaRd if available; otherwise a Higgins typical-variance fallback
# so a heterogeneity figure is always reported even under the heteroscedastic
# structure).
total_i2 <- function(res) {
    if (is.null(res) || is.null(res$model)) {
        return(NA_real_)
    }
    m <- res$model
    i2 <- tryCatch(orchaRd::i2_ml(m), error = function(e) NULL)
    if (!is.null(i2) && any(grepl("Total", names(i2)))) {
        return(as.numeric(i2[grep("Total", names(i2))[1]]))
    }
    vi <- res$data$vi
    wi <- 1 / vi
    k  <- length(vi)
    s2 <- (k - 1) * sum(wi) / (sum(wi)^2 - sum(wi^2))
    vc <- sum(c(m$sigma2, m$tau2, m$gamma2), na.rm = TRUE)
    100 * vc / (vc + s2)
}
i2_tax_hedges <- total_i2(res_tax_hedges)
i2_tax_or     <- total_i2(res_tax_or)

# Persist the fitted results + cascade diagnostics so they can be inspected
# interactively in an R session WITHOUT re-rendering the book, e.g.:
#   b <- readRDS("outputs/1_effect_size_calculation_pipeline/taxon_model_fits.rds")
#   b$hedges$spec ; b$hedges$diagnostics ; summary(b$hedges$model)
# (also consumed by scripts/1_effect_size_calculation_pipeline/diagnose_taxon_model.R)
tryCatch({
    tax_out_dir <- here::here("outputs", "1_effect_size_calculation_pipeline")
    dir.create(tax_out_dir, recursive = TRUE, showWarnings = FALSE)
    saveRDS(
        list(
            hedges = res_tax_hedges, or = res_tax_or,
            qm = list(hedges = qm_tax_hedges, or = qm_tax_or),
            i2 = list(hedges = i2_tax_hedges, or = i2_tax_or),
            saved_at = Sys.time()
        ),
        file.path(tax_out_dir, "taxon_model_fits.rds")
    )
}, error = function(e) warning("Could not save taxon_model_fits.rds: ", conditionMessage(e)))
```

### Fit Diagnostics {#sec-taxon-fit-diagnostics}

The model is fitted through a **fallback cascade**: the most flexible heteroscedastic structure is attempted first, and only if it *errors* does the next-simpler structure get tried (warnings do **not** trigger a fallback, but are captured and shown). The report below tells you exactly what happened for each effect-size class — which structures were attempted, whether each fit was clean / produced warnings / errored (with the captured message), how long it took, which structure was ultimately used, the per-group sample sizes, and the variance components of the chosen model (so you can see whether any group's variance collapsed to the zero boundary).

```{r}
#| label: taxon_fit_diagnostics
#| echo: false
#| message: false
#| warning: false
# Per-group sample sizes (context for unstable / zero variances).
group_counts <- function(res) {
    if (is.null(res) || is.null(res$data)) {
        return(NULL)
    }
    res$data %>%
        dplyr::group_by(tax_group) %>%
        dplyr::summarise(
            `k (effect sizes)` = dplyr::n(),
            studies = dplyr::n_distinct(identifierStudyId),
            species = dplyr::n_distinct(taxonomySpecies),
            .groups = "drop"
        ) %>%
        dplyr::rename(`Taxonomic group` = tax_group)
}

# Variance components of the chosen model, labelled by random-effect term.
# Labelling is STRUCTURE-AWARE (keyed on res$spec) because which variance lands
# in metafor's sigma2 / tau2 / gamma2 slots depends on which cascade structure
# was used:
#   * both-level het  -> tau2 = between-study (per group), gamma2 = residual
#                        (per group), sigma2 = species
#   * residual-only   -> tau2 = residual (per group), sigma2 = c(study, species)
#   * homoscedastic   -> sigma2 = c(study, effect-size residual, species)
chosen_varcomp <- function(res) {
    if (is.null(res) || is.null(res$model)) {
        return(NULL)
    }
    m   <- res$model
    glv <- levels(droplevels(factor(res$data$tax_group, levels = tax_levels)))
    out <- tibble::tibble(Component = character(), Variance = numeric())
    addc <- function(out, labs, vals) {
        if (is.null(vals) || length(vals) == 0) {
            return(out)
        }
        dplyr::bind_rows(out, tibble::tibble(Component = labs, Variance = round(as.numeric(vals), 5)))
    }
    spec <- if (is.null(res$spec)) "" else res$spec
    if (grepl("between-study \\+ residual", spec)) {
        out <- addc(out, paste0("Between-study τ² — ", glv[seq_along(m$tau2)]), m$tau2)
        out <- addc(out, paste0("Residual σ² — ", glv[seq_along(m$gamma2)]), m$gamma2)
        out <- addc(out, "Species σ²", m$sigma2)
    } else if (grepl("residual only", spec)) {
        out <- addc(out, paste0("Residual σ² — ", glv[seq_along(m$tau2)]), m$tau2)
        out <- addc(out, c("Between-study τ²", "Species σ²")[seq_along(m$sigma2)], m$sigma2)
    } else {
        out <- addc(out, c("Between-study τ²", "Residual σ² (effect size)", "Species σ²")[seq_along(m$sigma2)], m$sigma2)
    }
    out
}

# Console-style textual report (prints into the rendered document).
report_fit <- function(res, label) {
    cat("\n========================================================\n")
    cat(" FIT DIAGNOSTIC:", label, "\n")
    cat("========================================================\n")
    if (is.null(res)) {
        cat("  No result object (empty data?).\n")
        return(invisible())
    }
    cat("  Chosen structure :", res$spec, "\n")
    cat("  Model is NULL    :", is.null(res$model), "\n\n")
    cat("  Cascade log:\n")
    print(as.data.frame(res$diagnostics), row.names = FALSE, right = FALSE)
    gc <- group_counts(res)
    if (!is.null(gc)) {
        cat("\n  Per-group sample sizes:\n")
        print(as.data.frame(gc), row.names = FALSE, right = FALSE)
    }
    vc <- chosen_varcomp(res)
    if (!is.null(vc)) {
        cat("\n  Variance components (chosen model):\n")
        print(as.data.frame(vc), row.names = FALSE, right = FALSE)
        n_zero <- sum(vc$Variance < 1e-6, na.rm = TRUE)
        if (n_zero > 0) {
            cat("  NOTE:", n_zero, "variance component(s) at/near zero (boundary estimate).\n")
        }
    }
    cat("\n")
    invisible()
}

report_fit(res_tax_hedges, "Hedges' g")
report_fit(res_tax_or, "Log Odds Ratio")
```

```{r}
#| label: taxon_fit_diagnostics_table
#| echo: false
diag_combined <- dplyr::bind_rows(
    if (!is.null(res_tax_hedges)) dplyr::mutate(res_tax_hedges$diagnostics, Analysis = "Hedges' g"),
    if (!is.null(res_tax_or)) dplyr::mutate(res_tax_or$diagnostics, Analysis = "Log OR")
)
if (!is.null(diag_combined) && nrow(diag_combined) > 0) {
    diag_combined <- diag_combined %>%
        dplyr::select(Analysis, stage, structure, status, optimizer, seconds, n_warnings, detail)
}

DT::datatable(
    diag_combined,
    caption = "Fallback-cascade log: every candidate random-effects structure, whether it was attempted, its fit status (OK / OK with warnings / ERROR / not attempted), runtime in seconds, warning count, and the captured error or warning text. The first 'OK' row per analysis is the structure that was actually used.",
    options = list(dom = "t", pageLength = 12, columnDefs = list(list(className = "dt-left", targets = "_all"))),
    rownames = FALSE
)
```

### Model Output: Hedges' $g$

```{r}
#| label: taxon_hedges_summary
#| echo: false
if (!is.null(res_tax_hedges) && !is.null(res_tax_hedges$model)) {
    cat("Random-effects structure:", res_tax_hedges$spec, "\n\n")
    print(summary(res_tax_hedges$model))
} else {
    cat("No Hedges' g model was successfully fitted — see the Fit Diagnostics above.\n")
}
```

### Model Output: Log Odds Ratio

```{r}
#| label: taxon_or_summary
#| echo: false
if (!is.null(res_tax_or) && !is.null(res_tax_or$model)) {
    cat("Random-effects structure:", res_tax_or$spec, "\n\n")
    print(summary(res_tax_or$model))
} else {
    cat("No log OR model was successfully fitted — see the Fit Diagnostics above.\n")
}
```

### Group Means and Heterogeneity

```{r}
#| label: taxon_group_means_table
#| echo: false
group_means <- function(res, label) {
    if (is.null(res) || is.null(res$model)) {
        return(NULL)
    }
    mr <- orchaRd::mod_results(res$model, mod = "tax_group", group = "identifierStudyId")
    mr$mod_table %>%
        transmute(
            Analysis = label,
            `Taxonomic group` = name,
            Estimate = round(estimate, 3),
            `CI lower` = round(lowerCL, 3), `CI upper` = round(upperCL, 3),
            `PI lower` = round(lowerPR, 3), `PI upper` = round(upperPR, 3)
        )
}

means_tbl <- dplyr::bind_rows(
    group_means(res_tax_hedges, "Hedges' g"),
    group_means(res_tax_or, "Log OR")
)

DT::datatable(
    means_tbl,
    caption = "Estimated mean effect (with 95% confidence and prediction intervals) for each taxonomic group, from the heteroscedastic moderator models.",
    options = list(dom = "t", pageLength = 12, columnDefs = list(list(className = "dt-center", targets = "_all"))),
    rownames = FALSE
)
```

```{r}
#| label: taxon_variance_table
#| echo: false
# Pull the (group-specific) variance components out of a fitted model. Under the
# both-DIAG structure: tau2 = between-study variance per group, gamma2 = residual
# variance per group, sigma2 = species variance. Labels fall back gracefully if a
# simpler fallback structure was used.
# Reuse the structure-aware chosen_varcomp() (defined in the Fit Diagnostics
# chunk) so the labels stay correct whichever cascade structure was used.
variance_components <- function(res, label) {
    vc <- chosen_varcomp(res)
    if (is.null(vc) || nrow(vc) == 0) {
        return(NULL)
    }
    dplyr::mutate(vc, Analysis = label) %>% dplyr::select(Analysis, Component, Variance)
}

var_tbl <- dplyr::bind_rows(
    variance_components(res_tax_hedges, "Hedges' g"),
    variance_components(res_tax_or, "Log OR")
)

DT::datatable(
    var_tbl,
    caption = "Variance components from the heteroscedastic models. Under the both-DIAG structure the between-study and residual variances are estimated separately within each taxonomic group; the species component is shared.",
    options = list(dom = "t", pageLength = 20, columnDefs = list(list(className = "dt-center", targets = "_all"))),
    rownames = FALSE
)
```

```{r}
#| label: taxon_qm_helpers
#| include: false
# Inline helpers for the narrative below.
qm_p   <- function(a) if (is.null(a) || is.null(a$QMp)) NA_real_ else as.numeric(a$QMp)
qm_stat <- function(a) if (is.null(a) || is.null(a$QM)) NA_real_ else as.numeric(a$QM)
```

For Hedges' $g$, total heterogeneity is `r fmt(i2_tax_hedges, 1)`% and the among-group test is `r pstr(qm_p(qm_tax_hedges))` (Wald $Q_M$ = `r fmt(qm_stat(qm_tax_hedges), 2)`), so the difference among taxonomic groups is **`r sigw(qm_p(qm_tax_hedges))`**. For the log odds ratio, total heterogeneity is `r fmt(i2_tax_or, 1)`% and the among-group test is `r pstr(qm_p(qm_tax_or))` (Wald $Q_M$ = `r fmt(qm_stat(qm_tax_or), 2)`), a **`r sigw(qm_p(qm_tax_or))`** difference among groups. The group means and their intervals are tabulated above; the variance-component table shows the group-specific residual (within-study) variances, which is where the heteroscedastic structure lets variability differ across the three groups.

### Orchard Plot

```{r}
#| label: taxon_orchard_plots
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 9
plot_taxon_orchard <- function(res, xlab, title) {
    if (is.null(res) || is.null(res$model)) {
        return(NULL)
    }
    mr <- orchaRd::mod_results(res$model, mod = "tax_group", group = "identifierStudyId")
    oc <- orchaRd::orchard_plot(mr, xlab = xlab, angle = 0, g = TRUE, k = TRUE)
    # Replace each group's k-count label with "estimate [95% CI]\nk = K (studies)"
    # using the same helper as the overall plots. tryCatch guards against any
    # orchaRd layer-order change so the figure still renders if it fails.
    oc <- tryCatch(add_ci_labels(oc, mr), error = function(e) {
        warning("add_ci_labels (taxon) failed: ", conditionMessage(e)); oc
    })
    oc +
        ggplot2::scale_colour_manual(values = tax_palette, name = "Taxonomic group", aesthetics = c("colour", "fill")) +
        ggplot2::ggtitle(title) +
        ggplot2::theme(
            plot.title = ggplot2::element_text(face = "bold", hjust = 0.5),
            legend.position = "none"
        )
}

# Save a figure as a proper deliverable (PNG + PDF) in outputs/.
save_taxon_fig <- function(plot_obj, basename, width = 10, height = 9) {
    if (is.null(plot_obj)) {
        return(invisible())
    }
    tryCatch({
        d <- here::here("outputs", "1_effect_size_calculation_pipeline")
        dir.create(d, recursive = TRUE, showWarnings = FALSE)
        ggplot2::ggsave(file.path(d, paste0(basename, ".png")), plot_obj, width = width, height = height, dpi = 300, bg = "white")
        ggplot2::ggsave(file.path(d, paste0(basename, ".pdf")), plot_obj, width = width, height = height)
    }, error = function(e) warning("Could not save ", basename, ": ", conditionMessage(e)))
}

p_tax_h <- plot_taxon_orchard(res_tax_hedges, "Hedges' g", "Hedges' g by taxonomic group")
p_tax_o <- plot_taxon_orchard(res_tax_or, "Log Odds Ratio (lnOR)", "Log OR by taxonomic group")

# Save the individual panels too, in case they are wanted separately.
save_taxon_fig(p_tax_h, "taxon_moderator_orchard_hedges", width = 10, height = 5)
save_taxon_fig(p_tax_o, "taxon_moderator_orchard_or", width = 10, height = 5)

if (!is.null(p_tax_h) && !is.null(p_tax_o)) {
    fig_tax <- (p_tax_h / p_tax_o) +
        patchwork::plot_annotation(
            tag_levels = "a", tag_prefix = "(", tag_suffix = ")",
            title = "Mate-choice copying by taxonomic group (heteroscedastic moderator model)",
            theme = ggplot2::theme(plot.title = ggplot2::element_text(size = 15, face = "bold", hjust = 0.5))
        )
    print(fig_tax)
    # Combined stacked figure = the main deliverable.
    save_taxon_fig(fig_tax, "taxon_moderator_orchard", width = 10, height = 9)
} else {
    if (!is.null(p_tax_h)) print(p_tax_h)
    if (!is.null(p_tax_o)) print(p_tax_o)
}
```

---

## Heterogeneity Decomposition (Partial $I^2$) {#sec-heterogeneity}

Total heterogeneity ($I^2_{\text{Total}}$) quantifies the proportion of variance in the observed effect sizes that reflects true differences among effects rather than sampling error. In a multilevel model this total can be *partitioned* into the contribution of each random effect [@nakagawaMethodsTestingPublication2022]. Here the variance is split among four components: between-**study** differences, the residual **effect-size** (within-study) level, a **species** (non-phylogenetic) component, and a **phylogenetic** component that captures variance shared among species in proportion to their relatedness. We compute these partial $I^2$ values with `orchaRd::i2_ml()` for each uncorrected model.

```{r}
#| label: heterogeneity_i2_compute
#| message: false
#| warning: false
# Human-readable labels for the i2_ml components (names are prefixed with "I2_")
comp_labels <- c(
    Total                  = "Total",
    identifierStudyId      = "Study",
    identifierEffectSizeID = "Effect size (residual)",
    taxonomySpecies        = "Species (non-phylogenetic)",
    species_phylo          = "Phylogeny"
)

# Pull the partial I^2 vector from a fitted model into tidy long form
get_i2 <- function(res, effect, model) {
    if (is.null(res) || is.null(res$model)) {
        return(NULL)
    }
    v <- tryCatch(orchaRd::i2_ml(res$model), error = function(e) NULL)
    if (is.null(v)) {
        return(NULL)
    }
    tibble::tibble(
        Effect    = effect,
        Model     = model,
        Component = sub("^I2_", "", names(v)),
        I2        = as.numeric(v)
    )
}

i2_long <- dplyr::bind_rows(
    get_i2(res_hedges_new, "Hedges' g", "New Only"),
    get_i2(res_hedges_orig, "Hedges' g", "Original Only"),
    get_i2(res_hedges_comb, "Hedges' g", "Combined"),
    get_i2(res_or_new, "Log OR", "New Only"),
    get_i2(res_or_orig, "Log OR", "Original Only"),
    get_i2(res_or_comb, "Log OR", "Combined")
) %>%
    mutate(ComponentLabel = dplyr::recode(Component, !!!comp_labels))

# Lookup helper for inline narrative (returns NA if a component is absent)
i2get <- function(effect, model, component) {
    v <- i2_long$I2[i2_long$Effect == effect & i2_long$Model == model & i2_long$ComponentLabel == component]
    if (length(v) == 0 || all(is.na(v))) NA_real_ else v[1]
}
```

### Partial $I^2$ by Random Effect

```{r}
#| label: heterogeneity_i2_table
#| echo: false
i2_wide <- i2_long %>%
    mutate(Analysis = paste0(Effect, " — ", Model), I2 = round(I2, 1)) %>%
    select(Analysis, ComponentLabel, I2) %>%
    tidyr::pivot_wider(names_from = ComponentLabel, values_from = I2)

# Order columns consistently (only those that exist)
desired_cols <- c("Analysis", "Total", "Study", "Effect size (residual)", "Species (non-phylogenetic)", "Phylogeny")
i2_wide <- i2_wide %>% select(dplyr::any_of(desired_cols))

DT::datatable(
    i2_wide,
    caption = "Partial I² (%) for each random effect across the uncorrected multilevel models. Total is the overall proportion of variance due to heterogeneity (vs. sampling error); the remaining columns partition that heterogeneity among the study, residual effect-size, non-phylogenetic species, and phylogenetic components.",
    options = list(
        dom = "t",
        pageLength = 12,
        columnDefs = list(list(className = "dt-center", targets = "_all"))
    ),
    rownames = FALSE
)
```

### Interpreting the Heterogeneity {#sec-heterogeneity-narrative}

The combined models show that mate-choice copying effect sizes are **highly heterogeneous**: total $I^2$ reaches `r fmt(i2get("Hedges' g", "Combined", "Total"), 1)`% for Hedges' $g$ and `r fmt(i2get("Log OR", "Combined", "Total"), 1)`% for the log odds ratio. In other words, the large majority of the observed variation reflects genuine differences among effects rather than sampling error alone, so the pooled means should be read as averages over a broad distribution of true effects rather than as a single common effect.

Decomposing that heterogeneity is where the phylogeny becomes informative. For the combined Hedges' $g$ model, the **phylogenetic** component accounts for `r fmt(i2get("Hedges' g", "Combined", "Phylogeny"), 1)`% of the variance, the **non-phylogenetic species** component for `r fmt(i2get("Hedges' g", "Combined", "Species (non-phylogenetic)"), 1)`%, between-**study** differences for `r fmt(i2get("Hedges' g", "Combined", "Study"), 1)`%, and the residual **effect-size** level for `r fmt(i2get("Hedges' g", "Combined", "Effect size (residual)"), 1)`%. The combined log-odds-ratio model partitions similarly: phylogeny `r fmt(i2get("Log OR", "Combined", "Phylogeny"), 1)`%, species `r fmt(i2get("Log OR", "Combined", "Species (non-phylogenetic)"), 1)`%, study `r fmt(i2get("Log OR", "Combined", "Study"), 1)`%, and residual `r fmt(i2get("Log OR", "Combined", "Effect size (residual)"), 1)`%.

The size of the phylogenetic component is the key new result. A substantial value would indicate that closely related species tend to show similar degrees of mate-choice copying — i.e., the trait carries a phylogenetic signal and effects are not independent across the tree — which would caution against treating species as exchangeable and would motivate phylogenetically informed interpretation. A small phylogenetic component, by contrast, indicates that relatedness explains little of the among-species variation: most of the species-level differences are idiosyncratic (captured by the non-phylogenetic species term) rather than inherited, and the earlier non-phylogenetic models were not materially biased by ignoring the tree. The contrast between the phylogenetic and non-phylogenetic species components is therefore the most direct readout of how much shared ancestry, as opposed to species-specific biology, structures the data.

Two further patterns are worth noting. First, the residual effect-size level typically absorbs a large share of the variance, reflecting heterogeneity among individual trials within studies (different paradigms, demonstrators, and conditions). Second, comparing the New-Only, Original-Only, and Combined columns shows whether the new extraction introduces or dilutes heterogeneity relative to the legacy datasets; where the new and original subsets allocate variance differently, the combined estimate reflects a weighted blend of the two. Because this high heterogeneity is the rule rather than the exception, it justifies the moderator and sub-group analyses pursued elsewhere in the workflow, and it is the reason the pooled estimates are reported with prediction intervals in addition to confidence intervals.

---

## Sensitivity Analysis: Impact of the VCV Structure

To evaluate the robustness of our meta-analytic results, we conduct a sensitivity analysis comparing the main multilevel models (which incorporate a Variance-Covariance [VCV] matrix with an assumed within-study correlation $\rho = 0.5$) against simplified multilevel models that ignore this correlation (i.e., treating effect size variances as independent diagonal elements, $V = v_i$).

Ignoring within-study correlation can lead to improper weighting of studies with multiple effect sizes and can bias the standard errors and confidence intervals of the pooled estimates.

```{r}
#| label: vcv_sensitivity_analysis
#| message: false
#| warning: false
# 1. Define comparison function
compare_vcv_models <- function(dat, name, yi = "yi", vi = "vi", cluster_var = "identifierStudyId", obs_var = "identifierEffectSizeID", rho = 0.5) {
    dat2 <- dat %>%
        filter(!is.na(.data[[yi]]), !is.na(.data[[vi]]), .data[[vi]] > 0)

    if (nrow(dat2) == 0) {
        return(NULL)
    }

    # A. With VCV (rho = 0.5)
    VCV <- metafor::vcalc(
        vi = dat2[[vi]],
        cluster = dat2[[cluster_var]],
        obs = dat2[[obs_var]],
        rho = rho,
        data = dat2
    )
    mod_w_vcv <- metafor::rma.mv(
        yi = dat2[[yi]],
        V = VCV,
        random = list(
            ~ 1 | identifierStudyId,
            ~ 1 | identifierEffectSizeID,
            ~ 1 | taxonomySpecies
        ),
        test = "t",
        method = "REML",
        sparse = TRUE,
        data = dat2
    )

    # B. Without VCV (diagonal V = vi)
    mod_no_vcv <- metafor::rma.mv(
        yi = dat2[[yi]],
        V = dat2[[vi]],
        random = list(
            ~ 1 | identifierStudyId,
            ~ 1 | identifierEffectSizeID,
            ~ 1 | taxonomySpecies
        ),
        test = "t",
        method = "REML",
        sparse = TRUE,
        data = dat2
    )

    # Combine results
    data.frame(
        Analysis = name,
        Model_Type = c("With VCV (rho = 0.5)", "Without VCV (Independent)"),
        k = c(mod_w_vcv$k, mod_no_vcv$k),
        Estimate = c(mod_w_vcv$b[1], mod_no_vcv$b[1]),
        SE = c(mod_w_vcv$se[1], mod_no_vcv$se[1]),
        CI_Lower = c(mod_w_vcv$ci.lb[1], mod_no_vcv$ci.lb[1]),
        CI_Upper = c(mod_w_vcv$ci.ub[1], mod_no_vcv$ci.ub[1]),
        p_val = c(mod_w_vcv$pval[1], mod_no_vcv$pval[1])
    )
}

# 2. Run comparisons across all subsets
comp_hedges_new <- compare_vcv_models(db_hedges %>% filter(Dataset_Source == "New Extraction (Santos)"), "Hedges' g (New Only)")
comp_hedges_orig <- compare_vcv_models(db_hedges %>% filter(Dataset_Source == "Original (Davies 2020)"), "Hedges' g (Original Only)")
comp_hedges_comb <- compare_vcv_models(db_hedges, "Hedges' g (Combined)")

comp_or_new <- compare_vcv_models(db_or %>% filter(Dataset_Source == "New Extraction (Santos)"), "Log OR (New Only)")
comp_or_orig <- compare_vcv_models(db_or %>% filter(Dataset_Source == "Original (Jones & DuVal 2019)"), "Log OR (Original Only)")
comp_or_comb <- compare_vcv_models(db_or, "Log OR (Combined)")

# 3. Bind all comparisons together
vcv_comparison_results <- bind_rows(
    comp_hedges_new, comp_hedges_orig, comp_hedges_comb,
    comp_or_new, comp_or_orig, comp_or_comb
)
```

### Sensitivity Comparison Table

The interactive table below compares the pool estimate, standard error (SE), 95% Confidence Intervals, and p-values for both models across all subsets.

```{r}
#| label: vcv_comparison_table
#| echo: false
DT::datatable(
    vcv_comparison_results,
    options = list(
        pageLength = 12,
        dom = "t",
        columnDefs = list(list(className = "dt-center", targets = "_all"))
    ),
    rownames = FALSE
) %>%
    DT::formatRound(columns = c("Estimate", "SE", "CI_Lower", "CI_Upper"), digits = 4) %>%
    DT::formatSignif(columns = c("p_val"), digits = 4)
```

### Key Methodological Findings

1. **Marginal Shifts in Pooled Estimates:** 
   Treating effect sizes as completely independent (Without VCV) causes very minor changes to the overall pooled estimates. For the combined dataset, the Hedges' $g$ estimate shifts from $`r fmt(comp_hedges_comb$Estimate[1])`$ to $`r fmt(comp_hedges_comb$Estimate[2])`$, while the Log Odds Ratio (lnOR) estimate shifts from $`r fmt(comp_or_comb$Estimate[1])`$ to $`r fmt(comp_or_comb$Estimate[2])`$. This indicates that the main directional conclusions of the meta-analyses are robust to VCV specification.

2. **Standard Error Shifts:** 
   In this multilevel model specification (which partitions variance using nested random intercepts), omitting the VCV matrix and assuming independence leads to a *slight inflation* of the standard error and wider confidence intervals for the combined analyses:
   * **Hedges' $g$ (Combined):** SE increases from $`r fmt(comp_hedges_comb$SE[1])`$ to $`r fmt(comp_hedges_comb$SE[2])`$ (CI width increases from $`r fmt(comp_hedges_comb$CI_Upper[1] - comp_hedges_comb$CI_Lower[1])`$ to $`r fmt(comp_hedges_comb$CI_Upper[2] - comp_hedges_comb$CI_Lower[2])`$).
   * **Log Odds Ratio (Combined):** SE increases from $`r fmt(comp_or_comb$SE[1])`$ to $`r fmt(comp_or_comb$SE[2])`$ (CI width increases from $`r fmt(comp_or_comb$CI_Upper[1] - comp_or_comb$CI_Lower[1])`$ to $`r fmt(comp_or_comb$CI_Upper[2] - comp_or_comb$CI_Lower[2])`$).
   
   This demonstrates that accounting for known within-study covariance via `vcalc` yields more precise, correctly-weighted estimates.

3. **Impact on Statistical Significance:**
   For continuous outcomes (Hedges' $g$ Combined), the pooled estimate is `r sigw(comp_hedges_comb$p_val[1])` with the VCV matrix ($`r pstr(comp_hedges_comb$p_val[1])`$) and `r sigw(comp_hedges_comb$p_val[2])` without it ($`r pstr(comp_hedges_comb$p_val[2])`$). Binary outcomes (Log OR Combined) are `r sigw(comp_or_comb$p_val[1])` under both specifications ($`r pstr(comp_or_comb$p_val[1])`$ with VCV; $`r pstr(comp_or_comb$p_val[2])`$ without).

Therefore, we conclude that while the overall direction and magnitude of the estimates are stable, **incorporating the VCV matrix** ensures appropriate weighting, controlling for within-study correlation, and obtaining accurate standard errors and statistical tests.

---

## Session Info

```{r}
#| label: session_info
sessionInfo()
```
