---
title: "Overall Effect Size Calculation"
subtitle: "Multilevel Meta-Analysis of Mate-Choice Copying"
author: "Eduardo Santos & Collaborators"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    theme: cosmo
    self-contained: true
---

```{r}
#| label: setup
#| include: false
# Ensure pacman is loaded to manage dependencies
if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
pacman::p_load(
               readxl,
               tidyverse,
               metafor,
               orchaRd,
               patchwork,
               DT
               )
```

## Data Loading and Preprocessing

We read the extracted dataset from our Excel workbook. The script filters out any rows marked for removal and processes our calculated effect size columns.

```{r}
#| label: load_data
#| message: false
#| warning: false

# Define path to the Excel workbook relative to repository root
excel_path <- here::here("data", "1_effect_size_calculation_pipeline", "Data Extraction Mate Choice Meta Analysis.xlsx")

# Read EduardoExtraction sheet
db_raw <- readxl::read_excel(excel_path, sheet = "EduardoExtraction")

# Filter out rows flagged for removal and convert key columns
db <- db_raw %>%
  # Filter out rows marked for deletion (keep rows where RemoveFromDataset is NA or not "Yes")
  filter(is.na(RemoveFromDataset) | RemoveFromDataset != "Yes") %>%
  # Convert variables to factor/numeric structures
  mutate(across(where(is.character), as.factor)) %>%
  mutate(
    effectSizeCalculatedHedgesD = as.numeric(effectSizeCalculatedHedgesD),
    effectSizeCalculatedHedgesDVariance = as.numeric(effectSizeCalculatedHedgesDVariance),
    effectSizeCalculatedOddsRatio = as.numeric(effectSizeCalculatedOddsRatio),
    effectSizeCalculatedOddsRatioVariance = as.numeric(effectSizeCalculatedOddsRatioVariance)
  )

# Add log Odds Ratio and its variance for meta-analytic modeling
db <- db %>%
  mutate(
    logOddsRatio = log(effectSizeCalculatedOddsRatio),
    logOddsRatioVariance = effectSizeCalculatedOddsRatioVariance
  )
```

We fit multilevel random-effects meta-analytic models using the `metafor` package to estimate the overall effect sizes for:
1. **Hedges' $g$** (standardised mean difference for continuous/converted measurements).
2. **Log Odds Ratio ($\ln\text{OR}$)** (log-transformed odds ratio for binary/converted outcomes).

Because studies contribute multiple, non-independent trials (outcomes), we construct a variance–covariance (VCV) matrix using `vcalc()` to control for within-study correlation (assuming $r = 0.5$ for nested measurements where correlation is unreported).

---

## Model Fitting Function

This helper function handles VCV matrix construction and fits a multilevel random-effects model using `rma.mv()` in `metafor`. 

* **Random intercepts** are modeled for:
  * Study (`identifierStudyId`)
  * Individual trial/effect size (`identifierEffectSizeID`)
  * Species (`taxonomySpecies`) to control for taxonomic/species-level variation

```{r}
#| label: model_fitting_function

fit_overall_mv <- function(dat, yi, vi, cluster_var = "identifierStudyId", obs_var = "identifierEffectSizeID", rho = 0.5) {
  
  # Remove rows with missing effect size or variance estimates
  dat2 <- dat %>%
    filter(!is.na(.data[[yi]]), !is.na(.data[[vi]]), .data[[vi]] > 0)
  
  if (nrow(dat2) == 0) {
    message(paste("No complete cases found for", yi))
    return(NULL)
  }
  
  # 1. Build Variance-Covariance matrix to account for shared cohort/study errors
  VCV <- metafor::vcalc(
    vi = dat2[[vi]],
    cluster = dat2[[cluster_var]],
    obs = dat2[[obs_var]],
    rho = rho,
    data = dat2
  )
  
  # 2. Fit the Multi-Level Meta-Analytic Model
  mod <- metafor::rma.mv(
    yi = dat2[[yi]],
    V  = VCV,
    random = list(
      ~1 | identifierStudyId,
      ~1 | identifierEffectSizeID,
      ~1 | taxonomySpecies
    ),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = dat2
  )
  
  list(data = dat2, VCV = VCV, model = mod)
}
```

---

## Fit Overall Models

We run the multilevel meta-analysis for both Hedges' $g$ and Log Odds Ratio.

```{r}
#| label: fit_models
#| message: false
#| warning: false

res_hedges <- fit_overall_mv(db, yi = "effectSizeCalculatedHedgesD", vi = "effectSizeCalculatedHedgesDVariance")
res_log_or <- fit_overall_mv(db, yi = "logOddsRatio", vi = "logOddsRatioVariance")
```

---

## Results and Visualizations

::: {.panel-tabset}

### Hedges' $g$ (Standardised Mean Difference)

```{r}
#| label: hedges_results
#| message: false
#| warning: false

if (!is.null(res_hedges)) {
  # Model Summary
  print(summary(res_hedges$model))
  
  # Multi-level Heterogeneity (I^2)
  cat("\nHeterogeneity estimates (I^2):\n")
  print(orchaRd::i2_ml(res_hedges$model))
  
  # Orchard Plot
  p_hedges <- orchaRd::orchard_plot(
    res_hedges$model,
    group = "identifierStudyId",
    xlab = "Hedges' g",
    flip = TRUE
  ) +
    ggplot2::scale_x_discrete(labels = "Overall Hedges' g")
  
  p_hedges
} else {
  cat("Hedges' g model could not be fitted due to insufficient data.\n")
}
```

### Log Odds Ratio ($\ln\text{OR}$)

```{r}
#| label: or_results
#| message: false
#| warning: false

if (!is.null(res_log_or)) {
  # Model Summary
  print(summary(res_log_or$model))
  
  # Multi-level Heterogeneity (I^2)
  cat("\nHeterogeneity estimates (I^2):\n")
  print(orchaRd::i2_ml(res_log_or$model))
  
  # Orchard Plot
  p_or <- orchaRd::orchard_plot(
    res_log_or$model,
    group = "identifierStudyId",
    xlab = "Log Odds Ratio (lnOR)",
    flip = TRUE
  ) +
    ggplot2::scale_x_discrete(labels = "Overall lnOR")
  
  p_or
} else {
  cat("Log Odds Ratio model could not be fitted due to insufficient data.\n")
}
```

:::

---

## Session Info

```{r}
#| label: session_info
sessionInfo()
```
