---
title: "Overall Effect Size Calculation & Synthesis"
subtitle: "Meta-Analytic Update of Mate-Choice Copying (MCC)"
author: "Eduardo Santos & Collaborators"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    theme: cosmo
    self-contained: true
---

```{r}
#| label: setup
#| include: false
# Ensure pacman is loaded to manage dependencies
if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
pacman::p_load(
               readxl,
               tidyverse,
               metafor,
               orchaRd,
               patchwork,
               DT
               )
```

## Introduction & Project Update

This protocol implements the overall effect size synthesis for the updated Mate-Choice Copying (MCC) meta-analysis. 

Specifically, we integrate:
1. **The New Extractions:** Our newly extracted and verified dataset containing recent trials.
2. **Original Continuous Dataset (Davies et al. 2020):** Standardised mean differences (Hedges' $d$) analyzing continuous metrics.
3. **Original Binary Dataset (Jones & DuVal 2019):** Log Odds Ratio ($\ln\text{OR}$) analyzing binary choices.

The R script aligns and synthesizes these three sources into unified datasets, allowing comparison of meta-analytic summaries across three key subsets:
* **New Extractions Only:** Eduardo's new dataset.
* **Original Only:** Legacy data from the original publications.
* **Combined Dataset:** A complete synthesis of old and new trials to estimate the updated global effect sizes.

---

## Data Loading and Alignment

We read the three datasets, perform data type conversions, extract species epithets for taxonomic control, and align columns into unified Hedges' $g$ and Odds Ratio datasets.

```{r}
#| label: load_and_align_datasets
#| message: false
#| warning: false

# ---------------------------------------------------------------------------
# 1. Load Datasets
# ---------------------------------------------------------------------------

# Paths to legacy and new datasets
path_new <- here::here("data", "1_effect_size_calculation_pipeline", "Data Extraction Mate Choice Meta Analysis.xlsx")
path_davies <- here::here("data", "2_original_meta_analyses_datasets", "Davies_et_al_2020_Final_data.xlsx")
path_jones <- here::here("data", "2_original_meta_analyses_datasets", "Jones_DuVal_2019_data.CSV")

# A. Read new extraction dataset (EduardoExtraction sheet)
db_new_raw <- readxl::read_excel(path_new, sheet = "EduardoExtraction")

# B. Read Davies et al. 2020 (Original continuous dataset)
db_davies_raw <- readxl::read_excel(path_davies)

# C. Read Jones & DuVal 2019 (Original binary dataset)
db_jones_raw <- readr::read_csv(path_jones)

# ---------------------------------------------------------------------------
# 2. Process and Filter New Extraction
# ---------------------------------------------------------------------------
db_new <- db_new_raw %>%
  filter(is.na(RemoveFromDataset) | RemoveFromDataset != "Yes") %>%
  mutate(
    effectSizeCalculatedHedgesD = as.numeric(effectSizeCalculatedHedgesD),
    effectSizeCalculatedHedgesDVariance = as.numeric(effectSizeCalculatedHedgesDVariance),
    effectSizeCalculatedOddsRatio = as.numeric(effectSizeCalculatedOddsRatio),
    effectSizeCalculatedOddsRatioVariance = as.numeric(effectSizeCalculatedOddsRatioVariance),
    logOddsRatio = log(effectSizeCalculatedOddsRatio),
    logOddsRatioVariance = effectSizeCalculatedOddsRatioVariance
  )

# ---------------------------------------------------------------------------
# 3. Align Datasets for Hedges' g (Continuous SMD Analysis)
# ---------------------------------------------------------------------------

# A. Hedges' g from New Extraction
hedges_new <- db_new %>%
  filter(!is.na(effectSizeCalculatedHedgesD), !is.na(effectSizeCalculatedHedgesDVariance)) %>%
  select(
    yi = effectSizeCalculatedHedgesD,
    vi = effectSizeCalculatedHedgesDVariance,
    identifierStudyId,
    identifierEffectSizeID,
    taxonomySpecies
  ) %>%
  mutate(
    identifierStudyId = as.character(identifierStudyId),
    identifierEffectSizeID = as.character(identifierEffectSizeID),
    # Extract species epithet (last word) and make lowercase for cross-dataset mapping
    taxonomySpecies = str_to_lower(word(as.character(taxonomySpecies), -1)),
    Dataset_Source = "New Extraction (Santos)"
  )

# B. Hedges' g from Davies et al. 2020 (Original)
hedges_davies <- db_davies_raw %>%
  filter(!is.na(Hedges_d), !is.na(Variance), Variance > 0) %>%
  select(
    yi = Hedges_d,
    vi = Variance,
    identifierStudyId = Author,
    identifierEffectSizeID = Effect_size_no,
    taxonomySpecies = Species_latin
  ) %>%
  mutate(
    identifierStudyId = as.character(identifierStudyId),
    identifierEffectSizeID = as.character(identifierEffectSizeID),
    # Extract species epithet (last word) and make lowercase
    taxonomySpecies = str_to_lower(word(as.character(taxonomySpecies), -1)),
    Dataset_Source = "Original (Davies 2020)"
  )

# C. Unified Hedges' g Dataset
db_hedges <- bind_rows(hedges_new, hedges_davies) %>%
  mutate(across(c(identifierStudyId, taxonomySpecies, Dataset_Source), as.factor))

# ---------------------------------------------------------------------------
# 4. Align Datasets for Odds Ratio (Binary Analysis)
# ---------------------------------------------------------------------------

# A. Log Odds Ratio from New Extraction
or_new <- db_new %>%
  filter(!is.na(logOddsRatio), !is.na(logOddsRatioVariance)) %>%
  select(
    yi = logOddsRatio,
    vi = logOddsRatioVariance,
    identifierStudyId,
    identifierEffectSizeID,
    taxonomySpecies
  ) %>%
  mutate(
    identifierStudyId = as.character(identifierStudyId),
    identifierEffectSizeID = as.character(identifierEffectSizeID),
    taxonomySpecies = str_to_lower(word(as.character(taxonomySpecies), -1)),
    Dataset_Source = "New Extraction (Santos)"
  )

# B. Log Odds Ratio from Jones & DuVal 2019 (Original)
or_jones <- db_jones_raw %>%
  filter(!is.na(lnOR), !is.na(VlnOR), VlnOR > 0) %>%
  select(
    yi = lnOR,
    vi = VlnOR,
    identifierStudyId = study,
    identifierEffectSizeID = study_Exp,
    taxonomySpecies = Species
  ) %>%
  mutate(
    identifierStudyId = as.character(identifierStudyId),
    identifierEffectSizeID = as.character(identifierEffectSizeID),
    taxonomySpecies = str_to_lower(as.character(taxonomySpecies)),
    Dataset_Source = "Original (Jones & DuVal 2019)"
  )

# C. Unified Log Odds Ratio Dataset
db_or <- bind_rows(or_new, or_jones) %>%
  mutate(across(c(identifierStudyId, taxonomySpecies, Dataset_Source), as.factor))
```

---

## Model Fitting Function

This helper function handles the variance–covariance (VCV) matrix construction and fits a multilevel random-effects model using `rma.mv()` in `metafor`. 

* **Variance–Covariance:** We construct a VCV matrix via `vcalc()` to handle dependencies among multiple effect sizes within the same study (assuming $r = 0.5$ for nested measurements).
* **Random intercepts** are modeled for:
  * Study (`identifierStudyId`)
  * Individual trial/effect size (`identifierEffectSizeID`)
  * Species (`taxonomySpecies`) to control for taxonomic/species-level variation

```{r}
#| label: model_fitting_function

fit_overall_mv <- function(dat, yi = "yi", vi = "vi", cluster_var = "identifierStudyId", obs_var = "identifierEffectSizeID", rho = 0.5) {
  
  # Remove rows with missing values
  dat2 <- dat %>%
    filter(!is.na(.data[[yi]]), !is.na(.data[[vi]]), .data[[vi]] > 0)
  
  if (nrow(dat2) == 0) {
    message("No complete cases found.")
    return(NULL)
  }
  
  # 1. Build VCV Matrix
  VCV <- metafor::vcalc(
    vi = dat2[[vi]],
    cluster = dat2[[cluster_var]],
    obs = dat2[[obs_var]],
    rho = rho,
    data = dat2
  )
  
  # 2. Fit Multi-Level Random-Effects Model
  mod <- metafor::rma.mv(
    yi = dat2[[yi]],
    V  = VCV,
    random = list(
      ~1 | identifierStudyId,
      ~1 | identifierEffectSizeID,
      ~1 | taxonomySpecies
    ),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = dat2
  )
  
  list(data = dat2, VCV = VCV, model = mod)
}
```

---

## Multi-Level Synthesis Results

We fit the multilevel meta-analysis across all three database splits (New Only, Original Only, and Combined) for both effect size classes.

```{r}
#| label: fit_multilevel_models
#| message: false
#| warning: false

# ---------------------------------------------------------------------------
# HEDGES' G MODELS
# ---------------------------------------------------------------------------
res_hedges_new   <- fit_overall_mv(db_hedges %>% filter(Dataset_Source == "New Extraction (Santos)"))
res_hedges_orig  <- fit_overall_mv(db_hedges %>% filter(Dataset_Source == "Original (Davies 2020)"))
res_hedges_comb  <- fit_overall_mv(db_hedges)

# ---------------------------------------------------------------------------
# ODDS RATIO MODELS
# ---------------------------------------------------------------------------
res_or_new   <- fit_overall_mv(db_or %>% filter(Dataset_Source == "New Extraction (Santos)"))
res_or_orig  <- fit_overall_mv(db_or %>% filter(Dataset_Source == "Original (Jones & DuVal 2019)"))
res_or_comb  <- fit_overall_mv(db_or)
```

---

## Results and Visualizations

::: {.panel-tabset}

### Hedges' $g$ (Continuous outcomes)

::: {.panel-tabset}

#### New Extraction Only

This model isolates Eduardo's newly extracted dataset.

```{r}
#| label: hedges_new_summary
#| message: false
#| warning: false

if (!is.null(res_hedges_new)) {
  print(summary(res_hedges_new$model))
  cat("\nHeterogeneity estimates (I^2):\n")
  print(orchaRd::i2_ml(res_hedges_new$model))
  
  orchaRd::orchard_plot(res_hedges_new$model, group = "identifierStudyId", xlab = "Hedges' g", flip = TRUE) +
    ggplot2::scale_x_discrete(labels = "New Extraction")
}
```

#### Original Only (Davies 2020)

This model isolates legacy data from the original Davies et al. (2020) meta-analysis.

```{r}
#| label: hedges_orig_summary
#| message: false
#| warning: false

if (!is.null(res_hedges_orig)) {
  print(summary(res_hedges_orig$model))
  cat("\nHeterogeneity estimates (I^2):\n")
  print(orchaRd::i2_ml(res_hedges_orig$model))
  
  orchaRd::orchard_plot(res_hedges_orig$model, group = "identifierStudyId", xlab = "Hedges' g", flip = TRUE) +
    ggplot2::scale_x_discrete(labels = "Davies 2020")
}
```

#### Combined Analysis

This model synthesizes legacy data with new extractions for Hedges' $g$.

```{r}
#| label: hedges_comb_summary
#| message: false
#| warning: false

if (!is.null(res_hedges_comb)) {
  print(summary(res_hedges_comb$model))
  cat("\nHeterogeneity estimates (I^2):\n")
  print(orchaRd::i2_ml(res_hedges_comb$model))
  
  orchaRd::orchard_plot(res_hedges_comb$model, group = "identifierStudyId", xlab = "Hedges' g", flip = TRUE) +
    ggplot2::scale_x_discrete(labels = "Combined Hedges' g")
}
```

:::

### Odds Ratio (Binary outcomes)

::: {.panel-tabset}

#### New Extraction Only

This model isolates Eduardo's newly extracted dataset for Odds Ratios.

```{r}
#| label: or_new_summary
#| message: false
#| warning: false

if (!is.null(res_or_new)) {
  print(summary(res_or_new$model))
  cat("\nHeterogeneity estimates (I^2):\n")
  print(orchaRd::i2_ml(res_or_new$model))
  
  orchaRd::orchard_plot(res_or_new$model, group = "identifierStudyId", xlab = "Log Odds Ratio", flip = TRUE) +
    ggplot2::scale_x_discrete(labels = "New Extraction")
}
```

#### Original Only (Jones & DuVal 2019)

This model isolates legacy data from the original Jones & DuVal (2019) meta-analysis.

```{r}
#| label: or_orig_summary
#| message: false
#| warning: false

if (!is.null(res_or_orig)) {
  print(summary(res_or_orig$model))
  cat("\nHeterogeneity estimates (I^2):\n")
  print(orchaRd::i2_ml(res_or_orig$model))
  
  orchaRd::orchard_plot(res_or_orig$model, group = "identifierStudyId", xlab = "Log Odds Ratio", flip = TRUE) +
    ggplot2::scale_x_discrete(labels = "Jones & DuVal 2019")
}
```

#### Combined Analysis

This model synthesizes legacy data with new extractions for Odds Ratios.

```{r}
#| label: or_comb_summary
#| message: false
#| warning: false

if (!is.null(res_or_comb)) {
  print(summary(res_or_comb$model))
  cat("\nHeterogeneity estimates (I^2):\n")
  print(orchaRd::i2_ml(res_or_comb$model))
  
  orchaRd::orchard_plot(res_or_comb$model, group = "identifierStudyId", xlab = "Log Odds Ratio", flip = TRUE) +
    ggplot2::scale_x_discrete(labels = "Combined lnOR")
}
```

:::

:::

---

## Session Info

```{r}
#| label: session_info
sessionInfo()
```
