---
title: "Overall Effect Size Calculation & Synthesis"
subtitle: "Meta-Analytic Update of Mate-Choice Copying (MCC)"
author: "Eduardo Santos & Collaborators"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    theme: cosmo
    self-contained: true
---

```{r}
#| label: setup
#| include: false
# Ensure pacman is loaded to manage dependencies
if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
pacman::p_load(
               readxl,
               tidyverse,
               metafor,
               orchaRd,
               patchwork,
               DT
               )
```

## Introduction & Project Update

This protocol implements the overall effect size synthesis for the updated Mate-Choice Copying (MCC) meta-analysis. 

Specifically, we integrate:
1. **The New Extractions:** Our newly extracted and verified dataset containing recent trials.
2. **Original Continuous Dataset (Davies et al. 2020):** Standardised mean differences (Hedges' $d$) analyzing continuous metrics.
3. **Original Binary Dataset (Jones & DuVal 2019):** Log Odds Ratio ($\ln\text{OR}$) analyzing binary choices.

The R script aligns and synthesizes these three sources into unified datasets, allowing comparison of meta-analytic summaries across three key subsets:
* **New Extractions Only:** Eduardo's new dataset.
* **Original Only:** Legacy data from the original publications.
* **Combined Dataset:** A complete synthesis of old and new trials to estimate the updated global effect sizes.

---

## Data Loading and Alignment

We read the three datasets, perform data type conversions, extract species epithets for taxonomic control, and align columns into unified Hedges' $g$ and Odds Ratio datasets.

```{r}
#| label: load_and_align_datasets
#| message: false
#| warning: false

# ---------------------------------------------------------------------------
# 1. Load Datasets
# ---------------------------------------------------------------------------

# Paths to legacy and new datasets
path_new <- here::here("data", "1_effect_size_calculation_pipeline", "Data Extraction Mate Choice Meta Analysis.xlsx")
path_davies <- here::here("data", "2_original_meta_analyses_datasets", "Davies_et_al_2020_Final_data.xlsx")
path_jones <- here::here("data", "2_original_meta_analyses_datasets", "Jones_DuVal_2019_data.CSV")

# A. Read new extraction dataset (EduardoExtraction sheet)
db_new_raw <- readxl::read_excel(path_new, sheet = "EduardoExtraction")

# B. Read Davies et al. 2020 (Original continuous dataset)
db_davies_raw <- readxl::read_excel(path_davies, sheet = "Data")

# C. Read Jones & DuVal 2019 (Original binary dataset)
db_jones_raw <- readr::read_csv(path_jones)

# ---------------------------------------------------------------------------
# 2. Process and Filter New Extraction
# ---------------------------------------------------------------------------
db_new <- db_new_raw %>%
  filter(is.na(RemoveFromDataset) | RemoveFromDataset != "Yes") %>%
  mutate(
    effectSizeCalculatedHedgesD = as.numeric(effectSizeCalculatedHedgesD),
    effectSizeCalculatedHedgesDVariance = as.numeric(effectSizeCalculatedHedgesDVariance),
    effectSizeCalculatedOddsRatio = as.numeric(effectSizeCalculatedOddsRatio),
    effectSizeCalculatedOddsRatioVariance = as.numeric(effectSizeCalculatedOddsRatioVariance),
    logOddsRatio = log(effectSizeCalculatedOddsRatio),
    logOddsRatioVariance = effectSizeCalculatedOddsRatioVariance
  )

# ---------------------------------------------------------------------------
# 3. Align Datasets for Hedges' g (Continuous SMD Analysis)
# ---------------------------------------------------------------------------

# A. Hedges' g from New Extraction
hedges_new <- db_new %>%
  filter(!is.na(effectSizeCalculatedHedgesD), !is.na(effectSizeCalculatedHedgesDVariance)) %>%
  select(
    yi = effectSizeCalculatedHedgesD,
    vi = effectSizeCalculatedHedgesDVariance,
    identifierStudyId,
    identifierEffectSizeID,
    taxonomySpecies
  ) %>%
  mutate(
    identifierStudyId = as.character(identifierStudyId),
    identifierEffectSizeID = as.character(identifierEffectSizeID),
    # Extract species epithet (last word) and make lowercase for cross-dataset mapping
    taxonomySpecies = str_to_lower(word(as.character(taxonomySpecies), -1)),
    Dataset_Source = "New Extraction (Santos)"
  )

# B. Hedges' g from Davies et al. 2020 (Original)
hedges_davies <- db_davies_raw %>%
  mutate(
    Hedges_d_directional = as.numeric(Hedges_d_directional),
    Variance = as.numeric(Variance)
  ) %>%
  filter(!is.na(Hedges_d_directional), !is.na(Variance), Variance > 0) %>%
  select(
    yi = Hedges_d_directional,
    vi = Variance,
    identifierStudyId = Author,
    identifierEffectSizeID = Effect_size_no,
    taxonomySpecies = Species_latin
  ) %>%
  mutate(
    identifierStudyId = as.character(identifierStudyId),
    identifierEffectSizeID = as.character(identifierEffectSizeID),
    # Extract species epithet (last word) and make lowercase
    taxonomySpecies = str_to_lower(word(as.character(taxonomySpecies), -1)),
    Dataset_Source = "Original (Davies 2020)"
  )

# C. Unified Hedges' g Dataset
db_hedges <- bind_rows(hedges_new, hedges_davies) %>%
  mutate(across(c(identifierStudyId, taxonomySpecies, Dataset_Source), as.factor))

# ---------------------------------------------------------------------------
# 4. Align Datasets for Odds Ratio (Binary Analysis)
# ---------------------------------------------------------------------------

# A. Log Odds Ratio from New Extraction
or_new <- db_new %>%
  filter(!is.na(logOddsRatio), !is.na(logOddsRatioVariance)) %>%
  select(
    yi = logOddsRatio,
    vi = logOddsRatioVariance,
    identifierStudyId,
    identifierEffectSizeID,
    taxonomySpecies
  ) %>%
  mutate(
    identifierStudyId = as.character(identifierStudyId),
    identifierEffectSizeID = as.character(identifierEffectSizeID),
    taxonomySpecies = str_to_lower(word(as.character(taxonomySpecies), -1)),
    Dataset_Source = "New Extraction (Santos)"
  )

# B. Log Odds Ratio from Jones & DuVal 2019 (Original)
or_jones <- db_jones_raw %>%
  filter(!is.na(lnOR), !is.na(VlnOR), VlnOR > 0) %>%
  select(
    yi = lnOR,
    vi = VlnOR,
    identifierStudyId = study,
    identifierEffectSizeID = study_Exp,
    taxonomySpecies = Species
  ) %>%
  mutate(
    identifierStudyId = as.character(identifierStudyId),
    identifierEffectSizeID = as.character(identifierEffectSizeID),
    taxonomySpecies = str_to_lower(as.character(taxonomySpecies)),
    Dataset_Source = "Original (Jones & DuVal 2019)"
  )

# C. Unified Log Odds Ratio Dataset
db_or <- bind_rows(or_new, or_jones) %>%
  mutate(across(c(identifierStudyId, taxonomySpecies, Dataset_Source), as.factor))
```

---

## Effect Size Distribution (Density Plots) {#sec-density}

Before fitting the multi-level models, we inspect the distribution of our aligned effect sizes across the different subsets (New Extraction, Original Only, and Combined). 

The density plots below use the same split logic as the orchard plots, facet-wrapping by subset while coloring the distribution curves by their actual dataset source to check for shifts or outliers.

::: {.panel-tabset}

### Hedges' $g$ (Continuous SMD)

```{r}
#| label: hedges_density_plots
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 9

# Prepare density data for Hedges' g
db_hedges_density <- bind_rows(
  db_hedges %>% mutate(Group = as.character(Dataset_Source)),
  db_hedges %>% mutate(Group = "Combined")
) %>%
  mutate(Group = factor(Group, levels = c("New Extraction (Santos)", "Original (Davies 2020)", "Combined")))

# Plot densities
p_hedges_dens <- ggplot(db_hedges_density, aes(x = yi, fill = Dataset_Source, color = Dataset_Source)) +
  geom_density(alpha = 0.4, linewidth = 0.8) +
  geom_vline(xintercept = 0, linewidth = 1.0, linetype = "solid", color = "black") +
  geom_rug(alpha = 0.6) +
  facet_wrap(~ Group, ncol = 1) +
  scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
  scale_color_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
  labs(
    title = "Distribution of Hedges' g Effect Sizes",
    x = "Hedges' g",
    y = "Density"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    strip.text = element_text(face = "bold", size = 11),
    plot.title = element_text(face = "bold", hjust = 0.5)
  )

print(p_hedges_dens)
```

### Odds Ratio (Binary outcomes)

```{r}
#| label: or_density_plots
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 9

# Prepare density data for Odds Ratio (lnOR)
db_or_density <- bind_rows(
  db_or %>% mutate(Group = as.character(Dataset_Source)),
  db_or %>% mutate(Group = "Combined")
) %>%
  mutate(Group = factor(Group, levels = c("New Extraction (Santos)", "Original (Jones & DuVal 2019)", "Combined")))

# Plot densities
p_or_dens <- ggplot(db_or_density, aes(x = yi, fill = Dataset_Source, color = Dataset_Source)) +
  geom_density(alpha = 0.4, linewidth = 0.8) +
  geom_vline(xintercept = 0, linewidth = 1.0, linetype = "solid", color = "black") +
  geom_rug(alpha = 0.6) +
  facet_wrap(~ Group, ncol = 1) +
  scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal 2019)" = "#0072B2"), name = "Data Source") +
  scale_color_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal 2019)" = "#0072B2"), name = "Data Source") +
  labs(
    title = "Distribution of Log Odds Ratio (lnOR) Effect Sizes",
    x = "Log Odds Ratio (lnOR)",
    y = "Density"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    strip.text = element_text(face = "bold", size = 11),
    plot.title = element_text(face = "bold", hjust = 0.5)
  )

print(p_or_dens)
```

:::

---

## Model Fitting Function

This helper function handles the variance–covariance (VCV) matrix construction and fits a multilevel random-effects model using `rma.mv()` in `metafor`. 

* **Variance–Covariance:** We construct a VCV matrix via `vcalc()` to handle dependencies among multiple effect sizes within the same study (assuming $r = 0.5$ for nested measurements).
* **Random intercepts** are modeled for:
  * Study (`identifierStudyId`)
  * Individual trial/effect size (`identifierEffectSizeID`)
  * Species (`taxonomySpecies`) to control for taxonomic/species-level variation

```{r}
#| label: model_fitting_function

fit_overall_mv <- function(dat, yi = "yi", vi = "vi", cluster_var = "identifierStudyId", obs_var = "identifierEffectSizeID", rho = 0.5) {
  
  # Remove rows with missing values
  dat2 <- dat %>%
    filter(!is.na(.data[[yi]]), !is.na(.data[[vi]]), .data[[vi]] > 0)
  
  if (nrow(dat2) == 0) {
    message("No complete cases found.")
    return(NULL)
  }
  
  # 1. Build VCV Matrix
  VCV <- metafor::vcalc(
    vi = dat2[[vi]],
    cluster = dat2[[cluster_var]],
    obs = dat2[[obs_var]],
    rho = rho,
    data = dat2
  )
  
  # 2. Fit Multi-Level Random-Effects Model
  mod <- metafor::rma.mv(
    yi = dat2[[yi]],
    V  = VCV,
    random = list(
      ~1 | identifierStudyId,
      ~1 | identifierEffectSizeID,
      ~1 | taxonomySpecies
    ),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = dat2
  )
  
  list(data = dat2, VCV = VCV, model = mod)
}

# Helper function to fit bias-corrected multilevel models with a continuous moderator (Egger-type)
fit_corrected_mv <- function(dat, yi = "yi", vi = "vi", moderator = "vi", cluster_var = "identifierStudyId", obs_var = "identifierEffectSizeID", rho = 0.5) {
  
  # Remove rows with missing values
  dat2 <- dat %>%
    filter(!is.na(.data[[yi]]), !is.na(.data[[vi]]), .data[[vi]] > 0)
  
  if (nrow(dat2) == 0) {
    message("No complete cases found.")
    return(NULL)
  }
  
  if (moderator == "se" && !"se" %in% colnames(dat2)) {
    dat2 <- dat2 %>% mutate(se = sqrt(vi))
  }
  
  # 1. Build VCV Matrix
  VCV <- metafor::vcalc(
    vi = dat2[[vi]],
    cluster = dat2[[cluster_var]],
    obs = dat2[[obs_var]],
    rho = rho,
    data = dat2
  )
  
  # 2. Fit Multi-Level Egger Model using dynamic formula
  formula_str <- paste("~ 1 +", moderator)
  mod <- metafor::rma.mv(
    yi = dat2[[yi]],
    V  = VCV,
    mods = as.formula(formula_str),
    random = list(
      ~1 | identifierStudyId,
      ~1 | identifierEffectSizeID,
      ~1 | taxonomySpecies
    ),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = dat2
  )
  
  list(data = dat2, VCV = VCV, model = mod)
}
```

---

## Multi-Level Synthesis Results

We fit the multilevel meta-analysis across all three database splits (New Only, Original Only, and Combined) for both effect size classes.

```{r}
#| label: fit_multilevel_models
#| message: false
#| warning: false

# ---------------------------------------------------------------------------
# HEDGES' G MODELS (UNCORRECTED)
# ---------------------------------------------------------------------------
res_hedges_new   <- fit_overall_mv(db_hedges %>% filter(Dataset_Source == "New Extraction (Santos)"))
res_hedges_orig  <- fit_overall_mv(db_hedges %>% filter(Dataset_Source == "Original (Davies 2020)"))
res_hedges_comb  <- fit_overall_mv(db_hedges)

# HEDGES' G MODELS (BIAS-CORRECTED)
res_hedges_new_corr   <- fit_corrected_mv(db_hedges %>% filter(Dataset_Source == "New Extraction (Santos)"), moderator = "vi")
res_hedges_orig_corr  <- fit_corrected_mv(db_hedges %>% filter(Dataset_Source == "Original (Davies 2020)"), moderator = "vi")
res_hedges_comb_corr  <- fit_corrected_mv(db_hedges, moderator = "vi")

# ---------------------------------------------------------------------------
# ODDS RATIO MODELS (UNCORRECTED)
# ---------------------------------------------------------------------------
res_or_new   <- fit_overall_mv(db_or %>% filter(Dataset_Source == "New Extraction (Santos)"))
res_or_orig  <- fit_overall_mv(db_or %>% filter(Dataset_Source == "Original (Jones & DuVal 2019)"))
res_or_comb  <- fit_overall_mv(db_or)

# ODDS RATIO MODELS (BIAS-CORRECTED)
res_or_new_corr   <- fit_corrected_mv(db_or %>% filter(Dataset_Source == "New Extraction (Santos)"), moderator = "vi")
res_or_orig_corr  <- fit_corrected_mv(db_or %>% filter(Dataset_Source == "Original (Jones & DuVal 2019)"), moderator = "vi")
res_or_comb_corr  <- fit_corrected_mv(db_or, moderator = "vi")
```

---

## Results and Visualizations

::: {.panel-tabset}

### Hedges' $g$ (Continuous outcomes)

::: {.panel-tabset}

#### Uncorrected Overall Effect

This plot displays the standard multi-level meta-analytic summaries across the three splits (New Only, Original Only, and Combined) *without* adjusting for publication bias. Points represent individual effect sizes colored by their data source.

```{r}
#| label: hedges_unified_plot
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 7

if (!is.null(res_hedges_new) && !is.null(res_hedges_orig) && !is.null(res_hedges_comb)) {
  # 1. Extract S3 result objects using mod_results
  h_new <- orchaRd::mod_results(res_hedges_new$model, group = "identifierStudyId")
  h_orig <- orchaRd::mod_results(res_hedges_orig$model, group = "identifierStudyId")
  h_comb <- orchaRd::mod_results(res_hedges_comb$model, group = "identifierStudyId")

  # 2. Clone S3 result objects to lists for manual editing
  class(h_new) <- "list"
  h_new$mod_table$name <- "New Extraction (Santos)"
  h_new$data$moderator <- "New Extraction (Santos)"
  class(h_new) <- c("orchard", "data.frame")

  class(h_orig) <- "list"
  h_orig$mod_table$name <- "Original (Davies 2020)"
  h_orig$data$moderator <- "Original (Davies 2020)"
  class(h_orig) <- c("orchard", "data.frame")

  class(h_comb) <- "list"
  h_comb$mod_table$name <- "Combined"
  h_comb$data$moderator <- "Combined"
  class(h_comb) <- c("orchard", "data.frame")

  # 3. Merge model results
  sub_hedges <- orchaRd::submerge(h_new, h_orig, h_comb)

  # 4. Add point-level real source mapping
  class(sub_hedges) <- "list"
  sub_hedges$data$real_source <- c(
    rep("New Extraction (Santos)", nrow(h_new$data)),
    rep("Original (Davies 2020)", nrow(h_orig$data)),
    rep("New Extraction (Santos)", nrow(h_new$data)),
    rep("Original (Davies 2020)", nrow(h_orig$data))
  )
  class(sub_hedges) <- c("orchard", "data.frame")

  # 4. Generate Plot
  p_hedges <- orchaRd::orchard_plot(sub_hedges, xlab = "Hedges' g", flip = TRUE)

  # 5. Customize color mapping for points
  p_hedges$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
  p_hedges$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

  # 6. Apply themes & custom palette (colorblind friendly orange/blue)
  p_hedges <- p_hedges +
    ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
    ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
    ggplot2::ggtitle("Overall Synthesis — Hedges' g Estimates (Uncorrected)") +
    ggplot2::theme(legend.position = "right")

  print(p_hedges)
}
```

#### Bias-Corrected Overall Effect

This plot displays the small-study publication bias corrected meta-analytic summaries (evaluated at $\text{V} = 0$, representing the theoretical estimate for a study of infinite sample size/precision) using the two-step Egger-type meta-regression framework. Points represent the same raw effect sizes.

```{r}
#| label: hedges_corrected_plot
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 7

if (!is.null(res_hedges_new_corr) && !is.null(res_hedges_orig_corr) && !is.null(res_hedges_comb_corr)) {
  # 1. Extract S3 result objects using mod_results at vi = 0
  h_new_corr <- orchaRd::mod_results(res_hedges_new_corr$model, group = "identifierStudyId", at = list(vi = 0), mod = "1")
  h_orig_corr <- orchaRd::mod_results(res_hedges_orig_corr$model, group = "identifierStudyId", at = list(vi = 0), mod = "1")
  h_comb_corr <- orchaRd::mod_results(res_hedges_comb_corr$model, group = "identifierStudyId", at = list(vi = 0), mod = "1")

  # 2. Clone S3 result objects to lists for manual editing
  class(h_new_corr) <- "list"
  h_new_corr$mod_table$name <- "New Extraction (Santos)"
  h_new_corr$data$moderator <- "New Extraction (Santos)"
  class(h_new_corr) <- c("orchard", "data.frame")

  class(h_orig_corr) <- "list"
  h_orig_corr$mod_table$name <- "Original (Davies 2020)"
  h_orig_corr$data$moderator <- "Original (Davies 2020)"
  class(h_orig_corr) <- c("orchard", "data.frame")

  class(h_comb_corr) <- "list"
  h_comb_corr$mod_table$name <- "Combined"
  h_comb_corr$data$moderator <- "Combined"
  class(h_comb_corr) <- c("orchard", "data.frame")

  # 3. Merge model results
  sub_hedges_corr <- orchaRd::submerge(h_new_corr, h_orig_corr, h_comb_corr)

  # 4. Add point-level real source mapping
  class(sub_hedges_corr) <- "list"
  sub_hedges_corr$data$real_source <- c(
    rep("New Extraction (Santos)", nrow(h_new_corr$data)),
    rep("Original (Davies 2020)", nrow(h_orig_corr$data)),
    rep("New Extraction (Santos)", nrow(h_new_corr$data)),
    rep("Original (Davies 2020)", nrow(h_orig_corr$data))
  )
  class(sub_hedges_corr) <- c("orchard", "data.frame")

  # 4. Generate Plot
  p_hedges_corr <- orchaRd::orchard_plot(sub_hedges_corr, xlab = "Hedges' g (Bias Corrected)", flip = TRUE)

  # 5. Customize color mapping for points
  p_hedges_corr$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
  p_hedges_corr$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

  # 6. Apply themes & custom palette
  p_hedges_corr <- p_hedges_corr +
    ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
    ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
    ggplot2::ggtitle("Overall Synthesis — Hedges' g Estimates (Bias-Corrected)") +
    ggplot2::theme(legend.position = "right")

  print(p_hedges_corr)
}
```

#### Side-by-Side Comparison

We combine the uncorrected and small-study bias corrected orchard plots side-by-side using `patchwork` to directly evaluate the shifts in pooled estimates and prediction intervals when publication selection is accounted for.

```{r}
#| label: hedges_comparison_plot
#| message: false
#| warning: false
#| fig.width: 14
#| fig.height: 8

if (exists("p_hedges") && exists("p_hedges_corr")) {
  comparison_hedges <- (p_hedges + p_hedges_corr) + 
    patchwork::plot_annotation(
      title = "Comparison of Uncorrected vs. Bias-Corrected Synthesis (Hedges' g)",
      subtitle = "Bias-corrected estimates are conditioned on study variance = 0 (infinite precision)",
      theme = ggplot2::theme(
        plot.title = ggplot2::element_text(size = 16, face = "bold", hjust = 0.5),
        plot.subtitle = ggplot2::element_text(size = 12, italic = TRUE, hjust = 0.5)
      )
    )
  print(comparison_hedges)
}
```

#### Detailed Model Statistics

Here are the text outputs and fit statistics for the three database splits. Each split displays the uncorrected and small-study bias-corrected (Egger's variance-adjusted) model summaries:

::: {.panel-tabset}

##### New Only

::: {.panel-tabset}

###### Uncorrected
```{r}
#| label: hedges_new_stats
if (!is.null(res_hedges_new)) {
  print(summary(res_hedges_new$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_hedges_new$model))
}
```

###### Bias-Corrected
```{r}
#| label: hedges_new_corr_stats
if (!is.null(res_hedges_new_corr)) {
  print(summary(res_hedges_new_corr$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_hedges_new_corr$model))
}
```

:::

##### Original Only

::: {.panel-tabset}

###### Uncorrected
```{r}
#| label: hedges_orig_stats
if (!is.null(res_hedges_orig)) {
  print(summary(res_hedges_orig$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_hedges_orig$model))
}
```

###### Bias-Corrected
```{r}
#| label: hedges_orig_corr_stats
if (!is.null(res_hedges_orig_corr)) {
  print(summary(res_hedges_orig_corr$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_hedges_orig_corr$model))
}
```

:::

##### Combined

::: {.panel-tabset}

###### Uncorrected
```{r}
#| label: hedges_comb_stats
if (!is.null(res_hedges_comb)) {
  print(summary(res_hedges_comb$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_hedges_comb$model))
}
```

###### Bias-Corrected
```{r}
#| label: hedges_comb_corr_stats
if (!is.null(res_hedges_comb_corr)) {
  print(summary(res_hedges_comb_corr$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_hedges_comb_corr$model))
}
```

:::

:::

:::

### Odds Ratio (Binary outcomes)

::: {.panel-tabset}

#### Uncorrected Overall Effect

This plot displays the standard multi-level meta-analytic summaries across the three splits (New Only, Original Only, and Combined) *without* adjusting for publication bias. Points represent individual log odds ratios colored by their data source.

```{r}
#| label: or_unified_plot
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 7

if (!is.null(res_or_new) && !is.null(res_or_orig) && !is.null(res_or_comb)) {
  # 1. Extract S3 result objects using mod_results
  o_new <- orchaRd::mod_results(res_or_new$model, group = "identifierStudyId")
  o_orig <- orchaRd::mod_results(res_or_orig$model, group = "identifierStudyId")
  o_comb <- orchaRd::mod_results(res_or_comb$model, group = "identifierStudyId")

  # 2. Clone S3 result objects to lists for manual editing
  class(o_new) <- "list"
  o_new$mod_table$name <- "New Extraction (Santos)"
  o_new$data$moderator <- "New Extraction (Santos)"
  class(o_new) <- c("orchard", "data.frame")

  class(o_orig) <- "list"
  o_orig$mod_table$name <- "Original (Jones & DuVal 2019)"
  o_orig$data$moderator <- "Original (Jones & DuVal 2019)"
  class(o_orig) <- c("orchard", "data.frame")

  class(o_comb) <- "list"
  o_comb$mod_table$name <- "Combined"
  o_comb$data$moderator <- "Combined"
  class(o_comb) <- c("orchard", "data.frame")

  # 3. Merge model results
  sub_or <- orchaRd::submerge(o_new, o_orig, o_comb)

  # 4. Add point-level real source mapping
  class(sub_or) <- "list"
  sub_or$data$real_source <- c(
    rep("New Extraction (Santos)", nrow(o_new$data)),
    rep("Original (Jones & DuVal)", nrow(o_orig$data)),
    rep("New Extraction (Santos)", nrow(o_new$data)),
    rep("Original (Jones & DuVal)", nrow(o_orig$data))
  )
  class(sub_or) <- c("orchard", "data.frame")

  # 4. Generate Plot
  p_or <- orchaRd::orchard_plot(sub_or, xlab = "Log Odds Ratio (lnOR)", flip = TRUE)

  # 5. Customize color mapping for points
  p_or$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
  p_or$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

  # 6. Apply themes & custom palette (colorblind friendly orange/blue)
  p_or <- p_or +
    ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal)" = "#0072B2"), name = "Data Source") +
    ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal)" = "#0072B2"), name = "Data Source") +
    ggplot2::ggtitle("Overall Synthesis — Log Odds Ratio Estimates (Uncorrected)") +
    ggplot2::theme(legend.position = "right")

  print(p_or)
}
```

#### Bias-Corrected Overall Effect

This plot displays the small-study publication bias corrected meta-analytic summaries (evaluated at $\text{V} = 0$, representing the theoretical estimate for a study of infinite sample size/precision) using the two-step Egger-type meta-regression framework. Points represent the same raw effect sizes.

```{r}
#| label: or_corrected_plot
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 7

if (!is.null(res_or_new_corr) && !is.null(res_or_orig_corr) && !is.null(res_or_comb_corr)) {
  # 1. Extract S3 result objects using mod_results at vi = 0
  o_new_corr <- orchaRd::mod_results(res_or_new_corr$model, group = "identifierStudyId", at = list(vi = 0), mod = "1")
  o_orig_corr <- orchaRd::mod_results(res_or_orig_corr$model, group = "identifierStudyId", at = list(vi = 0), mod = "1")
  o_comb_corr <- orchaRd::mod_results(res_or_comb_corr$model, group = "identifierStudyId", at = list(vi = 0), mod = "1")

  # 2. Clone S3 result objects to lists for manual editing
  class(o_new_corr) <- "list"
  o_new_corr$mod_table$name <- "New Extraction (Santos)"
  o_new_corr$data$moderator <- "New Extraction (Santos)"
  class(o_new_corr) <- c("orchard", "data.frame")

  class(o_orig_corr) <- "list"
  o_orig_corr$mod_table$name <- "Original (Jones & DuVal 2019)"
  o_orig_corr$data$moderator <- "Original (Jones & DuVal 2019)"
  class(o_orig_corr) <- c("orchard", "data.frame")

  class(o_comb_corr) <- "list"
  o_comb_corr$mod_table$name <- "Combined"
  o_comb_corr$data$moderator <- "Combined"
  class(o_comb_corr) <- c("orchard", "data.frame")

  # 3. Merge model results
  sub_or_corr <- orchaRd::submerge(o_new_corr, o_orig_corr, o_comb_corr)

  # 4. Add point-level real source mapping
  class(sub_or_corr) <- "list"
  sub_or_corr$data$real_source <- c(
    rep("New Extraction (Santos)", nrow(o_new_corr$data)),
    rep("Original (Jones & DuVal)", nrow(o_orig_corr$data)),
    rep("New Extraction (Santos)", nrow(o_new_corr$data)),
    rep("Original (Jones & DuVal)", nrow(o_orig_corr$data))
  )
  class(sub_or_corr) <- c("orchard", "data.frame")

  # 4. Generate Plot
  p_or_corr <- orchaRd::orchard_plot(sub_or_corr, xlab = "Log Odds Ratio (lnOR) (Bias Corrected)", flip = TRUE)

  # 5. Customize color mapping for points
  p_or_corr$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
  p_or_corr$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

  # 6. Apply themes & custom palette
  p_or_corr <- p_or_corr +
    ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal)" = "#0072B2"), name = "Data Source") +
    ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal)" = "#0072B2"), name = "Data Source") +
    ggplot2::ggtitle("Overall Synthesis — Log Odds Ratio Estimates (Bias-Corrected)") +
    ggplot2::theme(legend.position = "right")

  print(p_or_corr)
}
```

#### Side-by-Side Comparison

We combine the uncorrected and small-study bias corrected orchard plots side-by-side using `patchwork` to directly evaluate the shifts in pooled estimates and prediction intervals when publication selection is accounted for.

```{r}
#| label: or_comparison_plot
#| message: false
#| warning: false
#| fig.width: 14
#| fig.height: 8

if (exists("p_or") && exists("p_or_corr")) {
  comparison_or <- (p_or + p_or_corr) + 
    patchwork::plot_annotation(
      title = "Comparison of Uncorrected vs. Bias-Corrected Synthesis (Odds Ratio)",
      subtitle = "Bias-corrected estimates are conditioned on study variance = 0 (infinite precision)",
      theme = ggplot2::theme(
        plot.title = ggplot2::element_text(size = 16, face = "bold", hjust = 0.5),
        plot.subtitle = ggplot2::element_text(size = 12, italic = TRUE, hjust = 0.5)
      )
    )
  print(comparison_or)
}
```

#### Detailed Model Statistics

Here are the text outputs and fit statistics for the three database splits. Each split displays the uncorrected and small-study bias-corrected (Egger's variance-adjusted) model summaries:

::: {.panel-tabset}

##### New Only

::: {.panel-tabset}

###### Uncorrected
```{r}
#| label: or_new_stats
if (!is.null(res_or_new)) {
  print(summary(res_or_new$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_or_new$model))
}
```

###### Bias-Corrected
```{r}
#| label: or_new_corr_stats
if (!is.null(res_or_new_corr)) {
  print(summary(res_or_new_corr$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_or_new_corr$model))
}
```

:::

##### Original Only

::: {.panel-tabset}

###### Uncorrected
```{r}
#| label: or_orig_stats
if (!is.null(res_or_orig)) {
  print(summary(res_or_orig$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_or_orig$model))
}
```

###### Bias-Corrected
```{r}
#| label: or_orig_corr_stats
if (!is.null(res_or_orig_corr)) {
  print(summary(res_or_orig_corr$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_or_orig_corr$model))
}
```

:::

##### Combined

::: {.panel-tabset}

###### Uncorrected
```{r}
#| label: or_comb_stats
if (!is.null(res_or_comb)) {
  print(summary(res_or_comb$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_or_comb$model))
}
```

###### Bias-Corrected
```{r}
#| label: or_comb_corr_stats
if (!is.null(res_or_comb_corr)) {
  print(summary(res_or_comb_corr$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_or_comb_corr$model))
}
```

:::

:::

:::

:::

---

## Sensitivity Analysis: Impact of the VCV Structure

To evaluate the robustness of our meta-analytic results, we conduct a sensitivity analysis comparing the main multilevel models (which incorporate a Variance-Covariance [VCV] matrix with an assumed within-study correlation $\rho = 0.5$) against simplified multilevel models that ignore this correlation (i.e., treating effect size variances as independent diagonal elements, $V = v_i$).

Ignoring within-study correlation can lead to improper weighting of studies with multiple effect sizes and can bias the standard errors and confidence intervals of the pooled estimates.

```{r}
#| label: vcv_sensitivity_analysis
#| message: false
#| warning: false

# 1. Define comparison function
compare_vcv_models <- function(dat, name, yi = "yi", vi = "vi", cluster_var = "identifierStudyId", obs_var = "identifierEffectSizeID", rho = 0.5) {
  dat2 <- dat %>%
    filter(!is.na(.data[[yi]]), !is.na(.data[[vi]]), .data[[vi]] > 0)
  
  if (nrow(dat2) == 0) return(NULL)
  
  # A. With VCV (rho = 0.5)
  VCV <- metafor::vcalc(
    vi = dat2[[vi]],
    cluster = dat2[[cluster_var]],
    obs = dat2[[obs_var]],
    rho = rho,
    data = dat2
  )
  mod_w_vcv <- metafor::rma.mv(
    yi = dat2[[yi]],
    V  = VCV,
    random = list(
      ~1 | identifierStudyId,
      ~1 | identifierEffectSizeID,
      ~1 | taxonomySpecies
    ),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = dat2
  )
  
  # B. Without VCV (diagonal V = vi)
  mod_no_vcv <- metafor::rma.mv(
    yi = dat2[[yi]],
    V  = dat2[[vi]],
    random = list(
      ~1 | identifierStudyId,
      ~1 | identifierEffectSizeID,
      ~1 | taxonomySpecies
    ),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = dat2
  )
  
  # Combine results
  data.frame(
    Analysis = name,
    Model_Type = c("With VCV (rho = 0.5)", "Without VCV (Independent)"),
    k = c(mod_w_vcv$k, mod_no_vcv$k),
    Estimate = c(mod_w_vcv$b[1], mod_no_vcv$b[1]),
    SE = c(mod_w_vcv$se[1], mod_no_vcv$se[1]),
    CI_Lower = c(mod_w_vcv$ci.lb[1], mod_no_vcv$ci.lb[1]),
    CI_Upper = c(mod_w_vcv$ci.ub[1], mod_no_vcv$ci.ub[1]),
    p_val = c(mod_w_vcv$pval[1], mod_no_vcv$pval[1])
  )
}

# 2. Run comparisons across all subsets
comp_hedges_new  <- compare_vcv_models(db_hedges %>% filter(Dataset_Source == "New Extraction (Santos)"), "Hedges' g (New Only)")
comp_hedges_orig <- compare_vcv_models(db_hedges %>% filter(Dataset_Source == "Original (Davies 2020)"), "Hedges' g (Original Only)")
comp_hedges_comb <- compare_vcv_models(db_hedges, "Hedges' g (Combined)")

comp_or_new  <- compare_vcv_models(db_or %>% filter(Dataset_Source == "New Extraction (Santos)"), "Log OR (New Only)")
comp_or_orig <- compare_vcv_models(db_or %>% filter(Dataset_Source == "Original (Jones & DuVal 2019)"), "Log OR (Original Only)")
comp_or_comb <- compare_vcv_models(db_or, "Log OR (Combined)")

# 3. Bind all comparisons together
vcv_comparison_results <- bind_rows(
  comp_hedges_new, comp_hedges_orig, comp_hedges_comb,
  comp_or_new, comp_or_orig, comp_or_comb
)
```

### Sensitivity Comparison Table

The interactive table below compares the pool estimate, standard error (SE), 95% Confidence Intervals, and p-values for both models across all subsets.

```{r}
#| label: vcv_comparison_table
#| echo: false

DT::datatable(
  vcv_comparison_results,
  options = list(
    pageLength = 12,
    dom = 't',
    columnDefs = list(list(className = 'dt-center', targets = "_all"))
  ),
  rownames = FALSE
) %>%
  DT::formatRound(columns = c("Estimate", "SE", "CI_Lower", "CI_Upper"), digits = 4) %>%
  DT::formatSignif(columns = c("p_val"), digits = 4)
```

### Key Methodological Findings

1. **Marginal Shifts in Pooled Estimates:** 
   Treating effect sizes as completely independent (Without VCV) causes very minor changes to the overall pooled estimates. For the combined dataset, the Hedges' $g$ estimate shifts from $0.1526$ to $0.1473$, while the Log Odds Ratio (lnOR) estimate shifts from $0.6147$ to $0.6530$. This indicates that the main directional conclusions of the meta-analyses are robust to VCV specification.

2. **Standard Error Shifts:** 
   In this multilevel model specification (which partitions variance using nested random intercepts), omitting the VCV matrix and assuming independence leads to a *slight inflation* of the standard error and wider confidence intervals for the combined analyses:
   * **Hedges' $g$ (Combined):** SE increases from $0.0780$ to $0.0816$ (CI width increases from $0.3074$ to $0.3217$).
   * **Log Odds Ratio (Combined):** SE increases from $0.1388$ to $0.1494$ (CI width increases from $0.5480$ to $0.5900$).
   
   This demonstrates that accounting for known within-study covariance via `vcalc` yields more precise, correctly-weighted estimates.

3. **Impact on Statistical Significance:** 
   The marginal significance test for continuous outcomes (Hedges' $g$ Combined) is sensitive to this specification. With the VCV matrix, the effect size is marginally non-significant ($p = 0.0517$). When omitting the VCV matrix, the $p$-value increases further to $p = 0.0724$. Conversely, binary outcomes (Log OR Combined) remain highly statistically significant ($p < 0.0001$) under both models.

Therefore, we conclude that while the overall direction and magnitude of the estimates are stable, **incorporating the VCV matrix is critical** for ensuring appropriate weighting, controlling for within-study correlation, and obtaining accurate standard errors and statistical tests.

---

## Session Info

```{r}
#| label: session_info
sessionInfo()
```
