---
title: "Overall Effect Size Calculation & Synthesis"
subtitle: "Meta-Analytic Update of Mate-Choice Copying (MCC)"
author: "Eduardo Santos & Collaborators"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    theme: cosmo
    self-contained: true
---

```{r}
#| label: setup
#| include: false
# Ensure pacman is loaded to manage dependencies
if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
pacman::p_load(
               readxl,
               tidyverse,
               metafor,
               orchaRd,
               patchwork,
               DT
               )
```

## Introduction & Project Update

This protocol implements the overall effect size synthesis for the updated Mate-Choice Copying (MCC) meta-analysis. 

Specifically, we integrate:
1. **The New Extractions:** Our newly extracted and verified dataset containing recent trials.
2. **Original Continuous Dataset (Davies et al. 2020):** Standardised mean differences (Hedges' $d$) analyzing continuous metrics.
3. **Original Binary Dataset (Jones & DuVal 2019):** Log Odds Ratio ($\ln\text{OR}$) analyzing binary choices.

The R script aligns and synthesizes these three sources into unified datasets, allowing comparison of meta-analytic summaries across three key subsets:
* **New Extractions Only:** Eduardo's new dataset.
* **Original Only:** Legacy data from the original publications.
* **Combined Dataset:** A complete synthesis of old and new trials to estimate the updated global effect sizes.

---

## Data Loading and Alignment

We read the three datasets, perform data type conversions, extract species epithets for taxonomic control, and align columns into unified Hedges' $g$ and Odds Ratio datasets.

```{r}
#| label: load_and_align_datasets
#| message: false
#| warning: false

# ---------------------------------------------------------------------------
# 1. Load Datasets
# ---------------------------------------------------------------------------

# Paths to legacy and new datasets
path_new <- here::here("data", "1_effect_size_calculation_pipeline", "Data Extraction Mate Choice Meta Analysis.xlsx")
path_davies <- here::here("data", "2_original_meta_analyses_datasets", "Davies_et_al_2020_Final_data.xlsx")
path_jones <- here::here("data", "2_original_meta_analyses_datasets", "Jones_DuVal_2019_data.CSV")

# A. Read new extraction dataset (EduardoExtraction sheet)
db_new_raw <- readxl::read_excel(path_new, sheet = "EduardoExtraction")

# B. Read Davies et al. 2020 (Original continuous dataset)
db_davies_raw <- readxl::read_excel(path_davies, sheet = "Data")

# C. Read Jones & DuVal 2019 (Original binary dataset)
db_jones_raw <- readr::read_csv(path_jones)

# ---------------------------------------------------------------------------
# 2. Process and Filter New Extraction
# ---------------------------------------------------------------------------
db_new <- db_new_raw %>%
  filter(is.na(RemoveFromDataset) | RemoveFromDataset != "Yes") %>%
  mutate(
    effectSizeCalculatedHedgesD = as.numeric(effectSizeCalculatedHedgesD),
    effectSizeCalculatedHedgesDVariance = as.numeric(effectSizeCalculatedHedgesDVariance),
    effectSizeCalculatedOddsRatio = as.numeric(effectSizeCalculatedOddsRatio),
    effectSizeCalculatedOddsRatioVariance = as.numeric(effectSizeCalculatedOddsRatioVariance),
    logOddsRatio = log(effectSizeCalculatedOddsRatio),
    logOddsRatioVariance = effectSizeCalculatedOddsRatioVariance
  )

# ---------------------------------------------------------------------------
# 3. Align Datasets for Hedges' g (Continuous SMD Analysis)
# ---------------------------------------------------------------------------

# A. Hedges' g from New Extraction
hedges_new <- db_new %>%
  filter(!is.na(effectSizeCalculatedHedgesD), !is.na(effectSizeCalculatedHedgesDVariance)) %>%
  select(
    yi = effectSizeCalculatedHedgesD,
    vi = effectSizeCalculatedHedgesDVariance,
    identifierStudyId,
    identifierEffectSizeID,
    taxonomySpecies
  ) %>%
  mutate(
    identifierStudyId = as.character(identifierStudyId),
    identifierEffectSizeID = as.character(identifierEffectSizeID),
    # Extract species epithet (last word) and make lowercase for cross-dataset mapping
    taxonomySpecies = str_to_lower(word(as.character(taxonomySpecies), -1)),
    Dataset_Source = "New Extraction (Santos)"
  )

# B. Hedges' g from Davies et al. 2020 (Original)
hedges_davies <- db_davies_raw %>%
  mutate(
    Hedges_d = as.numeric(Hedges_d),
    Variance = as.numeric(Variance)
  ) %>%
  filter(!is.na(Hedges_d), !is.na(Variance), Variance > 0) %>%
  select(
    yi = Hedges_d,
    vi = Variance,
    identifierStudyId = Author,
    identifierEffectSizeID = Effect_size_no,
    taxonomySpecies = Species_latin
  ) %>%
  mutate(
    identifierStudyId = as.character(identifierStudyId),
    identifierEffectSizeID = as.character(identifierEffectSizeID),
    # Extract species epithet (last word) and make lowercase
    taxonomySpecies = str_to_lower(word(as.character(taxonomySpecies), -1)),
    Dataset_Source = "Original (Davies 2020)"
  )

# C. Unified Hedges' g Dataset
db_hedges <- bind_rows(hedges_new, hedges_davies) %>%
  mutate(across(c(identifierStudyId, taxonomySpecies, Dataset_Source), as.factor))

# ---------------------------------------------------------------------------
# 4. Align Datasets for Odds Ratio (Binary Analysis)
# ---------------------------------------------------------------------------

# A. Log Odds Ratio from New Extraction
or_new <- db_new %>%
  filter(!is.na(logOddsRatio), !is.na(logOddsRatioVariance)) %>%
  select(
    yi = logOddsRatio,
    vi = logOddsRatioVariance,
    identifierStudyId,
    identifierEffectSizeID,
    taxonomySpecies
  ) %>%
  mutate(
    identifierStudyId = as.character(identifierStudyId),
    identifierEffectSizeID = as.character(identifierEffectSizeID),
    taxonomySpecies = str_to_lower(word(as.character(taxonomySpecies), -1)),
    Dataset_Source = "New Extraction (Santos)"
  )

# B. Log Odds Ratio from Jones & DuVal 2019 (Original)
or_jones <- db_jones_raw %>%
  filter(!is.na(lnOR), !is.na(VlnOR), VlnOR > 0) %>%
  select(
    yi = lnOR,
    vi = VlnOR,
    identifierStudyId = study,
    identifierEffectSizeID = study_Exp,
    taxonomySpecies = Species
  ) %>%
  mutate(
    identifierStudyId = as.character(identifierStudyId),
    identifierEffectSizeID = as.character(identifierEffectSizeID),
    taxonomySpecies = str_to_lower(as.character(taxonomySpecies)),
    Dataset_Source = "Original (Jones & DuVal 2019)"
  )

# C. Unified Log Odds Ratio Dataset
db_or <- bind_rows(or_new, or_jones) %>%
  mutate(across(c(identifierStudyId, taxonomySpecies, Dataset_Source), as.factor))
```

---

## Model Fitting Function

This helper function handles the variance–covariance (VCV) matrix construction and fits a multilevel random-effects model using `rma.mv()` in `metafor`. 

* **Variance–Covariance:** We construct a VCV matrix via `vcalc()` to handle dependencies among multiple effect sizes within the same study (assuming $r = 0.5$ for nested measurements).
* **Random intercepts** are modeled for:
  * Study (`identifierStudyId`)
  * Individual trial/effect size (`identifierEffectSizeID`)
  * Species (`taxonomySpecies`) to control for taxonomic/species-level variation

```{r}
#| label: model_fitting_function

fit_overall_mv <- function(dat, yi = "yi", vi = "vi", cluster_var = "identifierStudyId", obs_var = "identifierEffectSizeID", rho = 0.5) {
  
  # Remove rows with missing values
  dat2 <- dat %>%
    filter(!is.na(.data[[yi]]), !is.na(.data[[vi]]), .data[[vi]] > 0)
  
  if (nrow(dat2) == 0) {
    message("No complete cases found.")
    return(NULL)
  }
  
  # 1. Build VCV Matrix
  VCV <- metafor::vcalc(
    vi = dat2[[vi]],
    cluster = dat2[[cluster_var]],
    obs = dat2[[obs_var]],
    rho = rho,
    data = dat2
  )
  
  # 2. Fit Multi-Level Random-Effects Model
  mod <- metafor::rma.mv(
    yi = dat2[[yi]],
    V  = VCV,
    random = list(
      ~1 | identifierStudyId,
      ~1 | identifierEffectSizeID,
      ~1 | taxonomySpecies
    ),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = dat2
  )
  
  list(data = dat2, VCV = VCV, model = mod)
}
```

---

## Multi-Level Synthesis Results

We fit the multilevel meta-analysis across all three database splits (New Only, Original Only, and Combined) for both effect size classes.

```{r}
#| label: fit_multilevel_models
#| message: false
#| warning: false

# ---------------------------------------------------------------------------
# HEDGES' G MODELS
# ---------------------------------------------------------------------------
res_hedges_new   <- fit_overall_mv(db_hedges %>% filter(Dataset_Source == "New Extraction (Santos)"))
res_hedges_orig  <- fit_overall_mv(db_hedges %>% filter(Dataset_Source == "Original (Davies 2020)"))
res_hedges_comb  <- fit_overall_mv(db_hedges)

# ---------------------------------------------------------------------------
# ODDS RATIO MODELS
# ---------------------------------------------------------------------------
res_or_new   <- fit_overall_mv(db_or %>% filter(Dataset_Source == "New Extraction (Santos)"))
res_or_orig  <- fit_overall_mv(db_or %>% filter(Dataset_Source == "Original (Jones & DuVal 2019)"))
res_or_comb  <- fit_overall_mv(db_or)
```

---

## Results and Visualizations

::: {.panel-tabset}

### Hedges' $g$ (Continuous outcomes)

::: {.panel-tabset}

#### Unified Synthesis Plot

This plot combines all three estimates (New Extraction, Original legacy studies, and the Combined pool) side-by-side in a single panel. Points are colored individually by their true data source.

```{r}
#| label: hedges_unified_plot
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 7

if (!is.null(res_hedges_new) && !is.null(res_hedges_orig) && !is.null(res_hedges_comb)) {
  # 1. Extract S3 result objects using mod_results
  h_new <- orchaRd::mod_results(res_hedges_new$model, group = "identifierStudyId")
  h_orig <- orchaRd::mod_results(res_hedges_orig$model, group = "identifierStudyId")
  h_comb <- orchaRd::mod_results(res_hedges_comb$model, group = "identifierStudyId")

  # 2. Clone S3 result objects to lists for manual editing
  class(h_new) <- "list"
  h_new$mod_table$name <- "New Extraction (Santos)"
  h_new$data$moderator <- "New Extraction (Santos)"
  class(h_new) <- c("orchard", "data.frame")

  class(h_orig) <- "list"
  h_orig$mod_table$name <- "Original (Davies 2020)"
  h_orig$data$moderator <- "Original (Davies 2020)"
  class(h_orig) <- c("orchard", "data.frame")

  class(h_comb) <- "list"
  h_comb$mod_table$name <- "Combined"
  h_comb$data$moderator <- "Combined"
  class(h_comb) <- c("orchard", "data.frame")

  # 3. Merge model results
  sub_hedges <- orchaRd::submerge(h_new, h_orig, h_comb)

  # 4. Add point-level real source mapping
  class(sub_hedges) <- "list"
  sub_hedges$data$real_source <- c(
    rep("New Extraction (Santos)", nrow(h_new$data)),
    rep("Original (Davies 2020)", nrow(h_orig$data)),
    rep("New Extraction (Santos)", nrow(h_new$data)),
    rep("Original (Davies 2020)", nrow(h_orig$data))
  )
  class(sub_hedges) <- c("orchard", "data.frame")

  # 4. Generate Plot
  p_hedges <- orchaRd::orchard_plot(sub_hedges, xlab = "Hedges' g", flip = TRUE)

  # 5. Customize color mapping for points
  p_hedges$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
  p_hedges$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

  # 6. Apply themes & custom palette (colorblind friendly orange/blue)
  p_hedges <- p_hedges +
    ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
    ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
    ggplot2::ggtitle("Overall Synthesis — Hedges' g Estimates") +
    ggplot2::theme(legend.position = "right")

  print(p_hedges)
}
```

#### Detailed Model Statistics

Here are the text outputs and fit statistics for the three individual models:

::: {.panel-tabset}

##### New Only

```{r}
#| label: hedges_new_stats
if (!is.null(res_hedges_new)) {
  print(summary(res_hedges_new$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_hedges_new$model))
}
```

##### Original Only

```{r}
#| label: hedges_orig_stats
if (!is.null(res_hedges_orig)) {
  print(summary(res_hedges_orig$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_hedges_orig$model))
}
```

##### Combined

```{r}
#| label: hedges_comb_stats
if (!is.null(res_hedges_comb)) {
  print(summary(res_hedges_comb$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_hedges_comb$model))
}
```

:::

:::

### Odds Ratio (Binary outcomes)

::: {.panel-tabset}

#### Unified Synthesis Plot

This plot combines all three estimates (New Extraction, Original legacy studies, and the Combined pool) side-by-side in a single panel. Points are colored individually by their true data source.

```{r}
#| label: or_unified_plot
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 7

if (!is.null(res_or_new) && !is.null(res_or_orig) && !is.null(res_or_comb)) {
  # 1. Extract S3 result objects using mod_results
  o_new <- orchaRd::mod_results(res_or_new$model, group = "identifierStudyId")
  o_orig <- orchaRd::mod_results(res_or_orig$model, group = "identifierStudyId")
  o_comb <- orchaRd::mod_results(res_or_comb$model, group = "identifierStudyId")

  # 2. Clone S3 result objects to lists for manual editing
  class(o_new) <- "list"
  o_new$mod_table$name <- "New Extraction (Santos)"
  o_new$data$moderator <- "New Extraction (Santos)"
  class(o_new) <- c("orchard", "data.frame")

  class(o_orig) <- "list"
  o_orig$mod_table$name <- "Original (Jones & DuVal 2019)"
  o_orig$data$moderator <- "Original (Jones & DuVal 2019)"
  class(o_orig) <- c("orchard", "data.frame")

  class(o_comb) <- "list"
  o_comb$mod_table$name <- "Combined"
  o_comb$data$moderator <- "Combined"
  class(o_comb) <- c("orchard", "data.frame")

  # 3. Merge model results
  sub_or <- orchaRd::submerge(o_new, o_orig, o_comb)

  # 4. Add point-level real source mapping
  class(sub_or) <- "list"
  sub_or$data$real_source <- c(
    rep("New Extraction (Santos)", nrow(o_new$data)),
    rep("Original (Jones & DuVal)", nrow(o_orig$data)),
    rep("New Extraction (Santos)", nrow(o_new$data)),
    rep("Original (Jones & DuVal)", nrow(o_orig$data))
  )
  class(sub_or) <- c("orchard", "data.frame")

  # 4. Generate Plot
  p_or <- orchaRd::orchard_plot(sub_or, xlab = "Log Odds Ratio (lnOR)", flip = TRUE)

  # 5. Customize color mapping for points
  p_or$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
  p_or$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

  # 6. Apply themes & custom palette (colorblind friendly orange/blue)
  p_or <- p_or +
    ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal)" = "#0072B2"), name = "Data Source") +
    ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal)" = "#0072B2"), name = "Data Source") +
    ggplot2::ggtitle("Overall Synthesis — Log Odds Ratio Estimates") +
    ggplot2::theme(legend.position = "right")

  print(p_or)
}
```

#### Detailed Model Statistics

Here are the text outputs and fit statistics for the three individual models:

::: {.panel-tabset}

##### New Only

```{r}
#| label: or_new_stats
if (!is.null(res_or_new)) {
  print(summary(res_or_new$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_or_new$model))
}
```

##### Original Only

```{r}
#| label: or_orig_stats
if (!is.null(res_or_orig)) {
  print(summary(res_or_orig$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_or_orig$model))
}
```

##### Combined

```{r}
#| label: or_comb_stats
if (!is.null(res_or_comb)) {
  print(summary(res_or_comb$model))
  cat("\nHeterogeneity (I^2):\n")
  print(orchaRd::i2_ml(res_or_comb$model))
}
```

:::

:::

:::

---

## Session Info

```{r}
#| label: session_info
sessionInfo()
```
