---
title: "Publication Bias Diagnostics"
subtitle: "Small-Study Effects and Time-Lag Bias in Mate-Choice Copying (MCC) Meta-Analyses"
author: "Eduardo Santos & Collaborators"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    theme: cosmo
    self-contained: true
---

```{r}
#| label: setup
#| include: false
# Ensure pacman is loaded to manage dependencies
if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
pacman::p_load(
               readxl,
               tidyverse,
               metafor,
               orchaRd,
               patchwork,
               DT,
               scales
               )
```

## Introduction & Methodology

Publication bias is a critical threat to the validity of meta-analytic syntheses. If studies with statistically non-significant or small effects are systematically omitted from the literature, the pooled meta-analytic summary will be biased upward.

To thoroughly evaluate the presence of publication bias and report-level selection in our Mate-Choice Copying (MCC) dataset, we employ three complementary diagnostic approaches, keeping the exact multilevel modeling logic (e.g., nested random intercepts and taxonomy control) used in our main overall effect calculations:

1. **Visual Inspection (Funnel Plots):** We plot effect size estimates against their precision (inverse of the standard error, $1/\text{SE}$) to check for visual asymmetry around the overall mean estimate.
2. **Egger's Meta-Regression (Small-Study Effects):** We test whether study precision significantly moderates effect size. 
   * For **Hedges' $g$ (Continuous outcomes)**, we use an N-based precision proxy, $\sqrt{1/N_{\text{control}} + 1/N_{\text{treatment}}}$ (the square root of the inverse of the effective sample size), to prevent mathematical coupling between effect size and standard error.
   * For **Odds Ratio (Binary outcomes)**, because raw group-level sample sizes are not fully available in the original binary dataset, we use the standard error ($\text{SE} = \sqrt{v_i}$) as the moderator, controlling for taxonomic and study-level clustering.
3. **Decline Effect / Time-Lag Bias:** We test whether effect sizes shrink over time by adding publication year (mean-centered) to our models while simultaneously controlling for study precision.

---

## Data Loading and Alignment

We read, clean, and align the new extractions and original datasets, extracting sample sizes and publication years to construct our precision and time-lag moderators.

```{r}
#| label: load_and_align_datasets
#| message: false
#| warning: false

# Paths to datasets
path_new <- here::here("data", "1_effect_size_calculation_pipeline", "Data Extraction Mate Choice Meta Analysis.xlsx")
path_davies <- here::here("data", "2_original_meta_analyses_datasets", "Davies_et_al_2020_Final_data.xlsx")
path_jones <- here::here("data", "2_original_meta_analyses_datasets", "Jones_DuVal_2019_data.CSV")

# A. Read raw spreadsheets
db_new_raw <- readxl::read_excel(path_new, sheet = "EduardoExtraction")
db_davies_raw <- readxl::read_excel(path_davies, sheet = "Data")
db_jones_raw <- readr::read_csv(path_jones)

# B. Clean new extractions
db_new <- db_new_raw %>%
  filter(is.na(RemoveFromDataset) | RemoveFromDataset != "Yes") %>%
  mutate(
    effectSizeCalculatedHedgesD = as.numeric(effectSizeCalculatedHedgesD),
    effectSizeCalculatedHedgesDVariance = as.numeric(effectSizeCalculatedHedgesDVariance),
    effectSizeCalculatedOddsRatio = as.numeric(effectSizeCalculatedOddsRatio),
    effectSizeCalculatedOddsRatioVariance = as.numeric(effectSizeCalculatedOddsRatioVariance),
    logOddsRatio = log(effectSizeCalculatedOddsRatio),
    logOddsRatioVariance = effectSizeCalculatedOddsRatioVariance,
    identifierStudyYear = as.numeric(identifierStudyYear),
    controlN = as.numeric(controlN),
    treatmentN = as.numeric(treatmentN),
    effectSizesampleSize = as.numeric(effectSizesampleSize)
  )

# ---------------------------------------------------------------------------
# 1. Hedges' g Dataset Alignment (Continuous)
# ---------------------------------------------------------------------------
hedges_new <- db_new %>%
  filter(!is.na(effectSizeCalculatedHedgesD), !is.na(effectSizeCalculatedHedgesDVariance)) %>%
  select(
    yi = effectSizeCalculatedHedgesD,
    vi = effectSizeCalculatedHedgesDVariance,
    identifierStudyId,
    identifierEffectSizeID,
    taxonomySpecies,
    Year = identifierStudyYear,
    controlN,
    treatmentN,
    totalN = effectSizesampleSize
  ) %>%
  mutate(
    identifierStudyId = as.character(identifierStudyId),
    identifierEffectSizeID = as.character(identifierEffectSizeID),
    taxonomySpecies = str_to_lower(word(as.character(taxonomySpecies), -1)),
    Dataset_Source = "New Extraction (Santos)"
  )

hedges_davies <- db_davies_raw %>%
  mutate(
    Hedges_d_directional = as.numeric(Hedges_d_directional),
    Variance = as.numeric(Variance),
    Year = as.numeric(Year),
    N_control = as.numeric(N_control),
    N_treatment = as.numeric(N_treatment),
    Total_n_analysis = as.numeric(Total_n_analysis)
  ) %>%
  filter(!is.na(Hedges_d_directional), !is.na(Variance), Variance > 0) %>%
  select(
    yi = Hedges_d_directional,
    vi = Variance,
    identifierStudyId = Author,
    identifierEffectSizeID = Effect_size_no,
    taxonomySpecies = Species_latin,
    Year,
    controlN = N_control,
    treatmentN = N_treatment,
    totalN = Total_n_analysis
  ) %>%
  mutate(
    identifierStudyId = as.character(identifierStudyId),
    identifierEffectSizeID = as.character(identifierEffectSizeID),
    taxonomySpecies = str_to_lower(word(as.character(taxonomySpecies), -1)),
    Dataset_Source = "Original (Davies 2020)"
  )

db_hedges <- bind_rows(hedges_new, hedges_davies) %>%
  mutate(across(c(identifierStudyId, taxonomySpecies, Dataset_Source), as.factor))

# ---------------------------------------------------------------------------
# 2. Log Odds Ratio Dataset Alignment (Binary)
# ---------------------------------------------------------------------------
or_new <- db_new %>%
  filter(!is.na(logOddsRatio), !is.na(logOddsRatioVariance)) %>%
  select(
    yi = logOddsRatio,
    vi = logOddsRatioVariance,
    identifierStudyId,
    identifierEffectSizeID,
    taxonomySpecies,
    Year = identifierStudyYear
  ) %>%
  mutate(
    identifierStudyId = as.character(identifierStudyId),
    identifierEffectSizeID = as.character(identifierEffectSizeID),
    taxonomySpecies = str_to_lower(word(as.character(taxonomySpecies), -1)),
    Dataset_Source = "New Extraction (Santos)"
  )

or_jones <- db_jones_raw %>%
  filter(!is.na(lnOR), !is.na(VlnOR), VlnOR > 0) %>%
  select(
    yi = lnOR,
    vi = VlnOR,
    identifierStudyId = study,
    identifierEffectSizeID = study_Exp,
    taxonomySpecies = Species,
    Year = year
  ) %>%
  mutate(
    identifierStudyId = as.character(identifierStudyId),
    identifierEffectSizeID = as.character(identifierEffectSizeID),
    taxonomySpecies = str_to_lower(as.character(taxonomySpecies)),
    Dataset_Source = "Original (Jones & DuVal 2019)"
  )

db_or <- bind_rows(or_new, or_jones) %>%
  mutate(across(c(identifierStudyId, taxonomySpecies, Dataset_Source), as.factor))
```

---

## Diagnostics by Effect Size Class

::: {.panel-tabset}

### Hedges' $g$ (Continuous outcomes)

We fit multilevel models on the continuous combined dataset ($k = 223$) to run our diagnostics.

#### 1. Baseline Model & Funnel Plot

We run the intercept-only model to establish the global mean effect size and plot precision ($1/\text{SE}$) against the standardized residuals.

```{r}
#| label: hedges_baseline_and_funnel
#| message: false
#| warning: false
#| fig.width: 9
#| fig.height: 7

# Intercept-only model
VCV_hedges <- metafor::vcalc(vi = vi, cluster = identifierStudyId, obs = identifierEffectSizeID, rho = 0.5, data = db_hedges)

ma_hedges <- metafor::rma.mv(
  yi = yi, 
  V = VCV_hedges, 
  random = list(~1 | identifierStudyId, ~1 | identifierEffectSizeID, ~1 | taxonomySpecies), 
  test = "t", 
  method = "REML", 
  sparse = TRUE, 
  data = db_hedges
)

summary(ma_hedges)

# Funnel plot
metafor::funnel(
  ma_hedges,
  yaxis = "seinv",
  las = 1,
  digits = c(2, 1),
  xlab = "Standardised Hedges' g Effect Size",
  ylab = "Precision (Inverse of SE)",
  pch = 21,
  bg = alpha("black", 0.2),
  col = "black",
  cex = 1.2,
  main = "Funnel Plot for Continuous Outcomes (Hedges' g)"
)
```

#### 2. Egger's Meta-Regression (Small-Study Effects)

To perform Egger's regression rigorously on Hedges' $g$, we calculate the square root of the inverse of the effective sample size (`sqrt_inv_e_n`) to act as our precision proxy. If this moderator is statistically significant ($p < 0.05$), it indicates a significant small-study effect (smaller sample sizes associated with larger effect magnitudes).

```{r}
#| label: hedges_eggers_regression
#| message: false
#| warning: false

# Calculate precision proxy
db_hedges_egger <- db_hedges %>%
  mutate(
    controlN_eff = ifelse(is.na(controlN), totalN / 2, controlN),
    treatmentN_eff = ifelse(is.na(treatmentN), totalN / 2, treatmentN),
    sqrt_inv_e_n = sqrt((1 / controlN_eff) + (1 / treatmentN_eff))
  ) %>%
  filter(!is.na(sqrt_inv_e_n))

# Re-estimate VCV on this subset
VCV_hedges_egger <- metafor::vcalc(vi = vi, cluster = identifierStudyId, obs = identifierEffectSizeID, rho = 0.5, data = db_hedges_egger)

# Egger regression model
egger_hedges <- metafor::rma.mv(
  yi = yi, 
  V = VCV_hedges_egger, 
  mods = ~ 1 + sqrt_inv_e_n, 
  random = list(~1 | identifierStudyId, ~1 | identifierEffectSizeID, ~1 | taxonomySpecies), 
  test = "t", 
  method = "REML", 
  sparse = TRUE, 
  data = db_hedges_egger
)

summary(egger_hedges)
```

We visualize the Egger regression using a bubble plot. Effect sizes are plotted as circles scaled by precision, with the predicted regression line showing the relationship between sample size and effect size.

```{r}
#| label: hedges_egger_bubble_plot
#| message: false
#| warning: false
#| fig.width: 9
#| fig.height: 6

p_egger_hedges <- orchaRd::bubble_plot(
  egger_hedges, 
  mod = "sqrt_inv_e_n", 
  group = "identifierStudyId", 
  xlab = "Square root of inverse of effective sample size (Precision proxy)", 
  ylab = "Effect size (Hedges' g)"
)

# Color points by real source
p_egger_hedges$layers[[1]]$data$real_source <- db_hedges_egger$Dataset_Source
p_egger_hedges$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
p_egger_hedges$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

p_egger_hedges <- p_egger_hedges + 
  ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
  ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
  ggplot2::ggtitle("Egger's Meta-Regression (Hedges' g)") +
  ggplot2::theme_minimal()

print(p_egger_hedges)
```

> [!NOTE]
> The slope for `sqrt_inv_e_n` is **significant** ($\beta = 1.4313$, $SE = 0.6180$, $t = 2.316$, $p = 0.0215$), indicating clear evidence of small-study effects. Studies with smaller sample sizes tend to report larger, more positive Hedges' $g$ values.

#### 3. Decline Effect (Time-Lag Bias)

We analyze the decline effect by evaluating whether the publication year significantly predicts the effect size, while simultaneously controlling for the study sample size precision proxy. We mean-center the year so the intercept represents the effect in an average publication year.

```{r}
#| label: hedges_time_lag
#| message: false
#| warning: false

# Mean-center publication year
db_hedges_decline <- db_hedges_egger %>%
  mutate(Year_c = Year - mean(Year, na.rm = TRUE))

# VCV for decline subset
VCV_hedges_decline <- metafor::vcalc(vi = vi, cluster = identifierStudyId, obs = identifierEffectSizeID, rho = 0.5, data = db_hedges_decline)

# Time-lag decline model
decline_hedges <- metafor::rma.mv(
  yi = yi, 
  V = VCV_hedges_decline, 
  mods = ~ 1 + Year_c + sqrt_inv_e_n, 
  random = list(~1 | identifierStudyId, ~1 | identifierEffectSizeID, ~1 | taxonomySpecies), 
  test = "t", 
  method = "REML", 
  sparse = TRUE, 
  data = db_hedges_decline
)

summary(decline_hedges)
```

We construct the bubble plot showing the decline effect over years:

```{r}
#| label: hedges_decline_bubble_plot
#| message: false
#| warning: false
#| fig.width: 9
#| fig.height: 6

p_decline_hedges <- orchaRd::bubble_plot(
  decline_hedges, 
  mod = "Year_c", 
  group = "identifierStudyId", 
  xlab = "Publication Year (centered)", 
  ylab = "Effect size (Hedges' g)"
)

# Color points by real source
p_decline_hedges$layers[[1]]$data$real_source <- db_hedges_decline$Dataset_Source
p_decline_hedges$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
p_decline_hedges$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

p_decline_hedges <- p_decline_hedges +
  ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
  ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source")

# Relabel centered x-axis to actual calendar years
mean_year_hedges <- mean(db_hedges_egger$Year, na.rm = TRUE)
target_years_hedges <- seq(2000, 2026, by = 5)

p_decline_hedges_relabeled <- p_decline_hedges + 
  ggplot2::scale_x_continuous(
    name = "Publication Year",
    breaks = target_years_hedges - mean_year_hedges, 
    labels = target_years_hedges
  ) +
  ggplot2::ggtitle("Decline Effect / Time-Lag Bias (Hedges' g)") +
  ggplot2::theme_minimal()

print(p_decline_hedges_relabeled)
```

> [!NOTE]
> The publication year slope is **non-significant** ($\beta = -0.0104$, $SE = 0.0067$, $t = -1.5615$, $p = 0.1199$), indicating no statistically significant decline effect over time in continuous outcomes.

#### 4. Unified Diagnostic Dashboard (Hedges' $g$)

We combine the Egger regression and publication year time-lag bubble plots into a single visual panel using `patchwork`.

```{r}
#| label: hedges_unified_dashboard
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 12

dashboard_hedges <- (p_egger_hedges / p_decline_hedges_relabeled) +
  patchwork::plot_annotation(
    title = "Publication Bias Diagnostic Dashboard — Hedges' g",
    tag_levels = "A"
  ) & 
  ggplot2::theme(
    plot.title = ggplot2::element_text(size = 14, face = "bold", hjust = 0.5),
    plot.tag = ggplot2::element_text(size = 12, face = "bold")
  )

print(dashboard_hedges)
```

---

### Odds Ratio (Binary outcomes)

We fit multilevel models on the binary combined dataset ($k = 168$) to evaluate bias in Log Odds Ratio ($\ln\text{OR}$) studies.

#### 1. Baseline Model & Funnel Plot

We run the intercept-only model for Odds Ratio and inspect the funnel plot representing log Odds Ratio values plotted against their precision.

```{r}
#| label: or_baseline_and_funnel
#| message: false
#| warning: false
#| fig.width: 9
#| fig.height: 7

# Intercept-only model
VCV_or <- metafor::vcalc(vi = vi, cluster = identifierStudyId, obs = identifierEffectSizeID, rho = 0.5, data = db_or)

ma_or <- metafor::rma.mv(
  yi = yi, 
  V = VCV_or, 
  random = list(~1 | identifierStudyId, ~1 | identifierEffectSizeID, ~1 | taxonomySpecies), 
  test = "t", 
  method = "REML", 
  sparse = TRUE, 
  data = db_or
)

summary(ma_or)

# Funnel plot
metafor::funnel(
  ma_or,
  yaxis = "seinv",
  las = 1,
  digits = c(2, 1),
  xlab = "Standardised Log Odds Ratio (lnOR)",
  ylab = "Precision (Inverse of SE)",
  pch = 21,
  bg = alpha("black", 0.2),
  col = "black",
  cex = 1.2,
  main = "Funnel Plot for Binary Outcomes (Log Odds Ratio)"
)
```

#### 2. Egger's Meta-Regression (Small-Study Effects)

For Odds Ratio studies, because raw sample sizes are incomplete in the original binary dataset, we use the standard error ($\text{SE} = \sqrt{v_i}$) as our precision proxy. If the standard error moderator slope is statistically significant ($p < 0.05$), it represents a significant small-study effect.

```{r}
#| label: or_eggers_regression
#| message: false
#| warning: false

# Calculate standard error proxy
db_or_egger <- db_or %>%
  mutate(se = sqrt(vi)) %>%
  filter(!is.na(se))

# Re-estimate VCV on this subset
VCV_or_egger <- metafor::vcalc(vi = vi, cluster = identifierStudyId, obs = identifierEffectSizeID, rho = 0.5, data = db_or_egger)

# Egger regression model
egger_or <- metafor::rma.mv(
  yi = yi, 
  V = VCV_or_egger, 
  mods = ~ 1 + se, 
  random = list(~1 | identifierStudyId, ~1 | identifierEffectSizeID, ~1 | taxonomySpecies), 
  test = "t", 
  method = "REML", 
  sparse = TRUE, 
  data = db_or_egger
)

summary(egger_or)
```

We visualize the Egger regression for Odds Ratio using a bubble plot.

```{r}
#| label: or_egger_bubble_plot
#| message: false
#| warning: false
#| fig.width: 9
#| fig.height: 6

p_egger_or <- orchaRd::bubble_plot(
  egger_or, 
  mod = "se", 
  group = "identifierStudyId", 
  xlab = "Standard Error of Log Odds Ratio (se)", 
  ylab = "Effect size (lnOR)"
)

# Color points by real source
p_egger_or$layers[[1]]$data$real_source <- db_or_egger$Dataset_Source
p_egger_or$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
p_egger_or$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

p_egger_or <- p_egger_or + 
  ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal 2019)" = "#0072B2"), name = "Data Source") +
  ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal 2019)" = "#0072B2"), name = "Data Source") +
  ggplot2::ggtitle("Egger's Meta-Regression (Log Odds Ratio)") +
  ggplot2::theme_minimal()

print(p_egger_or)
```

> [!NOTE]
> The slope for the standard error is **extremely significant** ($\beta = 2.1129$, $SE = 0.4471$, $t = 4.7257$, $p < 0.0001$), indicating highly pronounced small-study effects in binary mate-choice copying outcomes. Studies with higher standard errors (smaller precision) report significantly larger log odds ratios.

#### 3. Decline Effect (Time-Lag Bias)

We test for a decline effect in the binary outcomes by regressing $\ln\text{OR}$ on centered publication year while simultaneously controlling for standard errors.

```{r}
#| label: or_time_lag
#| message: false
#| warning: false

# Mean-center publication year
db_or_decline <- db_or_egger %>%
  mutate(Year_c = Year - mean(Year, na.rm = TRUE))

# VCV for decline subset
VCV_or_decline <- metafor::vcalc(vi = vi, cluster = identifierStudyId, obs = identifierEffectSizeID, rho = 0.5, data = db_or_decline)

# Time-lag decline model
decline_or <- metafor::rma.mv(
  yi = yi, 
  V = VCV_or_decline, 
  mods = ~ 1 + Year_c + se, 
  random = list(~1 | identifierStudyId, ~1 | identifierEffectSizeID, ~1 | taxonomySpecies), 
  test = "t", 
  method = "REML", 
  sparse = TRUE, 
  data = db_or_decline
)

summary(decline_or)
```

We construct the bubble plot showing the decline effect for Odds Ratios:

```{r}
#| label: or_decline_bubble_plot
#| message: false
#| warning: false
#| fig.width: 9
#| fig.height: 6

p_decline_or <- orchaRd::bubble_plot(
  decline_or, 
  mod = "Year_c", 
  group = "identifierStudyId", 
  xlab = "Publication Year (centered)", 
  ylab = "Effect size (lnOR)"
)

# Color points by real source
p_decline_or$layers[[1]]$data$real_source <- db_or_decline$Dataset_Source
p_decline_or$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
p_decline_or$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

p_decline_or <- p_decline_or +
  ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal 2019)" = "#0072B2"), name = "Data Source") +
  ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal 2019)" = "#0072B2"), name = "Data Source")

# Relabel centered x-axis to actual calendar years
mean_year_or <- mean(db_or_egger$Year, na.rm = TRUE)
target_years_or <- seq(2000, 2026, by = 5)

p_decline_or_relabeled <- p_decline_or + 
  ggplot2::scale_x_continuous(
    name = "Publication Year",
    breaks = target_years_or - mean_year_or, 
    labels = target_years_or
  ) +
  ggplot2::ggtitle("Decline Effect / Time-Lag Bias (Log Odds Ratio)") +
  ggplot2::theme_minimal()

print(p_decline_or_relabeled)
```

> [!NOTE]
> The publication year slope is **statistically significant** ($\beta = -0.0282$, $SE = 0.0109$, $t = -2.5924$, $p = 0.0104$). This provides clear evidence of a **decline effect (Time-Lag Bias)** in binary outcomes—the reported strength of mate-choice copying has significantly decreased over the course of academic publication history.

#### 4. Unified Diagnostic Dashboard (Log Odds Ratio)

We combine the Egger regression and publication year time-lag bubble plots into a single visual panel using `patchwork`.

```{r}
#| label: or_unified_dashboard
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 12

dashboard_or <- (p_egger_or / p_decline_or_relabeled) +
  patchwork::plot_annotation(
    title = "Publication Bias Diagnostic Dashboard — Log Odds Ratio",
    tag_levels = "A"
  ) & 
  ggplot2::theme(
    plot.title = ggplot2::element_text(size = 14, face = "bold", hjust = 0.5),
    plot.tag = ggplot2::element_text(size = 12, face = "bold")
  )

print(dashboard_or)
```

:::

---

## Methodological Synthesis & Discussion

We compile the results of the small-study effect regressions and time-lag meta-regressions for both effect size classes below:

### Publication Bias Comparison Table

```{r}
#| label: summary_table_creation
#| echo: false

bias_summary <- data.frame(
  Effect_Size_Class = c("Hedges' g (Continuous)", "Hedges' g (Continuous)", "Log Odds Ratio (Binary)", "Log Odds Ratio (Binary)"),
  Diagnostic_Test = c("Egger's Regression (Small-Study)", "Time-Lag (Decline Effect)", "Egger's Regression (Small-Study)", "Time-Lag (Decline Effect)"),
  Moderator = c("sqrt_inv_e_n", "Year_c (centered)", "Standard Error (se)", "Year_c (centered)"),
  Slope_Beta = c(egger_hedges$b[2], decline_hedges$b[2], egger_or$b[2], decline_or$b[2]),
  Std_Error = c(egger_hedges$se[2], decline_hedges$se[2], egger_or$se[2], decline_or$se[2]),
  t_z_value = c(egger_hedges$zval[2], decline_hedges$zval[2], egger_or$zval[2], decline_or$zval[2]),
  p_value = c(egger_hedges$pval[2], decline_hedges$pval[2], egger_or$pval[2], decline_or$pval[2]),
  Evidence_of_Bias = c("Significant Bias (p = 0.0215)", "No Significant Bias (p = 0.1199)", "Highly Significant Bias (p < 0.0001)", "Significant Bias (p = 0.0104)")
)

DT::datatable(
  bias_summary,
  options = list(
    dom = 't',
    columnDefs = list(list(className = 'dt-center', targets = "_all"))
  ),
  rownames = FALSE
) %>%
  DT::formatRound(columns = c("Slope_Beta", "Std_Error", "t_z_value"), digits = 4) %>%
  DT::formatSignif(columns = c("p_value"), digits = 4)
```

### Main Takeaways

1. **Small-Study Effects Are Pervasive:** 
   Both Hedges' $g$ (continuous) and Log Odds Ratio (binary) datasets show statistically significant evidence of small-study effects. Smaller, less precise trials systematically report larger mate-choice copying effect sizes. This asymmetrical pattern is a classic signature of publication selection, where small-sample studies are more likely to be published or reported if they find large, hypothesis-supportive effects.
   
2. **Decline Effect (Time-Lag Bias) in Binary Outcomes:**
   We found robust evidence of a decline effect in binary mate-choice copying studies ($\beta = -0.0282$, $p = 0.0104$). Over the last two decades, the strength of reported mate-choice copying (as measured by Log Odds Ratios) has significantly decreased. This is typical in fields experiencing a "novelty effect," where initial breakthrough publications show large effect sizes, and subsequent rigorous replications or expanded testing reveal more moderate, realistic estimates.
   
3. **Continuous Outcomes Show More Stable Trends Over Time:**
   For Hedges' $g$ (continuous), although small-study effects are present, the publication year trend is non-significant ($\beta = -0.0104$, $p = 0.1199$). This suggests that the continuous reporting scales have been relatively more stable in magnitude over the history of the literature.

These diagnostic results suggest that while mate-choice copying is a genuine ecological phenomenon, the overall raw pooled estimates from both continuous and binary literatures may be moderately inflated due to selective reporting and publication biases. Future research should prioritize large-sample pre-registered replication studies to provide unbiased baselines.

---

## Session Info

::: callout-note
```{r}
#| label: session_info
sessionInfo()
```
:::
