---
title: "Publication Bias Diagnostics"
subtitle: "Small-Study Effects and Time-Lag Bias in Mate-Choice Copying (MCC) Meta-Analyses"
author: "Eduardo S. A. Santos & Collaborators"
date: today
bibliography: ../../references.bib
format:
  html:
    toc: true
    toc-depth: 3
    theme: cosmo
    self-contained: true
---

```{r}
#| label: setup
#| include: false
# Ensure a CRAN mirror is set so package installation works in non-interactive renders
if (is.null(getOption("repos")) || getOption("repos")["CRAN"] %in% c(NA, "@CRAN@")) {
    options(repos = c(CRAN = "https://cloud.r-project.org"))
}
# Ensure pacman is loaded to manage dependencies
if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
# orchaRd is hosted on GitHub (not CRAN); install from GitHub if missing.
# We require a recent build that ships `pub_bias_plot()` (Yang et al. 2024
# bias-robust correction), so we (re)install from GitHub if the function is
# not yet available in the locally installed version.
if (!requireNamespace("orchaRd", quietly = TRUE) ||
    !("pub_bias_plot" %in% getNamespaceExports("orchaRd"))) {
    pacman::p_load_gh("daniel1noble/orchaRd")
}
pacman::p_load(
    readxl,
    tidyverse,
    metafor,
    orchaRd,
    clubSandwich,
    patchwork,
    DT,
    scales
)

# Helper function to format numbers to a given precision and clean negative zeros
fmt <- function(x, digits = 2) {
    res <- sprintf(paste0("%.", digits, "f"), x)
    zero_pattern <- paste0("-0.", paste0(rep("0", digits), collapse = ""))
    res[res == zero_pattern] <- paste0("0.", paste0(rep("0", digits), collapse = ""))
    res
}

# Inline p-value formatter (reports "p < 0.0001" for very small p-values)
pstr <- function(p) {
    if (length(p) == 0 || is.na(p)) "p = NA" else if (p < 0.0001) "p < 0.0001" else paste0("p = ", fmt(p, 3))
}

# Significance-word helper: returns `yes` when p < 0.05, else `no`
sigw <- function(p, yes = "statistically significant", no = "non-significant") {
    if (length(p) == 0 || is.na(p)) "indeterminate" else if (p < 0.05) yes else no
}

# Directional-word helper based on the sign of a coefficient
dirw <- function(b, pos = "positive", neg = "negative") {
    if (length(b) == 0 || is.na(b)) "undetermined" else if (b >= 0) pos else neg
}

# ---------------------------------------------------------------------------
# Figure export helpers (manuscript-ready PDFs)
# ---------------------------------------------------------------------------
# All saved figures land in outputs/1_effect_size_calculation_pipeline/, the
# same convention used by overall_effect.qmd's save_taxon_fig().
fig_out_dir <- here::here("outputs", "1_effect_size_calculation_pipeline")
dir.create(fig_out_dir, recursive = TRUE, showWarnings = FALSE)

# Draw side-by-side marginal vs. conditional residual funnel plots (base
# graphics) with (a)/(b) panel tags. Reused for both on-screen display and
# PDF export so the rendered HTML and the saved file are identical.
draw_funnels <- function(resid_marg, resid_cond, vi,
                         xlab_marg = "Meta-analytic residuals (marginal)",
                         xlab_cond = "Meta-analytic residuals (conditional)") {
    op <- par(mfrow = c(1, 2))
    on.exit(par(op))

    metafor::funnel(
        x = resid_marg, vi = vi,
        yaxis = "seinv", digits = 2,
        level = c(0.1, 0.05, 0.01),
        shade = c("white", "gray55", "gray75"),
        las = 1, xlab = xlab_marg, legend = FALSE
    )
    title("Marginal Residual Funnel Plot", adj = 0.5, line = 2.5)
    mtext("(a)", side = 3, adj = 0, line = 2.5, font = 2, cex = 1.2)

    metafor::funnel(
        x = resid_cond, vi = vi,
        yaxis = "seinv", digits = 2,
        level = c(0.1, 0.05, 0.01),
        shade = c("white", "gray55", "gray75"),
        las = 1, xlab = xlab_cond, legend = FALSE
    )
    title("Conditional Residual Funnel Plot", adj = 0.5, line = 2.5)
    mtext("(b)", side = 3, adj = 0, line = 2.5, font = 2, cex = 1.2)
}

# Save a base-graphics drawing function to a PDF.
save_base_pdf <- function(basename, draw_fn, width = 12, height = 6) {
    tryCatch({
        grDevices::pdf(file.path(fig_out_dir, paste0(basename, ".pdf")), width = width, height = height)
        draw_fn()
        grDevices::dev.off()
    }, error = function(e) {
        if (length(grDevices::dev.list())) grDevices::dev.off()
        warning("Could not save ", basename, ": ", conditionMessage(e))
    })
}

# Save a ggplot/patchwork object to PDF (+ PNG for convenience).
save_gg_pdf <- function(plot_obj, basename, width = 10, height = 12) {
    if (is.null(plot_obj)) return(invisible())
    tryCatch({
        ggplot2::ggsave(file.path(fig_out_dir, paste0(basename, ".pdf")), plot_obj, width = width, height = height)
        ggplot2::ggsave(file.path(fig_out_dir, paste0(basename, ".png")), plot_obj, width = width, height = height, dpi = 300, bg = "white")
    }, error = function(e) warning("Could not save ", basename, ": ", conditionMessage(e)))
}
```

## Introduction & Methodology

Publication bias is a critical threat to the validity of meta-analytic syntheses. If studies with statistically non-significant or small effects are systematically omitted from the literature, the pooled meta-analytic summary will be biased upward.

To thoroughly evaluate the presence of publication bias and report-level selection in our Mate-Choice Copying (MCC) dataset, we employ three complementary diagnostic approaches, keeping the exact multilevel modeling logic (e.g., nested random intercepts and taxonomy control) used in our main overall effect calculations:

1. **Visual Inspection (Funnel Plots):** We plot effect size estimates against their precision (inverse of the standard error, $1/\text{SE}$) to check for visual asymmetry around the overall mean estimate.
2. **Egger's Meta-Regression (Small-Study Effects):** We test whether study precision significantly moderates effect size. Following the recommendations of @nakagawaMethodsTestingPublication2022 and Stanley [-@stanleyMetaregressionAnalysisEconomics2012; -@stanleyMetaregressionApproximationsReduce2014], we implement a **two-step approach**:
   * We first fit the Egger regression model using the standard error ($\text{SE} = \sqrt{v_i}$) as the precision proxy.
   * If the intercept of the standard error model is statistically significant, it indicates that the estimate is downwardly biased. In this case, we re-fit the model using the sampling variance ($\text{V} = v_i$) as the moderator to obtain a less biased estimate of the adjusted overall meta-analytic mean.
3. **Decline Effect / Time-Lag Bias:** We test whether effect sizes shrink over time by adding publication year (mean-centered) to our models while simultaneously controlling for study precision.
4. **Bias-Corrected Meta-Analytic Means (Method Comparison):** Finally, we estimate the *adjusted* overall mean under two complementary correction frameworks and compare them directly. The **current** approach follows @nakagawaMethodsTestingPublication2022: the intercept of a multilevel meta-regression on the sampling variance ($v_i$) is interpreted as the mean when sampling error is zero (i.e., an infinitely precise study). The **new** approach follows @yangRobustPointVariance2024: a *bias-robust* weighting scheme (a fixed-/common-effect GLS estimator that is less susceptible to selective reporting) is combined with **cluster-robust (CR2) variance estimation** to handle the non-independence of effect sizes within studies. We overlay both corrected means on an orchard plot using `orchaRd::pub_bias_plot()`.

---

## Data Loading and Alignment

We read, clean, and align the new extractions and original datasets, extracting sample sizes and publication years to construct our precision and time-lag moderators.

```{r}
#| label: load_and_align_datasets
#| message: false
#| warning: false
# Paths to datasets
path_new <- here::here("data", "1_effect_size_calculation_pipeline", "Data Extraction Mate Choice Meta Analysis.xlsx")
path_davies <- here::here("data", "original_meta_analyses_datasets", "Davies_et_al_2020_Final_data.xlsx")
path_jones <- here::here("data", "original_meta_analyses_datasets", "Jones_DuVal_2019_data.CSV")

# A. Read raw spreadsheets
db_new_raw <- readxl::read_excel(path_new, sheet = "EduardoExtraction")
db_davies_raw <- readxl::read_excel(path_davies, sheet = "Data")
db_jones_raw <- readr::read_csv(path_jones)

# B. Clean new extractions
db_new <- db_new_raw %>%
    filter(is.na(RemoveFromDataset) | RemoveFromDataset != "Yes") %>%
    mutate(
        effectSizeCalculatedHedgesD = as.numeric(effectSizeCalculatedHedgesD),
        effectSizeCalculatedHedgesDVariance = as.numeric(effectSizeCalculatedHedgesDVariance),
        effectSizeCalculatedOddsRatio = as.numeric(effectSizeCalculatedOddsRatio),
        effectSizeCalculatedOddsRatioVariance = as.numeric(effectSizeCalculatedOddsRatioVariance),
        logOddsRatio = log(effectSizeCalculatedOddsRatio),
        logOddsRatioVariance = effectSizeCalculatedOddsRatioVariance,
        identifierStudyYear = as.numeric(identifierStudyYear),
        controlN = as.numeric(controlN),
        treatmentN = as.numeric(treatmentN),
        effectSizesampleSize = as.numeric(effectSizesampleSize)
    )

# ---------------------------------------------------------------------------
# 1. Hedges' g Dataset Alignment (Continuous)
# ---------------------------------------------------------------------------
hedges_new <- db_new %>%
    filter(!is.na(effectSizeCalculatedHedgesD), !is.na(effectSizeCalculatedHedgesDVariance)) %>%
    select(
        yi = effectSizeCalculatedHedgesD,
        vi = effectSizeCalculatedHedgesDVariance,
        identifierStudyId,
        identifierEffectSizeID,
        taxonomySpecies,
        Year = identifierStudyYear,
        controlN,
        treatmentN,
        totalN = effectSizesampleSize
    ) %>%
    mutate(
        identifierStudyId = as.character(identifierStudyId),
        identifierEffectSizeID = as.character(identifierEffectSizeID),
        taxonomySpecies = str_to_lower(word(as.character(taxonomySpecies), -1)),
        Dataset_Source = "New Extraction (Santos)"
    )

hedges_davies <- db_davies_raw %>%
    mutate(
        Hedges_d_directional = as.numeric(Hedges_d_directional),
        Variance = as.numeric(Variance),
        Year = as.numeric(Year),
        N_control = as.numeric(N_control),
        N_treatment = as.numeric(N_treatment),
        Total_n_analysis = as.numeric(Total_n_analysis)
    ) %>%
    filter(!is.na(Hedges_d_directional), !is.na(Variance), Variance > 0) %>%
    select(
        yi = Hedges_d_directional,
        vi = Variance,
        identifierStudyId = Author,
        identifierEffectSizeID = Effect_size_no,
        taxonomySpecies = Species_latin,
        Year,
        controlN = N_control,
        treatmentN = N_treatment,
        totalN = Total_n_analysis
    ) %>%
    mutate(
        identifierStudyId = as.character(identifierStudyId),
        identifierEffectSizeID = as.character(identifierEffectSizeID),
        taxonomySpecies = str_to_lower(word(as.character(taxonomySpecies), -1)),
        Dataset_Source = "Original (Davies 2020)"
    )

db_hedges <- bind_rows(hedges_new, hedges_davies) %>%
    mutate(across(c(identifierStudyId, taxonomySpecies, Dataset_Source), as.factor))

# ---------------------------------------------------------------------------
# 2. Log Odds Ratio Dataset Alignment (Binary)
# ---------------------------------------------------------------------------
or_new <- db_new %>%
    filter(!is.na(logOddsRatio), !is.na(logOddsRatioVariance)) %>%
    select(
        yi = logOddsRatio,
        vi = logOddsRatioVariance,
        identifierStudyId,
        identifierEffectSizeID,
        taxonomySpecies,
        Year = identifierStudyYear
    ) %>%
    mutate(
        identifierStudyId = as.character(identifierStudyId),
        identifierEffectSizeID = as.character(identifierEffectSizeID),
        taxonomySpecies = str_to_lower(word(as.character(taxonomySpecies), -1)),
        Dataset_Source = "New Extraction (Santos)"
    )

or_jones <- db_jones_raw %>%
    filter(!is.na(lnOR), !is.na(VlnOR), VlnOR > 0) %>%
    select(
        yi = lnOR,
        vi = VlnOR,
        identifierStudyId = study,
        identifierEffectSizeID = study_Exp,
        taxonomySpecies = Species,
        Year = year
    ) %>%
    mutate(
        identifierStudyId = as.character(identifierStudyId),
        identifierEffectSizeID = as.character(identifierEffectSizeID),
        taxonomySpecies = str_to_lower(as.character(taxonomySpecies)),
        Dataset_Source = "Original (Jones & DuVal 2019)"
    )

db_or <- bind_rows(or_new, or_jones) %>%
    mutate(across(c(identifierStudyId, taxonomySpecies, Dataset_Source), as.factor))
```

---

## Study Inclusion & Exclusion {#sec-inclusion-pb}

The following tables document which effect sizes from the new extraction are included in the meta-analytic models and which have been removed, with the reason for each exclusion.

### Excluded Effect Sizes

```{r}
#| label: excluded_studies_table_pb
#| echo: false
db_excluded <- db_new_raw %>%
    filter(!is.na(RemoveFromDataset) & RemoveFromDataset == "Yes") %>%
    transmute(
        `Effect Size ID` = identifierEffectSizeID,
        `Study` = identifierStudyId,
        `Year` = identifierStudyYear,
        `Exclusion Reason` = ExclusionReason
    ) %>%
    arrange(`Effect Size ID`)

DT::datatable(
    db_excluded,
    caption = "Effect sizes removed from the meta-analytic models and their reasons for exclusion.",
    options = list(
        dom = "t",
        pageLength = 20,
        columnDefs = list(list(className = "dt-left", targets = "_all"))
    ),
    rownames = FALSE
)
```

### Included Effect Sizes

```{r}
#| label: included_studies_table_pb
#| echo: false
db_included <- db_new %>%
    transmute(
        `Effect Size ID` = identifierEffectSizeID,
        `Study` = identifierStudyId,
        `Year` = identifierStudyYear,
        `Species` = taxonomySpecies,
        `Data Type` = effectSizeDataTypeForCalculation,
        `Hedges' g` = round(effectSizeCalculatedHedgesD, 3),
        `OR` = round(effectSizeCalculatedOddsRatio, 3)
    ) %>%
    arrange(`Effect Size ID`)

DT::datatable(
    db_included,
    caption = "Effect sizes included in the meta-analytic models.",
    options = list(
        pageLength = 25,
        columnDefs = list(list(className = "dt-left", targets = "_all"))
    ),
    rownames = FALSE
)
```

---

## Diagnostics by Effect Size Class

::: {.panel-tabset}

### Hedges' $g$ (Continuous outcomes)

We fit multilevel models on the continuous combined dataset ($k = `r nrow(db_hedges)`$) to run our diagnostics.

#### 1. Baseline Model & Funnel Plot

We run the intercept-only model to establish the global mean effect size and plot precision ($1/\text{SE}$) against the standardized residuals.

```{r}
#| label: hedges_baseline_and_funnel
#| message: false
#| warning: false
#| fig.width: 12
#| fig.height: 6
# Intercept-only model
VCV_hedges <- metafor::vcalc(vi = vi, cluster = identifierStudyId, obs = identifierEffectSizeID, rho = 0.5, data = db_hedges)

ma_hedges <- metafor::rma.mv(
    yi = yi,
    V = VCV_hedges,
    random = list(~ 1 | identifierStudyId, ~ 1 | identifierEffectSizeID, ~ 1 | taxonomySpecies),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = db_hedges
)

summary(ma_hedges)

# Extract BLUPs for random effects
blups <- ranef(ma_hedges)
study_ef <- blups$identifierStudyId$intrcpt[match(db_hedges$identifierStudyId, row.names(blups$identifierStudyId))]
obs_ef <- blups$identifierEffectSizeID$intrcpt[match(db_hedges$identifierEffectSizeID, row.names(blups$identifierEffectSizeID))]
spec_ef <- blups$taxonomySpecies$intrcpt[match(db_hedges$taxonomySpecies, row.names(blups$taxonomySpecies))]

# Handle any missing values by replacing with 0
study_ef[is.na(study_ef)] <- 0
obs_ef[is.na(obs_ef)] <- 0
spec_ef[is.na(spec_ef)] <- 0

# Marginal residuals: observed - fixed-effect prediction
resid_marg <- db_hedges$yi - predict(ma_hedges)[[1]]

# Conditional residuals: observed - fixed - random intercepts (study, obs, species)
resid_cond <- db_hedges$yi - (predict(ma_hedges)[[1]] + study_ef + obs_ef + spec_ef)

# Side-by-side funnel plots of marginal vs. conditional residuals,
# with (a)/(b) panel tags. Drawn on-screen and saved to a manuscript PDF.
draw_funnels(resid_marg, resid_cond, db_hedges$vi)
save_base_pdf(
    "funnel_hedges",
    function() draw_funnels(resid_marg, resid_cond, db_hedges$vi)
)
```

::: {.callout-note title="Reading the funnel plots"}
Both panels plot each effect size against its precision ($1/\text{SE}$), but differ in what is removed before plotting:

* **(a) Marginal residuals** are the deviations of each observed effect from the *fixed-effect (intercept) prediction* — the pooled mean — so they still carry all of the random-effects heterogeneity (between-study, between-effect-size, and between-species). Asymmetry here — a shortage of small, imprecise studies on one side of the funnel — is the classic visual signature of small-study and publication bias. Because genuine heterogeneity can mimic or exaggerate that asymmetry, apparent gaps in (a) should be read cautiously and confirmed against the formal Egger's regression below.
* **(b) Conditional residuals** additionally subtract the estimated random intercepts (study, effect-size, and species BLUPs), so the heterogeneity captured by the multilevel structure is partialled out. A symmetric, evenly-filled funnel here means that once the nesting is accounted for there is little precision-related asymmetry left — i.e., the asymmetry seen in (a) is largely explained by the modelled random effects rather than by selective reporting. Any asymmetry that *persists* in (b) is the more credible visual indication of true small-study effects.
:::

#### 2. Egger's Meta-Regression (Small-Study Effects)

To perform Egger's regression on Hedges' $g$, we follow the two-step framework recommended by @nakagawaMethodsTestingPublication2022 and Stanley [-@stanleyMetaregressionAnalysisEconomics2012; -@stanleyMetaregressionApproximationsReduce2014]. We use the sampling standard error ($\text{SE} = \sqrt{v_i}$) as our precision proxy. If this moderator is statistically significant ($p < 0.05$), it indicates a significant small-study effect (larger standard errors / smaller sample sizes associated with larger effect magnitudes).

Because standard error model intercepts can be downwardly biased if they are significant, we apply a two-step correction: if the intercept of the standard error model is statistically significant, we re-fit the Egger's regression model using the sampling variance ($\text{V} = \text{vi}$) as the moderator to obtain a less biased estimate of the adjusted overall meta-analytic mean.

```{r}
#| label: hedges_eggers_regression
#| message: false
#| warning: false
# Calculate standard error precision proxy
db_hedges_egger <- db_hedges %>%
    mutate(se = sqrt(vi)) %>%
    filter(!is.na(se))

# Re-estimate VCV on this subset
VCV_hedges_egger <- metafor::vcalc(vi = vi, cluster = identifierStudyId, obs = identifierEffectSizeID, rho = 0.5, data = db_hedges_egger)

# Step 1: Egger regression model with standard error as moderator
egger_hedges <- metafor::rma.mv(
    yi = yi,
    V = VCV_hedges_egger,
    mods = ~ 1 + se,
    random = list(~ 1 | identifierStudyId, ~ 1 | identifierEffectSizeID, ~ 1 | taxonomySpecies),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = db_hedges_egger
)

summary(egger_hedges)

# Step 2: Since the intercept is significant, re-fit using variance (vi)
egger_hedges_v <- metafor::rma.mv(
    yi = yi,
    V = VCV_hedges_egger,
    mods = ~ 1 + vi,
    random = list(~ 1 | identifierStudyId, ~ 1 | identifierEffectSizeID, ~ 1 | taxonomySpecies),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = db_hedges_egger
)

summary(egger_hedges_v)
```

We visualize the Egger regression using a bubble plot. Effect sizes are plotted as circles scaled by precision, with the predicted regression line showing the relationship between sample size and effect size.

```{r}
#| label: hedges_egger_bubble_plot
#| message: false
#| warning: false
#| fig.width: 9
#| fig.height: 6
p_egger_hedges <- orchaRd::bubble_plot(
    egger_hedges,
    mod = "se",
    group = "identifierStudyId",
    xlab = "Standard Error of Hedges' g (se)",
    ylab = "Effect size (Hedges' g)"
)

# Color points by real source
p_egger_hedges$layers[[1]]$data$real_source <- db_hedges_egger$Dataset_Source
p_egger_hedges$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
p_egger_hedges$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

p_egger_hedges <- p_egger_hedges +
    ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
    ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
    ggplot2::ggtitle("Egger's Meta-Regression (Hedges' g)") +
    ggplot2::theme_minimal()

print(p_egger_hedges)
```

> [!NOTE]
> The slope for standard error (`se`) is **`r dirw(egger_hedges$b[2])`** ($\beta = `r fmt(egger_hedges$b[2])`$, $SE = `r fmt(egger_hedges$se[2])`$, $t = `r fmt(egger_hedges$zval[2])`$, $`r pstr(egger_hedges$pval[2])`$), indicating `r sigw(egger_hedges$pval[2], "clear evidence of small-study effects", "little evidence of small-study effects")`.
>
> **Two-Step Correction:** The intercept of the standard error model is `r sigw(egger_hedges$pval[1])` ($`r pstr(egger_hedges$pval[1])`$). Following the two-step recommendation of @nakagawaMethodsTestingPublication2022 and Stanley [-@stanleyMetaregressionAnalysisEconomics2012; -@stanleyMetaregressionApproximationsReduce2014], we re-fit Egger's regression using the sampling variance (`vi`) as the moderator. In this corrected model, the adjusted overall meta-analytic mean is estimated at $\beta_0 = `r fmt(egger_hedges_v$b[1])`$ ($SE = `r fmt(egger_hedges_v$se[1])`$, $t = `r fmt(egger_hedges_v$zval[1])`$, $`r pstr(egger_hedges_v$pval[1])`$).

#### 3. Decline Effect (Time-Lag Bias)

We analyze the decline effect by evaluating whether the publication year significantly predicts the effect size, while simultaneously controlling for the study sample size precision proxy. We mean-center the year so the intercept represents the effect in an average publication year.

```{r}
#| label: hedges_time_lag
#| message: false
#| warning: false
# Mean-center publication year
db_hedges_decline <- db_hedges_egger %>%
    mutate(Year_c = Year - mean(Year, na.rm = TRUE))

# VCV for decline subset
VCV_hedges_decline <- metafor::vcalc(vi = vi, cluster = identifierStudyId, obs = identifierEffectSizeID, rho = 0.5, data = db_hedges_decline)

# Time-lag decline model controlling for sampling variance (vi)
decline_hedges <- metafor::rma.mv(
    yi = yi,
    V = VCV_hedges_decline,
    mods = ~ 1 + Year_c + vi,
    random = list(~ 1 | identifierStudyId, ~ 1 | identifierEffectSizeID, ~ 1 | taxonomySpecies),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = db_hedges_decline
)

summary(decline_hedges)
```

We construct the bubble plot showing the decline effect over years:

```{r}
#| label: hedges_decline_bubble_plot
#| message: false
#| warning: false
#| fig.width: 9
#| fig.height: 6
p_decline_hedges <- orchaRd::bubble_plot(
    decline_hedges,
    mod = "Year_c",
    group = "identifierStudyId",
    xlab = "Publication Year (centered)",
    ylab = "Effect size (Hedges' g)"
)

# Color points by real source
p_decline_hedges$layers[[1]]$data$real_source <- db_hedges_decline$Dataset_Source
p_decline_hedges$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
p_decline_hedges$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

p_decline_hedges <- p_decline_hedges +
    ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
    ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source")

# Relabel centered x-axis to actual calendar years
mean_year_hedges <- mean(db_hedges_egger$Year, na.rm = TRUE)
target_years_hedges <- seq(2000, 2026, by = 5)

p_decline_hedges_relabeled <- p_decline_hedges +
    ggplot2::scale_x_continuous(
        name = "Publication Year",
        breaks = target_years_hedges - mean_year_hedges,
        labels = target_years_hedges
    ) +
    ggplot2::ggtitle("Decline Effect / Time-Lag Bias (Hedges' g)") +
    ggplot2::theme_minimal()

print(p_decline_hedges_relabeled)
```

> [!NOTE]
> The publication year slope is **`r sigw(decline_hedges$pval[2])`** ($\beta = `r fmt(decline_hedges$b[2])`$, $SE = `r fmt(decline_hedges$se[2])`$, $t = `r fmt(decline_hedges$zval[2])`$, $`r pstr(decline_hedges$pval[2])`$), indicating `r sigw(decline_hedges$pval[2], "evidence of a decline effect over time", "little evidence of a decline effect over time")` in continuous outcomes.

#### 4. Unified Diagnostic Dashboard (Hedges' $g$)

We combine the Egger regression and publication year time-lag bubble plots into a single visual panel using `patchwork`.

```{r}
#| label: hedges_unified_dashboard
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 12
dashboard_hedges <- (p_egger_hedges / p_decline_hedges_relabeled) +
    patchwork::plot_annotation(
        title = "Publication Bias Diagnostic Dashboard — Hedges' g",
        tag_levels = "a", tag_prefix = "(", tag_suffix = ")"
    ) &
    ggplot2::theme(
        plot.title = ggplot2::element_text(size = 14, face = "bold", hjust = 0.5),
        plot.tag = ggplot2::element_text(size = 12, face = "bold")
    )

print(dashboard_hedges)
save_gg_pdf(dashboard_hedges, "dashboard_eggers_timelag_hedges", width = 10, height = 12)
```

---

### Odds Ratio (Binary outcomes)

We fit multilevel models on the binary combined dataset ($k = `r nrow(db_or)`$) to evaluate bias in Log Odds Ratio ($\ln\text{OR}$) studies.

#### 1. Baseline Model & Funnel Plot

We run the intercept-only model for Odds Ratio and inspect the funnel plot representing log Odds Ratio values plotted against their precision.

```{r}
#| label: or_baseline_and_funnel
#| message: false
#| warning: false
#| fig.width: 12
#| fig.height: 6
# Intercept-only model
VCV_or <- metafor::vcalc(vi = vi, cluster = identifierStudyId, obs = identifierEffectSizeID, rho = 0.5, data = db_or)

ma_or <- metafor::rma.mv(
    yi = yi,
    V = VCV_or,
    random = list(~ 1 | identifierStudyId, ~ 1 | identifierEffectSizeID, ~ 1 | taxonomySpecies),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = db_or
)

summary(ma_or)

# Extract BLUPs for random effects
blups <- ranef(ma_or)
study_ef <- blups$identifierStudyId$intrcpt[match(db_or$identifierStudyId, row.names(blups$identifierStudyId))]
obs_ef <- blups$identifierEffectSizeID$intrcpt[match(db_or$identifierEffectSizeID, row.names(blups$identifierEffectSizeID))]
spec_ef <- blups$taxonomySpecies$intrcpt[match(db_or$taxonomySpecies, row.names(blups$taxonomySpecies))]

# Handle any missing values by replacing with 0
study_ef[is.na(study_ef)] <- 0
obs_ef[is.na(obs_ef)] <- 0
spec_ef[is.na(spec_ef)] <- 0

# Marginal residuals: observed - fixed-effect prediction
resid_marg <- db_or$yi - predict(ma_or)[[1]]

# Conditional residuals: observed - fixed - random intercepts (study, obs, species)
resid_cond <- db_or$yi - (predict(ma_or)[[1]] + study_ef + obs_ef + spec_ef)

# Side-by-side funnel plots of marginal vs. conditional residuals,
# with (a)/(b) panel tags. Drawn on-screen and saved to a manuscript PDF.
draw_funnels(resid_marg, resid_cond, db_or$vi)
save_base_pdf(
    "funnel_or",
    function() draw_funnels(resid_marg, resid_cond, db_or$vi)
)
```

::: {.callout-note title="Reading the funnel plots"}
Both panels plot each $\ln\text{OR}$ effect size against its precision ($1/\text{SE}$), but differ in what is removed before plotting:

* **(a) Marginal residuals** are the deviations of each observed effect from the *fixed-effect (intercept) prediction* — the pooled mean — so they still carry all of the random-effects heterogeneity (between-study, between-effect-size, and between-species). Asymmetry here — a shortage of small, imprecise studies on one side of the funnel — is the classic visual signature of small-study and publication bias. Because genuine heterogeneity can mimic or exaggerate that asymmetry, apparent gaps in (a) should be read cautiously and confirmed against the formal Egger's regression below.
* **(b) Conditional residuals** additionally subtract the estimated random intercepts (study, effect-size, and species BLUPs), so the heterogeneity captured by the multilevel structure is partialled out. A symmetric, evenly-filled funnel here means that once the nesting is accounted for there is little precision-related asymmetry left — i.e., the asymmetry seen in (a) is largely explained by the modelled random effects rather than by selective reporting. Any asymmetry that *persists* in (b) is the more credible visual indication of true small-study effects.
:::

#### 2. Egger's Meta-Regression (Small-Study Effects)

For Odds Ratio studies, we use the standard error ($\text{SE} = \sqrt{v_i}$) as our precision proxy in the two-step framework. If the standard error moderator slope is statistically significant ($p < 0.05$), it represents a significant small-study effect.

As with the continuous outcomes, we apply the two-step correction: if the intercept of the standard error model is statistically significant, we re-fit the Egger's regression model using the sampling variance ($\text{V} = \text{vi}$) as the moderator to obtain a less biased estimate of the adjusted overall meta-analytic mean.

```{r}
#| label: or_eggers_regression
#| message: false
#| warning: false
# Calculate standard error proxy
db_or_egger <- db_or %>%
    mutate(se = sqrt(vi)) %>%
    filter(!is.na(se))

# Re-estimate VCV on this subset
VCV_or_egger <- metafor::vcalc(vi = vi, cluster = identifierStudyId, obs = identifierEffectSizeID, rho = 0.5, data = db_or_egger)

# Step 1: Egger regression model with standard error as moderator
egger_or <- metafor::rma.mv(
    yi = yi,
    V = VCV_or_egger,
    mods = ~ 1 + se,
    random = list(~ 1 | identifierStudyId, ~ 1 | identifierEffectSizeID, ~ 1 | taxonomySpecies),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = db_or_egger
)

summary(egger_or)

# Step-2: Since the intercept is significant, re-fit using variance (vi)
egger_or_v <- metafor::rma.mv(
    yi = yi,
    V = VCV_or_egger,
    mods = ~ 1 + vi,
    random = list(~ 1 | identifierStudyId, ~ 1 | identifierEffectSizeID, ~ 1 | taxonomySpecies),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = db_or_egger
)

summary(egger_or_v)
```

We visualize the Egger regression for Odds Ratio using a bubble plot.

```{r}
#| label: or_egger_bubble_plot
#| message: false
#| warning: false
#| fig.width: 9
#| fig.height: 6
p_egger_or <- orchaRd::bubble_plot(
    egger_or,
    mod = "se",
    group = "identifierStudyId",
    xlab = "Standard Error of Log Odds Ratio (se)",
    ylab = "Effect size (lnOR)"
)

# Color points by real source
p_egger_or$layers[[1]]$data$real_source <- db_or_egger$Dataset_Source
p_egger_or$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
p_egger_or$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

p_egger_or <- p_egger_or +
    ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal 2019)" = "#0072B2"), name = "Data Source") +
    ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal 2019)" = "#0072B2"), name = "Data Source") +
    ggplot2::ggtitle("Egger's Meta-Regression (Log Odds Ratio)") +
    ggplot2::theme_minimal()

print(p_egger_or)
```

> [!NOTE]
> The slope for the standard error is **`r dirw(egger_or$b[2])`** ($\beta = `r fmt(egger_or$b[2])`$, $SE = `r fmt(egger_or$se[2])`$, $t = `r fmt(egger_or$zval[2])`$, $`r pstr(egger_or$pval[2])`$), indicating `r sigw(egger_or$pval[2], "pronounced small-study effects", "little evidence of small-study effects")` in binary mate-choice copying outcomes.
>
> **Two-Step Correction:** The intercept of the standard error model is `r sigw(egger_or$pval[1])` ($`r pstr(egger_or$pval[1])`$). Following the two-step framework, we re-fit the model using the sampling variance (`vi`) as the moderator to obtain a less biased estimate of the adjusted mean. In this corrected model, the adjusted overall log odds ratio is `r sigw(egger_or_v$pval[1], "significant", "non-significant")` at $\beta_0 = `r fmt(egger_or_v$b[1])`$ ($SE = `r fmt(egger_or_v$se[1])`$, $t = `r fmt(egger_or_v$zval[1])`$, $`r pstr(egger_or_v$pval[1])`$).

#### 3. Decline Effect (Time-Lag Bias)

We test for a decline effect in the binary outcomes by regressing $\ln\text{OR}$ on centered publication year while simultaneously controlling for standard errors.

```{r}
#| label: or_time_lag
#| message: false
#| warning: false
# Mean-center publication year
db_or_decline <- db_or_egger %>%
    mutate(Year_c = Year - mean(Year, na.rm = TRUE))

# VCV for decline subset
VCV_or_decline <- metafor::vcalc(vi = vi, cluster = identifierStudyId, obs = identifierEffectSizeID, rho = 0.5, data = db_or_decline)

# Time-lag decline model controlling for sampling variance (vi)
decline_or <- metafor::rma.mv(
    yi = yi,
    V = VCV_or_decline,
    mods = ~ 1 + Year_c + vi,
    random = list(~ 1 | identifierStudyId, ~ 1 | identifierEffectSizeID, ~ 1 | taxonomySpecies),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = db_or_decline
)

summary(decline_or)
```

We construct the bubble plot showing the decline effect for Odds Ratios:

```{r}
#| label: or_decline_bubble_plot
#| message: false
#| warning: false
#| fig.width: 9
#| fig.height: 6
p_decline_or <- orchaRd::bubble_plot(
    decline_or,
    mod = "Year_c",
    group = "identifierStudyId",
    xlab = "Publication Year (centered)",
    ylab = "Effect size (lnOR)"
)

# Color points by real source
p_decline_or$layers[[1]]$data$real_source <- db_or_decline$Dataset_Source
p_decline_or$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
p_decline_or$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

p_decline_or <- p_decline_or +
    ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal 2019)" = "#0072B2"), name = "Data Source") +
    ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal 2019)" = "#0072B2"), name = "Data Source")

# Relabel centered x-axis to actual calendar years
mean_year_or <- mean(db_or_egger$Year, na.rm = TRUE)
target_years_or <- seq(2000, 2026, by = 5)

p_decline_or_relabeled <- p_decline_or +
    ggplot2::scale_x_continuous(
        name = "Publication Year",
        breaks = target_years_or - mean_year_or,
        labels = target_years_or
    ) +
    ggplot2::ggtitle("Decline Effect / Time-Lag Bias (Log Odds Ratio)") +
    ggplot2::theme_minimal()

print(p_decline_or_relabeled)
```

> [!NOTE]
> The publication year slope is **`r sigw(decline_or$pval[2])`** ($\beta = `r fmt(decline_or$b[2])`$, $SE = `r fmt(decline_or$se[2])`$, $t = `r fmt(decline_or$zval[2])`$, $`r pstr(decline_or$pval[2])`$). `r if(decline_or$pval[2] < 0.05) "This provides evidence of a **decline effect (Time-Lag Bias)** in binary outcomes—the reported strength of mate-choice copying has decreased over the course of academic publication history." else "This indicates little evidence of a decline effect (Time-Lag Bias) in binary outcomes over the course of academic publication history."`

#### 4. Unified Diagnostic Dashboard (Log Odds Ratio)

We combine the Egger regression and publication year time-lag bubble plots into a single visual panel using `patchwork`.

```{r}
#| label: or_unified_dashboard
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 12
dashboard_or <- (p_egger_or / p_decline_or_relabeled) +
    patchwork::plot_annotation(
        title = "Publication Bias Diagnostic Dashboard — Log Odds Ratio",
        tag_levels = "a", tag_prefix = "(", tag_suffix = ")"
    ) &
    ggplot2::theme(
        plot.title = ggplot2::element_text(size = 14, face = "bold", hjust = 0.5),
        plot.tag = ggplot2::element_text(size = 12, face = "bold")
    )

print(dashboard_or)
save_gg_pdf(dashboard_or, "dashboard_eggers_timelag_or", width = 10, height = 12)
```

:::

---

## Bias-Corrected Meta-Analytic Means: Current vs. New Method

The diagnostics above *detect* small-study and time-lag bias. Here we turn to *correcting* the pooled mean, and we directly compare two correction philosophies on the same data.

* **Current method — Nakagawa et al. [-@nakagawaMethodsTestingPublication2022] regression correction.** We re-use the multilevel meta-regression that includes the sampling variance ($v_i$) as a moderator (the Step-2 models `egger_hedges_v` / `egger_or_v` fitted above). Its **intercept** estimates the meta-analytic mean extrapolated to $v_i = 0$ — i.e., the effect a hypothetical infinitely-precise (infinite-$n$) study would report, with the small-study component removed. This is the *"bias-corrected estimate"* in @yangRobustPointVariance2024's terminology and in `orchaRd`.
* **New method — Yang et al. [-@yangRobustPointVariance2024] bias-robust estimation.** A two-step procedure: (1) fit a **fixed-/common-effect** model, whose precision-based GLS weighting is far less inflated by selectively-reported small studies than the random-effects mean; then (2) apply **cluster-robust (CR2) variance estimation** (`clubSandwich`), clustering by study, so that the standard errors and confidence intervals remain valid under the non-independence of multiple effect sizes per study. This is the *"bias-robust estimate"*.

We compute both, overlay them on an orchard plot of the uncorrected multilevel mean with `orchaRd::pub_bias_plot()`, and tabulate the three estimates (uncorrected vs. the two corrections) side by side.

::: {.panel-tabset}

### Hedges' $g$ (Continuous outcomes)

```{r}
#| label: hedges_biasrobust_models
#| message: false
#| warning: false
# --- New method: Yang et al. (2024) bias-robust estimate ---------------------
# Step 1: common-/fixed-effect model (bias-robust GLS weighting).
fe_hedges <- metafor::rma(
    yi = yi,
    vi = vi,
    data = db_hedges_egger,
    method = "FE",
    test = "t"
)

# Step 2: cluster-robust (CR2) variance estimation, clustering by study,
# to account for non-independence of multiple effect sizes per study.
fe_hedges_robust <- metafor::robust(
    fe_hedges,
    cluster = db_hedges_egger$identifierStudyId,
    clubSandwich = TRUE
)

summary(fe_hedges_robust)

# `egger_hedges_v` (fitted above) is the current-method Nakagawa et al. (2022)
# vi-moderator model; its intercept is the bias-corrected mean at vi = 0.
```

```{r}
#| label: hedges_pub_bias_plot
#| message: false
#| warning: false
#| fig.width: 9
#| fig.height: 5
# Orchard plot of the uncorrected multilevel meta-analytic mean ...
oc_hedges <- orchaRd::orchard_plot(
    ma_hedges,
    group = "identifierStudyId",
    xlab = "Hedges' g",
    angle = 0
)

# ... with both corrected means overlaid:
#   red  = Yang et al. 2024 "Bias Robust"     (fixed-effect + CR2 robust)
#   blue = Nakagawa et al. 2022 "Bias Corrected" (vi-moderator intercept)
p_pubbias_hedges <- orchaRd::pub_bias_plot(
    oc_hedges,
    fe_model = fe_hedges_robust,
    v_model  = egger_hedges_v
) +
    ggplot2::ggtitle("Bias-Corrected Means — Hedges' g") +
    ggplot2::theme_minimal()

print(p_pubbias_hedges)
```

> [!NOTE]
> The uncorrected multilevel mean is $\hat\mu = `r fmt(ma_hedges$b[1])`$ (95% CI [`r fmt(ma_hedges$ci.lb[1])`, `r fmt(ma_hedges$ci.ub[1])`]). The **current** Nakagawa correction gives a bias-corrected mean of $`r fmt(egger_hedges_v$b[1])`$ (95% CI [`r fmt(egger_hedges_v$ci.lb[1])`, `r fmt(egger_hedges_v$ci.ub[1])`]), and the **new** Yang bias-robust estimate is $`r fmt(fe_hedges_robust$b[1])`$ (95% CI [`r fmt(fe_hedges_robust$ci.lb[1])`, `r fmt(fe_hedges_robust$ci.ub[1])`]).

### Odds Ratio (Binary outcomes)

```{r}
#| label: or_biasrobust_models
#| message: false
#| warning: false
# --- New method: Yang et al. (2024) bias-robust estimate ---------------------
fe_or <- metafor::rma(
    yi = yi,
    vi = vi,
    data = db_or_egger,
    method = "FE",
    test = "t"
)

fe_or_robust <- metafor::robust(
    fe_or,
    cluster = db_or_egger$identifierStudyId,
    clubSandwich = TRUE
)

summary(fe_or_robust)

# `egger_or_v` (fitted above) is the current-method Nakagawa et al. (2022)
# vi-moderator model; its intercept is the bias-corrected mean at vi = 0.
```

```{r}
#| label: or_pub_bias_plot
#| message: false
#| warning: false
#| fig.width: 9
#| fig.height: 5
oc_or <- orchaRd::orchard_plot(
    ma_or,
    group = "identifierStudyId",
    xlab = "Log Odds Ratio (lnOR)",
    angle = 0
)

p_pubbias_or <- orchaRd::pub_bias_plot(
    oc_or,
    fe_model = fe_or_robust,
    v_model  = egger_or_v
) +
    ggplot2::ggtitle("Bias-Corrected Means — Log Odds Ratio") +
    ggplot2::theme_minimal()

print(p_pubbias_or)
```

> [!NOTE]
> The uncorrected multilevel mean is $\hat\mu = `r fmt(ma_or$b[1])`$ (95% CI [`r fmt(ma_or$ci.lb[1])`, `r fmt(ma_or$ci.ub[1])`]) on the $\ln\text{OR}$ scale. The **current** Nakagawa correction gives $`r fmt(egger_or_v$b[1])`$ (95% CI [`r fmt(egger_or_v$ci.lb[1])`, `r fmt(egger_or_v$ci.ub[1])`]), and the **new** Yang bias-robust estimate is $`r fmt(fe_or_robust$b[1])`$ (95% CI [`r fmt(fe_or_robust$ci.lb[1])`, `r fmt(fe_or_robust$ci.ub[1])`]).

:::

### Side-by-Side Comparison Table

```{r}
#| label: bias_correction_comparison_table
#| echo: false
#| message: false
#| warning: false
# Helper to pull a one-row summary (estimate, SE, CI, p) from any rma/rma.mv/robust object
extract_mean <- function(model, label, method) {
    data.frame(
        Effect_Class = label,
        Method = method,
        Estimate = as.numeric(model$b[1]),
        SE = as.numeric(model$se[1]),
        CI_Lower = as.numeric(model$ci.lb[1]),
        CI_Upper = as.numeric(model$ci.ub[1]),
        p_value = as.numeric(model$pval[1]),
        stringsAsFactors = FALSE
    )
}

bias_correction_summary <- rbind(
    extract_mean(ma_hedges,         "Hedges' g (Continuous)",  "Uncorrected (Multilevel mean)"),
    extract_mean(egger_hedges_v,    "Hedges' g (Continuous)",  "Bias-Corrected — Nakagawa 2022 (current; vi-intercept)"),
    extract_mean(fe_hedges_robust,  "Hedges' g (Continuous)",  "Bias-Robust — Yang 2024 (new; FE + CR2)"),
    extract_mean(ma_or,             "Log Odds Ratio (Binary)", "Uncorrected (Multilevel mean)"),
    extract_mean(egger_or_v,        "Log Odds Ratio (Binary)", "Bias-Corrected — Nakagawa 2022 (current; vi-intercept)"),
    extract_mean(fe_or_robust,      "Log Odds Ratio (Binary)", "Bias-Robust — Yang 2024 (new; FE + CR2)")
)

# Shrinkage of each corrected mean relative to the uncorrected multilevel mean
uncorrected_lookup <- c(
    "Hedges' g (Continuous)"  = ma_hedges$b[1],
    "Log Odds Ratio (Binary)" = ma_or$b[1]
)
bias_correction_summary$Pct_Change_vs_Uncorrected <- ifelse(
    grepl("Uncorrected", bias_correction_summary$Method),
    NA_real_,
    100 * (bias_correction_summary$Estimate - uncorrected_lookup[bias_correction_summary$Effect_Class]) /
        abs(uncorrected_lookup[bias_correction_summary$Effect_Class])
)

DT::datatable(
    bias_correction_summary,
    caption = "Comparison of the uncorrected multilevel mean against the current (Nakagawa 2022) and new (Yang 2024) bias corrections. Estimates for the binary class are on the log-odds-ratio scale.",
    options = list(
        dom = "t",
        pageLength = 6,
        columnDefs = list(list(className = "dt-center", targets = "_all"))
    ),
    rownames = FALSE
) %>%
    DT::formatRound(columns = c("Estimate", "SE", "CI_Lower", "CI_Upper"), digits = 3) %>%
    DT::formatRound(columns = "Pct_Change_vs_Uncorrected", digits = 1) %>%
    DT::formatSignif(columns = "p_value", digits = 3)
```

> [!IMPORTANT]
> **How to read the comparison.** Both corrections should pull the pooled estimate *toward zero* relative to the uncorrected multilevel mean if small-study/selective-reporting bias is present; the `Pct_Change_vs_Uncorrected` column quantifies that shrinkage. The two methods answer slightly different questions. The **Nakagawa** intercept is a *model-based extrapolation* to a perfectly precise study and tends to give wide confidence intervals (it is an out-of-range prediction). The **Yang** bias-robust estimate stays within the observed data, down-weights heterogeneity-driven inflation through fixed-effect weighting, and uses cluster-robust CR2 standard errors that explicitly respect the within-study dependence — making it a more conservative, inference-valid sensitivity analysis. Agreement between the two corrections strengthens confidence in the adjusted conclusion; divergence flags that the corrected mean is sensitive to modelling assumptions.

---

## Methodological Synthesis & Discussion

We compile the results of the small-study effect regressions and time-lag meta-regressions for both effect size classes below:

### Publication Bias Comparison Table

```{r}
#| label: summary_table_creation
#| echo: false
# Dynamically classify the strength of evidence from each diagnostic p-value
classify_bias <- function(p) {
    if (length(p) == 0 || is.na(p)) {
        "Indeterminate (p = NA)"
    } else if (p < 0.0001) {
        "Highly Significant Bias (p < 0.0001)"
    } else if (p < 0.05) {
        paste0("Significant Bias (p = ", fmt(p, 4), ")")
    } else {
        paste0("No Significant Bias (p = ", fmt(p, 4), ")")
    }
}

p_vec <- c(egger_hedges$pval[2], decline_hedges$pval[2], egger_or$pval[2], decline_or$pval[2])

bias_summary <- data.frame(
    Effect_Size_Class = c("Hedges' g (Continuous)", "Hedges' g (Continuous)", "Log Odds Ratio (Binary)", "Log Odds Ratio (Binary)"),
    Diagnostic_Test = c("Egger's Regression (Small-Study)", "Time-Lag (Decline Effect)", "Egger's Regression (Small-Study)", "Time-Lag (Decline Effect)"),
    Moderator = c("Standard Error (se)", "Year_c (centered)", "Standard Error (se)", "Year_c (centered)"),
    Slope_Beta = c(egger_hedges$b[2], decline_hedges$b[2], egger_or$b[2], decline_or$b[2]),
    Std_Error = c(egger_hedges$se[2], decline_hedges$se[2], egger_or$se[2], decline_or$se[2]),
    t_z_value = c(egger_hedges$zval[2], decline_hedges$zval[2], egger_or$zval[2], decline_or$zval[2]),
    p_value = p_vec,
    Evidence_of_Bias = vapply(p_vec, classify_bias, character(1))
)

DT::datatable(
    bias_summary,
    options = list(
        dom = "t",
        columnDefs = list(list(className = "dt-center", targets = "_all"))
    ),
    rownames = FALSE
) %>%
    DT::formatRound(columns = c("Slope_Beta", "Std_Error", "t_z_value"), digits = 4) %>%
    DT::formatSignif(columns = c("p_value"), digits = 4)
```

### Main Takeaways

1. **Small-Study Effects:**
   The Egger's regressions indicate `r sigw(egger_hedges$pval[2])` small-study effects for Hedges' $g$ ($`r pstr(egger_hedges$pval[2])`$) and `r sigw(egger_or$pval[2])` small-study effects for Log Odds Ratio ($`r pstr(egger_or$pval[2])`$). Where present, smaller and less precise trials systematically report larger mate-choice copying effect sizes—a classic signature of publication selection, in which small-sample studies are more likely to be published or reported if they find large, hypothesis-supportive effects.

2. **Decline Effect (Time-Lag Bias) in Binary Outcomes:**
   The publication-year trend for binary outcomes is `r sigw(decline_or$pval[2])` ($\beta = `r fmt(decline_or$b[2])`$, $`r pstr(decline_or$pval[2])`$). `r if(decline_or$pval[2] < 0.05) "Over the last two decades, the reported strength of mate-choice copying (measured by Log Odds Ratios) has decreased, a pattern typical of fields experiencing a novelty effect in which early breakthrough publications report large effect sizes and subsequent rigorous replications reveal more moderate, realistic estimates." else "This indicates the reported strength of mate-choice copying (measured by Log Odds Ratios) has remained relatively stable across publication history."`

3. **Time Trend in Continuous Outcomes:**
   For Hedges' $g$ (continuous), the publication-year trend is `r sigw(decline_hedges$pval[2])` ($\beta = `r fmt(decline_hedges$b[2])`$, $`r pstr(decline_hedges$pval[2])`$). `r if(decline_hedges$pval[2] < 0.05) "This indicates a shift in the reported magnitude of continuous effects over the history of the literature." else "This suggests the continuous reporting scales have been relatively stable in magnitude over the history of the literature."`

4. **Bias-Corrected Means — Current vs. New Method:**
   We compared two correction frameworks on the adjusted pooled mean. For Hedges' $g$, the **current** Nakagawa [-@nakagawaMethodsTestingPublication2022] $v_i$-intercept correction yields $`r fmt(egger_hedges_v$b[1])`$ while the **new** Yang [-@yangRobustPointVariance2024] bias-robust (fixed-effect + CR2) estimate yields $`r fmt(fe_hedges_robust$b[1])`$ (uncorrected multilevel mean: $`r fmt(ma_hedges$b[1])`$). For Log Odds Ratio, the corresponding values are $`r fmt(egger_or_v$b[1])`$ (Nakagawa) and $`r fmt(fe_or_robust$b[1])`$ (Yang), against an uncorrected mean of $`r fmt(ma_or$b[1])`$. The Yang bias-robust estimate is generally the more conservative, inference-valid option because it respects within-study dependence through cluster-robust variance estimation and avoids the wide-interval extrapolation inherent in the regression-intercept approach; where the two corrections agree, the adjusted conclusion is robust to modelling choice.

These diagnostic results suggest that while mate-choice copying is a genuine ecological phenomenon, the overall raw pooled estimates from both continuous and binary literatures may be moderately inflated due to selective reporting and publication biases. Future research should prioritize large-sample pre-registered replication studies to provide unbiased baselines.

---

## Session Info

::: callout-note
```{r}
#| label: session_info
sessionInfo()
```
:::
