---
title: "Publication Bias Diagnostics"
subtitle: "Small-Study Effects and Time-Lag Bias in Mate-Choice Copying (MCC) Meta-Analyses"
author: "Eduardo S. A. Santos & Collaborators"
date: today
bibliography: ../../references.bib
format:
  html:
    toc: true
    toc-depth: 3
    theme: cosmo
    self-contained: true
---

```{r}
#| label: setup
#| include: false
# Ensure pacman is loaded to manage dependencies
if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
pacman::p_load(
    readxl,
    tidyverse,
    metafor,
    orchaRd,
    patchwork,
    DT,
    scales
)
```

## Introduction & Methodology

Publication bias is a critical threat to the validity of meta-analytic syntheses. If studies with statistically non-significant or small effects are systematically omitted from the literature, the pooled meta-analytic summary will be biased upward.

To thoroughly evaluate the presence of publication bias and report-level selection in our Mate-Choice Copying (MCC) dataset, we employ three complementary diagnostic approaches, keeping the exact multilevel modeling logic (e.g., nested random intercepts and taxonomy control) used in our main overall effect calculations:

1. **Visual Inspection (Funnel Plots):** We plot effect size estimates against their precision (inverse of the standard error, $1/\text{SE}$) to check for visual asymmetry around the overall mean estimate.
2. **Egger's Meta-Regression (Small-Study Effects):** We test whether study precision significantly moderates effect size. Following the recommendations of @nakagawaMethodsTestingPublication2022 and Stanley [-@stanleyMetaregressionAnalysisEconomics2012; -@stanleyMetaregressionApproximationsReduce2014], we implement a **two-step approach**:
   * We first fit the Egger regression model using the standard error ($\text{SE} = \sqrt{v_i}$) as the precision proxy.
   * If the intercept of the standard error model is statistically significant, it indicates that the estimate is downwardly biased. In this case, we re-fit the model using the sampling variance ($\text{V} = v_i$) as the moderator to obtain a less biased estimate of the adjusted overall meta-analytic mean.
3. **Decline Effect / Time-Lag Bias:** We test whether effect sizes shrink over time by adding publication year (mean-centered) to our models while simultaneously controlling for study precision.

---

## Data Loading and Alignment

We read, clean, and align the new extractions and original datasets, extracting sample sizes and publication years to construct our precision and time-lag moderators.

```{r}
#| label: load_and_align_datasets
#| message: false
#| warning: false
# Paths to datasets
path_new <- here::here("data", "1_effect_size_calculation_pipeline", "Data Extraction Mate Choice Meta Analysis.xlsx")
path_davies <- here::here("data", "2_original_meta_analyses_datasets", "Davies_et_al_2020_Final_data.xlsx")
path_jones <- here::here("data", "2_original_meta_analyses_datasets", "Jones_DuVal_2019_data.CSV")

# A. Read raw spreadsheets
db_new_raw <- readxl::read_excel(path_new, sheet = "EduardoExtraction")
db_davies_raw <- readxl::read_excel(path_davies, sheet = "Data")
db_jones_raw <- readr::read_csv(path_jones)

# B. Clean new extractions
db_new <- db_new_raw %>%
    filter(is.na(RemoveFromDataset) | RemoveFromDataset != "Yes") %>%
    mutate(
        effectSizeCalculatedHedgesD = as.numeric(effectSizeCalculatedHedgesD),
        effectSizeCalculatedHedgesDVariance = as.numeric(effectSizeCalculatedHedgesDVariance),
        effectSizeCalculatedOddsRatio = as.numeric(effectSizeCalculatedOddsRatio),
        effectSizeCalculatedOddsRatioVariance = as.numeric(effectSizeCalculatedOddsRatioVariance),
        logOddsRatio = log(effectSizeCalculatedOddsRatio),
        logOddsRatioVariance = effectSizeCalculatedOddsRatioVariance,
        identifierStudyYear = as.numeric(identifierStudyYear),
        controlN = as.numeric(controlN),
        treatmentN = as.numeric(treatmentN),
        effectSizesampleSize = as.numeric(effectSizesampleSize)
    )

# ---------------------------------------------------------------------------
# 1. Hedges' g Dataset Alignment (Continuous)
# ---------------------------------------------------------------------------
hedges_new <- db_new %>%
    filter(!is.na(effectSizeCalculatedHedgesD), !is.na(effectSizeCalculatedHedgesDVariance)) %>%
    select(
        yi = effectSizeCalculatedHedgesD,
        vi = effectSizeCalculatedHedgesDVariance,
        identifierStudyId,
        identifierEffectSizeID,
        taxonomySpecies,
        Year = identifierStudyYear,
        controlN,
        treatmentN,
        totalN = effectSizesampleSize
    ) %>%
    mutate(
        identifierStudyId = as.character(identifierStudyId),
        identifierEffectSizeID = as.character(identifierEffectSizeID),
        taxonomySpecies = str_to_lower(word(as.character(taxonomySpecies), -1)),
        Dataset_Source = "New Extraction (Santos)"
    )

hedges_davies <- db_davies_raw %>%
    mutate(
        Hedges_d_directional = as.numeric(Hedges_d_directional),
        Variance = as.numeric(Variance),
        Year = as.numeric(Year),
        N_control = as.numeric(N_control),
        N_treatment = as.numeric(N_treatment),
        Total_n_analysis = as.numeric(Total_n_analysis)
    ) %>%
    filter(!is.na(Hedges_d_directional), !is.na(Variance), Variance > 0) %>%
    select(
        yi = Hedges_d_directional,
        vi = Variance,
        identifierStudyId = Author,
        identifierEffectSizeID = Effect_size_no,
        taxonomySpecies = Species_latin,
        Year,
        controlN = N_control,
        treatmentN = N_treatment,
        totalN = Total_n_analysis
    ) %>%
    mutate(
        identifierStudyId = as.character(identifierStudyId),
        identifierEffectSizeID = as.character(identifierEffectSizeID),
        taxonomySpecies = str_to_lower(word(as.character(taxonomySpecies), -1)),
        Dataset_Source = "Original (Davies 2020)"
    )

db_hedges <- bind_rows(hedges_new, hedges_davies) %>%
    mutate(across(c(identifierStudyId, taxonomySpecies, Dataset_Source), as.factor))

# ---------------------------------------------------------------------------
# 2. Log Odds Ratio Dataset Alignment (Binary)
# ---------------------------------------------------------------------------
or_new <- db_new %>%
    filter(!is.na(logOddsRatio), !is.na(logOddsRatioVariance)) %>%
    select(
        yi = logOddsRatio,
        vi = logOddsRatioVariance,
        identifierStudyId,
        identifierEffectSizeID,
        taxonomySpecies,
        Year = identifierStudyYear
    ) %>%
    mutate(
        identifierStudyId = as.character(identifierStudyId),
        identifierEffectSizeID = as.character(identifierEffectSizeID),
        taxonomySpecies = str_to_lower(word(as.character(taxonomySpecies), -1)),
        Dataset_Source = "New Extraction (Santos)"
    )

or_jones <- db_jones_raw %>%
    filter(!is.na(lnOR), !is.na(VlnOR), VlnOR > 0) %>%
    select(
        yi = lnOR,
        vi = VlnOR,
        identifierStudyId = study,
        identifierEffectSizeID = study_Exp,
        taxonomySpecies = Species,
        Year = year
    ) %>%
    mutate(
        identifierStudyId = as.character(identifierStudyId),
        identifierEffectSizeID = as.character(identifierEffectSizeID),
        taxonomySpecies = str_to_lower(as.character(taxonomySpecies)),
        Dataset_Source = "Original (Jones & DuVal 2019)"
    )

db_or <- bind_rows(or_new, or_jones) %>%
    mutate(across(c(identifierStudyId, taxonomySpecies, Dataset_Source), as.factor))
```

---

## Diagnostics by Effect Size Class

::: {.panel-tabset}

### Hedges' $g$ (Continuous outcomes)

We fit multilevel models on the continuous combined dataset ($k = `r nrow(db_hedges)`$) to run our diagnostics.

#### 1. Baseline Model & Funnel Plot

We run the intercept-only model to establish the global mean effect size and plot precision ($1/\text{SE}$) against the standardized residuals.

```{r}
#| label: hedges_baseline_and_funnel
#| message: false
#| warning: false
#| fig.width: 12
#| fig.height: 6
# Intercept-only model
VCV_hedges <- metafor::vcalc(vi = vi, cluster = identifierStudyId, obs = identifierEffectSizeID, rho = 0.5, data = db_hedges)

ma_hedges <- metafor::rma.mv(
    yi = yi,
    V = VCV_hedges,
    random = list(~ 1 | identifierStudyId, ~ 1 | identifierEffectSizeID, ~ 1 | taxonomySpecies),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = db_hedges
)

summary(ma_hedges)

# Extract BLUPs for random effects
blups <- ranef(ma_hedges)
study_ef <- blups$identifierStudyId$intrcpt[match(db_hedges$identifierStudyId, row.names(blups$identifierStudyId))]
obs_ef <- blups$identifierEffectSizeID$intrcpt[match(db_hedges$identifierEffectSizeID, row.names(blups$identifierEffectSizeID))]
spec_ef <- blups$taxonomySpecies$intrcpt[match(db_hedges$taxonomySpecies, row.names(blups$taxonomySpecies))]

# Handle any missing values by replacing with 0
study_ef[is.na(study_ef)] <- 0
obs_ef[is.na(obs_ef)] <- 0
spec_ef[is.na(spec_ef)] <- 0

# Marginal residuals: observed - fixed-effect prediction
resid_marg <- db_hedges$yi - predict(ma_hedges)[[1]]

# Conditional residuals: observed - fixed - random intercepts (study, obs, species)
resid_cond <- db_hedges$yi - (predict(ma_hedges)[[1]] + study_ef + obs_ef + spec_ef)

# Side-by-side funnel plots of marginal vs. conditional residuals
par(mfrow = c(1, 2))
metafor::funnel(
    x = resid_marg, vi = db_hedges$vi,
    yaxis = "seinv", digits = 2,
    level = c(0.1, 0.05, 0.01),
    shade = c("white", "gray55", "gray75"),
    las = 1, xlab = "Meta-analytic residuals (marginal)",
    legend = FALSE
)
title("Marginal Residual Funnel Plot", adj = 0.5, line = 2.5)

metafor::funnel(
    x = resid_cond, vi = db_hedges$vi,
    yaxis = "seinv", digits = 2,
    level = c(0.1, 0.05, 0.01),
    shade = c("white", "gray55", "gray75"),
    las = 1, xlab = "Meta-analytic residuals (conditional)",
    legend = FALSE
)
title("Conditional Residual Funnel Plot", adj = 0.5, line = 2.5)
```

#### 2. Egger's Meta-Regression (Small-Study Effects)

To perform Egger's regression on Hedges' $g$, we follow the two-step framework recommended by @nakagawaMethodsTestingPublication2022 and Stanley [-@stanleyMetaregressionAnalysisEconomics2012; -@stanleyMetaregressionApproximationsReduce2014]. We use the sampling standard error ($\text{SE} = \sqrt{v_i}$) as our precision proxy. If this moderator is statistically significant ($p < 0.05$), it indicates a significant small-study effect (larger standard errors / smaller sample sizes associated with larger effect magnitudes).

Because standard error model intercepts can be downwardly biased if they are significant, we apply a two-step correction: if the intercept of the standard error model is statistically significant, we re-fit the Egger's regression model using the sampling variance ($\text{V} = \text{vi}$) as the moderator to obtain a less biased estimate of the adjusted overall meta-analytic mean.

```{r}
#| label: hedges_eggers_regression
#| message: false
#| warning: false
# Calculate standard error precision proxy
db_hedges_egger <- db_hedges %>%
    mutate(se = sqrt(vi)) %>%
    filter(!is.na(se))

# Re-estimate VCV on this subset
VCV_hedges_egger <- metafor::vcalc(vi = vi, cluster = identifierStudyId, obs = identifierEffectSizeID, rho = 0.5, data = db_hedges_egger)

# Step 1: Egger regression model with standard error as moderator
egger_hedges <- metafor::rma.mv(
    yi = yi,
    V = VCV_hedges_egger,
    mods = ~ 1 + se,
    random = list(~ 1 | identifierStudyId, ~ 1 | identifierEffectSizeID, ~ 1 | taxonomySpecies),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = db_hedges_egger
)

summary(egger_hedges)

# Step 2: Since the intercept is significant, re-fit using variance (vi)
egger_hedges_v <- metafor::rma.mv(
    yi = yi,
    V = VCV_hedges_egger,
    mods = ~ 1 + vi,
    random = list(~ 1 | identifierStudyId, ~ 1 | identifierEffectSizeID, ~ 1 | taxonomySpecies),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = db_hedges_egger
)

summary(egger_hedges_v)
```

We visualize the Egger regression using a bubble plot. Effect sizes are plotted as circles scaled by precision, with the predicted regression line showing the relationship between sample size and effect size.

```{r}
#| label: hedges_egger_bubble_plot
#| message: false
#| warning: false
#| fig.width: 9
#| fig.height: 6
p_egger_hedges <- orchaRd::bubble_plot(
    egger_hedges,
    mod = "se",
    group = "identifierStudyId",
    xlab = "Standard Error of Hedges' g (se)",
    ylab = "Effect size (Hedges' g)"
)

# Color points by real source
p_egger_hedges$layers[[1]]$data$real_source <- db_hedges_egger$Dataset_Source
p_egger_hedges$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
p_egger_hedges$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

p_egger_hedges <- p_egger_hedges +
    ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
    ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
    ggplot2::ggtitle("Egger's Meta-Regression (Hedges' g)") +
    ggplot2::theme_minimal()

print(p_egger_hedges)
```

> [!NOTE]
> The slope for standard error (`se`) is **strongly positive** ($\beta = `r sprintf("%.2f", egger_hedges$b[2])`$, $SE = `r sprintf("%.2f", egger_hedges$se[2])`$, $t = `r sprintf("%.2f", egger_hedges$zval[2])`$, $`r if(egger_hedges$pval[2] < 0.0001) "p < 0.0001" else paste0("p = ", sprintf("%.2f", egger_hedges$pval[2]))`$), indicating clear evidence of small-study effects. Studies with larger standard errors tend to report larger, more positive Hedges' $g$ values.
> 
> **Two-Step Correction:** Because the intercept of the standard error model is statistically significant ($`r if(egger_hedges$pval[1] < 0.0001) "p < 0.0001" else paste0("p = ", sprintf("%.2f", egger_hedges$pval[1]))`$), the intercept provides a downwardly biased estimate of the adjusted mean. Following the two-step recommendation of @nakagawaMethodsTestingPublication2022 and Stanley [-@stanleyMetaregressionAnalysisEconomics2012; -@stanleyMetaregressionApproximationsReduce2014], we re-fit Egger's regression using the sampling variance (`vi`) as the moderator. In this corrected model, the adjusted overall meta-analytic mean is estimated at $\beta_0 = `r sprintf("%.2f", egger_hedges_v$b[1])`$ ($SE = `r sprintf("%.2f", egger_hedges_v$se[1])`$, $t = `r sprintf("%.2f", egger_hedges_v$zval[1])`$, $`r if(egger_hedges_v$pval[1] < 0.0001) "p < 0.0001" else paste0("p = ", sprintf("%.2f", egger_hedges_v$pval[1]))`$).

#### 3. Decline Effect (Time-Lag Bias)

We analyze the decline effect by evaluating whether the publication year significantly predicts the effect size, while simultaneously controlling for the study sample size precision proxy. We mean-center the year so the intercept represents the effect in an average publication year.

```{r}
#| label: hedges_time_lag
#| message: false
#| warning: false
# Mean-center publication year
db_hedges_decline <- db_hedges_egger %>%
    mutate(Year_c = Year - mean(Year, na.rm = TRUE))

# VCV for decline subset
VCV_hedges_decline <- metafor::vcalc(vi = vi, cluster = identifierStudyId, obs = identifierEffectSizeID, rho = 0.5, data = db_hedges_decline)

# Time-lag decline model controlling for sampling variance (vi)
decline_hedges <- metafor::rma.mv(
    yi = yi,
    V = VCV_hedges_decline,
    mods = ~ 1 + Year_c + vi,
    random = list(~ 1 | identifierStudyId, ~ 1 | identifierEffectSizeID, ~ 1 | taxonomySpecies),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = db_hedges_decline
)

summary(decline_hedges)
```

We construct the bubble plot showing the decline effect over years:

```{r}
#| label: hedges_decline_bubble_plot
#| message: false
#| warning: false
#| fig.width: 9
#| fig.height: 6
p_decline_hedges <- orchaRd::bubble_plot(
    decline_hedges,
    mod = "Year_c",
    group = "identifierStudyId",
    xlab = "Publication Year (centered)",
    ylab = "Effect size (Hedges' g)"
)

# Color points by real source
p_decline_hedges$layers[[1]]$data$real_source <- db_hedges_decline$Dataset_Source
p_decline_hedges$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
p_decline_hedges$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

p_decline_hedges <- p_decline_hedges +
    ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source") +
    ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Davies 2020)" = "#0072B2"), name = "Data Source")

# Relabel centered x-axis to actual calendar years
mean_year_hedges <- mean(db_hedges_egger$Year, na.rm = TRUE)
target_years_hedges <- seq(2000, 2026, by = 5)

p_decline_hedges_relabeled <- p_decline_hedges +
    ggplot2::scale_x_continuous(
        name = "Publication Year",
        breaks = target_years_hedges - mean_year_hedges,
        labels = target_years_hedges
    ) +
    ggplot2::ggtitle("Decline Effect / Time-Lag Bias (Hedges' g)") +
    ggplot2::theme_minimal()

print(p_decline_hedges_relabeled)
```

> [!NOTE]
> The publication year slope is **non-significant** ($\beta = `r sprintf("%.2f", decline_hedges$b[2])`$, $SE = `r sprintf("%.2f", decline_hedges$se[2])`$, $t = `r sprintf("%.2f", decline_hedges$zval[2])`$, $`r if(decline_hedges$pval[2] < 0.0001) "p < 0.0001" else paste0("p = ", sprintf("%.2f", decline_hedges$pval[2]))`$), indicating little evidence of a decline effect over time in continuous outcomes.

#### 4. Unified Diagnostic Dashboard (Hedges' $g$)

We combine the Egger regression and publication year time-lag bubble plots into a single visual panel using `patchwork`.

```{r}
#| label: hedges_unified_dashboard
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 12
dashboard_hedges <- (p_egger_hedges / p_decline_hedges_relabeled) +
    patchwork::plot_annotation(
        title = "Publication Bias Diagnostic Dashboard — Hedges' g",
        tag_levels = "A"
    ) &
    ggplot2::theme(
        plot.title = ggplot2::element_text(size = 14, face = "bold", hjust = 0.5),
        plot.tag = ggplot2::element_text(size = 12, face = "bold")
    )

print(dashboard_hedges)
```

---

### Odds Ratio (Binary outcomes)

We fit multilevel models on the binary combined dataset ($k = `r nrow(db_or)`$) to evaluate bias in Log Odds Ratio ($\ln\text{OR}$) studies.

#### 1. Baseline Model & Funnel Plot

We run the intercept-only model for Odds Ratio and inspect the funnel plot representing log Odds Ratio values plotted against their precision.

```{r}
#| label: or_baseline_and_funnel
#| message: false
#| warning: false
#| fig.width: 12
#| fig.height: 6
# Intercept-only model
VCV_or <- metafor::vcalc(vi = vi, cluster = identifierStudyId, obs = identifierEffectSizeID, rho = 0.5, data = db_or)

ma_or <- metafor::rma.mv(
    yi = yi,
    V = VCV_or,
    random = list(~ 1 | identifierStudyId, ~ 1 | identifierEffectSizeID, ~ 1 | taxonomySpecies),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = db_or
)

summary(ma_or)

# Extract BLUPs for random effects
blups <- ranef(ma_or)
study_ef <- blups$identifierStudyId$intrcpt[match(db_or$identifierStudyId, row.names(blups$identifierStudyId))]
obs_ef <- blups$identifierEffectSizeID$intrcpt[match(db_or$identifierEffectSizeID, row.names(blups$identifierEffectSizeID))]
spec_ef <- blups$taxonomySpecies$intrcpt[match(db_or$taxonomySpecies, row.names(blups$taxonomySpecies))]

# Handle any missing values by replacing with 0
study_ef[is.na(study_ef)] <- 0
obs_ef[is.na(obs_ef)] <- 0
spec_ef[is.na(spec_ef)] <- 0

# Marginal residuals: observed - fixed-effect prediction
resid_marg <- db_or$yi - predict(ma_or)[[1]]

# Conditional residuals: observed - fixed - random intercepts (study, obs, species)
resid_cond <- db_or$yi - (predict(ma_or)[[1]] + study_ef + obs_ef + spec_ef)

# Side-by-side funnel plots of marginal vs. conditional residuals
par(mfrow = c(1, 2))
metafor::funnel(
    x = resid_marg, vi = db_or$vi,
    yaxis = "seinv", digits = 2,
    level = c(0.1, 0.05, 0.01),
    shade = c("white", "gray55", "gray75"),
    las = 1, xlab = "Meta-analytic residuals (marginal)",
    legend = FALSE
)
title("Marginal Residual Funnel Plot", adj = 0.5, line = 2.5)

metafor::funnel(
    x = resid_cond, vi = db_or$vi,
    yaxis = "seinv", digits = 2,
    level = c(0.1, 0.05, 0.01),
    shade = c("white", "gray55", "gray75"),
    las = 1, xlab = "Meta-analytic residuals (conditional)",
    legend = FALSE
)
title("Conditional Residual Funnel Plot", adj = 0.5, line = 2.5)
```

#### 2. Egger's Meta-Regression (Small-Study Effects)

For Odds Ratio studies, we use the standard error ($\text{SE} = \sqrt{v_i}$) as our precision proxy in the two-step framework. If the standard error moderator slope is statistically significant ($p < 0.05$), it represents a significant small-study effect. 

Because the intercept of the standard error model is statistically significant ($p = 0.0495$), we perform Step 2 of the two-step correction, re-fitting the Egger's regression model using the sampling variance ($\text{V} = \text{vi}$) as the moderator to obtain a less biased estimate of the adjusted overall meta-analytic mean.

```{r}
#| label: or_eggers_regression
#| message: false
#| warning: false
# Calculate standard error proxy
db_or_egger <- db_or %>%
    mutate(se = sqrt(vi)) %>%
    filter(!is.na(se))

# Re-estimate VCV on this subset
VCV_or_egger <- metafor::vcalc(vi = vi, cluster = identifierStudyId, obs = identifierEffectSizeID, rho = 0.5, data = db_or_egger)

# Step 1: Egger regression model with standard error as moderator
egger_or <- metafor::rma.mv(
    yi = yi,
    V = VCV_or_egger,
    mods = ~ 1 + se,
    random = list(~ 1 | identifierStudyId, ~ 1 | identifierEffectSizeID, ~ 1 | taxonomySpecies),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = db_or_egger
)

summary(egger_or)

# Step-2: Since the intercept is significant, re-fit using variance (vi)
egger_or_v <- metafor::rma.mv(
    yi = yi,
    V = VCV_or_egger,
    mods = ~ 1 + vi,
    random = list(~ 1 | identifierStudyId, ~ 1 | identifierEffectSizeID, ~ 1 | taxonomySpecies),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = db_or_egger
)

summary(egger_or_v)
```

We visualize the Egger regression for Odds Ratio using a bubble plot.

```{r}
#| label: or_egger_bubble_plot
#| message: false
#| warning: false
#| fig.width: 9
#| fig.height: 6
p_egger_or <- orchaRd::bubble_plot(
    egger_or,
    mod = "se",
    group = "identifierStudyId",
    xlab = "Standard Error of Log Odds Ratio (se)",
    ylab = "Effect size (lnOR)"
)

# Color points by real source
p_egger_or$layers[[1]]$data$real_source <- db_or_egger$Dataset_Source
p_egger_or$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
p_egger_or$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

p_egger_or <- p_egger_or +
    ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal 2019)" = "#0072B2"), name = "Data Source") +
    ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal 2019)" = "#0072B2"), name = "Data Source") +
    ggplot2::ggtitle("Egger's Meta-Regression (Log Odds Ratio)") +
    ggplot2::theme_minimal()

print(p_egger_or)
```

> [!NOTE]
> The slope for the standard error is **strongly positive** ($\beta = `r sprintf("%.2f", egger_or$b[2])`$, $SE = `r sprintf("%.2f", egger_or$se[2])`$, $t = `r sprintf("%.2f", egger_or$zval[2])`$, $`r if(egger_or$pval[2] < 0.0001) "p < 0.0001" else paste0("p = ", sprintf("%.2f", egger_or$pval[2]))`$), indicating pronounced small-study effects in binary mate-choice copying outcomes. Studies with higher standard errors (smaller precision) report  larger log odds ratios.
> 
> **Two-Step Correction:** Because the intercept in the standard error model is statistically significant ($`r if(egger_or$pval[1] < 0.0001) "p < 0.0001" else paste0("p = ", sprintf("%.2f", egger_or$pval[1]))`$), we re-fit the model using the sampling variance (`vi`) as the moderator to obtain a less biased estimate of the adjusted mean. In this corrected model, the adjusted overall log odds ratio is non-significant at $\beta_0 = `r sprintf("%.2f", egger_or_v$b[1])`$ ($SE = `r sprintf("%.2f", egger_or_v$se[1])`$, $t = `r sprintf("%.2f", egger_or_v$zval[1])`$, $`r if(egger_or_v$pval[1] < 0.0001) "p < 0.0001" else paste0("p = ", sprintf("%.2f", egger_or_v$pval[1]))`$).

#### 3. Decline Effect (Time-Lag Bias)

We test for a decline effect in the binary outcomes by regressing $\ln\text{OR}$ on centered publication year while simultaneously controlling for standard errors.

```{r}
#| label: or_time_lag
#| message: false
#| warning: false
# Mean-center publication year
db_or_decline <- db_or_egger %>%
    mutate(Year_c = Year - mean(Year, na.rm = TRUE))

# VCV for decline subset
VCV_or_decline <- metafor::vcalc(vi = vi, cluster = identifierStudyId, obs = identifierEffectSizeID, rho = 0.5, data = db_or_decline)

# Time-lag decline model controlling for sampling variance (vi)
decline_or <- metafor::rma.mv(
    yi = yi,
    V = VCV_or_decline,
    mods = ~ 1 + Year_c + vi,
    random = list(~ 1 | identifierStudyId, ~ 1 | identifierEffectSizeID, ~ 1 | taxonomySpecies),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = db_or_decline
)

summary(decline_or)
```

We construct the bubble plot showing the decline effect for Odds Ratios:

```{r}
#| label: or_decline_bubble_plot
#| message: false
#| warning: false
#| fig.width: 9
#| fig.height: 6
p_decline_or <- orchaRd::bubble_plot(
    decline_or,
    mod = "Year_c",
    group = "identifierStudyId",
    xlab = "Publication Year (centered)",
    ylab = "Effect size (lnOR)"
)

# Color points by real source
p_decline_or$layers[[1]]$data$real_source <- db_or_decline$Dataset_Source
p_decline_or$layers[[1]]$mapping$colour <- ggplot2::aes(colour = real_source)$colour
p_decline_or$layers[[1]]$mapping$fill <- ggplot2::aes(fill = real_source)$fill

p_decline_or <- p_decline_or +
    ggplot2::scale_colour_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal 2019)" = "#0072B2"), name = "Data Source") +
    ggplot2::scale_fill_manual(values = c("New Extraction (Santos)" = "#D55E00", "Original (Jones & DuVal 2019)" = "#0072B2"), name = "Data Source")

# Relabel centered x-axis to actual calendar years
mean_year_or <- mean(db_or_egger$Year, na.rm = TRUE)
target_years_or <- seq(2000, 2026, by = 5)

p_decline_or_relabeled <- p_decline_or +
    ggplot2::scale_x_continuous(
        name = "Publication Year",
        breaks = target_years_or - mean_year_or,
        labels = target_years_or
    ) +
    ggplot2::ggtitle("Decline Effect / Time-Lag Bias (Log Odds Ratio)") +
    ggplot2::theme_minimal()

print(p_decline_or_relabeled)
```

> [!NOTE]
> The publication year slope is **statistically significant** ($\beta = `r sprintf("%.2f", decline_or$b[2])`$, $SE = `r sprintf("%.2f", decline_or$se[2])`$, $t = `r sprintf("%.2f", decline_or$zval[2])`$, $`r if(decline_or$pval[2] < 0.0001) "p < 0.0001" else paste0("p = ", sprintf("%.2f", decline_or$pval[2]))`$). This provides evidence of a **decline effect (Time-Lag Bias)** in binary outcomes—the reported strength of mate-choice copying has significantly decreased over the course of academic publication history.

#### 4. Unified Diagnostic Dashboard (Log Odds Ratio)

We combine the Egger regression and publication year time-lag bubble plots into a single visual panel using `patchwork`.

```{r}
#| label: or_unified_dashboard
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 12
dashboard_or <- (p_egger_or / p_decline_or_relabeled) +
    patchwork::plot_annotation(
        title = "Publication Bias Diagnostic Dashboard — Log Odds Ratio",
        tag_levels = "A"
    ) &
    ggplot2::theme(
        plot.title = ggplot2::element_text(size = 14, face = "bold", hjust = 0.5),
        plot.tag = ggplot2::element_text(size = 12, face = "bold")
    )

print(dashboard_or)
```

:::

---

## Methodological Synthesis & Discussion

We compile the results of the small-study effect regressions and time-lag meta-regressions for both effect size classes below:

### Publication Bias Comparison Table

```{r}
#| label: summary_table_creation
#| echo: false
bias_summary <- data.frame(
    Effect_Size_Class = c("Hedges' g (Continuous)", "Hedges' g (Continuous)", "Log Odds Ratio (Binary)", "Log Odds Ratio (Binary)"),
    Diagnostic_Test = c("Egger's Regression (Small-Study)", "Time-Lag (Decline Effect)", "Egger's Regression (Small-Study)", "Time-Lag (Decline Effect)"),
    Moderator = c("Standard Error (se)", "Year_c (centered)", "Standard Error (se)", "Year_c (centered)"),
    Slope_Beta = c(egger_hedges$b[2], decline_hedges$b[2], egger_or$b[2], decline_or$b[2]),
    Std_Error = c(egger_hedges$se[2], decline_hedges$se[2], egger_or$se[2], decline_or$se[2]),
    t_z_value = c(egger_hedges$zval[2], decline_hedges$zval[2], egger_or$zval[2], decline_or$zval[2]),
    p_value = c(egger_hedges$pval[2], decline_hedges$pval[2], egger_or$pval[2], decline_or$pval[2]),
    Evidence_of_Bias = c("Highly Significant Bias (p < 0.0001)", "No Significant Bias (p = 0.8996)", "Highly Significant Bias (p < 0.0001)", "Significant Bias (p = 0.0104)")
)

DT::datatable(
    bias_summary,
    options = list(
        dom = "t",
        columnDefs = list(list(className = "dt-center", targets = "_all"))
    ),
    rownames = FALSE
) %>%
    DT::formatRound(columns = c("Slope_Beta", "Std_Error", "t_z_value"), digits = 4) %>%
    DT::formatSignif(columns = c("p_value"), digits = 4)
```

### Main Takeaways

1. **Small-Study Effects Are Pervasive:** 
   Both Hedges' $g$ (continuous) and Log Odds Ratio (binary) datasets show statistically significant evidence of small-study effects. Smaller, less precise trials systematically report larger mate-choice copying effect sizes. This asymmetrical pattern is a classic signature of publication selection, where small-sample studies are more likely to be published or reported if they find large, hypothesis-supportive effects.
   
2. **Decline Effect (Time-Lag Bias) in Binary Outcomes:**
   We found robust evidence of a decline effect in binary mate-choice copying studies ($\beta = `r sprintf("%.2f", decline_or$b[2])`$, $`r if(decline_or$pval[2] < 0.0001) "p < 0.0001" else paste0("p = ", sprintf("%.2f", decline_or$pval[2]))`$). Over the last two decades, the strength of reported mate-choice copying (as measured by Log Odds Ratios) has significantly decreased. This is typical in fields experiencing a "novelty effect," where initial breakthrough publications show large effect sizes, and subsequent rigorous replications or expanded testing reveal more moderate, realistic estimates.
   
3. **Continuous Outcomes Show More Stable Trends Over Time:**
   For Hedges' $g$ (continuous), although small-study effects are present, the publication year trend is non-significant ($\beta = `r sprintf("%.2f", decline_hedges$b[2])`$, $`r if(decline_hedges$pval[2] < 0.0001) "p < 0.0001" else paste0("p = ", sprintf("%.2f", decline_hedges$pval[2]))`$). This suggests that the continuous reporting scales have been relatively more stable in magnitude over the history of the literature.

These diagnostic results suggest that while mate-choice copying is a genuine ecological phenomenon, the overall raw pooled estimates from both continuous and binary literatures may be moderately inflated due to selective reporting and publication biases. Future research should prioritize large-sample pre-registered replication studies to provide unbiased baselines.

---

## Session Info

::: callout-note
```{r}
#| label: session_info
sessionInfo()
```
:::
