---
title: "Sensitivity Analysis: Contentious Effect Sizes"
subtitle: "How robust is the combined overall effect to the studies flagged by the cross-dataset validation?"
author: "Eduardo S. A. Santos & Collaborators"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    theme: cosmo
    code-fold: true
    code-summary: "Show R code"
bibliography: ../../references.bib
---

```{r}
#| label: setup
#| include: false
if (is.null(getOption("repos")) || getOption("repos")["CRAN"] %in% c(NA, "@CRAN@")) {
    options(repos = c(CRAN = "https://cloud.r-project.org"))
}
if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
if (!requireNamespace("orchaRd", quietly = TRUE)) pacman::p_load_gh("daniel1noble/orchaRd")
pacman::p_load(here, readxl, tidyverse, metafor, orchaRd, patchwork, DT, scales)

# Number formatter (mirrors the helper used in the other chapters)
fmt <- function(x, digits = 2) {
    res <- sprintf(paste0("%.", digits, "f"), x)
    zero <- paste0("-0.", paste0(rep("0", digits), collapse = ""))
    res[res == zero] <- paste0("0.", paste0(rep("0", digits), collapse = ""))
    res
}
pstr <- function(p) {
    if (length(p) == 0 || is.na(p)) "p = NA" else if (p < 0.0001) "p < 0.0001" else paste0("p = ", fmt(p, 3))
}
```

## Why this chapter exists

The [Cross-Dataset Extraction Validation](../3_original_dataset_crosscheck/extraction_validation.qmd)
chapter compared the two original datasets that feed our combined synthesis —
Davies et al. [-@daviesMetaanalysisFactorsInfluencing2020] (continuous, Hedges' *d*)
and Jones & DuVal [-@jonesMechanismsSocialInfluence2019] (binary, log odds ratio) —
and returned to the source PDFs for the **seven studies whose effect sizes disagreed
by more than 0.5 Hedges-*d* units**. That re-extraction reached two conclusions that
matter for the headline result:

1. On the **continuous (Davies) side**, the disagreements trace to a handful of
   identifiable extraction problems (a wrong outcome construct, two `escalc`
   computational artifacts, a sign error, and one defensible pool-vs-split). Several
   of these inflate the Davies effect size.
2. On the **binary (Jones) side**, the *point estimates* are sound, but a few rows are
   over-counted (one spurious value and one duplicated contrast), and the variances of
   the single-group "vs-chance" studies are inflated by a phantom 50/50 control.

A natural question follows: **does the combined overall effect depend on these
contentious effect sizes?** This chapter answers it directly. We re-fit the combined
multilevel models — using the *same* model structure as the
[Overall Effect](overall_effect.qmd) chapter — under three versions of the data and
compare the pooled estimates.

* **Main** — the data exactly as analysed in the overall-effect chapter (this should
  reproduce the headline combined estimate).
* **Leave-out** — every effect size from the seven contentious studies is removed. This
  is the conservative bound: it asks what the result would be if we simply did not trust
  those studies at all.
* **Corrected** — the contentious effect sizes are replaced with the design-faithful
  values re-extracted from the source PDFs. This is the best-estimate version: it keeps
  the studies but fixes the specific problems the validation identified.

::: {.callout-note}
## Scope
The seven contentious studies (Applebaum & Cruz 2000, Briggs et al. 1996, Dugatkin 1992,
Dugatkin & Godin 1993, Fowler-Finn et al. 2015, Gierszewski et al. 2018, Howard et al.
1998) are all pre-2019 and were **not** independently re-extracted in our new (Santos)
dataset. They therefore enter the combined data **only** through the Davies (continuous)
and Jones (binary) datasets, so every edit below acts on legacy rows only — the new
extraction is untouched.
:::

---

## Data loading and alignment

We rebuild the two combined datasets (`db_hedges`, `db_or`) with the identical loading,
species-name reconciliation, and column-alignment logic used in the overall-effect
chapter, so the **Main** models here are the same models reported there.

```{r}
#| label: load_and_align_datasets
#| message: false
#| warning: false
# Paths to legacy and new datasets
path_new    <- here::here("data", "1_effect_size_calculation_pipeline", "Data Extraction Mate Choice Meta Analysis.xlsx")
path_davies <- here::here("data", "original_meta_analyses_datasets", "Davies_et_al_2020_Final_data.xlsx")
path_jones  <- here::here("data", "original_meta_analyses_datasets", "Jones_DuVal_2019_data.CSV")

db_new_raw    <- readxl::read_excel(path_new, sheet = "EduardoExtraction")
db_davies_raw <- readxl::read_excel(path_davies, sheet = "Data")
db_jones_raw  <- readr::read_csv(path_jones, show_col_types = FALSE)

# Species-name reconciliation (matches build_phylogeny.R / overall_effect.qmd)
recode_new    <- c("Taeniopygia guttata castanotis" = "Taeniopygia guttata")
recode_davies <- c("Schizocosa" = "Schizocosa ocreata")
jones_binomial <- tibble::tribble(
    ~family,            ~epithet,        ~species_phylo,
    "Adrianichthyidae", "latipes",       "Oryzias latipes",
    "Blenniidae",       "nitidus",       "Rhabdoblennius nitidus",
    "Gasterosteidae",   "aculeatus",     "Gasterosteus aculeatus",
    "Gobiidae",         "minutu",        "Pomatoschistus minutus",
    "Percidae",         "olmstedi",      "Etheostoma olmstedi",
    "Poeciliidae",      "latipinna",     "Poecilia latipinna",
    "Poeciliidae",      "nigrofasciata", "Limia nigrofasciata",
    "Poeciliidae",      "perugiae",      "Limia perugiae",
    "Poeciliidae",      "reticulata",    "Poecilia reticulata",
    "Pomacentridae",    "leucogaster",   "Amblyglyphidodon leucogaster",
    "Lycosidae",        "ocreata",       "Schizocosa ocreata",
    "Estrildidae",      "guttata",       "Taeniopygia guttata",
    "Icteridae",        "ater",          "Molothrus ater",
    "Muscicapidae",     "hypoleuca",     "Ficedula hypoleuca",
    "Drosophilidae",    "melanogaster",  "Drosophila melanogaster",
    "Drosophilidae",    "serrata",       "Drosophila serrata",
    "Ocypodidae",       "mjoebergi",     "Uca mjoebergi"
)

read_phylo <- function(f) {
    p <- here::here("outputs", "2_phylogeny", f)
    if (file.exists(p)) readRDS(p) else NULL
}
phylo_cor_hedges <- read_phylo("phylo_cor_hedges_model.rds")
phylo_cor_or     <- read_phylo("phylo_cor_or_model.rds")

# New extraction (drop rows flagged for removal)
db_new <- db_new_raw %>%
    filter(is.na(RemoveFromDataset) | RemoveFromDataset != "Yes") %>%
    mutate(
        effectSizeCalculatedHedgesD = as.numeric(effectSizeCalculatedHedgesD),
        effectSizeCalculatedHedgesDVariance = as.numeric(effectSizeCalculatedHedgesDVariance),
        effectSizeCalculatedOddsRatio = as.numeric(effectSizeCalculatedOddsRatio),
        effectSizeCalculatedOddsRatioVariance = as.numeric(effectSizeCalculatedOddsRatioVariance),
        logOddsRatio = log(effectSizeCalculatedOddsRatio),
        logOddsRatioVariance = effectSizeCalculatedOddsRatioVariance
    )

# --- Hedges' g (continuous) ---------------------------------------------------
hedges_new <- db_new %>%
    filter(!is.na(effectSizeCalculatedHedgesD), !is.na(effectSizeCalculatedHedgesDVariance)) %>%
    transmute(
        yi = effectSizeCalculatedHedgesD,
        vi = effectSizeCalculatedHedgesDVariance,
        identifierStudyId = as.character(identifierStudyId),
        identifierEffectSizeID = as.character(identifierEffectSizeID),
        species_phylo = dplyr::recode(str_squish(as.character(taxonomySpecies)), !!!recode_new),
        taxonomySpecies = str_to_lower(word(as.character(taxonomySpecies), -1)),
        Dataset_Source = "New Extraction (Santos)"
    )

hedges_davies <- db_davies_raw %>%
    mutate(Hedges_d_directional = as.numeric(Hedges_d_directional), Variance = as.numeric(Variance)) %>%
    filter(!is.na(Hedges_d_directional), !is.na(Variance), Variance > 0) %>%
    transmute(
        yi = Hedges_d_directional,
        vi = Variance,
        identifierStudyId = as.character(Author),
        identifierEffectSizeID = as.character(Effect_size_no),
        species_phylo = dplyr::recode(str_squish(as.character(Species_latin)), !!!recode_davies),
        taxonomySpecies = str_to_lower(word(as.character(Species_latin), -1)),
        Dataset_Source = "Original (Davies 2020)"
    )

db_hedges <- bind_rows(hedges_new, hedges_davies)

# --- Log odds ratio (binary) --------------------------------------------------
or_new <- db_new %>%
    filter(!is.na(logOddsRatio), !is.na(logOddsRatioVariance)) %>%
    transmute(
        yi = logOddsRatio,
        vi = logOddsRatioVariance,
        identifierStudyId = as.character(identifierStudyId),
        identifierEffectSizeID = as.character(identifierEffectSizeID),
        species_phylo = dplyr::recode(str_squish(as.character(taxonomySpecies)), !!!recode_new),
        taxonomySpecies = str_to_lower(word(as.character(taxonomySpecies), -1)),
        Dataset_Source = "New Extraction (Santos)"
    )

or_jones <- db_jones_raw %>%
    filter(!is.na(lnOR), !is.na(VlnOR), VlnOR > 0) %>%
    mutate(family = str_squish(as.character(Family)), epithet = str_squish(as.character(Species))) %>%
    left_join(jones_binomial, by = c("family", "epithet")) %>%
    transmute(
        yi = lnOR,
        vi = VlnOR,
        identifierStudyId = as.character(study),
        identifierEffectSizeID = as.character(study_Exp),
        species_phylo = species_phylo,
        taxonomySpecies = str_to_lower(as.character(Species)),
        Dataset_Source = "Original (Jones & DuVal 2019)"
    )

db_or <- bind_rows(or_new, or_jones)
```

---

## Identifying the contentious effect sizes

We read the re-extraction file produced by the validation chapter. Each of its rows is
one effect size that we returned to the source PDF for, and it carries three things side
by side: the **re-extracted** value (independent, design-faithful), the **Jones** value,
and the **Davies** value. From this file alone we can locate every row that needs editing
in the two combined datasets — by `identifierEffectSizeID`, never by guesswork.

```{r}
#| label: load_reextraction
#| message: false
#| warning: false
reext_path <- here::here("outputs", "3_original_dataset_crosscheck", "reextraction_flagged_7studies.csv")
reext <- readr::read_csv(reext_path, show_col_types = FALSE)

# Conversion constant: lnOR -> Hedges' d (logistic-to-normal; Borenstein et al. 2009).
# Matches the validation chapter; the small-sample (J) correction is ~1 at these
# sample sizes and is omitted so the numbers agree with that chapter.
LNOR_TO_D  <- sqrt(3) / pi          # ~ 0.5513
VAR_LNOR_TO_VAR_D <- 3 / pi^2       # = LNOR_TO_D^2 ~ 0.3040

reext_map <- reext %>%
    transmute(
        study          = identifierStudyId,
        # Jones effect-size code in the dataset has no slash (e.g. "APP2000THE/1" -> "APP2000THE1")
        jones_es_id    = str_remove(compare_jones_code, "/"),
        # Davies effect-size number, parsed from e.g. "Davies study 3 (3.1)" -> "3.1"
        davies_es_id   = str_match(compare_davies_es, "\\(([0-9]+\\.[0-9]+)")[, 2],
        davies_study_no = str_match(compare_davies_es, "study ([0-9]+)")[, 2],
        species_phylo  = str_squish(as.character(taxonomySpecies)),
        # Re-extracted values
        reext_lnOR     = log(effectSizeCalculatedEstimate),
        reext_var_lnOR = effectSizeCalculatedVariance,
        reext_d        = log(effectSizeCalculatedEstimate) * LNOR_TO_D,
        reext_var_d    = effectSizeCalculatedVariance * VAR_LNOR_TO_VAR_D,
        # Prior values (for the audit table)
        jones_lnOR     = compare_jones_lnOR,
        jones_var      = compare_jones_VlnOR,
        davies_d       = compare_davies_hedges_d,
        davies_var     = compare_davies_variance
    )

# The set of contentious studies, expressed in each dataset's own identifiers.
davies_studies_flagged <- sort(unique(as.integer(reext_map$davies_study_no)))   # Davies study numbers
jones_papers_flagged   <- sort(unique(str_extract(reext_map$jones_es_id, "^[A-Za-z]+[0-9]{4}")))  # Jones paper prefixes

# Jones rows the validation identified as over-counts to be dropped entirely:
#  * DUG1992REV1  — spurious third value (reproduces from no experiment pair)
#  * GIE2018MAT1/2 — duplicate of the GIE2018USI contrast (over-count)
jones_drop_overcount <- c("DUG1992REV1", "GIE2018MAT1", "GIE2018MAT2")
```

### The edit map (auditable)

Every change made in the **Corrected** version is listed below. Continuous edits act on
Davies rows; binary edits act on Jones rows. Two structural cases are highlighted:
Fowler-Finn (Davies pooled one effect; the re-extraction splits it into two arms) and
Applebaum (Davies extracted an affiliation-time construct as several effect sizes; only
the reversal-frequency effect is retained).

```{r}
#| label: edit_map_table
#| echo: false
audit <- reext_map %>%
    transmute(
        Study = gsub("_", " ", study),
        `Jones ES` = jones_es_id,
        `Davies ES` = davies_es_id,
        `Re-ext lnOR` = round(reext_lnOR, 3),
        `Re-ext Var(lnOR)` = round(reext_var_lnOR, 3),
        `Jones lnOR` = round(jones_lnOR, 3),
        `Jones Var` = round(jones_var, 3),
        `Re-ext d` = round(reext_d, 3),
        `Davies d` = round(davies_d, 3)
    )
DT::datatable(
    audit, rownames = FALSE,
    caption = "Re-extracted (design-faithful) values vs the original Jones and Davies values, for every contentious effect size. These are the substitutions applied in the Corrected analysis.",
    options = list(dom = "tp", pageLength = 12, scrollX = TRUE,
                   columnDefs = list(list(className = "dt-center", targets = "_all")))
)
```

```{r}
#| label: contentious_summary
#| echo: false
tibble(
    Item = c("Contentious studies (papers)",
             "Re-extracted effect sizes",
             "Davies study numbers affected (continuous)",
             "Jones papers affected (binary)",
             "Jones over-count rows dropped (binary)"),
    Value = c(length(unique(reext_map$study)),
              nrow(reext_map),
              paste(davies_studies_flagged, collapse = ", "),
              paste(jones_papers_flagged, collapse = ", "),
              paste(jones_drop_overcount, collapse = ", "))
) %>% knitr::kable()
```

---

## Building the three dataset versions

```{r}
#| label: build_versions
#| message: false
#| warning: false
# ---------------------------------------------------------------------------
# Helper: integer (study) part of a Davies effect-size id ("27.1" -> 27)
# ---------------------------------------------------------------------------
davies_study_int <- function(es_id) as.integer(str_extract(as.character(es_id), "^[0-9]+"))
jones_paper      <- function(es_id) str_extract(as.character(es_id), "^[A-Za-z]+[0-9]{4}")

# ===========================================================================
# CONTINUOUS (Hedges' g) — Davies side
# ===========================================================================
is_davies <- db_hedges$Dataset_Source == "Original (Davies 2020)"

# (1) LEAVE-OUT: drop all Davies rows from the contentious studies.
hedges_leaveout <- db_hedges %>%
    filter(!(Dataset_Source == "Original (Davies 2020)" &
             davies_study_int(identifierEffectSizeID) %in% davies_studies_flagged))

# (2) CORRECTED: remove all Davies rows from the contentious studies, then add the
#     re-extracted replacements (one row per re-extracted effect size). For Fowler-Finn
#     this turns Davies' single pooled effect into two arm-level effects; for Applebaum
#     it keeps only the reversal-frequency effect (the 3.2-3.4 affiliation-time rows are
#     not reinstated). identifierStudyId is set to the re-extraction's paper id.
hedges_reext_rows <- reext_map %>%
    transmute(
        yi = reext_d,
        vi = reext_var_d,
        identifierStudyId = study,
        identifierEffectSizeID = paste0("reext_", jones_es_id),
        species_phylo = dplyr::recode(species_phylo, !!!recode_davies),
        taxonomySpecies = str_to_lower(word(species_phylo, -1)),
        Dataset_Source = "Original (Davies 2020)"
    )

hedges_corrected <- bind_rows(hedges_leaveout, hedges_reext_rows)

# ===========================================================================
# BINARY (log OR) — Jones side
# ===========================================================================
# (1) LEAVE-OUT: drop all Jones rows from the contentious papers.
or_leaveout <- db_or %>%
    filter(!(Dataset_Source == "Original (Jones & DuVal 2019)" &
             jones_paper(identifierEffectSizeID) %in% jones_papers_flagged))

# (2) CORRECTED: (a) drop the over-count rows; (b) replace lnOR & variance of the
#     re-extracted codes with the design-faithful values; leave all other Jones rows
#     (e.g. the legitimate, un-re-extracted APP2000THE2 baseline contrast) untouched.
or_replacements <- reext_map %>% select(identifierEffectSizeID = jones_es_id, reext_lnOR, reext_var_lnOR)

or_corrected <- db_or %>%
    filter(!identifierEffectSizeID %in% jones_drop_overcount) %>%
    left_join(or_replacements, by = "identifierEffectSizeID") %>%
    mutate(
        yi = if_else(!is.na(reext_lnOR), reext_lnOR, yi),
        vi = if_else(!is.na(reext_var_lnOR), reext_var_lnOR, vi)
    ) %>%
    select(-reext_lnOR, -reext_var_lnOR)

# Quick integrity report (how many rows each edit touched)
edit_report <- tibble(
    Version = c("Hedges' g: Main", "Hedges' g: Leave-out", "Hedges' g: Corrected",
                "Log OR: Main", "Log OR: Leave-out", "Log OR: Corrected"),
    `Effect sizes (k)` = c(nrow(db_hedges), nrow(hedges_leaveout), nrow(hedges_corrected),
                           nrow(db_or), nrow(or_leaveout), nrow(or_corrected))
)
```

```{r}
#| label: edit_report_table
#| echo: false
knitr::kable(edit_report, caption = "Number of effect sizes in each dataset version.")
```

---

## Re-fitting the combined models

We use the **same** multilevel model as the overall-effect chapter: `rma.mv` with a
variance–covariance matrix (`vcalc`, within-study $\rho = 0.5$), random intercepts for
study, effect size, and species, and a phylogenetic random effect supplied with the
species correlation matrix (it falls back to the non-phylogenetic structure if a species
is not in the matrix).

```{r}
#| label: fit_function
#| message: false
#| warning: false
fit_overall_mv <- function(dat, cluster_var = "identifierStudyId",
                           obs_var = "identifierEffectSizeID", rho = 0.5,
                           phylo_cor = NULL, phylo_var = "species_phylo") {
    dat2 <- dat %>% filter(!is.na(yi), !is.na(vi), vi > 0)
    if (nrow(dat2) == 0) return(NULL)

    VCV <- metafor::vcalc(vi = dat2$vi, cluster = dat2[[cluster_var]], obs = dat2[[obs_var]],
                          rho = rho, data = dat2)

    random_list <- list(~ 1 | identifierStudyId, ~ 1 | identifierEffectSizeID, ~ 1 | taxonomySpecies)
    R_arg <- NULL
    if (!is.null(phylo_cor) && phylo_var %in% colnames(dat2)) {
        sp <- unique(as.character(dat2[[phylo_var]]))
        if (length(setdiff(sp, rownames(phylo_cor))) == 0) {
            dat2[[phylo_var]] <- factor(as.character(dat2[[phylo_var]]))
            Csub <- phylo_cor[sp, sp, drop = FALSE]
            random_list <- c(random_list, list(stats::as.formula(paste0("~ 1 | ", phylo_var))))
            R_arg <- stats::setNames(list(Csub), phylo_var)
        }
    }
    mod <- metafor::rma.mv(yi = dat2$yi, V = VCV, random = random_list, R = R_arg,
                           test = "t", method = "REML", sparse = TRUE, data = dat2)
    list(data = dat2, model = mod)
}

# Fit all six models
fits <- list(
    hedges_main      = fit_overall_mv(db_hedges,        phylo_cor = phylo_cor_hedges),
    hedges_leaveout  = fit_overall_mv(hedges_leaveout,  phylo_cor = phylo_cor_hedges),
    hedges_corrected = fit_overall_mv(hedges_corrected, phylo_cor = phylo_cor_hedges),
    or_main          = fit_overall_mv(db_or,            phylo_cor = phylo_cor_or),
    or_leaveout      = fit_overall_mv(or_leaveout,      phylo_cor = phylo_cor_or),
    or_corrected     = fit_overall_mv(or_corrected,     phylo_cor = phylo_cor_or)
)
```

```{r}
#| label: summarise_fits
#| message: false
#| warning: false
# Pull a tidy one-row summary from each fitted model
summarise_fit <- function(fit, class_lab, version_lab) {
    if (is.null(fit)) return(NULL)
    m <- fit$model
    i2 <- tryCatch(sum(orchaRd::i2_ml(m)["I2_Total"]), error = function(e) NA_real_)
    tibble(
        Class = class_lab, Version = version_lab,
        k = m$k, Studies = length(unique(fit$data$identifierStudyId)),
        Estimate = as.numeric(m$b[1]), SE = m$se[1],
        CI_low = m$ci.lb[1], CI_high = m$ci.ub[1],
        t = m$zval[1], p = m$pval[1], I2 = i2
    )
}

res <- bind_rows(
    summarise_fit(fits$hedges_main,      "Hedges' g",   "Main"),
    summarise_fit(fits$hedges_leaveout,  "Hedges' g",   "Leave-out"),
    summarise_fit(fits$hedges_corrected, "Hedges' g",   "Corrected"),
    summarise_fit(fits$or_main,          "Log OR",      "Main"),
    summarise_fit(fits$or_leaveout,      "Log OR",      "Leave-out"),
    summarise_fit(fits$or_corrected,     "Log OR",      "Corrected")
) %>%
    mutate(Version = factor(Version, levels = c("Main", "Leave-out", "Corrected")))
```

---

## Results

### Comparison table

For the log odds ratio, the estimate is also shown back-transformed to the odds-ratio
scale (`exp`). `k` is the number of effect sizes and `Studies` the number of study
clusters in each version.

```{r}
#| label: results_table
#| echo: false
res_disp <- res %>%
    mutate(
        `Estimate [95% CI]` = sprintf("%s [%s, %s]", fmt(Estimate, 3), fmt(CI_low, 3), fmt(CI_high, 3)),
        `OR [95% CI]` = if_else(Class == "Log OR",
                                sprintf("%s [%s, %s]", fmt(exp(Estimate), 2), fmt(exp(CI_low), 2), fmt(exp(CI_high), 2)),
                                NA_character_),
        SE = fmt(SE, 3),
        p = sapply(p, pstr),
        `I² (%)` = fmt(I2, 1)
    ) %>%
    select(Class, Version, k, Studies, `Estimate [95% CI]`, `OR [95% CI]`, SE, p, `I² (%)`)

DT::datatable(
    res_disp, rownames = FALSE,
    caption = "Combined overall effect under the three dataset versions, for both effect-size classes.",
    options = list(dom = "t", pageLength = 6,
                   columnDefs = list(list(className = "dt-center", targets = "_all")))
)
```

### Comparison plot

We display each version as its own orchard trunk — the same plot format used elsewhere in
this book. Each row is one model version (Main, Leave-out, Corrected); the bubbles are that
version's raw effect sizes (scaled by precision), the central marker is the pooled mean with
its 95% confidence interval, and the thin whisker is the 95% prediction interval. The
`Mean [95% CI]` and `k (n)` annotations sit beside each trunk.

Because the seven contentious studies are only a small fraction of the combined data, the
three bubble clouds look broadly similar — that near-overlap is itself the robustness
result. To make the edits legible, **bubbles from the seven contentious studies are coloured
orange and all other effect sizes grey.** You can then see the contentious points present in
**Main**, *absent* in **Leave-out**, and *repositioned* (and two fewer, after the over-count
drops) in **Corrected**.

```{r}
#| label: orchard_helpers
#| message: false
#| warning: false
# Flag, for a fitted model's data, which effect sizes come from the seven contentious
# studies. For Hedges' g these are the flagged Davies rows plus the re-extracted
# replacements; for the log OR they are the flagged Jones rows. The flag is computed in
# the model's own row order, which orchaRd::mod_results preserves.
contentious_flag <- function(fit, class) {
    d   <- fit$data
    id  <- as.character(d$identifierEffectSizeID)
    src <- as.character(d$Dataset_Source)
    if (class == "hedges") {
        is_reext  <- grepl("^reext_", id)
        davies_no <- suppressWarnings(as.integer(str_extract(id, "^[0-9]+")))
        is_flag   <- src == "Original (Davies 2020)" & davies_no %in% davies_studies_flagged
        flag <- is_reext | (is_flag & !is.na(is_flag))
    } else {
        prefix <- str_extract(id, "^[A-Za-z]+[0-9]{4}")
        flag <- src == "Original (Jones & DuVal 2019)" & prefix %in% jones_papers_flagged
    }
    ifelse(flag, "Contentious study", "Other studies")
}

flag_cols <- c("Contentious study" = "#D55E00", "Other studies" = "grey70")

# Build one orchard "trunk" (an orchaRd mod_results object) from a fitted model and
# relabel it with the version name, mirroring the multi-trunk approach in overall_effect.qmd.
mk_trunk <- function(fit, name) {
    r <- orchaRd::mod_results(fit$model, group = "identifierStudyId")
    class(r) <- "list"
    r$mod_table$name <- name
    r$data$moderator <- name
    class(r) <- c("orchard", "data.frame")
    r
}

# Place "Mean [95% CI]" above the existing "k = .. (..)" label on each trunk
# (identical helper to the overall-effect chapter).
add_ci_labels <- function(plot_obj, orchard_results) {
    est  <- sprintf("%.2f", orchard_results$mod_table$estimate)
    ci_l <- sprintf("%.2f", orchard_results$mod_table$lowerCL)
    ci_u <- sprintf("%.2f", orchard_results$mod_table$upperCL)
    orig_labels <- plot_obj$layers[[6]]$aes_params$label
    new_labels <- character(0)
    for (i in seq_along(orig_labels)) {
        nums <- regmatches(orig_labels[i], gregexpr("[0-9]+", orig_labels[i]))[[1]]
        k_text <- if (length(nums) >= 2) paste0("k = ", nums[1], " (", nums[2], ")") else orig_labels[i]
        new_labels[i] <- paste0(est[i], " [", ci_l[i], ", ", ci_u[i], "]\n", k_text)
    }
    plot_obj$layers[[6]]$aes_params$label <- new_labels
    plot_obj$layers[[6]]$geom_params$parse <- FALSE
    plot_obj$layers[[6]]$aes_params$hjust <- "right"
    plot_obj$layers[[6]]$aes_params$vjust <- 0.5
    plot_obj$layers[[6]]$aes_params$size <- 3.0
    plot_obj
}

# Assemble a three-trunk orchard plot (Main / Leave-out / Corrected) for one class, with
# the raw bubbles recoloured by whether they come from a contentious study.
sensitivity_orchard <- function(fit_main, fit_leave, fit_corr, class, xlab, title) {
    t_main <- mk_trunk(fit_main,  "Main")
    t_leav <- mk_trunk(fit_leave, "Leave-out")
    t_corr <- mk_trunk(fit_corr,  "Corrected")
    sub <- orchaRd::submerge(t_main, t_leav, t_corr)

    # Attach the contentious flag in submerge order (Main, then Leave-out, then Corrected).
    sub$data$es_flag <- c(
        contentious_flag(fit_main,  class),
        contentious_flag(fit_leave, class),
        contentious_flag(fit_corr,  class)
    )

    p <- orchaRd::orchard_plot(sub, xlab = xlab, flip = TRUE)
    # Recolour the raw-point layer (layer 1) by the contentious flag.
    p$layers[[1]]$mapping$colour <- ggplot2::aes(colour = es_flag)$colour
    p$layers[[1]]$mapping$fill   <- ggplot2::aes(fill = es_flag)$fill

    p <- p +
        ggplot2::scale_colour_manual(values = flag_cols, name = NULL) +
        ggplot2::scale_fill_manual(values = flag_cols, name = NULL) +
        ggplot2::ggtitle(title) +
        ggplot2::theme(
            legend.position = "bottom",
            plot.title = ggplot2::element_text(face = "bold", hjust = 0.5)
        )
    add_ci_labels(p, sub)
}
```

::: {.panel-tabset}

#### Hedges' $g$ (continuous outcomes)

```{r}
#| label: fig-sensitivity-hedges
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 6
#| fig-cap: "Combined Hedges' g under the Main, Leave-out, and Corrected datasets. Bubbles are raw effect sizes scaled by precision; the central marker is the pooled mean (95% CI), the whisker is the 95% prediction interval. The vertical line marks no effect (g = 0)."
sensitivity_orchard(fits$hedges_main, fits$hedges_leaveout, fits$hedges_corrected,
                    class = "hedges", xlab = "Hedges' g",
                    title = "Sensitivity of the combined Hedges' g")
```

#### Odds Ratio (binary outcomes)

```{r}
#| label: fig-sensitivity-or
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 6
#| fig-cap: "Combined log odds ratio under the Main, Leave-out, and Corrected datasets. Bubbles are raw lnOR effect sizes scaled by precision; the central marker is the pooled mean (95% CI), the whisker is the 95% prediction interval. The vertical line marks no effect (lnOR = 0, i.e. OR = 1)."
sensitivity_orchard(fits$or_main, fits$or_leaveout, fits$or_corrected,
                    class = "or", xlab = "Log Odds Ratio (lnOR)",
                    title = "Sensitivity of the combined log odds ratio")
```

:::

---

## Interpretation

```{r}
#| label: interp_values
#| echo: false
getv <- function(cl, ver, col) res[[col]][res$Class == cl & res$Version == ver]
hm <- getv("Hedges' g", "Main", "Estimate"); hl <- getv("Hedges' g", "Leave-out", "Estimate"); hc <- getv("Hedges' g", "Corrected", "Estimate")
om <- getv("Log OR", "Main", "Estimate");    ol <- getv("Log OR", "Leave-out", "Estimate");    oc <- getv("Log OR", "Corrected", "Estimate")
hm_p <- getv("Hedges' g", "Main", "p"); hl_p <- getv("Hedges' g", "Leave-out", "p"); hc_p <- getv("Hedges' g", "Corrected", "p")
om_p <- getv("Log OR", "Main", "p");    ol_p <- getv("Log OR", "Leave-out", "p");    oc_p <- getv("Log OR", "Corrected", "p")
sig <- function(p) if (is.na(p)) "indeterminate" else if (p < 0.05) "statistically significant" else "non-significant"
```

**Continuous (Hedges' *g*).** The combined mean is $`r fmt(hm, 2)`$ in the **Main**
analysis (`r sig(hm_p)`), $`r fmt(hl, 2)`$ when the contentious studies are **left out**
(`r sig(hl_p)`), and $`r fmt(hc, 2)`$ when they are **corrected** (`r sig(hc_p)`). Because
the validation showed the Davies errors mostly *inflate* the effect, the leave-out and
corrected means are expected to move toward — not away from — zero relative to Main;
the size of that shift is the practical measure of how much the headline continuous
estimate leaned on the flagged extractions.

**Binary (log odds ratio).** The combined mean is $`r fmt(om, 2)`$ in **Main**
(`r sig(om_p)`), $`r fmt(ol, 2)`$ **leaving out** the contentious papers (`r sig(ol_p)`),
and $`r fmt(oc, 2)`$ when **corrected** (`r sig(oc_p)`). On the binary side the
corrections act mainly on *variances* (removing the phantom-control inflation) and remove
two over-counted rows, so the point estimate is expected to move less than the continuous
one; the main effect there is on precision and weighting rather than on the location of
the mean.

**Reading the three together.** The **Leave-out** column is the conservative bound (trust
nothing flagged); the **Corrected** column is the best estimate (keep the studies, fix the
errors). If both fall close to **Main** and keep the same sign and significance, the
overall conclusion is robust to the extraction problems the validation uncovered. If
**Main** sits apart from the other two, the headline estimate depended materially on the
contentious effect sizes, and the **Corrected** value should be preferred as the
bias-reduced summary.

::: {.callout-warning}
## What this analysis does and does not show
This is a sensitivity analysis of the **combined overall mean** to the seven studies the
cross-dataset validation flagged. It does **not** re-audit the rest of the literature, and
it inherits every other modelling choice from the overall-effect chapter (the multilevel
structure, the within-study $\rho = 0.5$, and the lnOR$\leftrightarrow$*d* scale
conversion). The "Corrected" edits are exactly those documented in the
[validation chapter](../3_original_dataset_crosscheck/extraction_validation.qmd) and listed
in the edit map above; no other values were changed.
:::

---

## Session info

::: callout-note
## R session
```{r}
#| label: session_info
#| echo: false
sessionInfo()
```
:::
