---
title: "Cross-Dataset Extraction Validation"
subtitle: "Davies et al. (2020) vs Jones & DuVal (2019) — shared effect sizes"
author: "Eduardo S. A. Santos & Collaborators"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    theme: cosmo
    code-fold: true
    code-summary: "Show R code"
---

```{r}
#| label: setup
#| include: false
if (is.null(getOption("repos")) || getOption("repos")["CRAN"] %in% c(NA, "@CRAN@")) {
    options(repos = c(CRAN = "https://cloud.r-project.org"))
}
if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
pacman::p_load(tidyverse, DT, patchwork, scales, knitr)

# number formatter (mirrors the helper used elsewhere in this project)
fmt <- function(x, digits = 2) {
    res <- sprintf(paste0("%.", digits, "f"), x)
    zero <- paste0("-0.", paste0(rep("0", digits), collapse = ""))
    res[res == zero] <- paste0("0.", paste0(rep("0", digits), collapse = ""))
    res
}
```

## What this document validates

We are updating a mate-choice-copying meta-analysis built on **two existing
datasets that overlap in the primary studies they cover but report different
effect-size metrics**:

* **Davies et al. (2020)** — Hedges' *d*
* **Jones & DuVal (2019)** — log odds ratio (lnOR)

Both are *standardized effect sizes*: different numerical scales for the same
underlying quantity — how strongly an observer's mate preference shifts toward
the male it saw being chosen — and one scale can be converted into the other.

Before pooling them we need to know: *for the studies both teams extracted, do
the effect sizes agree once placed on a common scale?* This is a
**method-comparison** problem (two "instruments" measuring the same underlying
effects), so — unlike a single-team re-extraction audit — we can apply formal
agreement statistics (correlation, Bland–Altman limits of agreement) on top of
the per-study qualitative review.

The matching itself (study-level crosswalk + within-study optimal effect-size
assignment) is performed by
[`crosscheck_davies_jones.py`](crosscheck_davies_jones.py); this report reads
its output and assesses agreement. Key methodological points carried over:

* **Scale conversion.** Jones lnOR is mapped to Hedges' *d* via the
  logistic-to-normal approximation $d = \text{lnOR}\times\sqrt{3}/\pi$
  (≈ 0.5513; Borenstein et al. 2009).
* **Absolute magnitudes.** The two datasets use opposite sign conventions
  (Davies' non-directional *d* carries the raw mean-difference sign; Jones'
  lnOR sign encodes copying direction), so agreement is assessed on $|d|$.
  Sign concordance is reported separately as a diagnostic.

**How this document is organised.** The analysis runs in five steps: (1) *match
studies* — find the studies both datasets extracted and pair up their individual
effect sizes; (2) *measure agreement* — quantify how closely the paired values
agree once on a common scale; (3) *classify disagreements* — for pairs that do
not agree, label *why*; (4) *re-extract the worst cases* — return to the original
PDFs and recompute the effect size from the raw data; and (5) *conclude*.
Throughout, "Jones" and "Davies" name the two source **datasets**, not the
primary studies they drew from. A *pair* is a single Jones effect size lined up
with the one Davies effect size that most plausibly measures the same thing in
the same study (the pairing is done automatically, choosing the alignment that
minimises the magnitude difference within each study).

```{r}
#| label: load
# Resolve the outputs dir whether rendered standalone (cwd = this file's dir)
# or as part of the book (cwd = project root, via execute-dir: project).
out_dir <- "outputs/3_original_dataset_crosscheck"
if (!dir.exists(out_dir)) out_dir <- file.path("..", "..", out_dir)
xwalk  <- read_csv(file.path(out_dir, "crosswalk_effect_sizes.csv"), show_col_types = FALSE)
shared <- read_csv(file.path(out_dir, "shared_studies.csv"), show_col_types = FALSE)
unmatched <- read_csv(file.path(out_dir, "unmatched_jones_studies.csv"), show_col_types = FALSE)

# study tag + agreement / structural-cause classification
xwalk <- xwalk %>%
    mutate(
        study      = paste0(author_prefix, year),
        agreement  = if_else(magnitude_match, "agree", "disagree"),
        es_structure = if_else(n_jones_es == n_davies_es,
                               "same ES count", "ES count differs"),
        cause = case_when(
            magnitude_match                      ~ "agree",
            n_jones_es != n_davies_es            ~ "ES-count / extraction structure differs",
            TRUE                                 ~ "same structure, magnitude conflict"
        )
    )
```

## Headline summary

```{r}
#| label: headline
n_shared <- nrow(shared)
n_pairs  <- nrow(xwalk)
n_match  <- sum(xwalk$magnitude_match)
n_sign   <- sum(xwalk$sign_agree)
tibble(
    Metric = c("Shared studies (in both datasets)",
               "Jones studies NOT in Davies",
               "Comparable effect-size pairs",
               "Pairs matching in magnitude (|Δd| ≤ 0.15 or ≤ 15%)",
               "Pairs with concordant sign"),
    Value  = c(n_shared,
               nrow(unmatched),
               n_pairs,
               sprintf("%d (%s%%)", n_match, fmt(100*n_match/n_pairs, 0)),
               sprintf("%d (%s%%)", n_sign,  fmt(100*n_sign/n_pairs, 0)))
) %>% kable()
```

**How to read this.** The two datasets share `r n_shared` primary studies, with a
further `r nrow(unmatched)` studies appearing only in Jones. Within the shared
studies their effect sizes line up into `r n_pairs` comparable pairs, of which
`r fmt(100*n_match/n_pairs,0)`% agree in magnitude (tolerance: 0.15 *d*-units or
15%). Sign agreement looks low (`r fmt(100*n_sign/n_pairs,0)`%), but that is
*expected* — the two datasets encode copying direction with opposite signs, so we
compare absolute magnitudes here and revisit sign separately later. The sections
that follow ask how serious the magnitude disagreements are, and what causes them.

## Method-comparison agreement statistics

With the effect sizes paired, we ask how well the two datasets agree *as
measurements of the same thing* — the classic question of whether two instruments
give the same reading. Three complementary tools answer it:

* **Correlation** (Pearson *r*, Spearman *ρ*) — do the two rise and fall together?
  (1 = perfectly in step.)
* **Regression slope** of Davies on Jones — is one dataset systematically larger
  than the other? (a slope of 1 = no scaling difference.)
* **Bland–Altman** — for each pair, the *difference* between the two values plotted
  against their *average*; the mean difference is the typical bias, and the 95%
  "limits of agreement" bound how far apart the two datasets usually fall.

```{r}
#| label: agreement-stats
pear <- cor.test(xwalk$abs_jones_d, xwalk$abs_davies_d, method = "pearson")
spear <- suppressWarnings(
    cor.test(xwalk$abs_jones_d, xwalk$abs_davies_d, method = "spearman"))
fit  <- lm(abs_davies_d ~ abs_jones_d, data = xwalk)

# same-ES-count subset (like-for-like extractions)
sub <- filter(xwalk, es_structure == "same ES count")
pear_sub <- cor.test(sub$abs_jones_d, sub$abs_davies_d)

# Bland-Altman on |d|
xwalk <- xwalk %>%
    mutate(ba_mean = (abs_jones_d + abs_davies_d)/2,
           ba_diff = abs_jones_d - abs_davies_d)
bias <- mean(xwalk$ba_diff); s <- sd(xwalk$ba_diff)
loa_lo <- bias - 1.96*s; loa_hi <- bias + 1.96*s

tibble(
    Statistic = c("Pearson r (all pairs)",
                  "Spearman ρ (all pairs)",
                  "Pearson r (same-ES-count studies only)",
                  "OLS slope (Davies ~ Jones)",
                  "OLS intercept",
                  "Bland–Altman mean bias (Jones − Davies, |d|)",
                  "95% limits of agreement"),
    Value = c(sprintf("%s (95%% CI %s–%s, %s)", fmt(pear$estimate,3),
                      fmt(pear$conf.int[1],2), fmt(pear$conf.int[2],2),
                      ifelse(pear$p.value < 1e-4, "p < 0.0001",
                             paste0("p = ", fmt(pear$p.value,4)))),
              fmt(spear$estimate,3),
              sprintf("%s (n = %d pairs)", fmt(pear_sub$estimate,3), nrow(sub)),
              fmt(coef(fit)[2],3),
              fmt(coef(fit)[1],3),
              sprintf("%s d-units", fmt(bias,3)),
              sprintf("%s to %s", fmt(loa_lo,2), fmt(loa_hi,2)))
) %>% kable()
```

`r if (bias < -0.05 || coef(fit)[2] > 1.1) sprintf("A negative mean bias (%s) and an OLS slope > 1 (%s) both indicate Davies tends to report **larger** magnitudes than the Jones-converted values, increasingly so for larger effects.", fmt(bias,2), fmt(coef(fit)[2],2)) else if (bias > 0.05 || coef(fit)[2] < 0.9) "A positive mean bias indicates Jones-converted values tend to exceed Davies." else "Slope near 1 and bias near 0 indicate no systematic scaling difference."`

### Agreement plots

In the **left** panel, perfect agreement would place every point on the dashed
diagonal; points above it are studies where Davies reports the larger magnitude.
In the **right** (Bland–Altman) panel, points scattered evenly around zero would
mean no systematic difference — a shift away from zero is bias, and points beyond
the dashed limits are the outlying disagreements that the rest of this document
investigates.

```{r}
#| label: fig-agreement
#| fig-width: 11
#| fig-height: 5
#| fig-cap: "Left: Jones-converted vs Davies |d| with the line of identity (dashed). Right: Bland–Altman plot (mean bias solid, 95% limits of agreement dashed). Points are coloured by whether the pair matches in magnitude; studies where the two teams extracted different numbers of effect sizes are ringed."
lim <- c(0, max(xwalk$abs_jones_d, xwalk$abs_davies_d) * 1.05)

p_scatter <- ggplot(xwalk, aes(abs_jones_d, abs_davies_d)) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey50") +
    geom_point(aes(colour = agreement, shape = es_structure), size = 2.6, alpha = .85) +
    geom_text(aes(label = study), size = 2.4, vjust = -0.8, check_overlap = TRUE) +
    scale_colour_manual(values = c(agree = "#2c7fb8", disagree = "#d95f0e")) +
    coord_equal(xlim = lim, ylim = lim) +
    labs(x = "Jones & DuVal  |d|  (lnOR × √3/π)", y = "Davies  |Hedges d|",
         colour = NULL, shape = NULL, title = "Effect-size agreement") +
    theme_minimal(base_size = 12) + theme(legend.position = "bottom")

p_ba <- ggplot(xwalk, aes(ba_mean, ba_diff)) +
    geom_hline(yintercept = 0, colour = "grey80") +
    geom_hline(yintercept = bias, colour = "#2c7fb8") +
    geom_hline(yintercept = c(loa_lo, loa_hi), linetype = "dashed", colour = "#d95f0e") +
    geom_point(aes(colour = agreement), size = 2.6, alpha = .85) +
    geom_text(aes(label = study), size = 2.4, vjust = -0.8, check_overlap = TRUE) +
    scale_colour_manual(values = c(agree = "#2c7fb8", disagree = "#d95f0e"), guide = "none") +
    labs(x = "Mean of the two |d|", y = "Jones − Davies  (|d|)",
         title = "Bland–Altman") +
    theme_minimal(base_size = 12)

p_scatter + p_ba
```

## Discrepancy classification

Borrowing the per-study "required action" structure from Sánchez-Tójar et al.'s
validation codebook, every pair is classified by *why* it does or does not
agree. The dominant structural cause is whether the two teams extracted the
**same number of effect sizes** from a paper.

```{r}
#| label: cause-table
xwalk %>% count(cause, name = "pairs") %>%
    mutate(`%` = fmt(100*pairs/sum(pairs), 0)) %>% kable()
```

The three causes mean very different things. *Agree* needs no action.
*ES-count / extraction structure differs* means the two teams split a study into
different numbers of effect sizes, so a clean one-to-one comparison is impossible
— a bookkeeping difference, not necessarily an error. *Same structure, magnitude
conflict* is the category that matters: the teams extracted the same number of
effect sizes yet still disagree, which signals a genuine extraction difference.
In the per-pair table below, the `behaviour` columns are the key tell — when a
Jones "reversal/random" contrast is paired against a Davies "affiliation/mating"
one, the two teams measured a *different outcome* from the same paper.

```{r}
#| label: dt-pairs
xwalk %>%
    transmute(study, jones_code, jones_behaviour,
              davies_study, davies_behaviour, davies_n,
              `Jones |d|` = round(abs_jones_d, 2),
              `Davies |d|` = round(abs_davies_d, 2),
              `Δ|d|` = round(abs_diff, 2),
              sign_agree, cause) %>%
    datatable(rownames = FALSE, filter = "top",
              options = list(pageLength = 15, scrollX = TRUE),
              caption = "Per-pair crosswalk. Note the behaviour columns: a Jones 'reversal/random' contrast paired against a Davies 'Affiliation/Mating' contrast flags a different OUTCOME extracted from the same paper.")
```

### Studies flagged for manual re-extraction

The clearest validation failures are pairs that **disagree despite both teams
extracting the same number of effect sizes** (so the disagreement is not merely
structural), plus any same-paper pair whose outcome construct differs.

```{r}
#| label: flag-table
xwalk %>%
    filter(!magnitude_match) %>%
    arrange(desc(abs(abs_diff))) %>%
    transmute(study, jones_behaviour, davies_behaviour,
              `Jones |d|` = round(abs_jones_d, 2),
              `Davies |d|` = round(abs_davies_d, 2),
              `Δ|d|` = round(abs_diff, 2),
              n_jones_es, n_davies_es, cause) %>%
    kable()
```

## Sign-concordance diagnostic

Because the sign conventions differ by construction, low sign agreement is
*expected* and is not itself a validation failure — but a breakdown helps
confirm the conventions are simply inverted rather than randomly assigned.

```{r}
#| label: sign-table
xwalk %>%
    summarise(`pairs` = n(),
              `sign agree` = sum(sign_agree),
              `% agree` = fmt(100*mean(sign_agree), 0)) %>%
    kable()
```

## Effect-size-count reconciliation

How the two teams partitioned each shared paper into effect sizes (a major
driver of the comparable-pair count, since only `min(nJ, nD)` pairs per study
can be aligned).

```{r}
#| label: count-recon
shared %>%
    transmute(study = paste0(author_prefix, year),
              jones_species, davies_species,
              n_jones_es, n_davies_es,
              n_compared,
              `unpaired` = abs(n_jones_es - n_davies_es)) %>%
    arrange(desc(unpaired)) %>%
    datatable(rownames = FALSE, options = list(pageLength = 15, scrollX = TRUE))
```

```{r}
#| label: count-totals
shared %>% summarise(
    `Jones effect sizes (shared studies)` = sum(n_jones_es),
    `Davies effect sizes (shared studies)` = sum(n_davies_es),
    `Comparable pairs (Σ min)` = sum(n_compared)) %>%
    pivot_longer(everything(), names_to = "Metric", values_to = "Count") %>%
    kable()
```

## Jones studies absent from Davies

Verified against **all** Davies studies (matching species epithet anywhere,
author prefix at any year, and year ± 1); none have a Davies counterpart.

```{r}
#| label: unmatched-table
unmatched %>%
    transmute(study = paste0(author_prefix, year), species,
              n_effect_sizes, reason) %>%
    kable()
```

## Re-extraction: what happens when we return to the source PDFs

The agreement analysis flags *where* the two datasets disagree, but not *which*
is correct. To resolve that for the worst cases, we pulled the primary-study
PDFs for every study whose effect-size pair disagreed by **> 0.5 Hedges-*d*
units** and re-extracted each effect size from the raw data, recomputing it as a
2×2 odds ratio — tabulating how many observers chose the demonstrated male
versus the alternative, in the treatment and (where one exists) in the control,
and taking the odds of one over the other — under the meta-analysis's own
conventions (treatment-vs-control where the study used a control group; the
observed proportion against 50% chance otherwise). This gives a third,
independent estimate per effect size — a tie-breaker.

```{r}
#| label: load-reext
reext <- read_csv(file.path(out_dir, "reextraction_flagged_7studies.csv"),
                  show_col_types = FALSE) %>%
    mutate(
        study     = gsub("_", " ", sub("^([^_]+_[0-9]+).*", "\\1", identifierStudyId)),
        label     = paste0(study, " · ", identifierExperimentId),
        reext_d   = log(effectSizeCalculatedEstimate) * 0.5513,   # ln(OR) -> Hedges d
        jones_d   = compare_jones_d,
        davies_d  = compare_davies_hedges_d,
        reext_abs = abs(reext_d), jones_abs = abs(jones_d), davies_abs = abs(davies_d)
    )
mj <- mean(abs(reext$reext_abs - reext$jones_abs))
md <- mean(abs(reext$reext_abs - reext$davies_abs))
```

### The three estimates side by side

```{r}
#| label: reext-table
reext %>%
    transmute(Study = label,
              `2×2 (A/B/C/D)` = paste(effectSizeSampleSizeA, effectSizeSampleSizeB,
                                      effectSizeSampleSizeC, effectSizeSampleSizeD, sep = "/"),
              `Re-extracted d` = round(reext_d, 2),
              `Jones d` = round(jones_d, 2),
              `Davies d` = round(davies_d, 2),
              `|re-ext − Jones|` = round(abs(reext_abs - jones_abs), 2),
              `|re-ext − Davies|` = round(abs(reext_abs - davies_abs), 2)) %>%
    datatable(rownames = FALSE, options = list(pageLength = 12, dom = "t", scrollX = TRUE),
              caption = "Re-extracted effect size vs the two prior extractions (Hedges d). Cells A/B = treatment chose-target / not; C/D = control or 50/50 reference.")
```

Across the `r nrow(reext)` re-extracted effect sizes the **point estimates** sit
almost exactly on the **Jones** values (mean |Δ*d*| = `r fmt(mj,2)`) and far from
the **Davies** values (mean |Δ*d*| = `r fmt(md,2)`). Two cautions stop this from
being read as "the extraction is validated against Jones":

* **Point-estimate agreement is expected, not corroborating.** These outcomes are
  discrete choice counts read directly from each PDF; any careful extractor
  building the same 2×2 obtains the same lnOR. Agreement here confirms we *read the
  papers the same way* — it is not independent evidence that either prior
  extraction is correct.
* **The original crosswalk selected the construction that matched Jones.** For each
  study it chose the 2×2 that reproduced Jones's variance, which makes the variance
  match circular. The next section removes that circularity: it re-derives every
  effect size from a rule fixed by the study design **before** any comparison to
  Jones, and reports count agreement and design agreement separately.

### Visualising the tie-break

```{r}
#| label: fig-reext
#| fig-width: 11
#| fig-height: 5.5
#| fig-cap: "Left: each flagged effect size on the |Hedges d| axis — re-extracted (independent, from the PDF) vs Jones vs Davies; the re-extracted point sits on Jones in every case, with Davies displaced. Right: absolute deviation of each prior extraction from the re-extracted value."
ord  <- reext %>% arrange(reext_abs) %>% pull(label)
long <- reext %>%
    select(label, Jones = jones_abs, Davies = davies_abs, `Re-extracted` = reext_abs) %>%
    pivot_longer(-label, names_to = "source", values_to = "absd") %>%
    mutate(label = factor(label, levels = ord))

p_db <- ggplot(long, aes(absd, label)) +
    geom_line(aes(group = label), colour = "grey75") +
    geom_point(aes(colour = source, shape = source), size = 3, alpha = .9) +
    scale_colour_manual(values = c(`Re-extracted` = "#1a9850", Jones = "#2c7fb8", Davies = "#d95f0e")) +
    labs(x = "|Hedges d|", y = NULL, colour = NULL, shape = NULL,
         title = "Re-extracted vs prior extractions") +
    theme_minimal(base_size = 11) + theme(legend.position = "bottom")

dev <- reext %>%
    transmute(label = factor(label, levels = ord),
              `vs Jones`  = abs(reext_abs - jones_abs),
              `vs Davies` = abs(reext_abs - davies_abs)) %>%
    pivot_longer(-label, names_to = "comparison", values_to = "dev")
p_dev <- ggplot(dev, aes(dev, label, fill = comparison)) +
    geom_col(position = "dodge") +
    scale_fill_manual(values = c(`vs Jones` = "#2c7fb8", `vs Davies` = "#d95f0e")) +
    labs(x = "|re-extracted − prior|  (Hedges d)", y = NULL, fill = NULL,
         title = "Deviation from re-extraction") +
    theme_minimal(base_size = 11) +
    theme(legend.position = "bottom", axis.text.y = element_blank())
p_db + p_dev
```

### Count agreement vs design agreement (and a correction to Jones)

To break the circularity we re-derive each effect size from a rule fixed by the
study design alone — *before* looking at Jones — and report two things separately:
**count agreement** (does our lnOR from the raw counts match Jones, i.e. did we
read the paper the same way) and **design agreement** (does our variance and N
match, i.e. did we model the design the same way). The rule has two branches:

* **Real control group** — the 2×2 has distinct treatment and control cells (e.g.
  *Dugatkin 1992*'s naïve non-observer group). A standard 2×2 odds ratio is valid:
  $\mathrm{Var}=1/A+1/B+1/C+1/D$, $N=A+B+C+D$.
* **Single group vs 50 % chance** — the "50/50 rule" sets $C=D=n/2$. The
  design-faithful estimator is the one-sample logit of the observed proportion:
  $\mathrm{Var}=1/A+1/B$, $N=A+B$. The 50/50 group is a *phantom* control: it
  invents $n$ observations that were never collected and inflates the variance by
  **exactly $4/n$** while doubling the reported $N$.

```{r}
#| label: design-rederivation
reext2 <- reext %>%
    mutate(
        A = effectSizeSampleSizeA, B = effectSizeSampleSizeB,
        C = effectSizeSampleSizeC, D = effectSizeSampleSizeD,
        design = if_else(C == D, "single group vs 50% chance", "real control group"),
        corr_var = if_else(design == "real control group",
                           1/A + 1/B + 1/C + 1/D, 1/A + 1/B),
        corr_N   = if_else(design == "real control group", A + B + C + D, A + B),
        jones_var    = compare_jones_VlnOR,
        count_agree  = abs(reext_d - jones_d) < 0.02,   # point estimate (lnOR × √3/π)
        design_agree = abs(corr_var - jones_var) < 0.01
    )

reext2 %>% summarise(
    `Effect sizes` = n(),
    `Count agreement (lnOR matches Jones)`      = sum(count_agree),
    `Design agreement (variance matches Jones)` = sum(design_agree)
) %>% pivot_longer(everything(), names_to = "Quantity", values_to = "n") %>%
    kable()
```

Count agreement is complete; design agreement holds **only** for the real-control
studies, where a 2×2 is the correct model. The single-group studies disagree by
construction — and the disagreement is Jones's $4/n$ phantom-control inflation:

```{r}
#| label: phantom-control
reext2 %>%
    filter(design == "single group vs 50% chance") %>%
    transmute(Study = label,
              `our Var (1/A+1/B)` = round(corr_var, 4),
              `+ 4/n`             = round(4/(A + B), 4),
              `our Var + 4/n`     = round(corr_var + 4/(A + B), 4),
              `Jones Var`         = round(jones_var, 4),
              `our N` = corr_N,
              Note = if_else(abs(corr_var + 4/(A + B) - jones_var) < 0.01,
                             "Jones inflated by phantom 50/50 control",
                             "Jones used a different null")) %>%
    kable()
```

For every single-group study except *Howard 1998*, `our Var + 4/n` reproduces
Jones's variance to the fourth decimal, confirming the inflation is the invented
control group rather than a different reading. *Howard 1998* is the exception
(Jones used a different null), and our estimate is preferred there. So on the
vs-chance studies the design-faithful re-derivation **corrects** Jones — smaller
variance, halved $N$ — rather than reproducing him: the earlier "matches Jones
exactly" was, for these studies, matching a 50/50-rule artifact. The corrected
$N$ also reframes Davies's $N$ errors below: Davies's $N=40$ for *Dugatkin 1993*
is wrong against the design-faithful $N=20$, not merely against Jones.

### Why Davies diverges — four failure modes

The re-extraction shows the discrepancies are not random noise or conversion
error; they trace to four identifiable problems, all on the Davies side:

1. **Different outcome construct.** For *Applebaum & Cruz 2000* and *Briggs et al.
   1996*, Davies extracted a secondary **affiliation-time** measure (often
   non-significant) rather than the **preference-reversal frequency** the studies
   analyse; for *Howard et al. 1998* Davies extracted a **male-size preference** —
   a different question entirely.
2. **Computational artifacts.** *Dugatkin 1992*'s Davies *d* = 3.732 is exactly
   **2 + √3**, the signature of a degenerate-2×2 `escalc` input; *Gierszewski et
   al. 2018*'s *d* = 2.43 is inflated by the structural zero cell at first choice.
3. **Sign error.** *Howard et al. 1998*: Davies records **+1.14** for a study that
   found **no copying** (females retained their original preference 15:5); the
   re-extraction and Jones both give **−0.61**.
4. **Pool-vs-split, not an error.** *Fowler-Finn et al. 2015* is the one case where
   neither is wrong: Davies pooled one effect (*d* 0.63) where Jones — and our
   re-extraction — split it into a copying arm (*d* 1.21) and a no-copying arm
   (*d* 0.10), matching the study's significant phenotype × age interaction.

A design subtlety surfaced during re-extraction: for *Dugatkin 1992* the copying
tests are correctly compared against the study's **naïve control group**, not
against chance. Under that (correct) design Davies's total **N = 40** is right —
only the *magnitude* (the 2 + √3 artifact) is wrong — while Jones's third recorded
value for this study (lnOR 2.234) is spurious: it reproduces from no pair of the
paper's six experiments and should be dropped.

## What to report in the manuscript

* **Re-extraction (the > 0.5 set).** Returning to the source PDFs for the
  `r nrow(reext)` effect sizes that disagreed most, the design-blind re-derivation
  reproduced the **point estimates** wherever the data are discrete counts (mean
  |Δ*d*| vs Jones = `r fmt(mj,2)`; vs Davies = `r fmt(md,2)`), **corrected** Jones
  on the single-group / vs-chance studies (a 4/n phantom-control variance inflation
  from the 50/50 rule, with halved *N*), and isolated **Davies** as the divergent
  extraction — via different outcome constructs, `escalc` artifacts, a sign error,
  and one defensible pool-vs-split. This is an independent re-derivation, **not** a
  validation against Jones: the point-estimate agreement is expected for discrete
  counts and the variance was no longer reverse-engineered to match.
* **Overlap.** Of `r nrow(shared) + nrow(unmatched)` Jones & DuVal studies,
  **`r n_shared` are shared** with Davies; `r nrow(unmatched)` are unique to
  Jones (`r sum(unmatched$n_effect_sizes)` effect sizes).
* **Agreement.** Across `r n_pairs` comparable effect-size pairs, Jones-converted
  and Davies magnitudes correlate at Pearson *r* = `r fmt(pear$estimate,2)`
  (`r fmt(100*n_match/n_pairs,0)`% within tolerance), rising to
  *r* = `r fmt(pear_sub$estimate,2)` among studies where both teams extracted
  the same number of effect sizes.
* **Direction of disagreement.** Mean Bland–Altman bias =
  `r fmt(bias,2)` *d*-units, indicating Davies reports
  `r if (bias < 0) "slightly larger" else "slightly smaller"` magnitudes on
  average; most large discrepancies trace to **different outcomes extracted**
  (behaviour-type columns) or **different effect-size partitioning**, not to the
  metric conversion.
```
